### 2021-02-08 main.tex
iffTeX export

%%
%\listfiles
%\documentclass[%
% reprint,%
%secnumarabic,%
% amssymb, amsmath,%
% aip,cha,%
%groupedaddress,%
%frontmatterverbose,
%]{revtex4-1}
\documentclass[twocolumn,amsmath,amssymb,showpacs,pre,nofootinbib,superscriptaddress]{revtex4-1} %,preprint,jcp

\bibstyle{apsrev}
%\usepackage{docs}%
\usepackage{bm}%
\usepackage{graphicx}
%\usepackage{subcaption}
\usepackage{tikz}
\usepackage{color}
\usepackage{xcolor}
\usepackage{physics}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}%
%\nofiles
\expandafter\ifx\csname package@font\endcsname\relax\else
 \expandafter\expandafter
 \expandafter\usepackage
 \expandafter\expandafter
 \expandafter{\csname package@font\endcsname}%
\fi
\hyphenation{title}
\newcommand{\JH}[1]{\textcolor{blue}{ JH: #1}}

\begin{document}

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}

\title{Switchable substrates and thin film morphologies}

\author{S. Zitz}
\email{s.zitz@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,\\
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
  \affiliation{Department of Chemical and Biological Engineering, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\author{A. Scagliarini}%
\email{andrea.scagliarini@cnr.it}
 \affiliation{Institute for Applied Mathematics "M. Picone" (IAc, \\
Consiglio Nazionale delle Ricerche,\\
Via dei Taurini 19, 00185 Rome, Italy}%
\affiliation{INFN, sezione Roma ``Tor Vergata'', via della Ricerca Scientifica 1, 00133 Rome, Italy}
\author{J. Harting}
\email{j.harting@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,\\
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
 \affiliation{Department of Chemical and Biological Engineering and Department of Physics, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\date{\today}

\begin{abstract}
We study the dynamics of an unstable thin film deposited on a switchable substrate.
The substrate is itself subject to a time varying function that affects the equilibrium contact angle $\theta_{\text{eq}}(\mathbf{x},t)$.
We show that in the case of the vanishing time dependence the patterning drives the dewetting process and all fluid is drained into regions of a small $\theta_{\text{eq}}(\mathbf{x})$ and forms droplets.
Interestingly enough with a time varying equilibrium contact angle we are able to stop the droplet fragmentation and are able to create ligament or rivulet states.
We show that these rivulet states mainly depend on two quantities, first being the wavelength of the pattern and second being the velocity of the substrate pattern.

\end{abstract}

\maketitle

\newcommand{\ts}{\textsuperscript}
%\tableofcontents

\section{Introduction}\label{sec:intro}
Thin film flows and especially droplets are an phenomenon we experience in our every day life.
Rather unknown is the fact that our body uses thin films to cover mucous membranes.

\begin{equation}\label{eq:thin_film}
    \partial_t h(\mathbf{x},t) = \nabla\cdot\left(M(h)\nabla p(\mathbf{x},t)\right),
\end{equation}

\section{Method}\label{sec:method}
In this section we describe the lattice Boltzmann method (LBM) we use to simulate thin film flows.
The model we use has been introduced in ref~\cite{PhysRevE.100.033313} and builds upon a class of lattice Boltzmann models originally developed to study the shallow water equations~\cite{Salmon:1999:0022-2402:503, zhou2004lattice, PhysRevE.65.036309}.

\subsection{Thin film lattice Boltzmann method}\label{subsec:LBM}
The internal degrees of freedom of this method are the distributions functions $f_{\alpha}(\mathbf{x},t)$.
They describe a distribution of fictitious particles, for this paper we consider the case of $D2Q9$ lattice.
Such there are nine velocity components ($Q9$) for a simulation in two dimensions ($D2$).
The dynamics of the thin film and thus the evolution of the distribution functions is given by
\begin{equation}\label{eq:LBE}
    \begin{split}
        &f_{\alpha}(\mathbf{x}+\mathbf{c}^{({\alpha})}\Delta t,t+\Delta t) = \\
        &\left(1 - \frac{\Delta t}{\tau}\right) f_i(\mathbf{x},t) + \frac{\Delta t}{\tau} f_{\alpha}^{(eq)}(\mathbf{x},t) + \Delta t\mathcal{S}_{\alpha},
    \end{split}
\end{equation}
where the $\mathbf{c}^{({\alpha})}$ are the $Q$ discrete lattice velocities that can be computed according to
\begin{equation}\label{eq:speeds}
\mathbf{c}^{({\alpha})}  =
\left\{
\begin{array}{ll}
(0,0) & \alpha = 0 \\
\left[\cos{\frac{(\alpha-1)\pi}{4}}, \sin{\frac{(\alpha-1)\pi}{4}} \right] &  \alpha=1,3,5,7 \\
\sqrt{2}\left[\cos{\frac{(\alpha-1)\pi}{4}}, \sin{\frac{(\alpha-1)\pi}{4}} \right] & \alpha=2,4,6,8
\end{array}
\right.,
\end{equation}
and $w$ are being the corresponding weights
\begin{equation}
w_{\alpha}  =
\left\{
\begin{array}{ll}
\frac{4}{9} & \alpha = 0 \\
\frac{1}{9} &  \alpha=1,3,5,7 \\
\frac{1}{36} & \alpha=2,4,6,8
\end{array}
\right..
\end{equation}
The last term to the right $\mathcal{S}_{\alpha}$ is a (generalized) source term.
The concrete implementation of the forces depend on the source term as~\cite{https://doi.org/10.1002/fld.4726}
\begin{equation}\label{eq:source_term}
    \mathcal{S}_{\alpha} = \frac{3w_{\alpha}}{\mathbf{c}_{\alpha}^2}\mathbf{c}_{\alpha}\cdot\mathbf{F}.
\end{equation}
We are able to match thin film dynamics with the usage of a force term comprising of two main components,
\begin{equation}\label{eq:forcesum}
    \mathbf{F} = \mathbf{F}_{\mbox{\tiny{fric}}} + \mathbf{F}_{\mbox{\tiny{fric}}}.
\end{equation}
The first being a friction force that includes a velocity boundary condition at the fluid substrate interface
\begin{equation}\label{eq:forcefric}
\mathbf{F}_{\mbox{\tiny{fric}}} = -\frac{6h\mu\mathbf{u}}{2h^2+6\delta h + 3\delta^2},
\end{equation}
where $\mu$ is the viscosity and $\delta$ is a socalled slip length~\cite{RevModPhys.69.931, PhysRevE.100.033313, PhysRevE.63.011208, munch_lubrication_2005}.
The film thickness $h$ is the zeroth moment of the distribution function $f_{\alpha}$, i.e. $h(\mathbf{x},t) = \sum_{\alpha}f_{\alpha}(\mathbf{x},t)$ see Eq.~(\ref{eq:LBE}). 
The second component is due to the film pressure and accounts for the minimization of the fluid vapor interface as well as for the wettability of the substrate,
\begin{equation}\label{eq:forcefilm}
    \mathbf{F}_{\mbox{\tiny{film}}} = -\frac{\gamma h}{\rho_0}\nabla(\Delta h - \Pi(h)), 
\end{equation}
with $\gamma$ being the surface tension and $\rho_0$ being the fluids density, which is $\rho_0 = 1$ and will be dropped in the following. 
Here we model the interaction between solid substrate and film with a disjoining pressure term~\cite{RevModPhys.69.931, RevModPhys.81.1131, PhysRevE.63.011208},
\begin{equation}\label{eq:disjoinpressure}
    \Pi(h) = \kappa(\theta_{\text{eq.}})\left[\left(\frac{h_{\ast}}{h}\right)^n - \left(\frac{h_{\ast}}{h}\right)^m\right],
\end{equation}
with $n$ and $m$ being integer exponents, $h_{\ast}$ defines the minimum of the potential and  
\begin{equation}\label{eq:kappacontactangle}
    \kappa(\theta_{\text{eq.}}) = (1-\cos\theta_{\text{eq.}}) \frac{(n-1)(m-1)}{(n-m)h_{\ast}},
\end{equation}
which therefor allows us to tune the fluids contact angle.

\subsection{Linear stability and characteristic quantities}\label{subsec:charq}
For further insights concerning the early time dynamics of the unstable film we perform a linear stability analysis.
Thus we expand Eq.~(\ref{eq:thin_film}) in terms of a small perturbation $\delta h(\mathbf{x},t) = h(\mathbf{x},t) - h_0$ with $h_0$ being the initial mean thickness value at $t=0$.
Rearanging in terms up to first order in $\delta h$ yields
\begin{equation}\label{eq:linearstability_realspace}
    \partial_t \delta h = \frac{h_0^3}{3\mu}\nabla^2\Big(\Pi(\delta h) - \gamma \nabla^2 \delta h\Big).
\end{equation}
Under Fourier transformation of $\delta h$ 
\begin{equation}\label{eq:fouriertrans}
    \delta h(\mathbf{x},t) = \int(2\pi)^{-2}\delta \tilde{h}(\mathbf{q},t)e^{i\mathbf{q}\cdot\mathbf{x}}\Diff2\mathbf{q},
\end{equation}
and upon inserting this in Eq.~(\ref{eq:linearstability_realspace}) we get
\begin{equation}\label{eq:linearstab_fourier}
    \partial_t \delta \tilde{h}(\mathbf{q},t) = \omega(\mathbf{q})\delta\tilde{h}(\mathbf{q},t),
\end{equation}
with dispersion relation $\omega(q)$, 
\begin{equation}\label{eq:dispersion}
    \omega(\mathbf{q}) = \frac{1-(\frac{q^2}{q_0^2}-1)^2}{t_0}.
\end{equation}
Both, the maximum of the dispersion relation $q_0$ and therefore the characteristic time $t_0$, are specific to the substrate the film is deposited on and can be calculated using~\cite{PhysRevLett.99.114503}
\begin{equation}\label{eq:q0_and_t0}
    q_0^2 = \frac{1}{2}\frac{\partial \Pi(h)}{\partial h}\bigg\rvert_{h=h_0}, \qquad t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}.
\end{equation}

\subsection{Time and spatially varying contact angles $\theta(\mathbf{x},t)$}\label{subsec:theta}
To get a spatially varying contact angle $\theta(\mathbf{x},t)$ we make use of the underlying lattice and assign different values for $\theta$ to different grid points. 
From now on we drop the subscript \textit{eq.} for convenience, such $\theta = \theta_{\text{eq.}}$ unless otherwise stated.

We consider two different patterns and three wavelengths for this manuscript.
First being the sinusoidal pattern with wavelengths $\lambda \in \{L, L/2, L/3\}$ which is given as
\begin{equation}\label{eq:sinetheta}
    \theta_1(x,y) = \theta_0 + \delta\theta\left[\sin\left(\frac{2\pi x}{\lambda}\right)\sin\left(\frac{2\pi y}{\lambda}\right)\right], 
\end{equation}
with $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$.
The second pattern is a triangle wave which accounts for a more realistic szenario, thus $\theta_2$ is given as
\begin{equation}\label{eq:triangle_wave}
    \theta_2(x,y) = \theta_0 + \frac{2\delta\theta}{\pi}\left[\sin^{-1}\left(\sin\left(\frac{2\pi x}{\lambda}\right)\right)\sin^{-1}\left(\sin\left(\frac{2\pi y}{\lambda}\right)\right)\right].
\end{equation}

Both patterns create a contact angle distribution with values within the interval $[10^{\circ}, 30^{\circ}]$ and by definition they are time independent.
For a time dependent dynamic we simply shift the pattern in a periodic manner, such that
\begin{equation}\label{eq:theta_shift}
    \theta(\mathbf{x},t+T) = \theta(\mathbf{x}-\mathbf{1}, t), 
\end{equation}
with $T$ being an arbitrary number of lattice Boltzmann time steps.

Due to convenience we artibute a velocity to the pattern which is simply the shift distance divided by the stepping time
\begin{equation}\label{eq:pattern_speed}
    v_{\theta} = \frac{\sqrt{2}}{T}.
\end{equation}
Since this velocity as well as the choice of $T$ is arbitrary we define a characteristic velocity $v_0$ similar to Eqs.~(\ref{eq:q0_and_t0}) as
\begin{equation}\label{eq:normvel}
    v_0 = \frac{\lambda}{t_0(\theta)},
\end{equation}
where we use the wavelength $\lambda$ as a characteristic distance and $t_0(30^{\circ})$ as a characteristic time scale.
Important to note is that $t_0$ is sensitive to the contact angle, and decreases with increasing angle.
Thus choosing the highest contact angle for the characteristic time scale means $t_0$

\subsection{Numerical system}
As outlined in Sec.~\ref{subsec:LBM} we simulate the evolution of the thin film with a lattice Boltzmann method.
For all our simulations we use periodic boundary conditions and a square of size $Lx = Ly = L = 512 \Delta x$ with $\Delta x = \Delta t = 1$.
The surface tension as well as the viscosity are constants and set to $\gamma = 0.01$ and $\mu = 1/6$.
Having this value of $\mu$ leaves us with a relaxation time $\tau = 1$.
To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer and a slip length $\delta = 1$. 
With this choice of parameters we are almost able simulate a no-slip boundary condition.
As for our disjoining pressure potential, we use $n=3, m=2$ and $h_{\ast} = 0.07$.
Inserting this quantities into Eq.~(\ref{eq:q0_and_t0}) with $h_0 = 1$ yields $q_0^2(30^{\circ}) \approx 1.68\cdot10^{-2}$ as well as $t_0(30^{\circ}) \approx 177000\Delta t$ which gives us a first glimpse into relevant time scales.

The initial condition of all our simulations is a film undulated with sinusoidal wave as
\begin{equation}\label{eq:hinitial}
    h(\mathbf{x},0) = h_0 \left[1 + \epsilon \left(\sin\left(\frac{2\pi x}{L}\right)\sin\left(\frac{2\pi y}{L}\right)\right)\right],
\end{equation}
with $\epsilon = 0.1$ and $h_0 = 1$.

\section{Results}\label{sec:results}
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{Figures/theta_0336.png}
    \caption{Grayscale image of the height film at $t\approx 2t_0(30^{\circ})$}
    \label{fig:gray_ligament}
\end{figure}
We now present our results and show how a time dependent wettability or pattern velocity alters the morphology of the film during the dewetting process.
In the following we consider three time scales.
First being the initial short time behaviour that ends with the rupture of the film.
After the film ruptures we observe a transition region that strongly depend on the pattern velocity, see Fig.~\ref{fig:gray_ligament}.
Lastly we discuss the morphologies of the dewetted film in the long time regime.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{Figures/makie_try2.png}
    \caption{Height of the film $h(\mathbf{x},t)$ with the contact angle $\theta(\mathbf{x})$ as contour.}
    \label{fig:handtheta}
\end{figure}
We proceed as follows, first we simulate the dewetting on a given pattern either sinusoidal- or triangular with $\lambda \in \{L, L/2, L/3\}$ without a time dependency in $\theta$.
In Fig.~\ref{fig:handtheta} we show $h(\mathbf{x},t)$ as surface plot and $\theta_2(\mathbf{x})$ as contour plot for $v_{\theta} = 0, \lambda = L/2$ in late stages of dewetting.
Subsequent simulations allow for $v_{\theta} \ne 0$.  
Based on $v_{\theta}$ we can modify the morphology of the dewetting film, in both the intermediate transition regime as well as the late time state.

\subsection{Early time behaviour}\label{subsec:earlyT}
Starting with Eq.~(\ref{eq:hinitial}) we have by definition an unstable film that will rupture.
Taking into account the addition of the wettability gradient we know that fluid will accumulate in regions of high surface energies or small contact angles.
This is shown in Fig.~\ref{fig:handtheta}, where all droplets are in a region of small contact angles.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{Figures/All_deltah.pdf}
    \caption{Evolution of the absolute difference in height, x-axis, up to $t\approx t_0(30^{\circ})$, y-axis, for different wavelengths and patterns.
    The upper series displays the data for $\theta_1$, see Eq.~(\ref{eq:sinetheta}), and $\lambda = L, L/2, L/3$ in (a,b,c) respectively.
    Below is the series for $\theta_2$, see Eq.~(\ref{eq:triangle_wave}).
    Different colors and line styles represent different pattern velocities, all of which are normalized with $v_0$, see Eq.~(\ref{eq:normvel}).}
    \label{fig:evolutin_deltah}
\end{figure*}
Naively one assumes in the early stages of the dewetting process that the perturbations grow exponentially~\cite{RevModPhys.69.931, RevModPhys.81.1131}.
This is not the case in our simulations as the initial conditions for $h$ as well as $\theta$ are the same, see Eq.~(\ref{eq:hinitial}, \ref{eq:sinetheta}).
We show this by measuring the time resolved difference between the maximum and the minimum of the height field as, 
\begin{equation}\label{eq:delta_h_measure}
    \Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\} - \min_{\mathbf{x}}\{h(\mathbf{x},t)\},    
\end{equation}
which is shown in Fig.~\ref{fig:evolutin_deltah}.
In the upper row we show the data for $\theta_1$ while in the lower we show data for $\theta_2$, see Eqs.~(\ref{eq:sinetheta}-\ref{eq:triangle_wave}).
Different linestyles and colors represent different $v_{\theta}$'s.
Each column shows a wavelength $\lambda$, such from the left to the right we have $\lambda = L, L/2, L/3$.
Clearly one can observe the dip in \ref{fig:evolutin_deltah}(a) as the initial perturbation flattens while flowing into regions of lower contact angle.
This behaviour however is only pronounced for the time independent pattern and the smallest $v_{\theta}$ we use.
Thus the film is not spinodally dewetting but rather nucleating in regions of high contact angles.

Interestingly we observe a correlation between the rupture time ($\tau_r$) and the speed ($v_{\theta}$) as well as the wavelength ($\lambda$) of the pattern.
In agreement with previous studies on time dependent wettability we observe a stabilizing effect due to $v_{\theta}$~\cite{suman2006dynamics}.

\subsection{Intermediate times}\label{subsec:interT}
After the initial rupture of the film we observe that the dewetting is driven by $\theta(\mathbf{x},t)$.

\subsection{Long time behaviour}\label{subsec:lateT}

\section{Summary and Conclusions}\label{sec:sum_conclu}

\begin{acknowledgements}
The authors acknowledge financial support by the Deutsche Forschungsgemeinschaft (DFG) within the Cluster of Excellence ''Engineering of Advanced Materials'' (project EXC 315) (Bridge Funding). 
\end{acknowledgements}

%\appendix
%\section{}\label{app:one}

\bibliography{Ref}

\end{document}

### 2021-09-27 main_AS.tex
Version 2 of the text with bib file

%%
%\listfiles
%\documentclass[%
% reprint,%
%secnumarabic,%
% amssymb, amsmath,%
% aip,cha,%
%groupedaddress,%
%frontmatterverbose,
%]{revtex4-1}
\documentclass[twocolumn,amsmath,amssymb,showpacs,pre,nofootinbib,superscriptaddress]{revtex4-1} %,preprint,jcp

\bibstyle{apsrev}
%\usepackage{docs}%
\usepackage{bm}%
\usepackage{graphicx}
%\usepackage{subcaption}
\usepackage{tikz}
\usepackage{color}
\usepackage{xcolor}
\usepackage{physics}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}%
%\nofiles
\expandafter\ifx\csname package@font\endcsname\relax\else
 \expandafter\expandafter
 \expandafter\usepackage
 \expandafter\expandafter
 \expandafter{\csname package@font\endcsname}%
\fi
\hyphenation{title}
\newcommand{\JH}[1]{\textcolor{blue}{ JH: #1}}

\begin{document}

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}

%\title{Switchable substrates and thin film morphologies}
%\title{Time varying contact angles and thin film morphologies}
\title{Morphologies of thin liquid film dewetting on time-dependent patterned substrates}

\author{S. Zitz}
\email{s.zitz@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,\\
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
  \affiliation{Department of Chemical and Biological Engineering, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\author{A. Scagliarini}%
\email{andrea.scagliarini@cnr.it}
 \affiliation{Institute for Applied Mathematics "M. Picone" (IAC), \\
Consiglio Nazionale delle Ricerche (CNR),\\
Via dei Taurini 19, 00185 Rome, Italy}%
\affiliation{INFN, sezione Roma ``Tor Vergata'', via della Ricerca Scientifica 1, 00133 Rome, Italy}
\author{J. Harting}
\email{j.harting@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,\\
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
 \affiliation{Department of Chemical and Biological Engineering and Department of Physics, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\date{\today}

\begin{abstract}
We study the dynamics of an unstable thin film deposited on a ``switchable`` substrate.
The substrate is itself subject to a time varying function that affects the contact angle $\theta(\mathbf{x},t)$.
We show that in the case of vanishing time dependence the patterning drives the dewetting morphology and that all fluid is drained into regions of small $\theta(\mathbf{x})$ and thus form droplets.
Interestingly enough with a time varying component to the pattern we are able to stop the droplet fragmentation and create pronounced ligament or rivulet morphologies.
We show that both the numbers of semistable rivulets as well as number of stable droplets can be controlled with the pattern wavelength ($\lambda$).
Finally we explain why the ligaments appear with a simple balance between capillary retraction and dynamic wetting.   
\end{abstract}

\maketitle

\newcommand{\ts}{\textsuperscript}
%\tableofcontents

\section{Introduction}\label{sec:intro}
Wet surfaces and droplets are part of our every-day experience and of several industrial processes,
such as coating, tribology, painting and printing,
to name but a few~\cite{gross1980fluid,szeri2010fluid,DERYCK1998278,doi:10.1146/annurev.fluid.31.1.347,DASILVASOBRINHO19991204,singh2010inkjet,jo2009evaluation}. 
%Being it the droplets condensed in the higher atmosphere and return to the ground as rain or the oil film in a pan during cooking. 
%While this two examples are illustrative there are many industrial applications that build on thin film flows. 
%Among them is the wetting of surfaces to reduce wear or put simply, lubrication~\cite{ReynoldsLubr, gross1980fluid, szeri2010fluid}. 
%Coating surfaces to change their physio-chemical properties is another widely used application~\cite{DERYCK1998278, doi:10.1146/annurev.fluid.31.1.347, DASILVASOBRINHO19991204}.
%On the other hand the need to control droplets is at the heart of many printing processes~\cite{singh2010inkjet, jo2009evaluation}.
Moreover, the continuously growing technological interest for lab-on-a-chip devices~\cite{C6LC00387G,Focke} as well as for
printable electronics~\cite{Kim_2005, Luechinger_2008}, whose efficiency relies crucially on a precise control of material deposition upon dewetting of
liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e. it is not inert but responds
dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{doi:10.1126/science.288.5471.1624, doi:10.1246/bcsj.20180076, xin2010reversibly}.
%More recently lab-on-a-chip devices have acquired much attention, to work they require precise droplet or
%film flow on small scales~\cite{stone2004engineering, darhuber2010planar}.
%One possibility for precise control is the local wettability.
%Light switchable substrates~\cite{WANG200718}, therefore substrates that change their wettability based on light illumination, open up a new% way to program this devices.

Spatial variations of the contact angle are long known and used in previous studies~\cite{PhysRevE.91.023013, Sommers_2006}, time dependency of the contact angle on the other hand is not as well studied~\cite{suman2006dynamics}. 
Switchable substrates are in fact a realization of a time dependent contact angle. 
Using both spatial as well as time dependent variations of the local contact angle leads to interesting phenomena, for example continuously sliding droplets as shown by Grawitter and Stark~\cite{D0SM02082F}.  
Our approach builds on the same idea but our method, see ref~\cite{PhysRevE.100.033313}, allows us to go beyond a single droplet and to simulate the full dewetting transition starting with a thin film and ending with droplets.
We use this to show that a time dependent component of the contact angle alter the dewetting morphology quite substantially.
In the limit of a small time dependency or slow evolution of the contact angle field we qualitatively reproduce the results from Grawitter and Stark, such sliding droplets.
Speeding up the evolution of the contact angle field introduces a morphological transition, which to the understanding of the authors have not been looked at so far.
Therefore the purpose of this manuscript is to increase the understanding of the dewetting process.
Highlighting the fact that a temporal and spatial variation of the contact angle leads to novel morphological structures.
Thus our finding offer a simple way to understand how time periodic switchable substrates could generate novel dewetting morphologies.

The manuscript is outlined as follows:
In section~\ref{sec:method} we discuss the numerical method we use to generate our results and introduce relevant scales.
We then present our results in section~\ref{sec:results}, where we discuss the impact of the time dependency on the contact angle and show that both stability as well as dewetting morphology is affected.
A short summary of our finding and a closing conclusion with outlook is given in section~\ref{sec:sum_conclu}.

\section{Numerical model and simulation details}\label{sec:method}
In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{ReynoldsLubr,RevModPhys.69.931,RevModPhys.81.1131} 
\begin{equation}\label{eq:thinfilm}
    \partial_t h(\mathbf{x},t) = \nabla\cdot\left(M_{\delta}(h)\nabla p(\mathbf{x},t)\right),
\end{equation}
by means of a recently developed method, based on a lattice Boltzmann
scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801}.
Eq.~(\ref{eq:thinfilm}) describes, in a lubrication approximation spirit, the evolution of the height field (film thickness) $h(\mathbf{x},t)$, denoting the location of the liquid/air interface;
the mobility function $M_{\delta}(h)$ depends on the velocity boundary condition at the substrate, parameterized by an effective slip length $\delta$, and reads
\begin{equation}\label{eq:mobility}
  M_{\delta}(h) = \frac{2h^3 + 6\delta h^2 + 3\delta^2h}{6\mu},
\end{equation}  
(which, for $\delta \rightarrow 0$, reduces to the no-slip form $h^3/(3\mu)$),
where $\mu$ is the fluid dynamic viscosity.
The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$. The disjoining pressure
$\Pi$ can be seen as (minus) the derivative, with respect to the film thickness, of an effective interfacial potential; as such, it contains the information on the
liquid/solid and solid/gas interactions and, hence, on the wettability, which is parameterized in terms of the contact angle, $\theta$ \cite{RevModPhys.81.739, SCHWARTZ1998173}. 
The expression adopted for $\Pi$ is:
\begin{equation}\label{eq:disjoinpressure}
  \Pi(h,\theta) = \frac{2\gamma}{h_{\ast}}(1-\cos(\theta(\mathbf{x},t)))\left[\left(\frac{h_{\ast}}{h}\right)^3 - \left(\frac{h_{\ast}}{h}\right)^2\right],
\end{equation}
where $h_{\ast}$, the height at which the disjoining pressure vanishes, sets the precursor layer thickness.
The time variability of the patterned substrate, then, enters the model precisely through the
disjoining pressure, by making the contact angle space and time dependent, $\theta = \theta(\mathbf{x},t)$.
In particular, the following sinusoidal form is employed:
\begin{equation}\label{eq:sinetheta}
    \theta(\mathbf{x},t) = \theta_0 + \delta\theta\left[\sin\left(\frac{2\pi (x+v_{\theta x}t)}{\lambda}\right)\sin\left(\frac{2\pi (y+v_{\theta y}t)}{\lambda}\right)\right], 
\end{equation}
with $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$; the pattern evolves in time as a plane wave.
We choose to fix the velocity direction to one diagonal, i.e.
$\mathbf{v}_{\theta} = (v_{\theta x},v_{\theta y}) = v_{\theta}(1/\sqrt{2},1/\sqrt{2})$.
Notice that, since a typical velocity is such that $v_{\theta} \Delta t \ll \Delta x$ (in one time step the wave would travel
a distance much smaller than a lattice spacing), the time update must be interpreted in an integer part sense, that is
the pattern is shifted by one $\Delta x$ every $1/v_{\theta x}$ time steps (and equivalently in the $y$-direction).
%The second pattern is a triangle wave which accounts for a different kind of pattern gradient.
%To the best of our knowledge contact angles can not be tuned precise enough to match a sinusoidal wave on %the scales we are interested in.
%Therefore we assume that a triangle wave with a linear gradient is more likely accessable in experiments, %thus $\theta_2$ is given as
%\begin{equation}\label{eq:triangle_wave}
%\begin{split}
%    \theta_2(x,y) &= \theta_0 + \\ &\frac{2\delta\theta}{\pi}\left[\sin^{-1}\left(\sin\left(\frac{2\pi x}{%\lambda}\right)\right)\sin^{-1}\left(\sin\left(\frac{2\pi y}{\lambda}\right)\right)\right].
%    \end{split}
%\end{equation}
%Both patterns create a contact angle distribution with values within the interval $[10^{\circ}, 30^{\circ}%]$ and by definition they are time independent.
%To have a time dependency we simply shift the pattern in a periodic manner, such that
%\begin{equation}\label{eq:theta_shift}
%    \theta(\mathbf{x},t+T) = \theta(\mathbf{x}-\mathbf{1}, t), 
%\end{equation}
%with $T$ being an arbitrary integer number of lattice Boltzmann time steps.
Hereafter, lengths will be meant expressed in units of the
mean film height, $h_0$ (which is constant in time, due to mass conservation) and pattern wavelength $\lambda$, whereas the charactestic
time scale is set by:
\begin{equation}\label{eq:t0}
t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4},
\end{equation}
where $\gamma$ is the surface tension
(the numerical values, in lattice Boltzmann units, are set to $\gamma = 0.01$ and $\mu=0.167$).
The thus defined $t_0$ is the inverse growth rate of the most unstable mode (of wavenumber $q_0$)
of a spinodally dewetting film (on a homogeneous substrate) \cite{Mecke_2005,PhysRevE.100.023108};
in our heterogeneous case, we define the wavenumber $q_0$ as 
$q_0^2=\frac{1}{2\gamma}\frac{\partial \Pi(h,\theta)}{\partial h}(h=h_0,\theta=\theta_0+\delta \theta)$.
At this point, a reference velocity can also be constructed as
\begin{equation}\label{eq:v0}
v_0 \equiv \frac{\lambda}{t_0}.
\end{equation}
%Due to convenience we attribute a velocity to the pattern which is the shift distance divided by the stepp%ing time
%\begin{equation}\label{eq:pattern_speed}
%    v_{\theta} = \frac{\sqrt{2}}{T}.
%\end{equation}
%Since this velocity as well as the choice of $T$ is arbitrary we define a characteristic velocity $v_0$ si%milar to Eqs.~(\ref{eq:q0_and_t0}) as
%\begin{equation}\label{eq:normvel}
%    v_0 = \frac{\lambda}{t_0(\theta)},
%\end{equation}
%where we use the wavelength $\lambda$ as a characteristic distance and $t_0(30^{\circ})$ as a characterist%ic time scale.
%Important to note is that $t_0$ is sensitive to the contact angle, and decreases with increasing angle.
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{Figures/No_vel_render.png}
    \caption{Rendered image of the film on the stationary contact angle pattern at late times ($t > 20t_0$).
    The contact angle field is color coded, areas where the angle is small (high) are shown in blue (yellow).}
    \label{fig:handtheta}
\end{figure}
%As outlined in Sec.~\ref{subsec:LBM} we simulate the evolution of the thin film with a lattice Boltzmann m%ethod.
All our simulations are run on a bi-periodic square domain of size $L \times L$ with $L = 512$.
%Having this value of $\mu$ leaves us with a relaxation time $\tau = 1$.
To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer thickness
$h_{\ast}=0.07$ and and a slip length $\delta = 1$
(see Eqs.~(\ref{eq:mobility})-(\ref{eq:disjoinpressure})). 
% With this choice of parameters we are almost able simulate a no-slip boundary condition.
%As for our disjoining pressure potential, see Eq.~(\ref{eq:disjoinpressure}), we use the power pair
%$n=3, m=2$ and $h_{\ast} = 0.07$.
%Inserting this quantities into Eq.~(\ref{eq:q0_and_t0}) with $h_0 = 1$ yields $q_0^2(30^{\circ}) \approx 1%.68\cdot10^{-2}$ as well as $t_0(30^{\circ}) \approx 177000\Delta t$ which gives us a first glimpse into r%elevant length and time scales.
The liquid film is initiliazed with a height field slightly perturbed around the mean valued $h_0=1$, i.e.:
\begin{equation}\label{eq:hinitial}
    h(\mathbf{x},0) = h_0 \left[1 + 0.1 \left(\sin\left(\frac{2\pi x}{L}\right)\sin\left(\frac{2\pi y}{L}\right)\right)\right].
\end{equation}
Various wavelengths, in the range $\lambda \in [L/9, L]$, and velocities, $v_{\theta} \in [0.1, 10]v_0$,
are considered for the wettability pattern, Eq.~(\ref{eq:sinetheta}). 
%We proceed as follows, first we simulate the dewetting on a given pattern either sinusoidal- or triangular% with  without a time dependency in $\theta$.
In Fig.~\ref{fig:handtheta} we show $h(\mathbf{x},t)$ (droplets) and $\theta(\mathbf{x})$ (color coded) for $v_{\theta} = 0$ (i.e. the stationary case) and $\lambda = L/2$, in late stages of dewetting.
As expected, droplets form in regions of small contact angles (blue) while the regions of high contact angles (yellow) dewet.
%In subsequent simulations we iterate through , see Eq.~(\ref{eq:normvel}) .  
% Based on $v_{\theta}$ we can modify the morphology of the dewetting film, in both the intermediate transition regime as well as the late time state.

\section{Results}\label{sec:results}

\subsection{Rupture times}\label{subsec:rupture}

We want to study how the rupture times depend on the parameters characterising the wettability pattern, namely the contact angle
wavelength, $\lambda$, and wave speed, $v_{\theta}$.
The film rupture time, $\tau_r$, is defined as the least $t$ such that $\{\exists\mathbf{x}: h(\mathbf{x},\tau_r)=h_{\ast}\}$ (that is, when the
free surface touches the substrate).
%Taking another look at Fig.~\ref{fig:deltah_v0_sine} reveals that the three wavelengths rupture at different times $\tau_r$.
%The most important values for $\tau_r$ can be found in Tab.~\ref{tab:rup_times}.
In Fig.~\ref{fig:model_rt} we report the rupture times, as a function of the wavelength, for the stationary pattern ($v_{\theta}=0$) and
for $v_{\theta}=10 v_0$. We observe that, in both cases, $\tau_r$ grows $\lambda$.
%Thus smaller values of $\lambda$ lead to an earlier rupture time.
%Which can be explained with a mismatch of length scales.
This fact can be explained, qualitatively, as follows.
There is a length scale set by Eq.~(\ref{eq:disjoinpressure}) ($\lambda_0 = 2\pi/q_0$) and one that we introduce with the substrates wavelength $\lambda$.
\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{Figures/Rupture_times_over_lambda_with_powerlaw.pdf}
    \caption{Distribution of rupture times $\tau_r$ normalized by $t_0$ for different pattern wavelengths $\lambda$.
    The bullets, stars correspond to data from our simulations with $v_{\theta} = 0, 10v_0$.
    The full blue and dashed orange line are generated using $f(x) = a~x^{\alpha}$, with $a=0.6,1.5$ respectively.
    }
    \label{fig:model_rt}
\end{figure}
Taking this into account we can model the rupture times for the stationary pattern with
\begin{equation}\label{eq:model_tr}
    \frac{\tau_r}{t_0} = \frac{1}{2\pi}\frac{L}{\lambda_0 i^D}\approx \xi\frac{\lambda}{\lambda_0},    
\end{equation}
where $i$ is the wavenumber of the substrates pattern ($\lambda = L/i$) and $D$ is the dimension, thus $D=2$.

Interestingly, the behaviour of $\tau_r$ vs $\lambda$ remains qualitatively the same also for the
time-dependent case, $v_{\theta} = 0.1v_0$, reported in Fig.~\ref{fig:model_rt} as blue squares.
%as the difference of rupture times between these two simulations is negligible,
%see Tab.~\ref{tab:rup_times}.
Nevertheless, larger $v_{\theta}$ introduce a shift towards longer rupture times, i.e. fast wettability
waves tend to stabilise the film.
%Next we look into the affects of the pattern velocity on the structure formation after the first ruptures.

\subsection{Morphologies}\label{subsec:morpho}
We now look into the long time dynamics, focusing, in particular on the characterisation of the
dewetting morphologies and how they are affected by the speed of the wettability wave.

\subsubsection{Stationary pattern}\label{subsec:no_vtheta}
%Starting with Eq.~(\ref{eq:hinitial}) we have by definition an unstable film that will rupture.
%This can be easily seen by the fact that the initial perturbation has the smallest possible wavenumber, $q(t=0)=1/L$.
%Taking into account the addition of the wettability gradient we know that fluid will accumulate in regions of high surface energies or
%small contact angles and flow away from regions of high contact angle.
%An illustrative results of this effect is shown in Fig.~\ref{fig:handtheta}, where all droplets are present in a region of
%small contact angles.
\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{Figures/Clusters_novel_sine_picutres.png}
    \includegraphics[width=0.45\textwidth]{Figures/delta_h_v0_sin_with_dash.pdf}
    \caption{TOP PANEL. Number of droplets, $N$, as a function of time, generated during the dewetting process on the patterned substrate
      given by Eq.~(\ref{eq:sinetheta}), with $v_{\theta}=0$ and $\lambda=L$ (blue circles), $\lambda=L/2$ (orange squares) and
      $\lambda=L/3$ (green stars). The three horizontal dashed lines indicate the number of minima of (\ref{eq:sinetheta}),
      which is $2\left(\frac{L}{\lambda}\right)^2$. In the insets, we show shanpshots of the stationary droplet state as a grayscale image
      of the film thickness $h(\mathbf{x},t_{\infty})$.
      BOTTOM PANEL. Evolution of $\Delta h$ with time.
      Different symbols and colors represent different pattern wavelengths.
      Dashed lines indicate solutions to the cap height $h_d$ given the volume is split equally among the droplets,
      see Eq.~(\ref{eq:height_guess}).}
    \label{fig:clusters_v0_sine}
\end{figure}
%This is in fact a consistency check of our simulations.
On the stationary substrate, after rupture all the fluid accumulates in droplets centered at contact angle minima.
Consequently, as we see from Fig.~\ref{fig:clusters_v0_sine} where we plot the number of droplets, $N$, in time, 
in the steady state ($t \gg t_0$) such number attains the value $N_{\infty} = 2(L/\lambda)^2$
(which equals the minima of (\ref{eq:sinetheta}), for $v_{\theta}=0$).
%We expect that in case of $v_{\theta} = 0$ all fluid accumulates in droplets inside the contact angle minima.
%While there are only two minima for $\lambda = L$ there are eight and eighteen minima for $\lambda = L/2$ and $\lambda = L/3$ respectively.
%This can be easily checked with a cluster analysis of the film thickness $h(\mathbf{x},t)$ which
%we shown in Fig.~\ref{fig:clusters_v0_sine}.
%The height $h(\mathbf{x},t_{\infty})$ is displayed as grayscale images and is linked to stationary state.
%Theoretically we expect therefor two, eight and eighteen clusters (droplets) which we mark with a vertical
%dashed line in the respective color.
%These values are consistent with our simulations.
Notice that the number of clusters converges faster for smaller pattern wavelengths, in line with the observation, reported
and justified in the previous section, that the characteristic dewetting time decreases with the wavelength.\\
%To some extend this is expected as the distance between contact angle minima is shorter for smaller wavelengths.

Another quantity that is easily accessible and helps to check the consistency of our simulations is the height of the droplets $h_d$.
Given that all the fluid is collected inside the droplets we can make an assumption of the droplets height based on its volume $V_d$  and contact angle $\theta_d$ 
\begin{equation}\label{eq:height_guess}
    h_d(V_d, \theta_d) = \left(\frac{3V_d(1-\cos(\theta_d))}{\pi(2+\cos(\theta_d))}\right)^{\frac{1}{3}},
\end{equation}
where we implicitly assume that all our droplets have a spherical cap shape.
Since our method does not allow $h=0$ we measure the droplets height as~\cite{PhysRevE.100.033313}
\begin{equation}\label{eq:delta_h_measure}
    \Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\} - \min_{\mathbf{x}}\{h(\mathbf{x},t)\}.
\end{equation}
Counting the minima in $\theta(\mathbf{x})$ for the different wavelengths $\lambda$ allows to calculate the volume of a single droplet, see Fig.~\ref{fig:clusters_v0_sine}.
Since the film dewets regions of high contact angle values we can safely assume that $\theta_d < 20^{\circ}$.
%\begin{figure}
%    \centering
%    \includegraphics[width=0.48\textwidth]{Figures/delta_h_v0_sin_with_dash.pdf}
%    \caption{Evolution of $\Delta h$ with time.
%    Different symbols and colors represent different pattern wavelengths.
%    Dashed lines indicate solutions to the cap height $h_d$ given the volume is split equally among the droplets,
%    see Eq.~(\ref{eq:height_guess}).}
%    \label{fig:deltah_v0_sine}
%\end{figure}
Measuring the height of the droplets is done in the lower panel of Fig.~\ref{fig:clusters_v0_sine}, where $\Delta h$ is the maximal difference in film thickness which is equivalent to the droplet height $h_d$.
Focusing on the late time behaviour ($t > 10t_0$) we identify plateauing values. 
Since the pattern is a continuous functions, see Eq.~(\ref{eq:sinetheta}), we are left with some degree of freedom for the theoretical $\theta_d$ values. 
As a guide to the eye we introduce three dashed lines.
The dashed lines are calculated using Eq.~(\ref{eq:height_guess}) with $V_d = (V/N_{\infty})$ and $\theta_d = (14^{\circ}, 16^{\circ}, 17^{\circ})$ for $\lambda=(L,L/2,L/3)$ respectively.
These fit values for $\theta_d$ agree well with the stationary droplet state of our simulations.
Thus we are fairly confident that our droplet states and thus the simulations capturing the correct physical state.

%\subsection{Stability}\label{subsec:time_dep_theta}
%\begin{table}[h!]
%  \begin{center}
%    \begin{tabular}{c|c|c|c}
%      \textbf{Pattern} & \textbf{Wavelength} & \textbf{Velocity } & \textbf{Rupture time}\\
%      $\theta(\mathbf{x})$ & $\lambda$ & $v/v_0 $ & $\tau_r/t_0$ \\
%      \hline
%      $\theta_1$ & L & 0 & 1.69\\
%      $\theta_1$ & L & 0.1 & 1.72\\
%      $\theta_1$ & L & 1 & 2.34\\
%      $\theta_1$ & L & 10 & 2.25\\
%      $\theta_1$ & L/2 & 0 & 0.39\\
%      $\theta_1$ & L/2 & 0.1 & 0.39\\
%      $\theta_1$ & L/2 & 1 & 0.59\\
%      $\theta_1$ & L/2 & 10 & 0.68\\
%      $\theta_1$ & L/3 & 0 & 0.23\\
%      $\theta_1$ & L/3 & 0.1 & 0.23\\
%      $\theta_1$ & L/3 & 1 & 0.25\\
%      $\theta_1$ & L/3 & 10 & 0.45\\
%      $\theta_2$ & L & 0 & 1.01\\
%      $\theta_2$ & L & 0.1 & 1.18\\
%      $\theta_2$ & L & 1 & 1.32\\
%      $\theta_2$ & L & 10 & 1.38\\
%      $\theta_2$ & L/2 & 0 & 0.48\\
%      $\theta_2$ & L/2 & 0.1 & 0.48\\
%      $\theta_2$ & L/2 & 1 & 0.7\\
%      $\theta_2$ & L/2 & 10 & 0.76\\
%      $\theta_2$ & L/3 & 0 & 0.39\\
%      $\theta_2$ & L/3 & 0.1 & 0.42\\
%      $\theta_2$ & L/3 & 1 & 0.68\\
%      $\theta_2$ & L/3 & 10 & 0.76\\
%    \end{tabular}
%    \caption{Rupture time values for different pattern wavelengths $\lambda$ and different substrate velocities $v_{\theta}$.
%    Due to sampling we have a systematic error of 0.03 in the rupture times $\tau_r$.}
%    \label{tab:rup_times}
%  \end{center}
%\end{table}

\subsubsection{Time-dependent pattern}\label{subsec:timedep}
Dewetting morphology is a rather broad topic~\cite{becker2003complex, RevModPhys.81.739, PhysRevLett.99.114503, craster2009dynamics, doi:10.1146/annurev-fluid-011212-140734, alizadeh_pahlavan_cueto-felgueroso_hosoi_mckinley_juanes_2018}, however we know that a simple patterned substrate will ultimately yield droplets in region of small contact angle and is likely to rupture in regions of high contact angle, as shown in Fig.~\ref{fig:handtheta}.
This is indeed what we observe for the simulation without pattern velocity ($v_{\theta} = 0$).
\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{Figures/q_2_different_v.pdf}
    \caption{Structure analysis of the pattern $\theta_1,~\lambda=L/2$ using the Minkowski structure metrics (MSM). 
    The dipole metric $q_2$ is plotted for different pattern velocities $v_{\theta}$, displayed with different colors and symbols.
    In grayscale we supply the film thickness that is associated with the $q_2$ values.
    }
    \label{fig:msm_q2}
\end{figure}
Increasing the pattern velocity $v_{\theta} > 0$ does affect the dewetting morphology quite substantially.
Having $v_{\theta} = 0.1 v_0$ we are still observing the formation of droplets, similar to $v_{\theta} = 0$ but instead of being stationary these droplets, depending on their mass, are transported with the contact angle minima.
A somewhat similar behaviour was recently described by Grawitter and Stark~\cite{D0SM02082F} where they have used a boundary integral method to simulate a droplet on a moving wettability step.

If the pattern velocity is further increased $v_{\theta} > 0.1 v_0$ we observe a transition towards ligament or rivulet like structures.
These structures are aligned with the update direction of the contact angle field, see Eq.~(\ref{eq:sinetheta}). 
Thus if
\begin{equation*}
    \theta(x,y,t+T) = \theta(x-1, y, t),    
\end{equation*}
ligaments form in x direction.
These ligaments are sensitive to $\lambda$, or more precisely to the contact angle minima since they attract fluid.
To classify the morphological differences in respect to $v_{\theta}$ we make use of \textbf{papaya2}~\cite{Schaller2020}.

We use the film thickness as input and compute different morphological metrics.
Among the most interesting ones are the Minkowski structure metrics (MSM) $q_s$.
They can be calculated using~\cite{doi:10.1063/1.4774084}
\begin{equation}
    q_s(K) = \frac{|\psi_s(K)|}{\psi_0(K)}, 
\end{equation}
where $K$ is the body to we extract morphological information from and $\psi_s$ are the irreducible Minskovski tensors (IMT)
\begin{equation}
    \psi_s(K) = \int_{0}^{2\pi}e^{is\phi}\rho_K(\phi) \dd\phi = \sum_{k=-\infty}^{\infty} L_k e^{is\phi_k},
\end{equation}
with $L_k$ being the set of normal vectors of the body $K$.
Within the various $q_s$ we are particularly interested in the dipole one $q_2$. 
With $q_2$ we can clearly distinguish between a droplet and a ligament structure, as we expect a much larger $q_2$ value for the later one which is what we see in Fig.~\ref{fig:msm_q2}.
Our simulations with $\lambda = L/2$ show a clear difference between $v_{\theta} \leq 0.1v_0$ and $v_{\theta}>0.1v_0$  in $q_2$.
Interestingly the $q_2$ values for $v_{\theta} = 0.1v_0$ are always larger than those of $v_{\theta} = 0$.
Even the smallest pattern velocity introduces a measurable deformation of the spherical cap shape.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{Figures/stab_lig_lam2_v2.pdf}
    \caption{Measurement of the stability of the ligaments in terms of $q_2$.
    The failure time is defined as the time when the value of $q_2$ is below a threshold value.
    If no ligaments formed at all, the failure time is set to 0.
    }
    \label{fig:stab_ligs_lam2}
\end{figure}
Moving the contact angle field faster creates metastable ligament states.
Metastable because this ligaments will eventually break and separate into droplets.
This is indicated in Fig.~\ref{fig:msm_q2} by the dip in $q_2$ for $v_{\theta} = 1, 10 v_0$ (green and violet curve).
The breakup is actually due to the varicose mode of the ligament~\cite{doi:10.1063/1.3211248, PhysRevE.77.061605}.
Interestingly the wavelength of this varicose mode is $\approx\lambda$ and therefore we only observe $N_{\infty}/2$ droplets after breakup.
An analysis of the stability of these ligaments can be found in Fig.~\ref{fig:stab_ligs_lam2}.
Two things can be seen in this plot, first being the fact that ligaments only form for $v_{\theta} > 0.1$.
On the other hand a saturation happens for large $v_{\theta}$ values. 
Meaning we do not observe an increase in the lifetime of a ligament for $v_{\theta} > 10$.
% One possibility for that behaviour is that due to fast update of the pattern the film only experiences a singular effective contact angle value.
Besides the fact that only $N_{\infty}/2$ droplets are formed after ligament failure we observe on top a different dynamic than the ones on slow or no pattern velocity.   
Instead of a continuous motion with the pattern the remnants of the ligament are in an oscillating pumping state.
In this state the droplets are to slow to follow the pattern, which leads to cycle of deformation and retraction driven by the pattern (see sup. material).
This behaviour is clearly captured by Fig.~\ref{fig:msm_q2}, see e.g. the green curve for $t > 15t_0$.
Nevertheless the formation of ligaments in the first place can be explained with a few relative simple assumptions.

This problem is associated with two time scales, first being $t_0(\theta)$ the time scale of the dewetting thin film, see Eq.~(\ref{eq:t0}).
After the rupture we identify another time scale, the capillary time scale $\tau_{\text{Ca}}$ that drives the rims away from the holes~\cite{Edwardse1600183},
\begin{equation}\label{eq:cap_time}
    \tau_{\text{Ca}} = \frac{9\mu l_0}{\gamma \theta^3(\mathbf{x})},
\end{equation}
where  $l_0$ is a characteristic length scale of the problem, we use $\lambda/2$ as two rims can travel at most $\lambda/2$ before they merge.
Knowing this time scales allows us to introduce characteristic velocities.
One of these two is $v_0$ which is given by Eq.~(\ref{eq:v0}) and used to normalize $v_{\theta}$.
The other one is set by $\tau_{\text{Ca}}$ and reads
\begin{equation}\label{eq:cap_speed}
    U_{Ca} =\frac{l_0}{\tau_{\text{Ca}}} = \frac{\gamma}{9\mu} \theta^3(\mathbf{x}), 
\end{equation}
where $U_{\text{Ca}}$ is the capillary speed.
The argument is as follows, if $U_{\text{Ca}}$ is large as compared to $v_{\theta}$ than the film retraction is faster than the transport due to the contact angle field and thus droplets from.
Which explains the cases of $v_{\theta} = 0$ and  $v_{\theta} = 0.1v_0$.
On the other hand if $v_{\theta}$ is larger than $U_{\text{Ca}}$ than the retracting film has to little time to form droplets and ends up in the metastable ligament state.
Taking a look at our simulations with $v_{\theta} = 1v_0$ and $v_{\theta} =  10v_0$ we clearly observe ligament states.

In fact the ratio 
\begin{equation}\label{eq:vel_ratio}
    \Gamma = \frac{v_{\theta}}{U_{\text{Ca}}} = \frac{3\lambda \chi h_0^3 q_0^4}{\Theta^3}, 
\end{equation}
where $\Theta = max(\theta(\mathbf{x}))$ and $v_{\theta} = \chi v_0$, allows to classify the dewetting morphology on a time and space dependent substrate pattern. 
% as
% \begin{equation}\label{eq:classify}
%     \Gamma~\begin{cases}
%     < 1\qquad ... \text{droplets} \\
%     > 1\qquad ... \text{ligaments} 
%     \end{cases}.
% \end{equation}
% This explains in fact what we observe and holds true independent of the wettability gradient, i.e. $\theta_1, \theta_2$.

Therefore by varying either $\chi$ or $\lambda$ for a fixed $\Theta$ one can change the intermediate dewetting morphology according to Eq.~(\ref{eq:vel_ratio}).
Interestingly but not surprising is the fact that the fluid transport is larger for $\Gamma > 1$ than for $\Gamma < 1$.

\section{Summary and Conclusions}\label{sec:sum_conclu}
To summarize, we have shown numerical simulations of a dewetting thin film on a patterned substrate.
Generating baseline experiments with time independent patterns of different wavelengths, we than move to a time and spatial varying contact angle field $\theta(\mathbf{x},t)$.

In the next step we define a characteristic pattern velocity $v_0$ which simply is the patterns wavelength $\lambda$ over a characteristic time $t_0$, see Eqs.~(\ref{eq:t0}, \ref{eq:v0}).
Experiments with increasing pattern velocity $v_{\theta} = \chi v_0$ with $\chi\in\{0,10\}$ show a pronounced change in dewetting morphology from well defined droplets to ligaments.
With the single balance between pattern velocity $v_{\theta}$ and capillary velocity $U_{\text{Ca}}$ we can identify a parameter that separates the two morphological regimes.

An outlook to this work is to study the transition from ligaments to droplets in a more detailed manner.
Looking at the edges $\Gamma =1$ with more suitable tools like e.g. bifurcation diagrams will likely be subject to future work.
But also the update scheme of the contact angle field allows for many applications.
Among them is the avenue of switchable substrates, where e.g. switching of $\theta(\mathbf{x})$ due to a light source is in the scope of this method.

\begin{acknowledgements}
The authors acknowledge financial support by the Deutsche Forschungsgemeinschaft (DFG) within the priority program SPP2171 ``Dynamic Wetting of Flexible, Adaptive, and Switchable Substrates'', project HA-4382/11.
\end{acknowledgements}

%\appendix
%\section{}\label{app:one}

\bibliography{Ref}

\end{document}

### 2021-09-27 Notebooks/structure_analysis.jl
Some more plots

@@ -5,7 +5,7 @@ using Markdown
 using InteractiveUtils
 
 # ╔═╡ 3832de4e-a82b-11eb-08b3-0184778c8ef7
-using DataFrames, CSV, Plots, HTTP, DataFramesMeta, Query, JDF, LsqFit
+using DataFrames, CSV, Plots, HTTP, DataFramesMeta, Query, LsqFit
 
 # ╔═╡ 7488404c-a534-4096-a0ea-c927acf352b4
 # Plots default fonts and font sizes, could add colors as well
@@ -58,6 +58,9 @@ all_data = CSV.File(HTTP.get("https://jugit.fz-juelich.de/s.zitz/timedependent_w
 
 # ╔═╡ 32f54d38-2b4d-4da0-b9f5-1a1c92654922
 combined = CSV.File(HTTP.get("https://jugit.fz-juelich.de/s.zitz/timedependent_wettability/-/raw/master/Data_CSV/morph_data_with_l2_v-2-4-6-8.csv").body) |> DataFrame
+
+# ╔═╡ cf78fc68-3c53-4907-b7f5-a27807dfece6
+linear_vel = CSV.File(HTTP.get("https://jugit.fz-juelich.de/s.zitz/timedependent_wettability/-/raw/master/Data_CSV/Direction_x_lam_2.csv").body) |> DataFrame
 
 # ╔═╡ 5d8b33f1-8bef-4d48-b607-8e883588fdda
 function t0(;hᵦ=0.07, γ=0.01, μ=1/6, θ=1/6)
@@ -213,6 +216,67 @@ begin
 	     grid = :none,						# grid variable
 		 title="Sine λ=$(lh)",
 		 legend=:topright)					# legend position
+end
+
+# ╔═╡ 3854a45f-efd2-4ba6-8f18-6909587799f5
+begin 
+	t_he = zeros(992)
+	q2s_lin = zeros(992, 4)
+	for i in enumerate([0.1, 1.0, 10.0])
+		lin_vel_01 = @linq linear_vel |>
+			where(:threshold .== thresh) |>
+			where(:vel_norm .== i[2]) |>
+			select(:q2, :t_norm)
+		t_he .= lin_vel_01.t_norm
+		replace!(lin_vel_01.q2, NaN => 0.0)
+		q2s_lin[:, i[1]] = lin_vel_01.q2
+		
+	end
[…]

### 2021-09-28 Notebooks/structure_analysis.jl
Minor changes

@@ -241,11 +241,11 @@ begin
 		 st = :samplemarkers, 				# some recipy stuff
 		 step = 50, 						# density of markers
 		 marker = (8, :auto, 0.6),			# marker size
-		 legendfontsize = 18,			# legend font size
-         tickfont = (16, "Arial"),	# tick font and size
+		 legendfontsize = 14,			# legend font size
+         tickfont = (14, "Arial"),	# tick font and size
          guidefont = (18, "Arial"),	# label font and size
 	     grid = :none,						# grid variable
-		 title="Sine λ=$(lh)",
+		 # title="Sine λ=$(lh)",
 		 legend=:topright)
 	plot!(t_he,
 		  q2s_lin[:,1],
@@ -273,9 +273,9 @@ end
 # ╔═╡ 417b81f3-ff68-4f84-b02e-ca58a9c48d1c
 begin
 	if os
-		savefig(plot_lin, "../Figures/q2_lam2_x-update.pdf")		
+		savefig(plot_lin, "../Figures/q2_lam2_x-update.svg")		
 	else
-		savefig(plot_lin, "..\\Figures\\q2_lam2_x-update.pdf")
+		savefig(plot_lin, "..\\Figures\\q2_lam2_x-update.svg")
 	end
 end
 
@@ -484,25 +484,56 @@ end
 # ╔═╡ 7255922b-0361-4975-9435-9353d1ed553d
 begin
 	vnames = ["vₜ=0" "vₜ=0.1v₀" "vₜ=1v₀" "vₜ=2v₀" "vₜ=4v₀" "vₜ=6v₀" "vₜ=8v₀" "vₜ=10v₀"]
-	lig_delth = plot(lig_deltas[:,9], lig_deltas[:,1], label=vnames[1])
-	# plot!(lig_deltas[:,9], lig_deltas[:,2], label=vnames[2])
-	plot!(lig_deltas[:,9], lig_deltas[:,3], label=vnames[3])
-	plot!(lig_deltas[:,9], lig_deltas[:,4], label=vnames[4])
-	plot!(lig_deltas[:,9], lig_deltas[:,5], label=vnames[5])
-	plot!(lig_deltas[:,9], lig_deltas[:,6], label=vnames[6])
-	plot!(lig_deltas[:,9], lig_deltas[:,7], label=vnames[7])
-	plot!(lig_deltas[:,9], lig_deltas[:,8], label=vnames[8])
[…]

### 2022-10-03 Reply_RefereeA.tex
Update

\documentclass[12pt,english]{article}

\usepackage{natbib}

\usepackage{graphics,graphicx,dcolumn,bm,fleqn,epic,eepic,float}
\usepackage{amssymb,amsmath,multirow,rotate,rotating,color}
\usepackage[utf8]{inputenc}

\usepackage[english]{babel}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\tikzset{fontscale/.style = {font=\relsize{#1}}}
\usetikzlibrary{calc}
\makeatother

\newcommand{\figref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\eqnref}[1]{Eq.~(\ref{eq:#1})}
\newcommand{\ts}{\textsuperscript}

\definecolor{tuered}{RGB}{214,0,74}

\newcommand{\todo}[1]{\textbf{\textcolor{tuered}{ TODO: #1}}}

\newcommand{\vectornorm}[1]{\left|\left|#1\right|\right|}

\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\uvec}[1]{\hat{\vec{#1}}}
\newcommand{\tensor}[1]{\mathbf{#1}}

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}

\newcommand{\JH}[1]{\textcolor{blue}{JH: #1}}
\begin{document}

% The referee PDF has been generated from r239 -- 'make referee' command.

\section*{Reply to Referee A}
We thank the Referee for his/her report on our manuscript. 
We are glad to read that he/she finds it of potential interest and in principle 
suitable of publication, if a thorough revision is performed. 
We substantially improved and revised the paper for the resubmission and hope that we were able to clarify the obscure points, following all the Referee's remarks.
Hereafter we provide answers to the various questions/comments raised. All changes in the revised version are reported in red colour.

\begin{itemize}

\item[ \textbf{\underline{Comment 1.}}]
{ 
I am unable to follow most of the arguments made with respect to the disjoining pressure. 
For me, one needs to integrate the disjoining pressure to arrive at an effective interface potential, and 
I do not really know how to go from the effective interface potential to the surface tension. 
The equation (3) therefore merits much more discussion.
}

\item[ \textbf{{Answer}}]
{
We did not give, in fact, any particular argument with respect to the disjoining pressure, since 
we assumed that it is a very well established concept employed in any theoretical and computational study of thin liquid film hydrodynamics in the lubrication approximation framework.
We acknowledge, though, that our assumption was probably erroneous and we give some more explanation. 
First of all, there is no need (and no way) to go from the disjoining pressure (or the interface potential) to the surface tension, since the two quantities carry complementary physical information. 
Let us, first, make a step back and underline a fundamental difference between the physics of {\it one free interface} and that of a {\it film}:
the latter is a layer of liquid confined between {\it two} interfaces. When the two interfaces are brought close enough to each other (the layer is thin), they interact.
The disjoining pressure stems from this interaction. Thermodynamically, it represents minus the derivative of the free energy difference, per unit surface, between 
the film and the bulk of the same liquid, or, equivalently, the pressure difference inside and outside the film 
\cite{Deryaguin1940,DeryaguinChuraev1978}.
In the case of wetting, there can be three types of interfaces: liquid/air, liquid/solid (the substrate) and, 
if the substrate is only partially wetted, solid/air, thus giving rise to three-phase contact lines. 
The usually called {\it surface tension} is the energy cost of forming a liquid/air interface of unit area and its only knowledge is, therefore, not enough in our case. 
The Young's condition \cite{Young1805,DeGennes1985} connects the surface tension with the analogous energies per unit
area associated with the solid/liquid ($\gamma_{\text{sl}}$) and solid/gas ($\gamma_{\text{sg}}$) interface as
\begin{equation}
\cos \theta = \frac{\gamma_{\text{sg}}-\gamma_{\text{sl}}}{\gamma},
\end{equation}
where $\theta$ is the contact angle.
So, there is no need (and no way) to go from the disjoining pressure (or the interface potential) to the surface tension, since the two quantities carry complementary physical information,
and both are required to determine the dynamics of dewetting.
Of course, on the other hand, the disjoining pressure depends also on the surface tension, which appears, in fact, in its expression (3). 
We think, though, that part of the confusion might be due to the fact that we do not define the symbol of the surface tension at its first occurrence, but only later. 
We have now fixed this and we apologize for it.
\JH{Do we need to change more in the paper?}
}

\item[ \textbf{\underline{Comment 2.}}] 
{
The same holds for the time scale. 
The viscous time scale is well defined, but necessitates a length scale in addition to the 
capillary velocity. It is unclear why the authors choose the length scale that they use in the paper.
Subsequently, they replace the surface tension by a complicated partial derivative of the 
disjoining pressure; here the same remark as above holds.
}

\item[ \textbf{{Answer}}]
{
We apologize if we have not been thorough in the discussion of the relevant scales, which indeed deserves some clarification. 
Thin film hydrodynamics is an intrinsically multiscale phenomenon, whereby there are at least two characteristic length scales: the mean film thickness (let us call it $h_0$)
and an horizontal length scale $\ell$ (which in our case can be, for example, the pattern wavelength, $\lambda$), where, typically $h_0 \ll \ell$. We chose to use the mean 
film thickness as a scale, because it is the only relevant length that is common to all the simulations.
In the same spirit, we chose as a reference time $t_0(\theta)$, which is not exactly the capillary time (though being related to it). 
$t_0$, as we explain in the text, is, in fact, the characteristic time of growth of the most unstable mode in the spinodal dewetting on a homogeneous
substrate of contact angle $\theta$. In spite of the substrate being patterned in our case, it is possible to define an effective $t_0$ in terms 
of the highest contact angle on the pattern, which is what determines the fastest instability, i.e. $t_0(\Theta)$, where 
$\Theta=\mbox{max}_{\mathbf{x}}\{\theta(\mathbf{x})\}= \theta_0 + \delta \theta$. We deem the so defined characteristic time as an appropriate scale,
since it does not depend on the pattern wavelength and speed. Actually, one can see that it captures well the time scales of the dewetting process
from figure 2, where it is shown that the rupture times are of the order of fraction to units of $t_0$.
We recognize, on the other hand that in this perspective the choice of $v_0 = \lambda/t_0$ as a reference velocity looks a bit awkward.
A more natural candidate would be, indeed, $v_0=h_0/t_0$. We have therefore decided to use the latter in the revised version of the manuscript.
We have also rephrased the corresponding paragraph in the {\it Method} section as follows: \\

\textcolor{red}{Hereafter, lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film. On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads 
$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$ \cite{Mecke_2005,PhysRevE.100.023108}; in our patterned case, we define 
$q_0^2=h_{\ast}^{-2}(1-\cos\Theta)f^{\prime}(h_0/h_{\ast})$, where 
$\Theta = \max_{\mathbf{x}} \theta(\mathbf{x}) = \theta_0 + \delta \theta$.
A velocity scale also then naturally arises, namely $v_0 = h_0/t_0$.}

}

\item[ \textbf{\underline{Comment 3.}}]
{ 
Incidentally, for a 'normal' fluid such as water, the capillary velocity is $\sim 70\mbox{m}/\mbox{s}$, 
so that very, very rapidly the fluid inertia becomes important. 
This is indeed what one usually observes for the formation or coalescence of drops, so that at the very least this should be discussed and the Reynolds number calculated.
}

\item[ \textbf{{Answer}}]
{
Such a high value corresponds to the ratio of (surface tension)/(dynamic viscosity) for water, i.e. $v = \gamma/\eta \approx 70 \mbox{m}/\mbox{s}$.
This expression, though, is only representative of the capillary velocity, again, for {\it free surface flows}, as, e.g., in the coalescence of 
two droplets in the bulk of another fluid. 
In the case wetting/dewetting, 
the phenomenology is complicated by the liquid/substrate (and vapour/substrate) interaction. 
In such a situation the relevant velocity scale is set by the characteristic speed of contact line 
motion which can be estimated from a viscocapillary balance as 
$\sim \varepsilon^3 \gamma/\eta$, where $\varepsilon = h_0/L \ll 1$ is the film thickness parameter 
\cite{RevModPhys.69.931,doi:10.1146/annurev-fluid-011212-140734}. 
The typical velocities in dewetting problems are, then, much smaller than the ratio $\gamma/\eta$.
Moreover, in many actual applications, as well as in the majority of experiments, the 
working fluids are in general much more viscous than water (e.g., polymeric liquids), thus making
the characteristic velocity even smaller.
For the dewetting of a millimetric film of micrometric thickness reatracting into a single droplet, for instance, experiments have measured that 
the typical velocity is $\sim 1 \mbox{mm}/\mbox{s}$, which gives a Reynolds number $Re \approx 10^{-3}$ \cite{Edwards2016}. 
In our problem, there are two relevant characteristic velocities: the capillary velocity, $U_{\text{cap}}$, and 
the pattern speed, $v_{\theta}$. When their ratio $\Gamma = v_{\theta}/U_{\text{cap}} < 1$, the capillary 
velocity dominates and fixes the Reynolds number which can be, then, evaluated as 
$Re = U_{\text{cap}}h_0/\nu$ \cite{RevModPhys.69.931}, whereas for $\Gamma > 1$, $Re = v_{\theta} h_0/\nu$. 
In any case, the Reynolds number stays always below $\approx 0.2$.
We have added the following comment at the end of the {\it Method} section:\\

\textcolor{red}{The Reynolds number, $Re$, is of the order of $Re \sim 10^{-2}$ on the 
static substrate and never exceeds the value $\approx 0.2$ in the time dependent case.}

}

\item[ \textbf{\underline{Comment 4.}}]
{ 
In addition, there is a lot of old literature that relates the dewetting to the disjoining pressure, and shows that the nucleation of a hole in a wetting film is much more difficult than the nucleation of such a wetting film; there are experiments on this by Rutledge and Taborek, theory by Schick and Taborek and Bausch, Blossey and Indekeu.
}

\item[ \textbf{{Answer}}]
{
We are well aware of the literature quoted by the Referee, but these are works 
about {\it prewetting}, namely the phenomenon of {\it nucleation} (condensation)
of films/droplets on solid surfaces confining a critical liquid-vapour mixture
{\it as a function of temperature}. So, we are a bit puzzled about what such literature may have to do with the {\it dewetting} of a film on a chemically patterned substrate. Here, in fact, we are modelling a thin liquid film which is 
thermodaynamically stable against evaporation (i.e. one can consider that 
we are at a - fixed - temperature below the critical one).
}

\item[ \textbf{\underline{Comment 5.}}]
{ 
The explanations also lack a discussion on time, length, and energy scales. 
I have no clue what the values mean ``the numerical values, in LB units, are set to $\gamma= 0.01$ and $\mu = 1/6$''?. 
}

\item[ \textbf{{Answer}}]
{
We agree that giving the values of surface tension and dynamic viscosity 
in lattice Boltzmann units was not particularly insightful and could, on 
the contrary, be misleading, so we have removed the reference to them.
\JH{But shouldn't we list them somewhere to make sure the paper is reproducible?}
Concerning length and time scales, as we have discussed in Comment 2, 
they are made dimensionless 
with the mean film height and the characteristic time $t_0$. These two quantities 
have a clear physical meaning and can, therefore, be used to compare with 
experimental setups. In particular, the expression of $t_0$, through its dependence 
on $q_0$ (the wavenumber of the most unstable mode) contains the details 
of the disjoining pressure, such as the Hamaker constant $A_H$. For instance, let us consider the problem studied in \cite{becker2003complex}, the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon 
dioxide substrate. The Hamaker constant for this system is 
$A_H  = 2.2 \times 10^{-20} \, \text{J}$ and the surface tension is 
$\gamma = 0.03 \, \text{N}/\text{m}$, such that, 
from equation (3) of \cite{becker2003complex} and 
from the definition of $q_0^2 = \frac{1}{2\gamma} \Pi^{\prime}(h_0)$, 
we get (neglecting the short range part of the interface potential)
$$
q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}.
$$
The characteristic time of the growth of the instability, then, reads \cite{PhysRevLett.99.114503} (the dynamic viscosity of polystyrene is 
$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$)
$$
t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}
$$
and it can be seen that indeed film rupture occurs within time 
scales of the order of $t_0$ ($\approx 2 t_0$) \cite{becker2003complex}.
We have reported these comments on comparison with actual experimental 
values in the new version of the paper.
}

\item[ \textbf{\underline{Comment 6.}}]
{ 
What is the disjoining pressure due to? 
Van der Waals forces? 
How thick is the precursor film, and why should the disjoining pressure be zero there?
Is there any comparison with experiments possible/feasible?
}

\item[ \textbf{{Answer}}]
{
As we have already discussed in the first answer (and as one can find 
in several reviews and undergraduate textbooks, but see, e.g., Craster 
and Matar \cite{craster2009dynamics}), the disjoining pressure encode 
any kind of interactions among the fluid molecules and the substrate, so 
also those of Van der Waals type. The precursor film is an equilibrium value for the thickness that emerges when the interface potential has also a short 
range repulsive term. The disjoining pressure is zero there by definition of equilibrium (or, equivalently, because that is a minimum of the interface potential). Concerning comparisons of a thin-film equation approach with 
experiments, there are tons of examples in the literature, such as 
\cite{becker2003complex,fetzer2005new,PhysRevLett.99.114503} to name but a few (see also \cite{craster2009dynamics,RevModPhys.81.739} for reviews). 
}

\end{itemize}

\bibliographystyle{abbrv}
\bibliography{Ref}

\end{document}

### 2022-10-03 Reply_RefereeB.tex
Update

\documentclass[12pt,english]{article}

\usepackage{natbib}

\usepackage{graphics,graphicx,dcolumn,bm,fleqn,epic,eepic,float}
\usepackage{amssymb,amsmath,multirow,rotate,rotating,color}
\usepackage[utf8]{inputenc}

\usepackage[english]{babel}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\tikzset{fontscale/.style = {font=\relsize{#1}}}
\usetikzlibrary{calc}
\makeatother

\newcommand{\figref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\eqnref}[1]{Eq.~(\ref{eq:#1})}
\newcommand{\ts}{\textsuperscript}

\definecolor{tuered}{RGB}{214,0,74}

\newcommand{\todo}[1]{\textbf{\textcolor{tuered}{ TODO: #1}}}

\newcommand{\vectornorm}[1]{\left|\left|#1\right|\right|}

\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\uvec}[1]{\hat{\vec{#1}}}
\newcommand{\tensor}[1]{\mathbf{#1}}

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}

\newcommand{\JH}[1]{\textcolor{blue}{JH: #1}}
\begin{document}

% The referee PDF has been generated from r239 -- 'make referee' command.

\section*{Reply to Referee B}

We thank the Referee for his/her thorough reading of our manuscript and detailed report.
We are happy to read that the referee acknowledges the amount of work we did and thinks it could be fitting for PRL after revision.
In what follows we provide detailed answers to all Referee's questions and we discuss all the changes made to the manuscript
(which are typeset in red colour).

\begin{itemize}

\item[ \textbf{\underline{Comment 1.}}]
{
The model disjoining pressure, Eq. 3 with a tail decaying as the
inverse squared power seems odd. I am more familiar with the expected
van der Waals tail with a decay as the inverse cube power. Is there
any reason for this choice?

\item[ \textbf{Answer}]
{
The Referee is right in pointing out that, indeed, the $h^{-3}$ decay 
for the attractive part of the disjoining pressure is more straightforward 
to understand as it can be derived directly from an intermolecular potential of Lennard-Jones type. Nevertheless, it is also true that other forms
can be encountered in the literature, depending on the actual interactions
\cite{PhysRevE.63.011208}. Here, though, our choice of the $(3,2)$ pair of exponents is motivated more by computational reasons rather than by the will of 
simulating a specific liquid with given physical properties.
With such choice, in fact, the characteristic dewetting time 
is shorter, thus making the overall computational cost more manageable.
In fact, we have seen in previous study~\cite{PhysRevE.104.034801} that
except for this shift of the time scales, the dewetting dynamics remains 
qualitatively the same. To confirm this observation, we have run 
a new simulation (for a single pair of pattern wavelength and speed) 
with the $(9,3)$ exponents and show the results in figure \ref{fig:rivulets_9_3}.
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{Figures/No_vel_render2.png}
    \caption{}
    \label{fig:rivulets9_3}
\end{figure}
We have also added a footnote in the paper commenting on this aspect.
%We thank the referee for pointing this out.
%It is true that there are variations concerning the power law model of %the disjoining pressure.
%A common one to which the referee refers to is $(n,m) = (9,3)$, the %choice we made for the manuscript is $(n,m) = (3,2)$.
%Both exponent pairs can be found in the literature with even more %variety~\cite{moulton_lega_2013,SCHWARTZ1998173,doi:10.1063/1.4828721,Br%asjen}.
%The pair $(9,3)$ is the one that can be motivated from functional %derivatives of a Lennard-Jones potential, understandably this could more %appealing to use.
%We have used both exponent pairs in previous works, see %e.g.~\cite{PhysRevE.104.034801}.
%In this manuscript we use $(3,2)$ to keep $t_0$ reasonably small and %thus the overall computational cost manageable.
%Using the pair $(9,3)$ does not alter the results, it effectively shifts %the value of $t_0$ by $\approx 10$.
}

\item[ \textbf{\underline{Comment 2.}}]
The second major concern refers to the choice of slip length. 
The authors study a fluid-substrate model that has a thin film equilibrium thickness of 
$0.07$ length units, but the slip length is one unit, hence about $10$ times larger than the equilibrium film thickness. 
This seems unexpected, in view of the small contact angle chosen for the system. 
Could the authors justify this choice?
}

\item[ \textbf{Answer}]
{
As the Referee correctly remarks, indeed small contact angles are typically associated to small slip lengths. From a phenomenological point of view, 
though, having $\delta \sim h_0$ is still commonly considered in the weak/intermediate slip regime, as proven, for instance, 
by the formation of satellite 
droplets following rivulet breakup~\cite{doi:10.1073/pnas.1820487116} or 
in the rim shape in dewetting by hole nucleation~\cite{fetzer2007quantifying, munch2005lubrication} (see also Fig.~\ref{fig:slip}).
\begin{figure}
    \centering
    \includegraphics [width=0.6\textwidth]{slip_measure.pdf}
    \caption{Dewetting rim profiles for various slip parameters $\delta$; notice the presence of a dimple behind the front, typical of 
    small to vanishing slip, which is preserved for slip lengths up to 
    $\delta \sim 10 h_0$.}
    \label{fig:slip}
\end{figure}
Our choice of this slip length value was dictated essentially by the need 
to keep numerically stability even in the strongly time dependent case 
(large pattern speed), yet remaining in a small slip regime. 
Nevertheless, motivated by the Referee's concern and in order to check that 
picture remains qualitatively the same, we have run a simulation with 
a slip length five times smaller. The results from the analysis of the Minkowski metric $q_2$, for two cases with low and high pattern speed, reveal that indeed
the droplet-to-rivulet transition is confirmed, with $\Gamma$ as the 
discriminating parameter (see figure \ref{fig:q2smallslip}).
\begin{figure}
    \centering
    %\includegraphics [width=0.6\textwidth]{slip_measure.pdf}
    \caption{}
    \label{fig:q2smallslip}
\end{figure}
We have commented on this when we give the value of the slip length, just before 
equation (5) of the revised manuscript.
%We thank the referee for bringing this up.
%Slip and disjoining pressure are an essential part of our model -- as for many %other thin film solvers.
%Here, they allow us to have moving contact lines with imposed wettability %features.
%We agree that the value may seem large at first glance.
%However, as Peschka et al. have shown, the generation of satellite droplets %during rivulet breakup, similar to what we observe, is a no-slip %feature~\cite{doi:10.1073/pnas.1820487116}.
%We observe the occurrence of satellite droplets in the initial dewetting %period.
%Increasing the slip length makes those droplets disappear.
%Another point to mention here is the dependency of the dewetting rim on the %slip.
%Small to vanishing slip induces a dimple behind the front, while for large slip %lengths this dimple vanishes~\cite{fetzer2007quantifying, %munch2005lubrication}.
%We checked this behaviour in Fig.~\ref{fig:slip} and found only a small %dependency for slip lengths $\delta < 10$.
%Spots with a thickness $h \leq h_{\ast}$ are considered dry spots, therefore %the term equilibrium thickness maybe exaturated.
}

\item[ \textbf{\underline{Comment 3.}}]
{
The relevance of the manuscript relies on the significance of the substrate's switchable properties. 
Particularly, it is assumed that the wetting properties can be modulated on space but also on time. 
The authors have mentioned in the introduction that this can be potentially done. 
However, it would be desirable that the authors relate their model of switchable substrate with some specific experimental realization, where the actual length and time scales are mentioned.
}

\item[ \textbf{{Answer}}]
{
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{simple_exp.pdf}
    \caption{Schematic realization of the experiment described in the manuscript.}
    \label{fig:experiment_illustration}
\end{figure}
We thank the Referee for bringing this up.
First, we like to note that experiments are not our field of expertise and therefore our envisioned realization could be flawed.

The idea we have would be to prepare a thin film ($h_0 \approx 4nm$ PS) on a light responsive substrate.
Dimensions of the substrate should at least be $2$ microns in $x$ and $y$, but larger is fine.
Arrange several light sources (diodes) in a lattice with a distance of $\lambda \approx 1\mu m$ between them, as displayed in Fig.~\ref{fig:experiment_illustration}.
It would be perfect if the structure on which the light sources are mounted can be moved in a periodic manner, e.g. wrapped around rolls.
On the other hand if the array of diodes is large enough, say $100\lambda$ it should be sufficient to move them continuously in one direction. 
\JH{Could one do this with a DMD as here https://doi.org/10.1021/jp301092y ? Then, one could use a transparent substrate and just do it from the bottom. I think a DMD + light responsive substrate could be a straightforward implementation of a time dependent wettability pattern.}

In Eq.~(5) we introduce two scales, $t_0$ and $v_0$.
From experiments of dewetting thin films we know that $t_0$ correlates with the rupture time~\cite{becker2003complex, PhysRevLett.99.114503}.
Considering the same experimental parameters as above, 
\begin{align*}
    t_0 &\approx 1800s,\qquad~~~\qquad \frac{2\pi}{q_0}\approx 400nm, \\
    \quad v_0 &= \frac{\lambda}{t_0} \approx \frac{1}{900}\frac{\mu m}{s},\quad \quad v_0^{\ast} = \frac{2\pi}{q_0 t_0} \approx \frac{4}{9}\frac{nm}{s},
\end{align*}
where we take into account that the rupture time is overestimated by about a factor of two.
The experiment therefore reduces to the movement of the diodes with a velocity between $[10^{-7}/900, 10^{-5}/900] m/s$. 
\JH{We need to comment on the experimental realization in the paper. Can we summarize this idea to 1-2 sentences?}
}

\item[ \textbf{\underline{Comment 4.}}]
{
An important parameter in the problem is the wave-vector $q_0$, which sets the correlation length in the problem. 
The authors should state what the actual value is relative to either the average film height, or the wave-length of modulations. 
Is there any reason to give the wave-length of modulations in units of the system size, i.e., are the results subject to finite size effects? Otherwise it appears more meaningful to describe the modulations in units of $q_0$ or the equilibrium film thickness.
}

\item[ \textbf{{Answer}}]
{
We thank the referee for this comment.
It is true that we use $q_0$ but never display its actual value.
Inserting numbers, we have 
\begin{equation*}
    q_0 \approx 0.13, \quad \lambda_0 = \frac{2\pi}{q_0} \approx 48.5 \approx \frac{L}{10},  
\end{equation*}
therefore $\lambda_0$ is usually smaller than $\lambda$.

No, these simulations do not suffer from finite size effects.
The honest reason for using multiples of the system size is the ease of the numerical realization.
Having periodic boundary conditions, a non-multiple wavelength of the system size would induce a discontinuity.
Given the complexity of the problem, we would like to have as much control of the pattern as possible.

The arbitrary choice we made had however helped us to understand where the boundary between rivulet and droplets appears.
In Eq.~(11) we use
\begin{equation*}
    \Gamma = \frac{v_{\theta}}{U_{\theta}} = \frac{\chi v_0}{U_{\theta}},
\end{equation*}
where rivulets appear for $\chi > 1$.
Substituting the wavelength dependency of $v_0$ with $q_0$,
\begin{equation*}
    v_0^{\ast} = \frac{\lambda_0}{t_0} 
\end{equation*}
and computing $\Gamma$ in terms of $v_0^{\ast}$ yields
\begin{equation*}
    \Gamma = \frac{v_0^{\ast}}{U_{\theta}} = \frac{6\pi h_0^3 q_0^3}{\Theta^3}.
\end{equation*}
Relating both with each other yields
\begin{equation*}
    \frac{\chi 3\lambda h_0^3 q_0^4}{\Theta^3} = \frac{6\pi h_0^3 q_0^3}{\Theta^3},
\end{equation*}
and with most of the terms canceling one obtains
\begin{equation*}
    \lambda = \frac{2\pi}{q_0\chi}, 
\end{equation*}
which in fact states that if the wavelength $\chi^{-1}2\pi/q_0 < \lambda$ rivulets should form. 
\JH{Change in paper?}
}

\item[ \textbf{\underline{Comment 5.}}]
{
The thin film equation relies on a local approximation for the interface potential and disjoining pressure. 
Notice that to first order in non-local effects, the surface tension entering the thin film equation adopts a film thickness dependence which could potentially affect the dynamics. 
Particularly, as far as the parallel correlation length $q_0$ is referred, the non-local effects provide an upper bound equal to the inverse bulk correlation length. 
This could be a significant effect in polymer thin films (c.f. Benet et al. jpc-c 118, 22079, 2014 and the related discussion on dynamical effects Pahlavan et al., jfm, 845, 642, 2018).
}

\item[ \textbf{{Answer}}]
{
That is indeed a very interesting point.
In fact, we know the second reference very well from our own work~\cite{PhysRevE.104.034801}.
It is an appealing model, and we look forward to discussing it in further details in future publications.
We therefore added the following to the conclusion 
\\
\textcolor{red}{...}
}

\end{itemize}

\bibliographystyle{abbrv}
\bibliography{Ref}

\end{document}

### 2022-10-03 Reply_RefereeC.tex
Update

\documentclass[12pt,english]{article}

\usepackage{natbib}

\usepackage{graphics,graphicx,dcolumn,bm,fleqn,epic,eepic,float}
\usepackage{amssymb,amsmath,multirow,rotate,rotating,color}
\usepackage[utf8]{inputenc}

\usepackage[english]{babel}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\tikzset{fontscale/.style = {font=\relsize{#1}}}
\usetikzlibrary{calc}
\makeatother

\newcommand{\figref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\eqnref}[1]{Eq.~(\ref{eq:#1})}
\newcommand{\ts}{\textsuperscript}

\definecolor{tuered}{RGB}{214,0,74}

\newcommand{\todo}[1]{\textbf{\textcolor{tuered}{ TODO: #1}}}

\newcommand{\vectornorm}[1]{\left|\left|#1\right|\right|}

\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\uvec}[1]{\hat{\vec{#1}}}
\newcommand{\tensor}[1]{\mathbf{#1}}

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}

\newcommand{\JH}[1]{\textcolor{blue}{JH: #1}}
\begin{document}

% The referee PDF has been generated from r239 -- 'make referee' command.

\section*{Reply to Referee C}
We thank the Referee for his/her report on our manuscript.
We are glad to read that he/she finds the study scientifically sound and credible and 
the manuscript well written. Based on his/her important remarks, we have 
prepared a revised version which we hope will be more convincing also for what concerns
its impact and interest for the readership of Physical Review Letters.
Hereafter detailed answers to the various questions raised are reported. All changes 
in the text are edited in red colour, for the sake of clarity.

\begin{itemize}

\item[ \textbf{\underline{Comment 1.}}]
{ 
However, I feel that the
results presented lack a sufficient breadth and impact so as to
warrant publication in Physical Review Letters - provided that one has
a suitable substrate, I do not see how would the transport of a fluid
by way of transient rivulets surpass the transport by way of droplets
in practice.
}

\item[ \textbf{{Answer}}]
{ 
We say nowhere that the rivulets would transport fluid more efficiently. Actually 
we do not talk about transport at all. The focus, here, is not on sliding or similar force-driven processes, but 
on dewetting and, in particular, on the different dewetting {\it morphologies}. 
It is true, on the other hand, that rivulets are metastable, but, as 
we showed, their lifetime can be extended increasing the pattern speed
(therefore enabling treatment, e.g. curing, of the substrate to 
fix a given pattern).
Concerning the breadth and impact to warrant publication in Physical Review Letter, as we think we stressed extensively in the introduction,
the problem of controlling dewetting, besides being of fundamental interest in Fluid Dynamics (which represents an important part of the 
PRL readership), it is crucially relevant for applications in 
nanoprinting (including printable electronics). 
This is witnessed by the various publications in PRL on this and closely related problems. To cite but a few recent ones: J.C. Fern\'andez-Toledano et al, "How wettability controls nanoprinting",
{\it Phys. Rev. Lett.} {\bf 124} 224503 (2020); A.A. Pahlavan et al,
"Evaporation of binary-mixture liquid droplets: the formation of picoliter pancakelike shapes", {\it Phys. Rev. Lett.} {\bf 127}, 024501 (2021); C. Clavaud et al, "Modification of the fluctuation dynamics of ultrathin wetting films", {\it Phys. Rev. Lett.} {\bf 126}, 228004 (2021).
\JH{We have to make sure to strongly doubt the competence of Ref C in the letter to the editor! I guess this will be the only chance for us to get the paper through/}
}

\item[ \textbf{\underline{Comment 2.}}]
{ 
The core of the manuscript is a numerical study which includes a
number of model parameters and specific choices of their values, say
$\theta_0$, $ \delta \theta$, $\delta$, $h_0$, 
the square pattern itself, the
orientation of the velocity of the contact angle, etc. Naturally,
these choices need to be made, but drawing conclusions from just a few
sets is, in my opinion, premature.
For example, I imagine that the rivulets will not form if $\delta\theta$ were too small, or too small relative to $\theta_0$. 
The authors focused on the importance of the speed of the pattern and on its wavelength but they left most other parameters fixed to a single value, the reasons for their choice being unclear.
}

\item[ \textbf{{Answer}}]
{ 
The Referee is right. Indeed there are of course several parameters 
that may influence the physics of the problem. In this (first) contribution, though,
our aim is to single out the factor determining the novel effect found (the 
droplet-rivulet duality), namely
the pattern time dependence (and, hence, the 'wave' speed). 
One may expect of course that there is a dependence on $\delta \theta$ and $\theta_0$, but this would apply to the static case. 
A systematic study of the full parameter space would then deserve a dedicated, 
extended, work that goes somehow beyond the goal of the present paper and a typical PRL contribution.
Nevertheless, we agree with the Referee that it is appropriate to add further details, therefore we commented in the text 
on the choice of the checkerboard pattern and of $\theta_0$ and we have added 
results at changing $\delta \theta$ in a Supplemental Material section.

{\bf ANDREA: Stefan, we do have these data, right? I think to remember that you ran the simulations}

%The observation of the referee is correct.
%Indeed, the difference in contact angle amounts to a net force on the liquid.
%This force can be large enough to actuate %droplets~\cite{doi:10.1021/acs.langmuir.5b02335}.

%The thin film equation is valid only in the regime where $\theta \ll \pi/2$.
%Choosing $\theta_0 = 20^{\circ}$ and a $\Delta\theta$ of the same magnitude %therefore ensures that we are well within the predictability of our model.
%Therefore we find that our value of $\theta_0$ is a good compromise to satisfying %the constraints.

%Most of our results are given in terms of $t_0$ and $q_0$.
%Both contain parameters we choose to set up our numerical experiments. 
%We supply an equation to calculate them and have no evidence to believe that our theoretical assumptions are incorrect.
%Could the referee please elaborate why he/she thinks we overlook something?

%\textcolor{pyorange}{Stefan}: I would put the results we have from the x-only %velocity run here.
%They kind a answer what happens if $\Delta\theta$ becomes larger (rivulets life for %a very short amount of time only).
}

\item[ \textbf{\underline{Comment 3.}}]
{ 
Also unclear is how does the substrate patterning scale $2\pi/q_\theta$ compare with the scale of homogeneous dewetting $2\pi/q_0$.
I tried hard to understand whether the different cases of $\lambda$ discussed in the paper agree with $2\pi/q_0$ but I was unable to do so. 
How would the film behave if the dynamics of the film were controlled primarily by spinodal dewetting and not by the contact-angle pattern?
At the same time, if the two characteristic lengthscales ($=$ that of the spinodal dewetting and that of the patterned substrate) are suitably synchronized, perhaps rivulets can be rendered permanent rather than transient? Or is the dewetting, once initiated, immanently irreversible even for a
time-dependent contact-angle pattern like the one studied in LM17723?
}

\item[ \textbf{{Answer}}]
{
We apologize for having lacked thoroughness not providing the numerical 
value of $q_0$ (of which we reported the expression, though) to compare with the 
other characteristic scales. Plugging in the values of $\Theta = \theta_0 + \delta \theta$, 
$h_0$ and $h_{\ast}$ we get $q_0 \approx 0.13 h_0^{-1}$, whence 
$\lambda_0 = 2\pi/q_0 \approx 50 h_0$, so it corresponds approximately to the 
shortest pattern wavelength considered. The pattern-driven instability overwhelms the spinodal instability because it is active even on an initially
flat film 
(i.e., where the initial perturbation $|\delta h_0| \rightarrow 0$) 
\cite{KonnurPRL2000,KarguptaLangmuir2000,KarguptaPRL2001}; ideally one would need to have a vanishing 
contact angle mismatch in order to see the spinodal dewetting, namely $\delta \theta \sim \delta h_0/h_0$,
which is also somehow obvious, but it would eventually deprive the pattern of meaning. 
The emergence of rivulets is observed even for $\lambda \approx \lambda_0$, as we show in a new figure
(in the Supplemental Material), but their nature is still metastable.

{\bf ANDREA: Stefan your input is needed here, too. Did you make the simulations/figures, with the 
properly defined $\Gamma$ for $\lambda \approx 2 \pi/q_0$?}
}
%We are sorry for creating confusion with the choice of our variables.
%Let us call $\lambda_0 = 2\pi/q_0$ and rephrase Eq.~(11) in terms of $\lambda_0$ %using $v_0^{\ast}$,
%\begin{equation*}
%    v_0^{\ast} = \frac{\lambda_0}{t_0}, 
%\end{equation*}
%computing $\Gamma$ in terms of $v_0^{\ast}$ yields
%\begin{equation*}
%    \Gamma = \frac{v_0^{\ast}}{U_{\theta}} = \frac{6\pi h_0^3 q_0^3}{\Theta^3}.
%\end{equation*}
%Setting this result equal with Eq.~(11) in the manuscript we have
%\begin{equation*}
%    \frac{\chi 3\lambda h_0^3 q_0^4}{\Theta^3} = \frac{6\pi h_0^3 q_0^3}{\Theta^3},
%\end{equation*}
%with most of the terms canceling
%\begin{equation*}
%    \lambda = \frac{2\pi}{q_0\chi}, 
%\end{equation*}
%which in fact state that if the wavelength $\chi^{-1}2\pi/q_0 < \lambda$ rivulets %should form. 
%}

%\item[ \textbf{\underline{Comment 4.}}]
%{ 
%How would the film behave if the dynamics of the film were controlled primarily by %spinodal dewetting and not by the contact-angle pattern?
%}

%\item[ \textbf{{Answer}}]
%{ 
%We thank the referee for raising this question.
%The straightforward answer is we can reproduce the structure factor of a spinodally %dewetting thin film which is given by
%\begin{equation*}
%    S(q,t) = S_0 e^{2\omega(q)t},
%\end{equation*}
%where is $S_0$ is a constant factor and 
%\begin{equation*}
%    \omega(q) = \frac{1}{t_0}\left[2\left(\frac{q}{q_0}\right)^2 - %\left(\frac{q}{q_0}\right)^4\right],
%\end{equation*}
%}
%which can be found in ref.~\cite{PhysRevE.104.034801}, Fig~1.
%After the generation of droplets, we observe a coarsening scaling according to a %powerlaw.

%\item[ \textbf{\underline{Comment 5.}}]
%{ 
%At the same time, if the two characteristic lengthscales ($=$ that of the spinodal %dewetting and that of the patterned substrate) are suitably synchronized, perhaps %rivulets can be rendered permanent rather than transient? Or is the dewetting, once %initiated, immanently irreversible even for a
%time-dependent contact-angle pattern like the one studied in LM17723?
%}

%\item[ \textbf{{Answer}}]
%{ 
%This is an interesting comment.
%To our understanding, this is not the case as shown in Fig.~5 of the manuscript.
%We observe a saturation like effect for high pattern velocity.
%Independent of the velocities, the rivulet admits a varicose instability, which %within our study was not possible to overcome.
%We hypothesize that another substrate pattern could ensure indefinite rivulet %stability.
%However as we do not have an intuition, testing N-patterns would be beyond the scope %of this work.

%For further details on the scale synchronization, we would point towards the answer %to Comment 4.
%}

\item[ \textbf{\underline{Comment 4.}}]
{ 
Many of the numerical results are supported by scaling estimates. I
like these arguments, but the agreement is hardly convincing. For
example, the linear and the square scaling laws in Fig. 2 each cover a
small range (much less than a tenfold variation) and so it is hard to
see whether they really apply. 
The same goes for the rivulet lifetime in Fig. 5 where the trend is OK but most likely a scaling law different from $\log(v_{\theta})$ would do as well in the range of $\tau_{riv}$ covered by the numerical data.
}

\item[ \textbf{{Answer}}]
{ 
We are well aware that the parameter ranges are not too wide, but 
of course our method (as any other) has to cope with limitations. 
For example, the range of achievable wavelengths is limited from below 
by the need to stay in the lubrication regime, and from above by the size of the system (i.e. by computing resources). We could actually increase the system size and thus push further the upper limit of $\lambda$, but this would only extend the $\lambda^2$ scaling range. {\bf ANDREA: May we try one run on a bigger 
system? Being 1d it shouldn't take too long, should it? It would be also interesting to try a smaller $\Theta$: my handwaving argument on comparing times tells that it should extend a bit the linear, $\sim \lambda$,  scaling range.}
As for the rivulet lifetimes, analogously, for reasons intrinsic to the stability of the numerical method, much higher pattern speeds cannot be reached 
(in particular one has to comply with the condition $\Gamma \ll 5 \times 10^2$).
We tend to disagree, instead, with the comment that any other scaling 
law would be equivalently good. Essentially because the logarithm scaling is 
not a fit, it is the result of a derivation based on plausible theoretical arguments, that 
is 'OK' with the numerical data.
We think, anyhow, that the Referees' concerns are legitimate and we have, therefore, added 
comments of caution about the scaling laws in the revised text.
Let us remark, though, that even in real microfluidic systems, 
one cannot expect too wild (and wide) variation of lengths and velocities.
In fact, we would be glad if our pioneering study might motivate further work 
in this direction, hopefully confirming our findings on an extended parameter 
range. Of course, that is more likely to happen if our message is conveyed
by a contribution in a high impact journal such as PRL. 
}

\end{itemize}

\bibliographystyle{abbrv}
\bibliography{Ref}

\end{document}

### 2022-10-03 Response.tex
Update

\documentclass[12pt,english]{article}

\usepackage{natbib}

\usepackage{graphics,graphicx,dcolumn,bm,fleqn,epic,eepic,float}
\usepackage{amssymb,amsmath,multirow,rotate,rotating,color}
\usepackage[utf8]{inputenc}

\usepackage[english]{babel}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\tikzset{fontscale/.style = {font=\relsize{#1}}}
\usetikzlibrary{calc}
\makeatother

\newcommand{\figref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\eqnref}[1]{Eq.~(\ref{eq:#1})}
\newcommand{\ts}{\textsuperscript}

\definecolor{tuered}{RGB}{214,0,74}

\newcommand{\todo}[1]{\textbf{\textcolor{tuered}{ TODO: #1}}}

\newcommand{\vectornorm}[1]{\left|\left|#1\right|\right|}

\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\uvec}[1]{\hat{\vec{#1}}}
\newcommand{\tensor}[1]{\mathbf{#1}}

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}

\newcommand{\JH}[1]{\textcolor{blue}{JH: #1}}
\begin{document}

% The referee PDF has been generated from r239 -- 'make referee' command.

\section*{Reply to Referee A}
We thank the Referee for his/her report on our manuscript. 
We are glad to read that he/she finds it of potential interest and in principle 
suitable of publication, if a thorough revision is performed. 
This is what we did in the new version, where we tried to clarify the obscure points, following all the Referee's remarks.
Hereafter we provided answers to the various questions/comments raised. All changes in the revised 
version are reported in red colour.

\begin{itemize}

\item[ \textbf{\underline{Comment 1.}}]
{ 
I am unable to follow most of the arguments made with respect to the disjoining pressure. 
For me, one needs to integrate the disjoining pressure to arrive at an effective interface potential, and 
I do not really know how to go from the effective interface potential to the surface tension. 
The equation (3) therefore merits much more discussion.
}

\item[ \textbf{{Answer}}]
{
We did not give, in fact, any particular argument with respect to the disjoining pressure, since 
we assumed that it is a very well established concept employed in any theoretical and computational study of thin liquid film hydrodynamic in the lubrication approximation framework.
We acknowledge, though, that our assumption was probably erroneous and we give some more explanation. 
First of all, there is no need (and no way) to go from the disjoining pressure (or the interface potential) to the surface tension, since the two quantities carry complementary physical information. 
Let us, first, make a step back and underline a fundamental difference between the physics of {\it one free interface} and that of a {\it film}:
the latter is a layer of liquid confined between {\it two} interfaces. When the two interfaces are brought close enough to each other (the layer is thin), they interact.
The disjoining pressure stems from the interaction. Thermodynamically, it represents minus the derivative of the free energy difference, per unit surface, between 
the film and the bulk of the same liquid, or, equivalently, the pressure difference inside and outside the film 
\cite{Deryaguin1936,DeryaguinChuraev1978}.
In the case of wetting, there can be three types of interfaces: liquid/air, liquid/solid (the substrate) and, 
if the substrate is only partially wetted, solid/air, thus giving rise to three-phase contact lines. 
The usually called {\it surface tension} is the energy cost of forming a liquid/air interface of unit area and its only knowledge is, therefore, not enough in our case. 
The Young's condition (Young, 1805, de Gennes, 1985) connects the surface tension with the analogous energies per unit
area associated with the solid/liquid ($\gamma_{\text{sl}}$) and solid/gas ($\gamma_{\text{sg}}$) interface as
\begin{equation}
\cos \theta = \frac{\gamma_{\text{sg}}-\gamma_{\text{sl}}}{\gamma},
\end{equation}
where $\theta$ is the contact angle.
So, there is no need (and no way) to go from the disjoining pressure (or the interface potential) to the surface tension, since the two quantities carry complementary physical information,
and both are required to determine the dynamics of dewetting.
Of course, on the other hand, the disjoining pressure depends also on the surface tension, which appears, in fact, in its expression (3). 
We think, though, that part of the confusion might be due to the fact that we do not define the symbol of the surface tension at its first occurrence, but only later. 
We have now fixed this and we apologize for it.
}

\item[ \textbf{\underline{Comment 2.}}] 
{
The same holds for the time scale. 
The viscous time scale is well defined, but necessitates a length scale in addition to the 
capillary velocity. It is unclear why the authors choose the length scale that they use in the paper.
Subsequently, they replace the surface tension by a complicated partial derivative of the 
disjoining pressure; here the same remark as above holds.
}

\item[ \textbf{{Answer}}]
{
We apologize if we have not been thorough in the discussion of the relevant scales, which indeed deserves some clarification. 
Thin film hydrodynamics is an intrinsically multiscale phenomenon, whereby there are at least two characteristic length scales: the mean film thickness (let us call it $h_0$)
and an horizontal length scale $\ell$ (which in our case can be, for example, the pattern wavelength, $\lambda$), where, typically $h_0 \ll \ell$. We chose to use the mean 
film thickness as a scale, because it is the only relevant length that is common to all the simulations.
In the same spirit, we chose as a reference time, $t_0(\theta)$, which is not exactly the capillary time (though being related to it). 
$t_0$, as we explain in the text, is, in fact, the characteristic time of growth of the most unstable mode in the spinodal dewetting on a homogeneous
substrate of contact angle $\theta$. In spite of the substrate being patterned in our case, it is possible to define an effective $t_0$ in terms 
of the highest contact angle on the pattern, which is what determines the fastest instability, i.e. $t_0(\Theta)$, where 
$\Theta=\mbox{max}_{\mathbf{x}}\{\theta(\mathbf{x})\}= \theta_0 + \delta \theta$. We deem the so defined characteristic time as an appropriate scale,
since it does not depend on the pattern wavelength and speed. Actually, one can see that it captures well the time scales of the dewetting process
from figure 2, where it is shown that the rupture times are of the order of fraction to units of $t_0$.
We recognize, on the other hand, that, in this spirit it is unclear the reason of the choice of $v_0 = \lambda/t_0$ as a reference velocity.
A more natural candidate would be, indeed, $v_0=h_0/t_0$. We have therefore decided to use the latter in the revised version of the manuscript.
We have also rephrased the corresponding paragraph in the {\it Method} section as follows: \\

\textcolor{red}{}

}

\item[ \textbf{\underline{Comment 3.}}]
{ 
Incidentally, for a 'normal' fluid such as water, the capillary velocity is $\sim 70\mbox{m}/\mbox{s}$, 
so that very, very rapidly the fluid inertia becomes important. 
This is indeed what one usually observes for the formation or coalescence of drops, so that at the very least this should be discussed and the Reynolds number calculated.
}

\item[ \textbf{{Answer}}]
{
Such a high value corresponds to the ratio of (surface tension)/(dynamic viscosity) for water, i.e. $v = \gamma/\eta \approx 70 \mbox{m}/\mbox{s}$.
This expression, though, is only representative of the capillary velocity, again, for {\it free surface flows}, as, e.g., in the coalescence of 
two droplets in the bulk of another fluid. In the case of thin liquid films, the capillary velocity is still proportional to the ratio 
$\gamma/\eta$, but by a factor $\varepsilon^3$, where $\varepsilon = h_0/L$ is the film thichkness parameter, which is typically $\varepsilon \ll 1$.
For the dewetting of a millimetric film of micrometric thickness reatracting into a single droplet, experiments have measured that 
the typical velocity is $\sim 1 \mbox{mm}/\mbox{s}$, which gives a Reynolds number $Re \approx 10^{-3}$ \cite{Edwards}.
}

\item[ \textbf{\underline{Comment 4.}}]
{ 
In addition, there is a lot of old literature that relates the dewetting to the disjoining pressure, and shows that the nucleation of a hole in a wetting film is much more difficult than the nucleation of such a wetting film; there are experiments on this by Rutledge and Taborek, theory by Schick and Taborek and Bausch, Blossey and Indekeu.
}

\item[ \textbf{{Answer}}]
{
We thank the referee for bringing this to our attention.
However, we have hard time understanding what he/she means with ``\textit{... and shows that the nucleation of a hole in a wetting film is much more difficult than the nucleation of such a wetting film}''
}

\item[ \textbf{\underline{Comment 8.}}]
{ 
The explanations also lack a discussion on time, length, and energy scales. 
}

\item[ \textbf{{Answer}}]
{
We find this comment odd as we provide suitable time and length scales, derived from the disjoining pressure.
A natural energy scale emerges as the spreading parameter $\mathcal{S}$ which can be derived from the disjoining pressure as well. 
We are very sorry if the referee dislikes the model of the disjoining pressure, but various publications can be found using this model, see answer to Comment 2.
}

\item[ \textbf{\underline{Comment 9.}}]
{ 
I have no clue what the values mean ``the numerical values, in LB units, are set to $\gamma= 0.01$ and  $\mu = 1/6$''. 
}

\item[ \textbf{{Answer}}]
{
We are sorry to hear that these numbers have led to confusion.
For completeness of the manuscript, we provide the parameters (and the code, see arXiv) we used to generate data.
The referee is right to question these numbers, and we agree that their values seem arbitrary.
Using the lattice Boltzmann method, it can be shown that the choice of $\mu = 1/6$ is the most stable for a single relaxation time collision operator~\cite{kruger2017lattice}.

To make sense of them, we scale them with characteristic quantities of a dewetting thin film, therefore $q_0$ and $t_0$.
}

\item[ \textbf{\underline{Comment 10.}}]
{ 
What is the disjoining pressure due to? 
Van der Waals forces? 
How thick is the precursor film, and why should the disjoining pressure be zero there?
Is there any comparison with experiments possible/feasible?
}

\item[ \textbf{{Answer}}]
{
The disjoining pressure allows us to have an effective three phase contact line with the correct wetting behavior and without diverging pressure.
We agree that this a purely numerical problem that does not appear in experiments, although it is possible to relate the disjoing pressure to the Hamaker constant,
\begin{equation*}
    \mathcal{A} = 6\pi h^3\kappa(\theta),
\end{equation*}
using 
\begin{equation*}
    \kappa(\theta) = \frac{2\gamma}{h_{\ast}}(1-\cos(\theta(\mathbf{x},t))),
\end{equation*}
which is non powerlaw part of the disjoining pressure.

Advancing three phase contact lines are a topic by its own, and we would point the referee towards relevant literature, as this discussion would be beyond the scope of the problem at hand.

The referee is correct in assuming it includes Van der Waals forces.
This is another very interesting comment, as far as the authors can tell there is no common sense on the existence of the precursor layer.
There are experimental studied that claim its existence, but with a thickness of only a few nanometers.
In our model it mainly ensures that the pressure near the three phase contact line is well behaved.
There is actually a lot of comparison with experiments see refs.~\cite{becker2003complex, fetzer2007thermal, doi:10.1073/pnas.1820487116} or e.g. the review from Bonn et al.~\cite{RevModPhys.81.739}.
We would further refer to the answer of Comment 2.
}

\end{itemize}

\section*{Reply to Referee B}

We thank the Referee for his/her thorough reading of our manuscript and detailed report.
We are happy to read that the referee acknowledges the amount of work we did and thinks it could be fitting for PRL after revision.
In what follows we provide detailed answers to all Referee's questions and we discuss all the changes made to the manuscript
(which are written in the text in red colour).

\begin{itemize}

\item[ \textbf{\underline{Comment 1.}}]
{
The model disjoining pressure, Eq. 3 with a tail decaying as the inverse squared power seems odd. 
I am more familiar with the expected van der Waals tail with a decay as the inverse cube power. 
Is there any reason for this choice?

\item[ \textbf{Answer}]
{
We thank the referee for pointing this out.
It is true that there are variations concerning the powerlaw model of the disjoining pressure.
A common one to which the referee refers to is $(n,m) = (9,3)$, the choice we made for the manuscript is $(n,m) = (3,2)$.
Both exponent pairs can be found in the literature with even more variety~\cite{moulton_lega_2013,SCHWARTZ1998173,doi:10.1063/1.4828721,brasjen2013dewetting}.
The pair $(9,3)$ is the one that can be motivated from functional derivatives of a Lenard-Jones potential, understandably this could more appealing to use.
We have used both exponent pairs in previous works, see e.g.~\cite{PhysRevE.104.034801}.
In this manuscript we used $(3,2)$ to keep $t_0$ reasonably small and thus the overall computational cost manageable.
Using the pair $(9,3)$ does not alter the results, it effectively shifts the value of $t_0$ by $\approx 10$.
}

\item[ \textbf{\underline{Comment 3.}}]
The second major concern refers to the choice of slip length. 
The authors study a fluid-substrate model that has a thin film equilibrium thickness of 0.07 length units, but the slip length is one unit, hence about 10 times larger than the equilibrium film thickness. 
This seems unexpected, in view of the small contact angle chosen for the system. 
Could the authors justify this choice?
}

\item[ \textbf{Answer}]
{
\begin{figure}
    \centering
    \includegraphics [width=0.6\textwidth]{slip_measure.pdf}
    \caption{Dewetting rim profiles for various slip parameters $\delta$.}
    \label{fig:slip}
\end{figure}

We thank the referee for bringing this up and are happy to talk about this issue.
Slip and disjoining pressure are an essential part of our model, and many other thin film solvers.
Here they allow us to have moving contact lines with imposed wettability features.

We agree that the value may seem large at first glance.
As Peschka et al. have shown, the creation of satellite droplets during rivulet breakup, similar to what we observe, is a no-slip feature~\cite{doi:10.1073/pnas.1820487116}.
We observe the creation of satellite droplets in the initial dewetting.
Increasing the slip length makes those droplets disappear.

Another point to mention here is the slip dependency of the dewetting rim.
Small to vanishing slip induce a dimple behind the front, while for large slip values this dimple vanishes~\cite{fetzer2007quantifying, munch2005contact}.
We checked this behaviour in Fig.~\ref{fig:slip} and found only a small dependency for slip lengths $\delta < 10$.

Spots with thickness $h \leq h_{\ast}$ are considered dry spots, therefore the term equilibrium thickness maybe exaturated.
}

\item[ \textbf{\underline{Comment 4.}}]
{
The relevance of the manuscript relies on the significance of the substrate's switchable properties. 
Particularly, it is assumed that the wetting properties can be modulated on space but also on time. 
The authors have mentioned in the introduction that this can be potentially done. 
However, it would be desirable that the authors relate their model of switchable substrate with some specific experimental realization, where the actual length and time scales are mentioned.
}

\item[ \textbf{{Answer}}]
{
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{simple_exp.pdf}
    \caption{Schematic realization of the experiment described in the manuscript.}
    \label{fig:experiment_illustration}
\end{figure}
We thank the referee for bringing this up.
First, we like to note that experiments are not our field of expertise and therefore our envisioned realization could be flawed.

The idea we have would be to prepare a thin film ($h_0 \approx 4nm$ PS) on a light responsive substrate.
Dimensions of the substrate should at least be $2$ microns in $x$ and $y$, but larger is fine.
Arrange several light sources (diodes) in a lattice with a distance of $\lambda \approx 1\mu m$ between them, as displayed in Fig.~\ref{fig:experiment_illustration}.
It would be perfect if the structure on which the light sources are mounted can be moved in a periodic manner, e.g. wrapped around rolls.
On the other hand if the array of diodes is large enough, say $100\lambda$ it should be sufficient to move them continuously in one direction.

In Eq.~(5) we introduce two scales, $t_0$ and $v_0$.
From experiments of dewetting thin films we know that $t_0$ correlates with the rupture time~\cite{becker2003complex, fetzer2007thermal}.
Considering the same experimental parameters as above, 
\begin{align*}
    t_0 &\approx 1800s,\qquad~~~\qquad \frac{2\pi}{q_0}\approx 400nm, \\
    \quad v_0 &= \frac{\lambda}{t_0} \approx \frac{1}{900}\frac{\mu m}{s},\quad \quad v_0^{\ast} = \frac{2\pi}{q_0 t_0} \approx \frac{4}{9}\frac{nm}{s},
\end{align*}
where we take into account that the rupture time is overestimated by about a factor of two.
The experiment therefore reduces to the movement of the diodes with a velocity between $[10^{-7}/900, 10^{-5}/900] m/s$. 
}

\item[ \textbf{\underline{Comment 5.}}]
{
An important parameter in the problem is the wave-vector $q_0$, which sets the correlation length in the problem. 
The authors should state what the actual value is relative to either the average film height, or the wave-length of modulations. 
Is there any reason to give the wave-length of modulations in units of the system size, i.e., are the results subject to finite size effects? Otherwise it appears more meaningful to describe the modulations in units of $q_0$ or the equilibrium film thickness.
}

\item[ \textbf{{Answer}}]
{
We thank the referee for this comment.
It is true that we use $q_0$ but never display its actual value.
Inserting numbers, we have 
\begin{equation*}
    q_0 \approx 0.13, \quad \lambda_0 = \frac{2\pi}{q_0} \approx 48.5 \approx \frac{L}{10},  
\end{equation*}
therefore $\lambda_0$ is usually smaller than $\lambda$.

No, these simulations do not suffer from finite size effects.
The honest reason for using multiples of the system size is the ease of the numerical realization.
Having periodic boundary conditions, a non-multiple wavelength of the system size would induce a discontinuity.
Given the complexity of the problem, we would like to have as much control of the pattern as possible.

The arbitrary choice we made had however helped us to understand where the boundary between rivulet and droplets appear.
In Eq.~(11) we use
\begin{equation*}
    \Gamma = \frac{v_{\theta}}{U_{\theta}} = \frac{\chi v_0}{U_{\theta}},
\end{equation*}
where rivulets appear for $\chi > 1$.
Substituting the wavelength dependence of $v_0$ with $q_0$,
\begin{equation*}
    v_0^{\ast} = \frac{\lambda_0}{t_0}, 
\end{equation*}
computing $\Gamma$ in terms of $v_0^{\ast}$ yields
\begin{equation*}
    \Gamma = \frac{v_0^{\ast}}{U_{\theta}} = \frac{6\pi h_0^3 q_0^3}{\Theta^3},
\end{equation*}
relating both with each other yields
\begin{equation*}
    \frac{\chi 3\lambda h_0^3 q_0^4}{\Theta^3} = \frac{6\pi h_0^3 q_0^3}{\Theta^3},
\end{equation*}
with most of the terms canceling one gets
\begin{equation*}
    \lambda = \frac{2\pi}{q_0\chi}, 
\end{equation*}
which in fact state that if the wavelength $\chi^{-1}2\pi/q_0 < \lambda$ rivulets should form. 
}

\item[ \textbf{\underline{Comment 6.}}]
{
The thin film equation relies on a local approximation for the interface potential and disjoining pressure. 
Notice that to first order in non-local effects, the surface tension entering the thin film equation adopts a film thickness dependence which could potentially affect the dynamics. 
Particularly, as far as the parallel correlation length $q_0$ is referred, the non-local effects provide an upper bound equal to the inverse bulk correlation length. 
This could be a significant effect in polymer thin films (c.f. Benet et al. jpc-c 118, 22079, 2014 and the related discussion on dynamical effects Pahlavan et al., jfm, 845, 642, 2018).
}

\item[ \textbf{{Answer}}]
{
That is indeed a very interesting point.
In fact, we know the second reference very well from our own work~\cite{PhysRevE.104.034801}.
It is an appealing model, and we look forward to discussing it in further details in future publications.
We therefore added the following to the conclusion 
\\
\textcolor{red}{Never gonna give you up\\
Never gonna let you down\\
Never gonna run around and desert you\\
Never gonna make you cry\\
Never gonna say goodbye\\
Never gonna tell a lie and hurt you}
}

\end{itemize}

\section*{Reply to Referee C}
We thank the Referee for his/her report on our manuscript.
It is a pleasure to read that he/she acknowledges the credibility of the work and finds our writing appealing.
We are happy to see the he/she likes the interpretation of our results in terms of scaling arguments.
In what follows we provide detailed answers to all Referee's questions and we discuss all the changes made to the manuscript
(which are written in the text in red colour).

\begin{itemize}

\item[ \textbf{\underline{Comment 1.}}]
{ 
The topic may be potentially relevant for microfluidic and related applications journals, e.g., Physical Review E or Physical Review Fluids.
}

\item[ \textbf{{Answer}}]
{ 
We think this comment is judgmental, and it relates neither to the quality of the manuscript nor to the findings of the manuscript.
It is a statement about personal preference and belittles work which was published in Physical Review Letters on dewetting~\cite{fetzer2007thermal}, microfludic~\cite{PhysRevLett.110.048303} and thin film morphologies~\cite{PhysRevLett.84.931}.
Given the novelty of the dewetting interaction with the modelled ``switchable substrate'' we truly think that our findings fit the scope of Physical Review Letters.
}

\item[ \textbf{\underline{Comment 2.}}]
{ 
However, I feel that the results presented lack a sufficient breadth and impact so as to warrant publication in Physical Review Letters - provided that one has a suitable substrate, I do not see how would the transport of a fluid by way of transient rivulets surpass the transport by way of droplets in practice - and thus I cannot recommend the manuscript for publication.
}

\item[ \textbf{{Answer}}]
{ 
We understand that the referee expresses his/her point of view and acknowledge it.
However, we find it odd that he/she would disregard a result because there is no immediate use case that comes to his/her mind.
Scientific work is creative work, solving unsolved problems to deepen our understanding of the world.
To the best of our knowledge, this manuscript describes an effect that has not been observed so far and is accompanied by theoretical justification.

The referee raises the point that transport using rivulets can not surpass transport by droplets.
We have to disagree with this comment.
Transporting droplets on a patterned substrate is a difficult task, with many hideous effects such as contact line pinning or pearling/coalescence~\cite{PhysRevLett.119.204501}.
The contact line of a rivulet on the other hand is simpler can can only if at all move along the flow direction.
Therefore, the authors think that there can be cases for which transport via rivulet could be favorable.

Apart from the transport problem, controlling the morphology of small volumes of fluids is a highly desired application.
From a purely theoretical point of view it is an interesting question of how to overcome the surface energy minimization due to droplet fragmentation. 
Printing integrated circuits only by the means of a dewetting thin film, without having to use lithography, sounds for example very appealing.
What we want to say is that a singular opinion may not cover the whole space of possible applications, be it in a scientific set up or an industrial.
}

\item[ \textbf{\underline{Comment 3.}}]
{ 
For example, I imagine that the rivulets will not form if $\delta\theta$ were too small, or too small relative to $\theta_0$. 
The authors focused on the importance of the speed of the pattern and on its wavelength but they left most other parameters fixed to a single value, the reasons for their choice being unclear.
}

\item[ \textbf{{Answer}}]
{ 
The observation of the referee is correct.
Indeed, the difference in contact angle amounts to a net force on the liquid.
This force can be large enough to actuate droplets~\cite{doi:10.1021/acs.langmuir.5b02335}.

The thin film equation is valid only in the regime where $\theta \ll \pi/2$.
Choosing $\theta_0 = 20^{\circ}$ and a $\Delta\theta$ of the same magnitude therefore ensures that we are well within the predictability of our model.
Therefore we find that our value of $\theta_0$ is a good compromise to satisfying the constraints.

Most of our results are given in terms of $t_0$ and $q_0$.
Both contain parameters we choose to set up our numerical experiments. 
We supply an equation to calculate them and have no evidence to believe that our theoretical assumptions are incorrect.
Could the referee please elaborate why he/she thinks we overlook something?

\textcolor{pyorange}{Stefan}: I would put the results we have from the x-only velocity run here.
They kind a answer what happens if $\Delta\theta$ becomes larger (rivulets life for a very short amount of time only).
}

\item[ \textbf{\underline{Comment 4.}}]
{ 
Also unclear is how does the substrate patterning scale $2\pi/q_\theta$ compare with the scale of homogeneous dewetting $2\pi/q_0$.
I tried hard to understand whether the different cases of $\lambda$ discussed in the paper agree with $2\pi/q_0$ but I was unable to do so. 
}

\item[ \textbf{{Answer}}]
{ 
We are sorry for creating confusion with the choice of our variables.
Let us call $\lambda_0 = 2\pi/q_0$ and rephrase Eq.~(11) in terms of $\lambda_0$ using $v_0^{\ast}$,
\begin{equation*}
    v_0^{\ast} = \frac{\lambda_0}{t_0}, 
\end{equation*}
computing $\Gamma$ in terms of $v_0^{\ast}$ yields
\begin{equation*}
    \Gamma = \frac{v_0^{\ast}}{U_{\theta}} = \frac{6\pi h_0^3 q_0^3}{\Theta^3}.
\end{equation*}
Setting this result equal with Eq.~(11) in the manuscript we have
\begin{equation*}
    \frac{\chi 3\lambda h_0^3 q_0^4}{\Theta^3} = \frac{6\pi h_0^3 q_0^3}{\Theta^3},
\end{equation*}
with most of the terms canceling
\begin{equation*}
    \lambda = \frac{2\pi}{q_0\chi}, 
\end{equation*}
which in fact state that if the wavelength $\chi^{-1}2\pi/q_0 < \lambda$ rivulets should form. 
}

\item[ \textbf{\underline{Comment 5.}}]
{ 
How would the film behave if the dynamics of the film were controlled primarily by spinodal dewetting and not by the contact-angle pattern?
}

\item[ \textbf{{Answer}}]
{ 
We thank the referee for raising this question.
The straightforward answer is we can reproduce the structure factor of a spinodally dewetting thin film which is given by
\begin{equation*}
    S(q,t) = S_0 e^{2\omega(q)t},
\end{equation*}
where is $S_0$ is a constant factor and 
\begin{equation*}
    \omega(q) = \frac{1}{t_0}\left[2\left(\frac{q}{q_0}\right)^2 - \left(\frac{q}{q_0}\right)^4\right],
\end{equation*}
}
which can be found in ref.~\cite{PhysRevE.104.034801}, Fig~1.
After the generation of droplets, we observe a coarsening scaling according to a powerlaw.

\item[ \textbf{\underline{Comment 6.}}]
{ 
At the same time, if the two characteristic lengthscales (= that of the spinodal dewetting and that of the patterned substrate) are suitably synchronized, perhaps rivulets can be rendered permanent rather than transient?
}

\item[ \textbf{{Answer}}]
{ 
This is an interesting comment.
To our understanding, this is not the case as shown in Fig.~5 of the manuscript.
We observe a saturation like effect for high pattern velocity.
Independent of the velocities, the rivulet admits a varicose instability, which within our study was not possible to overcome.
We hypothesize that another substrate pattern could ensure indefinite rivulet stability.
However as we do not have an intuition, testing N-patterns would be beyond the scope of this work.

For further details on the scale synchronization, we would point towards the answer to Comment 4.
}

\item[ \textbf{\underline{Comment 7.}}]
{ 
Or is the dewetting, once initiated, immanently irreversible even for a time-dependent contact-angle pattern like the one studied in LM17723?
}

\item[ \textbf{{Answer}}]
{ 
That is a very interesting question.
It could be possible, assuming
\begin{equation*}
    \theta(\mathbf{x},t)|_{t > 5t_0} = 0.    
\end{equation*}
We don't think that it is reversible for a non fully wetting substrate.

}

\item[ \textbf{\underline{Comment 8.}}]
{ 
For example, the linear and the square scaling laws in Fig. 2 each cover a small range (much less than a tenfold variation) and so it is hard to see whether they really apply.
}

\item[ \textbf{{Answer}}]
{ 
We agree with the referee and are sorry for the lack of data.
On the upper end we are limited by the total domain of our simulation, otherwise we lose periodicity.
We would like to keep periodicity otherwise sharp wetting gradients appear, see ref.~\cite{PhysRevE.104.034801} for their influence.
On the lower end, we need sufficiently many lattice points to resolve $\sin(x)$.
This is why we comprimise and display two theoretically valid powerlaws in less than tenfold variation.
}

\item[ \textbf{\underline{Comment 9.}}]
{ 
The same goes for the rivulet lifetime in Fig. 5 where the trend is OK but most likely a scaling law different from $\log(v_{\theta})$ would do as well in the range of $\tau_{riv}$ covered by the numerical data.
}

\item[ \textbf{{Answer}}]
{ 
We agree that another powerlaw could fit the data as well.
On the lower end we are bound to the appearance of rivulets, thus $v_{\theta} > 1$ or $v_0^{\ast} < \lambda$.
While in the upper end we observe a saturation explained in the answer to Comment 6.

Theoretically, this seems to be the easiest explanation to our data.
In Comment 3 the referee correctly observes that we have a well-defined protocol where we only change wavelength and pattern velocity.
Any powerlaw that does not correlate with the pattern wavelength would therefore not capture our observations.
}

\end{itemize}

\bibliographystyle{abbrv}
\bibliography{Ref}

\end{document}

### 2022-10-03 Old_version/main.tex
Update

%%
%\listfiles
%\documentclass[%
% reprint,%
%secnumarabic,%
% amssymb, amsmath,%
% aip,cha,%
%groupedaddress,%
%frontmatterverbose,
%]{revtex4-1}
\documentclass[twocolumn,amsmath,amssymb,showpacs,pre,nofootinbib,superscriptaddress]{revtex4-1} %,preprint,jcp

\bibstyle{apsrev}
%\usepackage{docs}%
\usepackage{bm}%
\usepackage{graphicx}
%\usepackage{subcaption}
\usepackage{tikz}
\usepackage{color}
\usepackage{xcolor}
\usepackage{physics}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}%
%\nofiles
\expandafter\ifx\csname package@font\endcsname\relax\else
 \expandafter\expandafter
 \expandafter\usepackage
 \expandafter\expandafter
 \expandafter{\csname package@font\endcsname}%
\fi
\hyphenation{title}
\newcommand{\JH}[1]{\textcolor{blue}{ JH: #1}}

\begin{document}

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}

%\title{Switchable substrates and thin film morphologies}
\title{Spatiotemporal varying contact angles and thin film morphologies}

\author{S. Zitz}
\email{s.zitz@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,\\
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
  \affiliation{Department of Chemical and Biological Engineering, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\author{A. Scagliarini}%
\email{andrea.scagliarini@cnr.it}
 \affiliation{Institute for Applied Mathematics "M. Picone" (IAC), \\
Consiglio Nazionale delle Ricerche (CNR),\\
Via dei Taurini 19, 00185 Rome, Italy}%
\affiliation{INFN, sezione Roma ``Tor Vergata'', via della Ricerca Scientifica 1, 00133 Rome, Italy}
\author{J. Harting}
\email{j.harting@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,\\
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
 \affiliation{Department of Chemical and Biological Engineering and Department of Physics, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\date{\today}

\begin{abstract}
We study the dynamics of an unstable thin film deposited on a ``switchable`` substrate.
The substrate is itself subject to a time varying function that affects the equilibrium contact angle $\theta_{\text{eq}}(\mathbf{x},t)$.
We show that in the case of vanishing time dependence the patterning drives the dewetting morphology and that all fluid is drained into regions of small $\theta_{\text{eq}}(\mathbf{x})$ and thus form droplets.
Interestingly enough with a time varying component to the pattern we are able to stop the droplet fragmentation and create pronounced ligament or rivulet morphologies.
We show that both the numbers of semistable rivulets as well as number of stable droplets can be controlled with the pattern wavelength ($\lambda$).
Finally we explain why the ligaments appear with a simple balance between capillary retraction and dynamic wetting.   
\end{abstract}

\maketitle

\newcommand{\ts}{\textsuperscript}
%\tableofcontents

\section{Introduction}\label{sec:intro}
Thin film flows and especially droplets are an phenomenon we experience in our every day life.
Being it the droplets condensed in the higher atmosphere and return to the ground as rain or the oil film in a pan during cooking. 
While this two examples are illustrative there are many industrial applications that build on thin film flows. 
Among them is the wetting of surfaces to reduce wear or put simply, lubrication~\cite{ReynoldsLubr, gross1980fluid, szeri2010fluid}. 
Coating surfaces to change their physio-chemical properties is another widely used application~\cite{DERYCK1998278, doi:10.1146/annurev.fluid.31.1.347, DASILVASOBRINHO19991204}.
On the other hand the need to control droplets is at the heart of many printing processes~\cite{singh2010inkjet, jo2009evaluation}.

More recently lab-on-a-chip devices have acquired much attention~\cite{C6LC00387G,Focke}, to work they require precise droplet or film flow on small scales~\cite{stone2004engineering, darhuber2010planar}.
One possibility for precise control is the local wettability.
Light switchable substrates~\cite{WANG200718}, therefore substrates that change their wettability based on light illumination, open up a new way to program this devices. 
From a theoretical point of view the motion of the fluid can be computed using the thin film equation~\cite{RevModPhys.69.931, RevModPhys.81.1131, RevModPhys.81.739, ReynoldsLubr}
\begin{equation}\label{eq:thin_film}
    \partial_t h(\mathbf{x},t) = \nabla\cdot\left(M(h)\nabla p(\mathbf{x},t)\right),
\end{equation}
where $h(\mathbf{x},t)$ is a time and spatial varying film thickness, $M(h)$ is a mobility function and $p(\mathbf{x},t)$ is a pressure term that drives the flow. 
The function $M(h)$ contains information about the velocity boundary condition at $h=0$, using no-slip yields $M(h) = h^3/(3\mu)$.
More interesting however is the pressure function $p(\mathbf{x},t)$ as it allows to model the wettability of a given substrate.
We therefore introduce a time and spatial dependent contact angle to control the local wettability.

Spatial variations of the contact angle are long known and used in previous studies~\cite{PhysRevE.91.023013, Sommers_2006}, time dependency of the contact angle on the other hand is not as well studied~\cite{suman2006dynamics}. 
Switchable substrates are in fact a realization of a time dependent contact angle. 
Using both spatial as well as time dependent variations of the local contact angle leads to interesting phenomena, for example continuously sliding droplets as shown by Grawitter and Stark~\cite{D0SM02082F}.  
Our approach builds on the same idea but our method, see ref~\cite{PhysRevE.100.033313}, allows us to go beyond a single droplet and to simulate the full dewetting transition starting with a thin film and ending with droplets.
We use this to show that a time dependent component of the contact angle alter the dewetting morphology quite substantially.
In the limit of a small time dependency or slow evolution of the contact angle field we qualitatively reproduce the results from Grawitter and Stark, such sliding droplets.
Speeding up the evolution of the contact angle field introduces a morphological transition, which to the understanding of the authors have not been looked at so far.
Therefore the purpose of this manuscript is to increase the understanding of the dewetting process.
Highlighting the fact that a temporal and spatial variation of the contact angle leads to novel morphological structures.
Thus our finding offer a simple way to understand how time periodic switchable substrates could generate novel dewetting morphologies.

The manuscript is outlined as follows:
In section~\ref{sec:method} we discuss the numerical method we use to generate our results and introduce relevant scales.
We then present our results in section~\ref{sec:results}, where we discuss the impact of the time dependency on the contact angle and show that both stability as well as dewetting morphology is affected.
A short summary of our finding and a closing conclusion with outlook is given in section~\ref{sec:sum_conclu}.

\section{Method}\label{sec:method}
In this section we describe the lattice Boltzmann method (LBM) we use.
The LBM is a meso scale approach that builds on the Boltzmann equation to simulate fluid flows~\cite{krueger2017}. 
It is widely applied to many kinds of flow problems, among them are for instance multi phase and multi component flows~\cite{shan1993lattice,PhysRevE.73.047701}, flows through porous media~\cite{liu2016multiphase} and shallow water flows~\cite{Salmon:1999:0022-2402:503, zhou2004lattice, PhysRevE.65.036309}.
The model we use has been introduced in ref.~\cite{PhysRevE.100.033313} and builds upon a class of lattice Boltzmann models originally developed to study the shallow water equations~\cite{Salmon:1999:0022-2402:503, PhysRevE.65.036309}.

\subsection{Thin film lattice Boltzmann method}\label{subsec:LBM}
Since the model has been developed and tested in previous work, ref.~\cite{PhysRevE.100.033313}, we just outline the most important features.
The internal degrees of freedom of this method are the distributions functions $f_{\alpha}(\mathbf{x},t)$.
These distribution functions describe the movement of fictitious particles on a lattice, in this case on a $D2Q9$ lattice.
Meaning there are nine velocity components ($Q9$), directions in which a particle could move, for a simulation in two dimensions ($D2$).
The dynamics of the thin film and thus the evolution of the distribution functions is given by
\begin{equation}\label{eq:LBE}
    \begin{split}
        &f_{\alpha}(\mathbf{x}+\mathbf{c}^{({\alpha})}\Delta t,t+\Delta t) = \\
        &\left(1 - \frac{\Delta t}{\tau}\right) f_i(\mathbf{x},t) + \frac{\Delta t}{\tau} f_{\alpha}^{(eq)}(\mathbf{x},t) + \Delta t\mathcal{S}_{\alpha},
    \end{split}
\end{equation}
where the $\mathbf{c}^{({\alpha})}$ are the $Q$ discrete lattice velocities that can be computed according to
\begin{equation}\label{eq:speeds}
\mathbf{c}^{({\alpha})}  =
\left\{
\begin{array}{ll}
(0,0) & \alpha = 0 \\
\left[\cos{\frac{(\alpha-1)\pi}{4}}, \sin{\frac{(\alpha-1)\pi}{4}} \right] &  \alpha=1,3,5,7 \\
\sqrt{2}\left[\cos{\frac{(\alpha-1)\pi}{4}}, \sin{\frac{(\alpha-1)\pi}{4}} \right] & \alpha=2,4,6,8
\end{array}
\right.,
\end{equation}
and $w$ are being the corresponding weights
\begin{equation}
w_{\alpha}  =
\left\{
\begin{array}{ll}
\frac{4}{9} & \alpha = 0 \\
\frac{1}{9} &  \alpha=1,3,5,7 \\
\frac{1}{36} & \alpha=2,4,6,8
\end{array}
\right..
\end{equation}
The last term to the right $\mathcal{S}_{\alpha}$ is a (generalized) source term.
The concrete implementation of the forces depend on the source term as~\cite{https://doi.org/10.1002/fld.4726}
\begin{equation}\label{eq:source_term}
    \mathcal{S}_{\alpha} = \frac{3w_{\alpha}}{\mathbf{c}_{\alpha}^2}\mathbf{c}_{\alpha}\cdot\mathbf{F}.
\end{equation}
We are able to match thin film dynamics with the usage of a force term comprising of two main components,
\begin{equation}\label{eq:forcesum}
    \mathbf{F} = \mathbf{F}_{\mbox{\tiny{fric}}} + \mathbf{F}_{\mbox{\tiny{film}}}.
\end{equation}
The first being a friction force that includes a velocity boundary condition at the fluid substrate interface
\begin{equation}\label{eq:forcefric}
\mathbf{F}_{\mbox{\tiny{fric}}} = -\frac{6h\mu\mathbf{u}}{2h^2+6\delta h + 3\delta^2},
\end{equation}
where $\mu$ is the viscosity and $\delta$ is a socalled slip length~\cite{RevModPhys.69.931, PhysRevE.100.033313, PhysRevE.63.011208, munch_lubrication_2005}.
The film thickness $h$ is the zeroth moment of the distribution function $f_{\alpha}$, i.e. $h(\mathbf{x},t) = \sum_{\alpha}f_{\alpha}(\mathbf{x},t)$ see Eq.~(\ref{eq:LBE}). 
The second component is due to the film pressure $p_{\text{film}}$ and accounts for the minimization of the fluid vapor interface as well as for the wettability of the substrate,
\begin{equation}\label{eq:pressure_here}
    p_{\text{film}} = -\gamma (\Delta h - \Pi(h)),
\end{equation}
which results in the addition of the force $\mathbf{F}_{\mbox{\tiny{film}}}$
\begin{equation}\label{eq:forcefilm}
    \mathbf{F}_{\mbox{\tiny{film}}} = -\frac{h}{\rho_0}\nabla(p_{\text{film}}), 
\end{equation}
with $\gamma$ being the (constant) surface tension and $\rho_0$ being the fluids density, which is $\rho_0 = 1$ and will be dropped in the following. 
Here we model the interaction between solid substrate and film with a disjoining pressure term~\cite{RevModPhys.69.931, RevModPhys.81.1131, PhysRevE.63.011208},
\begin{equation}\label{eq:disjoinpressure}
    \Pi(h) = \kappa(\theta_{\text{eq.}})\left[\left(\frac{h_{\ast}}{h}\right)^n - \left(\frac{h_{\ast}}{h}\right)^m\right],
\end{equation}
with $n$ and $m$ being integer exponents, $h_{\ast}$ defines the minimum of the potential and  
\begin{equation}\label{eq:kappacontactangle}
    \kappa(\theta_{\text{eq.}}) = (1-\cos\theta_{\text{eq.}}) \frac{(n-1)(m-1)}{(n-m)h_{\ast}},
\end{equation}
which therefor allows us to tune the fluids contact angle.

\subsection{Linear stability and characteristic quantities}\label{subsec:charq}
For further insights concerning the early time dynamics of the unstable film we perform a linear stability analysis.
Thus we expand Eq.~(\ref{eq:thin_film}) in terms of a small perturbation $\delta h(\mathbf{x},t) = h(\mathbf{x},t) - h_0$ with $h_0$ being the initial mean thickness value at $t=0$.
Rearanging in terms up to first order in $\delta h$ yields
\begin{equation}\label{eq:linearstability_realspace}
    \partial_t \delta h = \frac{h_0^3}{3\mu}\nabla^2\Big(\Pi(\delta h) - \gamma \nabla^2 \delta h\Big).
\end{equation}
Under Fourier transformation of $\delta h$ 
\begin{equation}\label{eq:fouriertrans}
    \delta h(\mathbf{x},t) = \int(2\pi)^{-2}\delta \tilde{h}(\mathbf{q},t)e^{i\mathbf{q}\cdot\mathbf{x}}\Diff2\mathbf{q},
\end{equation}
and upon inserting this in Eq.~(\ref{eq:linearstability_realspace}) we get
\begin{equation}\label{eq:linearstab_fourier}
    \partial_t \delta \tilde{h}(\mathbf{q},t) = \omega(\mathbf{q})\delta\tilde{h}(\mathbf{q},t),
\end{equation}
with dispersion relation $\omega(q)$, 
\begin{equation}\label{eq:dispersion}
    \omega(\mathbf{q}) = \frac{1-(\frac{q^2}{q_0^2}-1)^2}{t_0}.
\end{equation}
Both, the maximum of the dispersion relation $q_0$ and therefore the characteristic time $t_0$, are specific to the substrate the film is deposited on and can be calculated using~\cite{PhysRevLett.99.114503}
\begin{equation}\label{eq:q0_and_t0}
    q_0^2 = \frac{1}{2}\frac{\partial \Pi(h)}{\partial h}\bigg\rvert_{h=h_0}, \qquad t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}.
\end{equation}

\subsection{Time and spatially varying contact angles $\theta(\mathbf{x},t)$}\label{subsec:theta}
To get a spatially varying contact angle $\theta(\mathbf{x},t)$ we make use of the underlying lattice and assign different values for $\theta$ to different grid points. 
From now on we drop the subscript \textit{eq.} for convenience, such $\theta = \theta_{\text{eq.}}$ unless otherwise stated.

We consider two different patterns and three wavelengths in this manuscript.
First being the sinusoidal pattern with wavelength $\lambda = L/i$ and wavenumbers $i=[1,2,3]$ which is given as
\begin{equation}\label{eq:sinetheta}
    \theta_1(x,y) = \theta_0 + \delta\theta\left[\sin\left(\frac{2\pi x}{\lambda}\right)\sin\left(\frac{2\pi y}{\lambda}\right)\right], 
\end{equation}
with $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$.
The second pattern is a triangle wave which accounts for a different kind of pattern gradient.
To the best of our knowledge contact angles can not be tuned precise enough to match a sinusoidal wave on the scales we are interested in.
Therefore we assume that a triangle wave with a linear gradient is more likely accessable in experiments, thus $\theta_2$ is given as
\begin{equation}\label{eq:triangle_wave}
\begin{split}
    \theta_2(x,y) &= \theta_0 + \\ &\frac{2\delta\theta}{\pi}\left[\sin^{-1}\left(\sin\left(\frac{2\pi x}{\lambda}\right)\right)\sin^{-1}\left(\sin\left(\frac{2\pi y}{\lambda}\right)\right)\right].
    \end{split}
\end{equation}

Both patterns create a contact angle distribution with values within the interval $[10^{\circ}, 30^{\circ}]$ and by definition they are time independent.
To have a time dependency we simply shift the pattern in a periodic manner, such that
\begin{equation}\label{eq:theta_shift}
    \theta(\mathbf{x},t+T) = \theta(\mathbf{x}-\mathbf{1}, t), 
\end{equation}
with $T$ being an arbitrary integer number of lattice Boltzmann time steps.

Due to convenience we attribute a velocity to the pattern which is the shift distance divided by the stepping time
\begin{equation}\label{eq:pattern_speed}
    v_{\theta} = \frac{\sqrt{2}}{T}.
\end{equation}
Since this velocity as well as the choice of $T$ is arbitrary we define a characteristic velocity $v_0$ similar to Eqs.~(\ref{eq:q0_and_t0}) as
\begin{equation}\label{eq:normvel}
    v_0 = \frac{\lambda}{t_0(\theta)},
\end{equation}
where we use the wavelength $\lambda$ as a characteristic distance and $t_0(30^{\circ})$ as a characteristic time scale.
Important to note is that $t_0$ is sensitive to the contact angle, and decreases with increasing angle.

\subsection{Numerical system}
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{Figures/makie_try2.png}
    \caption{Film thickness $h(\mathbf{x},t)$ at a stationary state. 
    The contact angle $\theta(\mathbf{x})$ is shown as contour.}
    \label{fig:handtheta}
\end{figure}
As outlined in Sec.~\ref{subsec:LBM} we simulate the evolution of the thin film with a lattice Boltzmann method.
For all our simulations we use periodic boundary conditions and a square of size $Lx = Ly = L = 512 \Delta x$ with $\Delta x = \Delta t = 1$.
The surface tension as well as the viscosity are constants and set to $\gamma = 0.01$ and $\nu|_{\rho_0=1} = \mu = 1/6$.
Having this value of $\mu$ leaves us with a relaxation time $\tau = 1$.
To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer and a slip parameter $\delta = 1$. 
% With this choice of parameters we are almost able simulate a no-slip boundary condition.
As for our disjoining pressure potential, see Eq.~(\ref{eq:disjoinpressure}), we use the power pair $n=3, m=2$ and $h_{\ast} = 0.07$.
Inserting this quantities into Eq.~(\ref{eq:q0_and_t0}) with $h_0 = 1$ yields $q_0^2(30^{\circ}) \approx 1.68\cdot10^{-2}$ as well as $t_0(30^{\circ}) \approx 177000\Delta t$ which gives us a first glimpse into relevant length and time scales.

The initial condition of all our simulations is a film undulated with sinusoidal wave as
\begin{equation}\label{eq:hinitial}
    h(\mathbf{x},0) = h_0 \left[1 + \epsilon \left(\sin\left(\frac{2\pi x}{L}\right)\sin\left(\frac{2\pi y}{L}\right)\right)\right],
\end{equation}
with $\epsilon = 0.1$ and $h_0 = 1$.

We proceed as follows, first we simulate the dewetting on a given pattern either sinusoidal- or triangular with $\lambda \in \{L, L/2, L/3\}$ without a time dependency in $\theta$.
In Fig.~\ref{fig:handtheta} we show $h(\mathbf{x},t)$ as surface plot and $\theta_2(\mathbf{x})$ as contour plot for $v_{\theta} = 0, \lambda = L/2$ in late stages of dewetting.
Droplets form in regions of small contact angles while the regions of high contact angles dewet.
In subsequent simulations we iterate through $v_{\theta} \in [0.1, 1, 10]v_0$, see Eq.~(\ref{eq:normvel}) .  
% Based on $v_{\theta}$ we can modify the morphology of the dewetting film, in both the intermediate transition regime as well as the late time state.

\section{Results}\label{sec:results}
We now present our results and show how a time dependent wettability or pattern velocity alters the dynamics of a dewetting thin film.
In the following we address the growth of the initial perturbation $\delta h$ and the associated stability of the film.
Therefor we take a look at the rupture times and show that a pattern velocity $v_{\theta} > 0$ does have a stabilizing effect.

After rupture we discuss the morphological transitions the film undergoes.
Therefor the transition from a dewetting film to a droplet state.
This transition is heavily influenced by the concrete choice of $v_{\theta}$ and we observe a strong a alignment with the moving direction of the pattern.

Before we move to the case of $v_{\theta} > 0$ we like to draw a base line with simulation done with a stationary wettability profile.

\subsection{Stationary pattern}\label{subsec:no_vtheta}
Starting with Eq.~(\ref{eq:hinitial}) we have by definition an unstable film that will rupture.
This can be easily seen by the fact that the initial perturbation has the smallest possible wavenumber, $q(t=0)=1/L$.
Taking into account the addition of the wettability gradient we know that fluid will accumulate in regions of high surface energies or small contact angles and flow away from regions of high contact angle.
An illustrative results of this effect is shown in Fig.~\ref{fig:handtheta}, where all droplets are present in a region of small contact angles.

\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{Figures/Clusters_novel_sine_picutres.png}
    \caption{Connected component analysis of the film thickness for different wavelengths $\lambda$ on $\theta_1$ ($\theta_2$ matches qualitatively).
    On the x-axis we display the number of clusters while on the y-axis we show the time normalized by $t_0$.
    Different colors and symbols display different wavelengths $\lambda$.
    The three dashed lines indicate the number of minima the pattern offers.
    Above the data we show the stationary droplet state as a grayscale image of the film thickness $h(\mathbf{x},t_{\infty})$.}
    \label{fig:clusters_v0_sine}
\end{figure}
This is in fact a consistency check of our simulations.
We expect that in case of $v_{\theta} = 0$ all fluid accumulates in droplets inside the contact angle minima.
While there are only two minima for $\lambda = L$ there are eight and eighteen minima for $\lambda = L/2$ and $\lambda = L/3$ respectively.
This can be easily checked with a cluster analysis of the film thickness $h(\mathbf{x},t)$ which we shown in Fig.~\ref{fig:clusters_v0_sine}.
The height $h(\mathbf{x},t_{\infty})$ is displayed as grayscale images and is linked to stationary state.
Theoretically we expect therefor two, eight and eighteen clusters (droplets) which we mark with a vertical dashed line in the respective color.
These values are consistent with our simulations.
Interesting is the fact that the number of clusters converge faster for smaller wavelengths.
To some extend this is expected as the distance between contact angle minima is shorter for smaller wavelengths.

Another quantity that is easily accessible and helps to check the consistency of our simulations is the height of the droplets $h_d$.
Given that all the fluid is collected inside the droplets we can make an assumption of the droplets height based on its volume $V_d$  and contact angle $\theta_d$ 
\begin{equation}\label{eq:height_guess}
    h_d(V_d, \theta_d) = \left(\frac{3V_d(1-\cos(\theta_d)}{\pi(2+\cos(\theta_d))}\right)^{\frac{1}{3}},
\end{equation}
where we implicitly assume that all our droplets have a spherical cap shape.
Since our method does not allow $h=0$ we measure the droplets height as~\cite{PhysRevE.100.033313}
\begin{equation}\label{eq:delta_h_measure}
    \Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\} - \min_{\mathbf{x}}\{h(\mathbf{x},t)\}.
\end{equation}
Counting the minima in $\theta(\mathbf{x})$ for the different wavelengths $\lambda$ allows to calculate the volume of a single droplet, see Fig.~\ref{fig:clusters_v0_sine}.
Since the film dewets regions of high contact angle values we can safely assume that $\theta_d < 20^{\circ}$.
\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{Figures/delta_h_v0_sin_with_dash.pdf}
    \caption{Evolution of $\Delta h$ with time.
    Different symbols and colors represent different pattern wavelengths.
    Dashed lines indicate solutions to the cap height $h_d$ given the volume is split equally among the droplets, see Eq.~(\ref{eq:height_guess}).}
    \label{fig:deltah_v0_sine}
\end{figure}
Measuring the height of the droplets is done in Fig.~\ref{fig:deltah_v0_sine}, where $\Delta h$ is the maximal difference in film thickness which is equivalent to the droplet height $h_d$.
Focusing on the late time behaviour ($t > 10t_0$) we identify plateauing values. 
Since the pattern is a continuous functions, see Eq.~(\ref{eq:sinetheta}), we are left with some degree of freedom for the theoretical $\theta_d$ values. 
As a guide to the eye we introduce three dashed lines.
The dashed lines are calculated using Eq.~(\ref{eq:height_guess}) with $V_d = (V/2, V/8, V/18)$ and $\theta_d = (14^{\circ}, 16^{\circ}, 17^{\circ})$ for $\lambda=(L,L/2,L/3)$ respectively.
The values of $\theta_d$ are chosen to fit with the stationary droplet state of our simulations.
Thus we are fairly confident that our droplet states and thus the simulations capturing the correct physical state.

\subsection{Stability}\label{subsec:time_dep_theta}
\begin{table}[h!]
  \begin{center}
    \begin{tabular}{c|c|c|c}
      \textbf{Pattern} & \textbf{Wavelength} & \textbf{Velocity } & \textbf{Rupture time}\\
      $\theta(\mathbf{x})$ & $\lambda$ & $v/v_0 $ & $\tau_r/t_0$ \\
      \hline
      $\theta_1$ & L & 0 & 1.69\\
      $\theta_1$ & L & 0.1 & 1.72\\
      $\theta_1$ & L & 1 & 2.34\\
      $\theta_1$ & L & 10 & 2.25\\
      $\theta_1$ & L/2 & 0 & 0.39\\
      $\theta_1$ & L/2 & 0.1 & 0.39\\
      $\theta_1$ & L/2 & 1 & 0.59\\
      $\theta_1$ & L/2 & 10 & 0.68\\
      $\theta_1$ & L/3 & 0 & 0.23\\
      $\theta_1$ & L/3 & 0.1 & 0.23\\
      $\theta_1$ & L/3 & 1 & 0.25\\
      $\theta_1$ & L/3 & 10 & 0.45\\
      $\theta_2$ & L & 0 & 1.01\\
      $\theta_2$ & L & 0.1 & 1.18\\
      $\theta_2$ & L & 1 & 1.32\\
      $\theta_2$ & L & 10 & 1.38\\
      $\theta_2$ & L/2 & 0 & 0.48\\
      $\theta_2$ & L/2 & 0.1 & 0.48\\
      $\theta_2$ & L/2 & 1 & 0.7\\
      $\theta_2$ & L/2 & 10 & 0.76\\
      $\theta_2$ & L/3 & 0 & 0.39\\
      $\theta_2$ & L/3 & 0.1 & 0.42\\
      $\theta_2$ & L/3 & 1 & 0.68\\
      $\theta_2$ & L/3 & 10 & 0.76\\
    \end{tabular}
    \caption{Rupture time values for different pattern wavelengths $\lambda$ and different substrate velocities $v_{\theta}$.
    Due to sampling we have a systematic error of 0.03 in the rupture times $\tau_r$.}
    \label{tab:rup_times}
  \end{center}
\end{table}
Taking another look at Fig.~\ref{fig:deltah_v0_sine} reveals that the three wavelengths rupture at different times $\tau_r$.
The most important values for $\tau_r$ can be found in Tab.~\ref{tab:rup_times}.
We observe two features, first being that rupture time $\tau_r$ correlates with the wavelength $\lambda$.
Thus smaller values of $\lambda$ lead to an earlier rupture time.
Which can be explained with a mismatch of length scales.
There is a length scale set by Eq.~(\ref{eq:q0_and_t0}) ($\lambda_0 = 2\pi/q_0$) and one that we introduce with the substrates wavelength $\lambda$.
\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{Figures/RT_with_v10_inkscape.pdf}
    \caption{Distribution of rupture times $\tau_r$ normalized by $t_0$ for different pattern wavelengths $\lambda$.
    The bullets, stars correspond to data from our simulations with $v_{\theta} = 0, 10v_0$.
    Eq.~(\ref{fig:model_rt}) with $D=2$ is displayed by the blue line.
    }
    \label{fig:model_rt}
\end{figure}
Taking this into account we can model the rupture times for the stationary pattern with
\begin{equation}\label{eq:model_tr}
    \frac{\tau_r}{t_0} = \frac{1}{2\pi}\frac{L}{\lambda_0 i^D}\approx \xi\frac{\lambda}{\lambda_0},    
\end{equation}
where $i$ is the wavenumber of the substrates pattern ($\lambda = L/i$) and $D$ is the dimension, thus $D=2$.

Interestingly Eq.~(\ref{eq:model_tr}) is able to describe the data for $v_{\theta} = 0.1v_0$ as the difference of rupture times between these two simulations is negligible, see Tab.~\ref{tab:rup_times}.
Larger $v_{\theta}$ introduce a shift towards longer rupture times, therefore increasing the stability of the film.
Next we look into the affects of the pattern velocity on the structure formation after the first ruptures.

\subsection{Morphologies}\label{subsec:morphme}
Dewetting morphology is a rather broad topic~\cite{becker2003complex, RevModPhys.81.739, PhysRevLett.99.114503, craster2009dynamics, doi:10.1146/annurev-fluid-011212-140734, alizadeh_pahlavan_cueto-felgueroso_hosoi_mckinley_juanes_2018}, however we know that a simple patterned substrate will ultimately yield droplets in region of small contact angle and is likely to rupture in regions of high contact angle, as shown in Fig.~\ref{fig:handtheta}.
This is indeed what we observe for the simulation without pattern velocity ($v_{\theta} = 0$).
\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{Figures/q_2_different_v.pdf}
    \caption{Structure analysis of the pattern $\theta_1,~\lambda=L/2$ using the Minkowski structure metrics (MSM). 
    The dipole metric $q_2$ is plotted for different pattern velocities $v_{\theta}$, displayed with different colors and symbols.
    In grayscale we supply the film thickness that is associated with the $q_2$ values.
    }
    \label{fig:msm_q2}
\end{figure}
Increasing the pattern velocity $v_{\theta} > 0$ does affect the dewetting morphology quite substantially.
Having $v_{\theta} = 0.1 v_0$ we are still observing the formation of droplets, similar to $v_{\theta} = 0$ but instead of being stationary these droplets, depending on their mass, are transported with the contact angle minima.
A somewhat similar behaviour was recently described by Grawitter et al.~\cite{D0SM02082F} where they have used a boundary integral method to simulate a droplet on a moving wettability step.

If the pattern velocity is further increased $v_{\theta} > 0.1 v_0$ we observe a transition towards ligament or rivulet like structures.
These structures are aligned with the update direction of the contact angle field, see Eq.~(\ref{eq:theta_shift}). 
Thus if
\begin{equation*}
    \theta(x,y,t+T) = \theta(x-1, y, t),    
\end{equation*}
ligaments form in x direction.
These ligaments are sensitive to $\lambda$, or more precisely to the contact angle minima since they attract fluid.
To classify the morphological differences in respect to $v_{\theta}$ we make use of the software \textbf{papaya2}~\cite{Schaller2020}.

We use the film thickness as input and compute different morphological metrics.
Among the most interesting ones are the Minkowski structure metrics (MSM) $q_s$.
They can be calculated using~\cite{doi:10.1063/1.4774084}
\begin{equation}
    q_s(K) = \frac{|\psi_s(K)|}{\psi_0(K)}, 
\end{equation}
where $K$ is the body to we extract morphological information from and $\psi_s$ are the irreducible Minskovski tensors (IMT)
\begin{equation}
    \psi_s(K) = \int_{0}^{2\pi}e^{is\phi}\rho_K(\phi) \dd\phi = \sum_{k=-\infty}^{\infty} L_k e^{is\phi_k},
\end{equation}
with $L_k$ being the set of normal vectors of the body $K$.
Within the various $q_s$ we are particularly interested in the dipole one $q_2$. 
With $q_2$ we can clearly distinguish between a droplet and a ligament structure, as we expect a much larger $q_2$ value for the ligament state which is what we see in Fig.~\ref{fig:msm_q2}.
Our simulations with $\lambda = L/2$ show a clear difference between $v_{\theta} \leq 0.1v_0$ and $v_{\theta}>0.1v_0$  in $q_2$.
Interestingly the $q_2$ values for $v_{\theta} = 0.1v_0$ are always larger than those of $v_{\theta} = 0$.
The difference is due to the distortion of the droplets surface area as it is transported with the pattern.

Moving the contact angle field faster creates metastable ligament states.
Metastable because this ligaments will eventually break and separate into droplets, as indicated by the dip in $q_2$ for $v_{\theta} = 1, 10 v_0$ in Fig.~\ref{fig:msm_q2}.
Nevertheless their formation in the first place can be explained with a few relative simple assumptions.

This problem is associated with two time scales, first being $t_0(\theta)$ the time scale of the dewetting thin film, see Eq.~(\ref{eq:q0_and_t0}).
After the rupture we identify another time scale, the capillary time scale $\tau_{\text{Ca}}$ that drives the rims away from the holes~\cite{Edwardse1600183},
\begin{equation}\label{eq:cap_time}
    \tau_{\text{Ca}} = \frac{9\mu l_0}{\gamma \theta^3(\mathbf{x})},
\end{equation}
where  $l_0$ is a characteristic length scale of the problem, we use $\lambda/2$ as two rims can travel at most $\lambda/2$ before they merge.
Knowing this time scales allows us to introduce characteristic velocities.
One of these two is $v_0$ which is given by Eq.~(\ref{eq:normvel}) and used to normalize $v_{\theta}$.
The other one is set by $\tau_{\text{Ca}}$ and reads
\begin{equation}\label{eq:cap_speed}
    U_{Ca} =\frac{l_0}{\tau_{\text{Ca}}} = \frac{\gamma}{9\mu} \theta^3(\mathbf{x}), 
\end{equation}
where $U_{\text{Ca}}$ is the capillary speed.
The argument is as follows, if $U_{\text{Ca}}$ is large as compared to $v_{\theta}$ than the film retraction is faster than the transport due to the contact angle field and thus droplets from.
Which explains the cases of $v_{\theta} = 0, 0.1v_0$.
On the other hand if $v_{\theta}$ is larger than $U_{\text{Ca}}$ than the retracting film has to little time to create stable droplets and ends up in the metastable ligament state.
Taking a look at our simulations with $v_{\theta} = 1, 10v_0$ we clearly observe ligament states.

In fact the ratio 
\begin{equation}\label{eq:vel_ratio}
    \Gamma = \frac{v_{\theta}}{U_{\text{Ca}}} = \frac{3\lambda \chi h_0^3 q_0^4}{\Theta^3}, 
\end{equation}
where $\Theta = max(\theta(\mathbf{x}))$ and $v_{\theta} = \chi v_0$, allows to classify the dewetting morphology on a time and space dependent substrate pattern as
\begin{equation}\label{eq:classify}
    \Gamma~\begin{cases}
    < 1\qquad ... \text{droplets} \\
    > 1\qquad ... \text{ligaments} 
    \end{cases}.
\end{equation}
This explains in fact what we observe and holds true independent of the wettability gradient, i.e. $\theta_1, \theta_2$.

Therefore by varying either $\chi$ or $\lambda$ for a fixed $\Theta$ one can change the intermediate dewetting morphology according to the equation above.
Interestingly but not surprising is the fact that the fluid transport is larger for $\Gamma > 1$ than for $\Gamma < 1$.

\section{Summary and Conclusions}\label{sec:sum_conclu}
To summarize, we have shown numerical simulations of a dewetting thin film on a patterned substrate.
Generating baseline experiments with time independent patterns of different wavelengths, we than move to a time and spatial varying contact angle field $\theta(\mathbf{x},t)$.

In the next step we define a characteristic pattern velocity $v_0$ which simply is the patterns wavelength $\lambda$ over a characteristic time $t_0$, see Eqs.~(\ref{eq:q0_and_t0}, \ref{eq:normvel}).
Experiments with increasing pattern velocity $v_{\theta} = \chi v_0$ with $\chi\in\{0,10\}$ show a pronounced change in dewetting morphology from well defined droplets to ligaments.
With the single balance between pattern velocity $v_{\theta}$ and capillary velocity $U_{\text{Ca}}$ we can identify a parameter that separates the two morphological regimes.

An outlook to this work is to study the transition from ligaments to droplets in a more detailed manner.
Looking at the edges $\Gamma =1$ with more suitable tools like e.g. bifurcation diagramms will likely be subject to future work.
But also the update scheme of the contact angle field allows for many applications.
Among them is the avenue of switchable substrates, where e.g. switching of $\theta(\mathbf{x})$ due to a light source is in the scope of this method.

\begin{acknowledgements}
The authors acknowledge financial support by the Deutsche Forschungsgemeinschaft (DFG) within the priority program SPP2171 ``Dynamic Wetting of Flexible, Adaptive, and Switchable Substrates'', project HA-4382/11.
\end{acknowledgements}

%\appendix
%\section{}\label{app:one}

\bibliography{Ref}

\end{document}

### 2022-10-03 main_AS.tex
Update

@@
-%%
-%\listfiles
-%\documentclass[%
-% reprint,%
-%secnumarabic,%
-% amssymb, amsmath,%
-% aip,cha,%
-%groupedaddress,%
-%frontmatterverbose,
-%]{revtex4-1}
-\documentclass[twocolumn,amsmath,amssymb,showpacs,pre,nofootinbib,superscriptaddress]{revtex4-1} %,preprint,jcp
+\documentclass[twocolumn,amsmath,amssymb,showpacs,prl,superscriptaddress]{revtex4-1} %,preprint,jcp,nofootinbib
 
 \bibstyle{apsrev}
-%\usepackage{docs}%
 \usepackage{bm}%
 \usepackage{graphicx}
-%\usepackage{subcaption}
 \usepackage{tikz}
@@
 \usepackage[colorlinks=true,linkcolor=blue]{hyperref}%
-%\nofiles
 \expandafter\ifx\csname package@font\endcsname\relax\else
@@
 \definecolor{pyred}{HTML}{D62728}
+\definecolor{jlblue}{rgb}{0.0,0.6056031611752245,0.9786801175696073}
+\definecolor{jlorange}{rgb}{0.8888735002725198,0.43564919034818994,0.2781229361419438}
 \newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}
 
-%\title{Switchable substrates and thin film morphologies}
-%\title{Time varying contact angles and thin film morphologies}
-\title{Morphologies of thin liquid film dewetting on time-dependent patterned substrates}
+\title{Controlling the dewetting morphologies of thin liquid films by switchable substrates}
 
 \author{S. Zitz}
-\email{s.zitz@fz-juelich.de}
+\email{zitz@ruc.dk}
+ %Changed order: Work done while Stefan worked in Nuremberg
  \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
-  Forschungszentrum J\"ulich,\\
+  Forschungszentrum J\"ulich,
   F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
   \affiliation{Department of Chemical and Biological Engineering, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
+  \affiliation{IMFUFA, Department of Science and Environment,\\ 
+Roskilde University, Postbox 260, DK-4000 Roskilde, Denmark}%
 \author{A. Scagliarini}%
 \email{andrea.scagliarini@cnr.it}
- \affiliation{Institute for Applied Mathematics "M. Picone" (IAC), \\
+ \affiliation{Institute for Applied Mathematics "M. Picone" (IAC), 
 Consiglio Nazionale delle Ricerche (CNR),\\
@@
  \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
-  Forschungszentrum J\"ulich,\\
+  Forschungszentrum J\"ulich,
   F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
@@
 \begin{abstract}
-We study the dynamics of an unstable thin film deposited on a ``switchable`` substrate.
-The substrate is itself subject to a time varying function that affects the contact angle $\theta(\mathbf{x},t)$.
-We show that in the case of vanishing time dependence the patterning drives the dewetting morphology and that all fluid is drained into regions of small $\theta(\mathbf{x})$ and thus form droplets.
-Interestingly enough with a time varying component to the pattern we are able to stop the droplet fragmentation and create pronounced ligament or rivulet morphologies.
-We show that both the numbers of semistable rivulets as well as number of stable droplets can be controlled with the pattern wavelength ($\lambda$).
-Finally we explain why the ligaments appear with a simple balance between capillary retraction and dynamic wetting.   
+\noindent Switchable and adaptive substrates emerged as valuable tools for the control of wetting and actuation of droplet motion.
+Here we report a computational study of the dynamics of an unstable thin liquid film deposited on a switchable substrate, modelled with a space and time varying contact angle.
+With a static pattern, all the fluid is drained into droplets located around contact angle minima, whereas for a sufficiently large rate of wettability variation a state consisting of metastable rivulets is observed.
+A criterion discriminating whether rivulets can be observed or not is identified in terms of a single dimensionless parameter.
+Finally, we show and explain theoretically how the film rupture times, droplet shape and rivulet life time depend on the pattern wavelength and speed.  
 \end{abstract}
@@
 \newcommand{\ts}{\textsuperscript}
-%\tableofcontents
 
-\section{Introduction}\label{sec:intro}
-Wet surfaces and droplets are part of our every-day experience and of several industrial processes,
-such as coating, tribology, painting and printing,
-to name but a few~\cite{gross1980fluid,szeri2010fluid,DERYCK1998278,doi:10.1146/annurev.fluid.31.1.347,DASILVASOBRINHO19991204,singh2010inkjet,jo2009evaluation}. 
-%Being it the droplets condensed in the higher atmosphere and return to the ground as rain or the oil film in a pan during cooking. 
-%While this two examples are illustrative there are many industrial applications that build on thin film flows. 
-%Among them is the wetting of surfaces to reduce wear or put simply, lubrication~\cite{ReynoldsLubr, gross1980fluid, szeri2010fluid}. 
-%Coating surfaces to change their physio-chemical properties is another widely used application~\cite{DERYCK1998278, doi:10.1146/annurev.fluid.31.1.347, DASILVASOBRINHO19991204}.
-%On the other hand the need to control droplets is at the heart of many printing processes~\cite{singh2010inkjet, jo2009evaluation}.
-Moreover, the continuously growing technological interest for lab-on-a-chip devices~\cite{C6LC00387G,Focke} as well as for
-printable electronics~\cite{Kim_2005, Luechinger_2008}, whose efficiency relies crucially on a precise control of material deposition upon dewetting of
-liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e. it is not inert but responds
-dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{doi:10.1126/science.288.5471.1624, doi:10.1246/bcsj.20180076, xin2010reversibly}.
-%More recently lab-on-a-chip devices have acquired much attention, to work they require precise droplet or
-%film flow on small scales~\cite{stone2004engineering, darhuber2010planar}.
-%One possibility for precise control is the local wettability.
-%Light switchable substrates~\cite{WANG200718}, therefore substrates that change their wettability based on light illumination, open up a new% way to program this devices.
+\noindent {\it Introduction}.
+Wet surfaces and droplets are part of our every-day experience and of numerous industrial processes including coating, tribology, painting and printing, to name but a few~\cite{gross1980fluid,szeri2010fluid,DERYCK1998278,doi:10.1146/annurev.fluid.31.1.347,DASILVASOBRINHO19991204,singh2010inkjet,jo2009evaluation,Wijshoff2010}.
+Moreover, the continuously growing technological interest for lab-on-a-chip devices~\cite{C6LC00387G,Focke} as well as for printable electronics~\cite{Kim_2005, Luechinger_2008} and printable photovoltaics~\cite{Brabec2001,RH20}, whose efficiency relies crucially on a precise control of material deposition upon (de-)wetting of liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e.
+it is not inert but responds dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{ButtEtAl_Langmuir2018,GuoGuo_RSCAdv2016}.
+Several realizations of switchable and adaptive substrates have been proposed~\cite{LiuChem2005, XiCSR2010}, involving smart materials such as polymer brushes~\cite{CohenStuartEtAl_NatMat2010,ayres2007stimuli,YongEtAl_Mat2018}, thermal-responsive hydrogels~\cite{ChenEtAl_SM2010}, light-responsive molecules and microstructures~\cite{IchimuraEtAl_Science2000,DelormeEtAl_Langmuir2005,OscuratoEtAl_AAMI2017} or processes such as electrowetting~\cite{MugeleEtAl_JPCM2005}.
+Probably the simplest, yet non-trivial, modelling of a switchable substrate can be realized by a space and time dependent wettability pattern~\cite{GrawitterStark1}. 
+While a consistent body of theoretical/computational work has been devoted to processes on static heterogeneous substrates, though, the time dependent case is still almost unexplored, with few relevant exceptions focusing on single droplet spreading and sliding~\cite{GrawitterStark1,GrawitterStark2,ThieleHartmann} or limited to an analysis of the linear regime~\cite{suman2006dynamics}.
 
-Spatial variations of the contact angle are long known and used in previous studies~\cite{PhysRevE.91.023013, Sommers_2006}, time dependency of the contact angle on the other hand is not as well studied~\cite{suman2006dynamics}.
-Switchable substrates are in fact a realization of a time dependent contact angle. 
-Using both spatial as well as time dependent variations of the local contact angle leads to interesting phenomena, for example continuously sliding droplets as shown by Grawitter and Stark~\cite{D0SM02082F}.
-Our approach builds on the same idea but our method, see ref~\cite{PhysRevE.100.033313}, allows us to go beyond a single droplet and to simulate the full dewetting transition starting with a thin film and ending with droplets.
-We use this to show that a time dependent component of the contact angle alter the dewetting morphology quite substantially.
-In the limit of a small time dependency or slow evolution of the contact angle field we qualitatively reproduce the results from Grawitter and Stark, such sliding droplets.
-Speeding up the evolution of the contact angle field introduces a morphological transition, which to the understanding of the authors have not been looked at so far.
-Therefore the purpose of this manuscript is to increase the understanding of the dewetting process.
-Highlighting the fact that a temporal and spatial variation of the contact angle leads to novel morphological structures.
-Thus our finding offer a simple way to understand how time periodic switchable substrates could generate novel dewetting morphologies.
-
-The manuscript is outlined as follows:
-In section~\ref{sec:method} we discuss the numerical method we use to generate our results and introduce relevant scales.
-We then present our results in section~\ref{sec:results}, where we discuss the impact of the time dependency on the contact angle and show that both stability as well as dewetting morphology is affected.
-A short summary of our finding and a closing conclusion with outlook is given in section~\ref{sec:sum_conclu}.
-
-\section{Numerical model and simulation details}\label{sec:method}
-In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{ReynoldsLubr,RevModPhys.69.931,RevModPhys.81.1131} 
+In this Letter, we study, by means of numerical simulations, the full dewetting dynamics of a thin liquid film deposited over a substrate with a time varying wettability pattern, from rupture to the long time morphology.
+We identify two regimes where the rupture times grow with the pattern wavelength either linearly (on the static pattern) or attain a constant value (in the time dependent case), for short wavelengths, and approach a quadratic law as the wavelength increases.
+These observations are then explained theoretically.
+We show that, by tuning the rate of change (the ``speed'') of the underlying pattern, it is possible to control, to some extent, the dewetting morphology.
+In particular, for large enough pattern speeds, we detect a metastable state, where the film retracts into metastable rivulets, eventually breaking up into multiple droplets.
+We introduce a control parameter that discriminates whether rivulets or just droplets (as in the static situation)
+can be observed and we propose a phenomenological argument that justifies the logarithmic dependence of the rivulets life time on the pattern speed.\\
+\\
+\noindent {\it Method.} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{ReynoldsLubr,RevModPhys.69.931,RevModPhys.81.1131}
 \begin{equation}\label{eq:thinfilm}
@@
 \end{equation}
-by means of a recently developed method, based on a lattice Boltzmann
-scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801}.
-Eq.~(\ref{eq:thinfilm}) describes, in a lubrication approximation spirit, the evolution of the height field (film thickness) $h(\mathbf{x},t)$, denoting the location of the liquid/air interface;
-the mobility function $M_{\delta}(h)$ depends on the velocity boundary condition at the substrate, parameterized by an effective slip length $\delta$, and reads
+by means of a recently developed method, based on a lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801}.
+Eq.~(\ref{eq:thinfilm}) describes, in a lubrication approximation spirit, the evolution of the height field (film thickness) $h(\mathbf{x},t)$, denoting the location of the liquid/air interface.
+The mobility function
 \begin{equation}\label{eq:mobility}
-  M_{\delta}(h) = \frac{2h^3 + 6\delta h^2 + 3\delta^2h}{6\mu},
+  M_{\delta}(h) = \frac{2h^3 + 6\,\delta\, h^2 + 3\delta^2h}{6\mu},
 \end{equation}  
-(which, for $\delta \rightarrow 0$, reduces to the no-slip form $h^3/(3\mu)$),
-where $\mu$ is the fluid dynamic viscosity.
-The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$. The disjoining pressure
-$\Pi$ can be seen as (minus) the derivative, with respect to the film thickness, of an effective interfacial potential; as such, it contains the information on the
-liquid/solid and solid/gas interactions and, hence, on the wettability, which is parameterized in terms of the contact angle, $\theta$ \cite{RevModPhys.81.739, SCHWARTZ1998173}. 
-The expression adopted for $\Pi$ is:
+depends on the velocity boundary condition at the substrate, parameterized by an effective slip length $\delta$
+(for $\delta \rightarrow 0$ it reduces to the no-slip form $h^3/(3\mu)$). Here, $\mu$ is the fluid dynamic viscosity.
+The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$, 
+\textcolor{red}{where $\gamma$ is the surface tension}.
+The disjoining pressure $\Pi$ can be seen as (minus) the derivative, with respect to the film thickness, of an effective interfacial potential.
+As such, it contains the information on the liquid/solid and solid/gas interactions and, hence, on the wettability, which is parameterized in terms of the contact angle $\theta$~\cite{RevModPhys.81.739, SCHWARTZ1998173}.
+The expression adopted for $\Pi$ is
 \begin{equation}\label{eq:disjoinpressure}
-  \Pi(h,\theta) = \frac{2\gamma}{h_{\ast}}(1-\cos(\theta(\mathbf{x},t)))\left[\left(\frac{h_{\ast}}{h}\right)^3 - \left(\frac{h_{\ast}}{h}\right)^2\right],
+\textcolor{red}{\Pi(h,\theta) = \frac{2\gamma}{h_{\ast}}(1-\cos(\theta(\mathbf{x},t)))
+  f\left(\frac{h}{h_{\ast}}\right),}
 \end{equation}
-where $h_{\ast}$, the height at which the disjoining pressure vanishes, sets the precursor layer thickness.
-The time variability of the patterned substrate, then, enters the model precisely through the
-disjoining pressure, by making the contact angle space and time dependent, $\theta = \theta(\mathbf{x},t)$.
-In particular, the following sinusoidal form is employed:
+\textcolor{red}{where $f(\xi)=\xi^{-3} - \xi^{-2}$} and $h_{\ast}$ is the height at which the disjoining pressure vanishes and sets the precursor layer thickness.
+The time variation of the patterned substrate then enters the model precisely through the disjoining pressure, by making the contact angle space and time dependent, i.e.
+$\theta = \theta(\mathbf{x},t)$.
+In particular, we employ the sinusoidal form
 \begin{equation}\label{eq:sinetheta}
-    \theta(\mathbf{x},t) = \theta_0 + \delta\theta\left[\sin\left(\frac{2\pi (x+v_{\theta x}t)}{\lambda}\right)\sin\left(\frac{2\pi (y+v_{\theta y}t)}{\lambda}\right)\right], 
+   \!\! \theta(\mathbf{x},t) = \theta_0 + \delta\theta\left[\sin\left(q_{\theta} (x+v_{\theta x}t)\right)\sin\left(q_{\theta}(y+v_{\theta y}t)\right)\right],\! 
 \end{equation}
-with $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$; the pattern evolves in time as a plane wave.
-We choose to fix the velocity direction to one diagonal, i.e.
-$\mathbf{v}_{\theta} = (v_{\theta x},v_{\theta y}) = v_{\theta}(1/\sqrt{2},1/\sqrt{2})$.
-Notice that, since a typical velocity is such that $v_{\theta} \Delta t \ll \Delta x$ (in one time step the wave would travel
-a distance much smaller than a lattice spacing), the time update must be interpreted in an integer part sense, that is
-the pattern is shifted by one $\Delta x$ every $1/v_{\theta x}$ time steps (and equivalently in the $y$-direction).
-%The second pattern is a triangle wave which accounts for a different kind of pattern gradient.
-%To the best of our knowledge contact angles can not be tuned precise enough to match a sinusoidal wave on %the scales we are interested in.
-%Therefore we assume that a triangle wave with a linear gradient is more likely accessable in experiments, %thus $\theta_2$ is given as
-%\begin{equation}\label{eq:triangle_wave}
-%\begin{split}
-%    \theta_2(x,y) &= \theta_0 + \\ &\frac{2\delta\theta}{\pi}\left[\sin^{-1}\left(\sin\left(\frac{2\pi x}{%\lambda}\right)\right)\sin^{-1}\left(\sin\left(\frac{2\pi y}{\lambda}\right)\right)\right].
-%    \end{split}
-%\end{equation}
-%Both patterns create a contact angle distribution with values within the interval $[10^{\circ}, 30^{\circ}%]$ and by definition they are time independent.
-%To have a time dependency we simply shift the pattern in a periodic manner, such that
-%\begin{equation}\label{eq:theta_shift}
-%    \theta(\mathbf{x},t+T) = \theta(\mathbf{x}-\mathbf{1}, t), 
-%\end{equation}
-%with $T$ being an arbitrary integer number of lattice Boltzmann time steps.
-Hereafter, lengths will be meant expressed in units of the
-mean film height, $h_0$ (which is constant in time, due to mass conservation) and pattern wavelength $\lambda$, whereas the charactestic
-time scale is set by:
-\begin{equation}\label{eq:t0}
-t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4},
-\end{equation}
-where $\gamma$ is the surface tension
-(the numerical values, in lattice Boltzmann units, are set to $\gamma = 0.01$ and $\mu=0.167$).
-The thus defined $t_0$ is the inverse growth rate of the most unstable mode (of wavenumber $q_0$)
-of a spinodally dewetting film (on a homogeneous substrate) \cite{Mecke_2005,PhysRevE.100.023108};
-in our heterogeneous case, we define the wavenumber $q_0$ as 
-$q_0^2=\frac{1}{2\gamma}\frac{\partial \Pi(h,\theta)}{\partial h}(h=h_0,\theta=\theta_0+\delta \theta)$.
-At this point, a reference velocity can also be constructed as
-\begin{equation}\label{eq:v0}
-v_0 \equiv \frac{\lambda}{t_0}.
-\end{equation}
-%Due to convenience we attribute a velocity to the pattern which is the shift distance divided by the stepp%ing time
-%\begin{equation}\label{eq:pattern_speed}
-%    v_{\theta} = \frac{\sqrt{2}}{T}.
-%\end{equation}
-%Since this velocity as well as the choice of $T$ is arbitrary we define a characteristic velocity $v_0$ si%milar to Eqs.~(\ref{eq:q0_and_t0}) as
-%\begin{equation}\label{eq:normvel}
-%    v_0 = \frac{\lambda}{t_0(\theta)},
-%\end{equation}
-%where we use the wavelength $\lambda$ as a characteristic distance and $t_0(30^{\circ})$ as a characterist%ic time scale.
-%Important to note is that $t_0$ is sensitive to the contact angle, and decreases with increasing angle.
+where $q_{\theta} = 2\pi/\lambda$, i.e.~the pattern evolves in time as a plane wave. 
+We fix the velocity direction to one diagonal, namely $\mathbf{v}_{\theta} = (v_{\theta x},v_{\theta y}) = v_{\theta}(1/\sqrt{2},-1/\sqrt{2})$ (we will return later on the importance of this choice), and we set $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$~\footnote{Notice that, since a typical velocity is such that $v_{\theta} \Delta t \ll \Delta x$ (in one LB time step $\Delta t$ the wave would travel a distance much smaller than a lattice spacing $\Delta x$),
+the time update needs to be interpreted in an integer part sense, that is the pattern is shifted by one $\Delta x$ every $1/v_{\theta x}$ time steps (and equivalently in the $y$-direction).}.
+\textcolor{red}{Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film \cite{Mecke_2005}.
+On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads
+$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$ \cite{Mecke_2005,PhysRevE.100.023108}. In our patterned case, we define 
+$q_0^2=h_{\ast}^{-2}(1-\cos\Theta)f^{\prime}(h_0/h_{\ast})$, where 
+$\Theta = \max_{\mathbf{x}} \theta(\mathbf{x}) = \theta_0 + \delta \theta$.
+A velocity scale then also naturally arises, namely $v_0 = h_0/t_0$.}
 \begin{figure}
     \centering
-    \includegraphics[width=0.45\textwidth]{Figures/No_vel_render.png}
-    \caption{Rendered image of the film on the stationary contact angle pattern at late times ($t > 20t_0$).
-    The contact angle field is color coded, areas where the angle is small (high) are shown in blue (yellow).}
+    \includegraphics[width=0.45\textwidth]{Figures/No_vel_render2.png}
+    \caption{Stationary film thickness field ($t>20t_0$) showing the formation of droplets. The color map indicates the contact angle pattern 
+    (Eq.~(\ref{eq:sinetheta}) with $v_{\theta}=0$), with lower (higher) values coded in light blue (yellow).
+    }
     \label{fig:handtheta}
 \end{figure}
-%As outlined in Sec.~\ref{subsec:LBM} we simulate the evolution of the thin film with a lattice Boltzmann m%ethod.
-All our simulations are run on a bi-periodic square domain of size $L \times L$ with $L = 512$.
-%Having this value of $\mu$ leaves us with a relaxation time $\tau = 1$.
-To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer thickness
-$h_{\ast}=0.07$ and and a slip length $\delta = 1$
-(see Eqs.~(\ref{eq:mobility})-(\ref{eq:disjoinpressure})). 
-% With this choice of parameters we are almost able simulate a no-slip boundary condition.
-%As for our disjoining pressure potential, see Eq.~(\ref{eq:disjoinpressure}), we use the power pair
-%$n=3, m=2$ and $h_{\ast} = 0.07$.
-%Inserting this quantities into Eq.~(\ref{eq:q0_and_t0}) with $h_0 = 1$ yields $q_0^2(30^{\circ}) \approx 1%.68\cdot10^{-2}$ as well as $t_0(30^{\circ}) \approx 177000\Delta t$ which gives us a first glimpse into r%elevant length and time scales.
-The liquid film is initiliazed with a height field slightly perturbed around the mean valued $h_0=1$, i.e.:
+All our simulations are run on a bi-periodic square domain of size $L \times L$ with 
+$L/h_0 = 512 $.
+To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$ (see Eqs.~(\ref{eq:mobility})-(\ref{eq:disjoinpressure})).
+The liquid film is initialized with a height field slightly perturbed around the mean value $h_0$, i.e.
 \begin{equation}\label{eq:hinitial}
@@
 \end{equation}
-Various wavelengths, in the range $\lambda \in [L/9, L]$, and velocities, $v_{\theta} \in [0.1, 10]v_0$,
-are considered for the wettability pattern, Eq.~(\ref{eq:sinetheta}). 
-%We proceed as follows, first we simulate the dewetting on a given pattern either sinusoidal- or triangular% with  without a time dependency in $\theta$.
+Various wavelengths, in the range \textcolor{red}{$\lambda/h_0 \in [50, 512]$}, and velocities, $v_{\theta}/v_0 \in [0, 5 \times 10^3]$, are considered for the wettability pattern, Eq.~(\ref{eq:sinetheta}).
+\textcolor{red}{A comment on the limits of the wavelengths range is noteworthy at this point.
+The wavelengths are obviously limited from above by the size of the computational domain.
+On the other extreme, small values are prevented by
+the limited capability of accomodating droplets of too large height, $h_d$, on too
+small patches, since the local contact angle would possibly exceed the largest substrate contact angle, such that the patterning 
+itself would loose meaning. In fact, if the patches were much smaller than 
+the intrinsic size the droplets would have when formed on a homogeneous substrate
+(fixed by the spinodal characteristic length, $\lambda_s$), the confinement by the 
+pattern would not be effective~\cite{KarguptaPRL2001}
+Besides, the very same lubrication approximation 
+requires small contact angles. 
+We keep, therefore, $\lambda \geq \lambda_s \approx 70 h_0$.}
+%\textcolor{red}{Hereafter, we will omit 
+%to write explicitly the division by $h_0$, $t_0$ and $v_0$ that will be, then, 
+%surmised, unless else specified.}
+\textcolor{red}{The Reynolds number, $Re$, is of the order of $Re \sim 10^{-2}$ on the 
+static substrate and never exceeds the value $\approx 0.2$ in the time dependent case.}
 In Fig.~\ref{fig:handtheta} we show $h(\mathbf{x},t)$ (droplets) and $\theta(\mathbf{x})$ (color coded) for $v_{\theta} = 0$ (i.e.
-the stationary case) and $\lambda = L/2$, in late stages of dewetting.
-As expected, droplets form in regions of small contact angles (blue) while the regions of high contact angles (yellow) dewet.
-%In subsequent simulations we iterate through , see Eq.~(\ref{eq:normvel}) .  
-% Based on $v_{\theta}$ we can modify the morphology of the dewetting film, in both the intermediate transition regime as well as the late time state.
+the static case) and \textcolor{red}{$\lambda = 256 h_0$}\footnote{For a better visualization we take twice the domain length $L$ and periodically continue the image.}, in late stages of dewetting.
+As expected, droplets form in regions of small contact angles (blue) while the regions of high contact angles (yellow) dewet.\\
 
-\section{Results}\label{sec:results}
-
-\subsection{Rupture times}\label{subsec:rupture}
-
-We want to study how the rupture times depend on the parameters characterising the wettability pattern, namely the contact angle
-wavelength, $\lambda$, and wave speed, $v_{\theta}$.
-The film rupture time, $\tau_r$, is defined as the least $t$ such that $\{\exists\mathbf{x}: h(\mathbf{x},\tau_r)=h_{\ast}\}$ (that is, when the
-free surface touches the substrate).
-%Taking another look at Fig.~\ref{fig:deltah_v0_sine} reveals that the three wavelengths rupture at different times $\tau_r$.
-%The most important values for $\tau_r$ can be found in Tab.~\ref{tab:rup_times}.
-In Fig.~\ref{fig:model_rt} we report the rupture times, as a function of the wavelength, for the stationary pattern ($v_{\theta}=0$) and
-for $v_{\theta}=10 v_0$. We observe that, in both cases, $\tau_r$ grows $\lambda$.
-%Thus smaller values of $\lambda$ lead to an earlier rupture time.
-%Which can be explained with a mismatch of length scales.
-This fact can be explained, qualitatively, as follows.
-There is a length scale set by Eq.~(\ref{eq:disjoinpressure}) ($\lambda_0 = 2\pi/q_0$) and one that we introduce with the substrates wavelength $\lambda$.
+\noindent {\it Results.} We start our analysis studying, first, how the rupture times depend on the parameters characterising the wettability pattern, namely the wavelength of the contact angle variation, $\lambda$, and wave speed, $v_{\theta}$.
+The film rupture time, $\tau_r$, is defined as the least $t$ such that $h(\mathbf{x},\tau_r)=h_{\ast}$ (that is, when the free surface "touches" the substrate).
+In Fig.~\ref{fig:model_rt} we report the rupture times, as a function of the wavelength, for stationary ($v_{\theta}=0$) and time-dependent ($v_{\theta}=10 v_0$) patterns. 
+It is conveyed that, overall, rupture occurs earlier on the static substrate, suggesting that the time variation tends to stabilize the film, in agreement with linear stability analysis results~\cite{suman2006dynamics}.
+In the stationary case we observe that $\tau_r$ grows linearly with $\lambda$, for short wavelengths, and quadratically for longer $\lambda$.
+Also for $v_{\theta}=10$ the rupture times tend to approach the $\lambda^2$ scaling at large $\lambda$, whereas they tend to saturate to a constant value for short wavelengths.
+These facts can be explained, qualitatively, as follows. 
+Let us first notice that the film rupture consists of two processes, each with its associated time scale: the growth of unstable interface perturbation with rate $t_{\theta}^{-1}$, and the retraction of liquid from more hydrophobic regions, occurring in a characteristic time $t_R$.
+For longer pattern wavelengths, the dewetting instability amplifies more slowly, such that it is reasonable to take it as the process determining the rupture time, $\tau_r \sim t_{\theta}$.
+In this case, from the linearized thin-film equation (in one spatial dimension, for simplicity), obtained setting $h=h_0 + \delta h$ with $\delta h \ll h_0$, 
+we can easily see that the exponential growth of the height perturbation is affected by the wettability pattern (variable contact angle) in such a way that $\partial_t (\delta h) \propto (\partial_x^2 (\partial_h\Pi(h_0))) \delta h$.
+Therefore, since the characteristic time $t_{\theta}$ can be estimated dimensionally as $t_{\theta} \sim \delta h/\dot{(\delta h)}$, the rupture time should go as
+\begin{multline}\label{eq:taur_l2}
+    \tau_r \sim t_{\theta} \sim  \delta h/\dot{(\delta h)} \propto \frac{3\mu}{h_0^3}(\partial_x^2 (\partial_h\Pi (h_0)))^{-1} \sim \\
+    t_0 \left(\frac{q_{\theta}}{q_0}\right)^{-2} \propto t_0 q_0^2 \lambda^2.
+\end{multline}
+Conversely, for fast growths ($t_{\theta} \ll t_R$), retraction dominates and fixes the time scale, $\tau_r \sim t_R$. 
+The latter is related to the time the liquid takes to flow out of regions of high contact angle, whose size is $\sim \lambda$. Hence we have 
+\begin{equation}\label{eq:taur_l1}
+ \tau_r \sim \tau_R \propto \frac{1}{U_{\Theta}}\lambda,
+\end{equation}
+where $U_{\Theta}$ is the retraction speed $U_{\Theta} = \frac{\gamma \Theta^3}{9\mu}$~\cite{Edwards2016}, with $\Theta = \max_{\mathbf{x}}\theta(\mathbf{x})$.
 \begin{figure}
     \centering
-    \includegraphics[width=0.48\textwidth]{Figures/Rupture_times_over_lambda_with_powerlaw.pdf}
-    \caption{Distribution of rupture times $\tau_r$ normalized by $t_0$ for different pattern wavelengths $\lambda$.
-    The bullets, stars correspond to data from our simulations with $v_{\theta} = 0, 10v_0$.
-    The full blue and dashed orange line are generated using $f(x) = a~x^{\alpha}$, with $a=0.6,1.5$ respectively.
+    \includegraphics[width=0.45\textwidth]{Figures/low_lambda_rupture.pdf}
+    %\includegraphics[width=0.45\textwidth]{Figures/Rupture_times_Gamma.pdf}
+    % \includegraphics[width=0.45\textwidth]{Figures/rupture_times_all_lambda_v0v10.pdf}
+    \caption{Rupture times, $\tau_r$, as a function of the pattern wavelength $\lambda$, for $v_{\theta}=0$ (\textcolor{jlblue}{$\bullet$}) and $v_{\theta}=10 v_0$ 
+    (\textcolor{jlorange}{$\star$}).
+    The continuous and dashed lines indicate the linear, $\sim \lambda$, and quadratic, $\sim \lambda^2$, scaling laws, respectively.
         }
@@
 \end{figure}
-Taking this into account we can model the rupture times for the stationary pattern with
-\begin{equation}\label{eq:model_tr}
-    \frac{\tau_r}{t_0} = \frac{1}{2\pi}\frac{L}{\lambda_0 i^D}\approx \xi\frac{\lambda}{\lambda_0},    
-\end{equation}
-where $i$ is the wavenumber of the substrates pattern ($\lambda = L/i$) and $D$ is the dimension, thus $D=2$.
+Interestingly, this phenomenology is qualitatively supported also by the time-dependent case, $v_{\theta} = 10v_0$ (reported in Fig.~\ref{fig:model_rt} as orange stars), \textcolor{red}{but for a smaller prefactor in the linear regime, due to the fact that the pattern wave velocity is large enough to dominate over $U_{\Theta}$ and sets the retraction speed, such that the characteristic time becomes $\tau_R \sim \lambda/v_{\theta}$.}
+%\begin{equation}\label{eq:taur_saturation}
+% \tau_r \sim \tau_R \sim \frac{\lambda}{v_{\theta}} \propto %\frac{\lambda}{v_0} = \lambda\frac{t_0}{\lambda} \sim \mbox{const}.
+%\end{equation}
 
-Interestingly, the behaviour of $\tau_r$ vs $\lambda$ remains qualitatively the same also for the
-time-dependent case, $v_{\theta} = 0.1v_0$, reported in Fig.~\ref{fig:model_rt} as blue squares.
-%as the difference of rupture times between these two simulations is negligible,
-%see Tab.~\ref{tab:rup_times}.
-Nevertheless, larger $v_{\theta}$ introduce a shift towards longer rupture times, i.e. fast wettability
-waves tend to stabilise the film.
-%Next we look into the affects of the pattern velocity on the structure formation after the first ruptures.
-
-\subsection{Morphologies}\label{subsec:morpho}
-We now look into the long time dynamics, focusing, in particular on the characterisation of the
-dewetting morphologies and how they are affected by the speed of the wettability wave.
-
-\subsubsection{Stationary pattern}\label{subsec:no_vtheta}
-%Starting with Eq.~(\ref{eq:hinitial}) we have by definition an unstable film that will rupture.
-%This can be easily seen by the fact that the initial perturbation has the smallest possible wavenumber, $q(t=0)=1/L$.
-%Taking into account the addition of the wettability gradient we know that fluid will accumulate in regions of high surface energies or
-%small contact angles and flow away from regions of high contact angle.
-%An illustrative results of this effect is shown in Fig.~\ref{fig:handtheta}, where all droplets are present in a region of
-%small contact angles.
+We now focus on the long time dynamics and in particular on the characterization of the dewetting morphologies and how they are affected by the speed of the wettability wave.
 \begin{figure}
     \centering
-    \includegraphics[width=0.4\textwidth]{Figures/Clusters_novel_sine_picutres.png}
-    \includegraphics[width=0.45\textwidth]{Figures/delta_h_v0_sin_with_dash.pdf}
-    \caption{TOP PANEL. Number of droplets, $N$, as a function of time, generated during the dewetting process on the patterned substrate
-      given by Eq.~(\ref{eq:sinetheta}), with $v_{\theta}=0$ and $\lambda=L$ (blue circles), $\lambda=L/2$ (orange squares) and
-      $\lambda=L/3$ (green stars). The three horizontal dashed lines indicate the number of minima of (\ref{eq:sinetheta}),
-      which is $2\left(\frac{L}{\lambda}\right)^2$. In the insets, we show shanpshots of the stationary droplet state as a grayscale image
-      of the film thickness $h(\mathbf{x},t_{\infty})$.
-      BOTTOM PANEL. Evolution of $\Delta h$ with time.
-      Different symbols and colors represent different pattern wavelengths.
-      Dashed lines indicate solutions to the cap height $h_d$ given the volume is split equally among the droplets,
-      see Eq.~(\ref{eq:height_guess}).}
+    \includegraphics[width=0.45\textwidth]{Figures/static_pattern_test.pdf}
+    \caption{MAIN PANEL: Time evolution of the height fluctuations, $\Delta h(t)$, during the dewetting process on the patterned substrate given by       
+    Eq.~(\ref{eq:sinetheta})
+    with $v_{\theta}= 0$ and $\lambda=L$ (blue circles), $\lambda=L/2$ (orange squares) and $\lambda=L/3$ (green stars).
+    The dashed lines indicate the geometrically expected values of the droplet height $h_d$, assuming monodispersity and a perfect spherical cap shape. 
+    INSET: Number of droplets, $N(t)$, as a function of time. The three horizontal dashed lines indicate the number of minima of Eq.~(\ref{eq:sinetheta}),
+      which is $2\left(\frac{L}{\lambda}\right)^2$. In the insets, we display snapshots of the stationary droplet states as grey-scale images
+      of the film thickness field, $h(\mathbf{x},t)$.
+      }
     \label{fig:clusters_v0_sine}
 \end{figure}
-%This is in fact a consistency check of our simulations.
 On the stationary substrate, after rupture all the fluid accumulates in droplets centered at contact angle minima.
-Consequently, as we see from Fig.~\ref{fig:clusters_v0_sine} where we plot the number of droplets, $N$, in time, 
-in the steady state ($t \gg t_0$) such number attains the value $N_{\infty} = 2(L/\lambda)^2$
-(which equals the minima of (\ref{eq:sinetheta}), for $v_{\theta}=0$).
-%We expect that in case of $v_{\theta} = 0$ all fluid accumulates in droplets inside the contact angle minima.
-%While there are only two minima for $\lambda = L$ there are eight and eighteen minima for $\lambda = L/2$ and $\lambda = L/3$ respectively.
-%This can be easily checked with a cluster analysis of the film thickness $h(\mathbf{x},t)$ which
-%we shown in Fig.~\ref{fig:clusters_v0_sine}.
-%The height $h(\mathbf{x},t_{\infty})$ is displayed as grayscale images and is linked to stationary state.
-%Theoretically we expect therefor two, eight and eighteen clusters (droplets) which we mark with a vertical
-%dashed line in the respective color.
-%These values are consistent with our simulations.
-Notice that the number of clusters converges faster for smaller pattern wavelengths, in line with the observation, reported
-and justified in the previous section, that the characteristic dewetting time decreases with the wavelength.\\
-%To some extend this is expected as the distance between contact angle minima is shorter for smaller wavelengths.
-
-Another quantity that is easily accessible and helps to check the consistency of our simulations is the height of the droplets $h_d$.
-Given that all the fluid is collected inside the droplets we can make an assumption of the droplets height based on its volume $V_d$  and contact angle $\theta_d$ 
-\begin{equation}\label{eq:height_guess}
-    h_d(V_d, \theta_d) = \left(\frac{3V_d(1-\cos(\theta_d))}{\pi(2+\cos(\theta_d))}\right)^{\frac{1}{3}},
+Consequently, as we see from the inset of Fig.~\ref{fig:clusters_v0_sine}, where we plot the number of droplets $N(t)$ in time\footnote{A droplet is identified by the set (``cluster'') of points, in the plane, constituting each of the connected components of the set $\{\mathbf{x} \in [0,L]^2 | h(\mathbf{x},t) \geq h_{\ast}$\}; the clusters are determined by means of an algorithm of Hoshen-Kopelman type~\cite{HK}.}, in the steady state ($t \gg t_0$) $N(t)$ attains the value $N_{\infty} = 2(L/\lambda)^2$ (reported as horizontal lines), which equals the minima of Eq.~(\ref{eq:sinetheta}), for $v_{\theta}=0$.
+Notice that the number of droplets converges faster for smaller pattern wavelengths, in line with the observation reported and justified in the previous section that the characteristic dewetting time decreases with the wavelength.\\
+As a further consistency check, we follow the evolution of the quantity
+\begin{equation}\label{eq:deltah}
+  \Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\}-\min_{\mathbf{x}}\{h(\mathbf{x},t)\}.
 \end{equation}
-where we implicitly assume that all our droplets have a spherical cap shape.
-Since our method does not allow $h=0$ we measure the droplets height as~\cite{PhysRevE.100.033313}
-\begin{equation}\label{eq:delta_h_measure}
-    \Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\} - \min_{\mathbf{x}}\{h(\mathbf{x},t)\}.
-\end{equation}
-Counting the minima in $\theta(\mathbf{x})$ for the different wavelengths $\lambda$ allows to calculate the volume of a single droplet, see Fig.~\ref{fig:clusters_v0_sine}.
-Since the film dewets regions of high contact angle values we can safely assume that $\theta_d < 20^{\circ}$.
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=0.48\textwidth]{Figures/delta_h_v0_sin_with_dash.pdf}
-%    \caption{Evolution of $\Delta h$ with time.
-%    Different symbols and colors represent different pattern wavelengths.
-%    Dashed lines indicate solutions to the cap height $h_d$ given the volume is split equally among the droplets,
-%    see Eq.~(\ref{eq:height_guess}).}
-%    \label{fig:deltah_v0_sine}
-%\end{figure}
-Measuring the height of the droplets is done in the lower panel of Fig.~\ref{fig:clusters_v0_sine}, where $\Delta h$ is the maximal difference in film thickness which is equivalent to the droplet height $h_d$.
-Focusing on the late time behaviour ($t > 10t_0$) we identify plateauing values. 
-Since the pattern is a continuous functions, see Eq.~(\ref{eq:sinetheta}), we are left with some degree of freedom for the theoretical $\theta_d$ values. 
-As a guide to the eye we introduce three dashed lines.
-The dashed lines are calculated using Eq.~(\ref{eq:height_guess}) with $V_d = (V/N_{\infty})$ and $\theta_d = (14^{\circ}, 16^{\circ}, 17^{\circ})$ for $\lambda=(L,L/2,L/3)$ respectively.
-These fit values for $\theta_d$ agree well with the stationary droplet state of our simulations.
-Thus we are fairly confident that our droplet states and thus the simulations capturing the correct physical state.
-
-%\subsection{Stability}\label{subsec:time_dep_theta}
-%\begin{table}[h!]
-%  \begin{center}
-%    \begin{tabular}{c|c|c|c}
-%      \textbf{Pattern} & \textbf{Wavelength} & \textbf{Velocity } & \textbf{Rupture time}\\
-%      $\theta(\mathbf{x})$ & $\lambda$ & $v/v_0 $ & $\tau_r/t_0$ \\
-%      \hline
-%      $\theta_1$ & L & 0 & 1.69\\
-%      $\theta_1$ & L & 0.1 & 1.72\\
-%      $\theta_1$ & L & 1 & 2.34\\
-%      $\theta_1$ & L & 10 & 2.25\\
-%      $\theta_1$ & L/2 & 0 & 0.39\\
-%      $\theta_1$ & L/2 & 0.1 & 0.39\\
-%      $\theta_1$ & L/2 & 1 & 0.59\\
-%      $\theta_1$ & L/2 & 10 & 0.68\\
-%      $\theta_1$ & L/3 & 0 & 0.23\\
-%      $\theta_1$ & L/3 & 0.1 & 0.23\\
-%      $\theta_1$ & L/3 & 1 & 0.25\\
-%      $\theta_1$ & L/3 & 10 & 0.45\\
-%      $\theta_2$ & L & 0 & 1.01\\
-%      $\theta_2$ & L & 0.1 & 1.18\\
-%      $\theta_2$ & L & 1 & 1.32\\
-%      $\theta_2$ & L & 10 & 1.38\\
-%      $\theta_2$ & L/2 & 0 & 0.48\\
-%      $\theta_2$ & L/2 & 0.1 & 0.48\\
-%      $\theta_2$ & L/2 & 1 & 0.7\\
-%      $\theta_2$ & L/2 & 10 & 0.76\\
-%      $\theta_2$ & L/3 & 0 & 0.39\\
-%      $\theta_2$ & L/3 & 0.1 & 0.42\\
-%      $\theta_2$ & L/3 & 1 & 0.68\\
-%      $\theta_2$ & L/3 & 10 & 0.76\\
-%    \end{tabular}
-%    \caption{Rupture time values for different pattern wavelengths $\lambda$ and different substrate velocities $v_{\theta}$.
-%    Due to sampling we have a systematic error of 0.03 in the rupture times $\tau_r$.}
-%    \label{tab:rup_times}
-%  \end{center}
-%\end{table}
-
-\subsubsection{Time-dependent pattern}\label{subsec:timedep}
-Dewetting morphology is a rather broad topic~\cite{becker2003complex, RevModPhys.81.739, PhysRevLett.99.114503, craster2009dynamics, doi:10.1146/annurev-fluid-011212-140734, alizadeh_pahlavan_cueto-felgueroso_hosoi_mckinley_juanes_2018}, however we know that a simple patterned substrate will ultimately yield droplets in region of small contact angle and is likely to rupture in regions of high contact angle, as shown in Fig.~\ref{fig:handtheta}.
-This is indeed what we observe for the simulation without pattern velocity ($v_{\theta} = 0$).
+In the steady state, since the droplets are essentially monodisperse, this observable represents a measure of the mean droplet height $h_d$.
+Assuming that the droplet shape is a spherical cap, it can be estimated as $h_d =\left(\frac{3V_d(1-\cos(\theta_d))}{\pi(2+\cos(\theta_d))}\right)^{1/3}$, where $V_d = \frac{h_0 \lambda^2}{2}$ is the droplet volume and $\theta_d$ is the local contact angle (i.e.
+the value at the contact line).
+This, in turn, depends on $h_d$, through Eq.~(\ref{eq:sinetheta}), thus making the one above an implicit equation to be solved numerically for $h_d = h_d(\lambda)$.
+$\Delta h(t)$ is reported in Fig.~\ref{fig:clusters_v0_sine} for three different wavelengths, $\lambda = \{L/3,L/2,L \}$, together with the theoretically expected $h_d(\lambda)$ (depicted with lines), showing excellent agreement.\\
 \begin{figure}
     \centering
-    \includegraphics[width=0.48\textwidth]{Figures/q_2_different_v.pdf}
-    \caption{Structure analysis of the pattern $\theta_1,~\lambda=L/2$ using the Minkowski structure metrics (MSM). 
-    The dipole metric $q_2$ is plotted for different pattern velocities $v_{\theta}$, displayed with different colors and symbols.
-    In grayscale we supply the film thickness that is associated with the $q_2$ values.
+    \includegraphics[width=0.45\textwidth]{Figures/lam_2_inset.pdf}
+    \caption{Time evolution of the second order Minkowski structure metric, $q_2(t)$ (see text for details), for different $\Gamma$ values (see Eq.~(\ref{eq:vel_ratio})), on a substrate with pattern wavelength $\lambda=L/2$.
+    The grey-scale insets supply snapshots of the film thickness field associated to the various curves.
     }
@@
 \end{figure}
-Increasing the pattern velocity $v_{\theta} > 0$ does affect the dewetting morphology quite substantially.
-Having $v_{\theta} = 0.1 v_0$ we are still observing the formation of droplets, similar to $v_{\theta} = 0$ but instead of being stationary these droplets, depending on their mass, are transported with the contact angle minima.
-A somewhat similar behaviour was recently described by Grawitter and Stark~\cite{D0SM02082F} where they have used a boundary integral method to simulate a droplet on a moving wettability step.
+A time-dependent pattern affects the dewetting morphology quite substantially.
+For $v_{\theta} = 0.1 v_0$ we still observe the formation of droplets, similarly to the stationary case ($v_{\theta} = 0$).
+However, these are transported with the contact angle minima, reproducing a somehow similar behaviour recently described in a numerical study of a droplet on a moving wettability step~\cite{GrawitterStark1}.
+If the pattern speed is further increased, for $v_{\theta} = v_0$ we observe the development of rivulet-like structures, aligned with $\mathbf{v}_{\theta}$. 
+The film, in fact, while dewetting in the direction normal to the pattern velocity, is exposed, in the direction of the velocity, to a periodic potential with alternating minima and saddle points, which partially (as we will see) stabilizes the film over ``preferential'' lanes along the diagonals.
+This makes the chosen velocity direction, $(1/\sqrt{2},-1/\sqrt{2})$ (or, equivalently, the orthogonal one $(1/\sqrt{2},1/\sqrt{2})$), optimal for the formation of rivulets.
 
-If the pattern velocity is further increased $v_{\theta} > 0.1 v_0$ we observe a transition towards ligament or rivulet like structures.
-These structures are aligned with the update direction of the contact angle field, see Eq.~(\ref{eq:sinetheta}). 
-Thus if
-\begin{equation*}
-    \theta(x,y,t+T) = \theta(x-1, y, t),    
-\end{equation*}
-ligaments form in x direction.
-These ligaments are sensitive to $\lambda$, or more precisely to the contact angle minima since they attract fluid.
-To classify the morphological differences in respect to $v_{\theta}$ we make use of \textbf{papaya2}~\cite{Schaller2020}.
-
-We use the film thickness as input and compute different morphological metrics.
-Among the most interesting ones are the Minkowski structure metrics (MSM) $q_s$.
-They can be calculated using~\cite{doi:10.1063/1.4774084}
-\begin{equation}
-    q_s(K) = \frac{|\psi_s(K)|}{\psi_0(K)}, 
-\end{equation}
-where $K$ is the body to we extract morphological information from and $\psi_s$ are the irreducible Minskovski tensors (IMT)
-\begin{equation}
-    \psi_s(K) = \int_{0}^{2\pi}e^{is\phi}\rho_K(\phi) \dd\phi = \sum_{k=-\infty}^{\infty} L_k e^{is\phi_k},
-\end{equation}
-with $L_k$ being the set of normal vectors of the body $K$.
-Within the various $q_s$ we are particularly interested in the dipole one $q_2$. 
-With $q_2$ we can clearly distinguish between a droplet and a ligament structure, as we expect a much larger $q_2$ value for the later one which is what we see in Fig.~\ref{fig:msm_q2}.
-Our simulations with $\lambda = L/2$ show a clear difference between $v_{\theta} \leq 0.1v_0$ and $v_{\theta}>0.1v_0$  in $q_2$.
-Interestingly the $q_2$ values for $v_{\theta} = 0.1v_0$ are always larger than those of $v_{\theta} = 0$.
-Even the smallest pattern velocity introduces a measurable deformation of the spherical cap shape.
-
+In order to better characterize the various morphologies we apply the theory of Minkowski's functionals. 
+In particular, we employ the second order Minkowski structure metric, $q_2$~\cite{doi:10.1063/1.4774084, Schaller2020}, which can be computed from a Voronoi tessellation of the set of discrete points $(x_i, y_i)$ on the 2D lattice, such that the height field lies above a certain threshold\footnote{The expression is: $q_2 = \frac{1}{N}\sum_j \frac{1}{P_j}|\sum_k L^{(j)}_k e^{2i\phi^{(j)}_k}|$, where the inner sum runs over the edges of length $L^{(j)}_k$, of the $j$-th Voronoi cell, whose perimeter is $P_j$, and $\phi^{(j)}_k$ is the polar angle of the normal to the $k$-th edge.
+The outer sum represents an ensemble average over the $N$ points in the set.}.
+The $q_2$ metric quantifies the degree of anisotropy of the dewetting morphology, so it takes relatively large values if the structures formed display a preferential direction.
+Measuring $q_2$ then enables us to clearly distinguish between the formation of droplets and rivulets: much larger $q_2$ values are attained for the latter type of structure, as we can see in Fig.~\ref{fig:msm_q2}.
+%Our simulations with $\lambda = L/2$ show a clear difference between $v_{\theta} \leq 0.1v_0$ and $v_{\theta}>0.1v_0$  in $q_2$.
+We observe, on the other hand, that such rivulets are metastable and eventually break up into droplets, as indicated by the collapse of $q_2$ at later times. 
+Notice, though, that the $q_2$ signal for any $v_{\theta} >0$ always stays above the one for the static case, suggesting that even the smallest pattern velocity introduces a sizeable deformation of the spherical cap shape.
 \begin{figure}
     \centering
-    \includegraphics[width=0.45\textwidth]{Figures/stab_lig_lam2_v2.pdf}
-    \caption{Measurement of the stability of the ligaments in terms of $q_2$.
-    The failure time is defined as the time when the value of $q_2$ is below a threshold value.
-    If no ligaments formed at all, the failure time is set to 0.
+    % \includegraphics[width=0.45\textwidth]{Figures/Lig_stab_inset_delta_h.pdf}
+    \includegraphics[width=0.45\textwidth]{Figures/Gamma_riv_lifetime.pdf}
+    \caption{MAIN PANEL: Rivulets life-times, $\tau_{\text{riv}}$, for various $\Gamma$ values.
+    The dashed line is plotted as a guide to the eye to highlight the logarithmic dependence, in agreement with 
+    the theoretical prediction, Eq.~(\ref{eq:rivlt}).
+    INSET: Height fluctuations, $\Delta h(t)$, vs time, computed along the rivulet axis, for three different $\Gamma$ values.
     }
@@
 \end{figure}
-Moving the contact angle field faster creates metastable ligament states.
-Metastable because this ligaments will eventually break and separate into droplets.
-This is indicated in Fig.~\ref{fig:msm_q2} by the dip in $q_2$ for $v_{\theta} = 1, 10 v_0$ (green and violet curve).
-The breakup is actually due to the varicose mode of the ligament~\cite{doi:10.1063/1.3211248, PhysRevE.77.061605}.
-Interestingly the wavelength of this varicose mode is $\approx\lambda$ and therefore we only observe $N_{\infty}/2$ droplets after breakup.
-An analysis of the stability of these ligaments can be found in Fig.~\ref{fig:stab_ligs_lam2}.
-Two things can be seen in this plot, first being the fact that ligaments only form for $v_{\theta} > 0.1$.
-On the other hand a saturation happens for large $v_{\theta}$ values. 
-Meaning we do not observe an increase in the lifetime of a ligament for $v_{\theta} > 10$.
-% One possibility for that behaviour is that due to fast update of the pattern the film only experiences a singular effective contact angle value.
-Besides the fact that only $N_{\infty}/2$ droplets are formed after ligament failure we observe on top a different dynamic than the ones on slow or no pattern velocity.   
-Instead of a continuous motion with the pattern the remnants of the ligament are in an oscillating pumping state.
-In this state the droplets are to slow to follow the pattern, which leads to cycle of deformation and retraction driven by the pattern (see sup. material).
-This behaviour is clearly captured by Fig.~\ref{fig:msm_q2}, see e.g. the green curve for $t > 15t_0$.
-Nevertheless the formation of ligaments in the first place can be explained with a few relative simple assumptions.
+The breakup is the result of a varicose mode of the rivulet~\cite{doi:10.1063/1.3211248, PhysRevE.77.061605}, whose wavelength is $\approx\lambda$, such that only $N_{\infty}/2$ droplets are counted after breakup.
+These droplets show a peculiar dynamics, characterized by a periodic sequence of spreading and retraction, driven by the pattern, that we dub "pumping state" (see movie file ligament\_formation\_and\_breakup.mp4 in the Supp.
+Mat.).
 
-This problem is associated with two time scales, first being $t_0(\theta)$ the time scale of the dewetting thin film, see Eq.~(\ref{eq:t0}).
-After the rupture we identify another time scale, the capillary time scale $\tau_{\text{Ca}}$ that drives the rims away from the holes~\cite{Edwardse1600183},
-\begin{equation}\label{eq:cap_time}
-    \tau_{\text{Ca}} = \frac{9\mu l_0}{\gamma \theta^3(\mathbf{x})},
+We argue that the emergence of rivulets is controlled by the competition of two characteristic velocities: the pattern wave speed, $v_{\theta}$, and the retraction speed, $U_{\Theta}$, introduced in Eq.~(\ref{eq:taur_l1}).
+If $U_{\Theta}$ is large as compared to $v_{\theta}$, the film retraction is faster than the transport due to the contact angle field and thus droplets from.
+However, if $v_{\theta}$ is larger than $U_{\Theta}$, then the retracting film has too little time to form droplets and ends up in the metastable rivulet state.
+\textcolor{red}{It appears, therefore, natural to consider the ratio of these two velocities,
+$\Gamma \equiv v_{\theta}/U_{\theta}$, as the discriminating parameter.}
+%Recalling the expression (\ref{eq:t0}) for the reference velocity $v_0$, %we define the parameter 
+%\begin{equation}\label{eq:vel_ratio}
+%    \Gamma = \frac{v_{\theta}}{U_{\Theta}} = \frac{3\lambda h_0^3 %q_0^4}{\Theta^3}\chi 
+%\end{equation}
+%as the ratio of these two velocities,
+%where $\chi \equiv v_{\theta}/v_0$. 
+We see from Fig.~\ref{fig:msm_q2} that indeed rivulets form only for $\Gamma > 1$. 
+Moreover, the larger $\Gamma$, the more stable the rivulets are; in other words, the rivulet life-time, $\tau_{\text{riv}}$, that can be conventionally taken as the time at which the drop of $q_2$ occurs, grows with $v_{\theta}$ (see Fig.~\ref{fig:stab_ligs_lam2}).
+The rivulet itself is, in fact, prone to dewetting, with the liquid accumulating over patches around contact angle minima.
+However, as the pattern moves, the instability is tamed due to configurations whereby higher contact angle regions underlie height field maxima, thus tending to revert the fluid flow.
+Heuristically speaking, this means that, if we define $\Delta h(t)$ as in Eq.~(\ref{eq:deltah}), but restricted on the rivulet axis, this should grow exponentially (with a certain growth rate) only when the system is in the unstable configuration.
+Namely, $\Delta h(t) \propto e^{\alpha t}$ (see inset of Fig.~\ref{fig:stab_ligs_lam2}) with a prefactor proportional to the time spent by the rivulet in such a configuration, which goes as $\sim \lambda/v_{\theta}$, therefore $\Delta h(t) \sim \alpha (\lambda/v_{\theta})e^{\alpha t}$.
+The rivulet life-time can be seen as the rupture time of the structure along its axis, hence such that $\Delta h (\tau_{\text{riv}}) \sim h_0$ \cite{PhysRevE.104.034801}, which yields
+\begin{equation}\label{eq:rivlt}
+    \tau_{\text{riv}} \sim \log(v_{\theta}) \sim \log(\Gamma). 
 \end{equation}
-where  $l_0$ is a characteristic length scale of the problem, we use $\lambda/2$ as two rims can travel at most $\lambda/2$ before they merge.
-Knowing this time scales allows us to introduce characteristic velocities.
-One of these two is $v_0$ which is given by Eq.~(\ref{eq:v0}) and used to normalize $v_{\theta}$.
-The other one is set by $\tau_{\text{Ca}}$ and reads
-\begin{equation}\label{eq:cap_speed}
-    U_{Ca} =\frac{l_0}{\tau_{\text{Ca}}} = \frac{\gamma}{9\mu} \theta^3(\mathbf{x}), 
-\end{equation}
-where $U_{\text{Ca}}$ is the capillary speed.
-The argument is as follows, if $U_{\text{Ca}}$ is large as compared to $v_{\theta}$ than the film retraction is faster than the transport due to the contact angle field and thus droplets from.
-Which explains the cases of $v_{\theta} = 0$ and  $v_{\theta} = 0.1v_0$.
-On the other hand if $v_{\theta}$ is larger than $U_{\text{Ca}}$ than the retracting film has to little time to form droplets and ends up in the metastable ligament state.
-Taking a look at our simulations with $v_{\theta} = 1v_0$ and $v_{\theta} =  10v_0$ we clearly observe ligament states.
+This logarithmic dependence is indeed observed in the numerical data as shown in Fig.~\ref{fig:stab_ligs_lam2}.
 
-In fact the ratio 
-\begin{equation}\label{eq:vel_ratio}
-    \Gamma = \frac{v_{\theta}}{U_{\text{Ca}}} = \frac{3\lambda \chi h_0^3 q_0^4}{\Theta^3}, 
-\end{equation}
-where $\Theta = max(\theta(\mathbf{x}))$ and $v_{\theta} = \chi v_0$, allows to classify the dewetting morphology on a time and space dependent substrate pattern. 
-% as
-% \begin{equation}\label{eq:classify}
-%     \Gamma~\begin{cases}
-%     < 1\qquad ... \text{droplets} \\
-%     > 1\qquad ... \text{ligaments} 
-%     \end{cases}.
-% \end{equation}
-% This explains in fact what we observe and holds true independent of the wettability gradient, i.e. $\theta_1, \theta_2$.
+\noindent {\it Conclusions.}
+We have presented results, from numerical simulations and theoretical analysis, on the dewetting of thin liquid films on a switchable substrate, modelled with a space and time periodically varying contact angle in the thin-film equation.
+Studying how the stability of the film depends on the underlying static pattern, we found that the rupture times grow linearly with the pattern wavelength, for short wavelengths, and quadratically in the long wavelength limit.
+In the time-dependent case, the rupture times are in general longer, indicating an induced greater film stability, and, while the quadratic growth is preserved at long wavelengths, a plateauing behaviour is observed as the wavelength decreases.
+A theoretical explanation has been provided for all these various regimes.
+Furthermore, we showed that, at increasing the wettability wave speed, a transition occurs in the dewetting morphology from a multi-droplet to a metastable multi-rivulet state. 
+A dimensionless parameter, $\Gamma$, controlling the transition has been identified in the ratio of the pattern speed and the typical film retraction speed over the substrate. 
+The rivulets life-time itself grows with the pattern speed, displaying a logarithmic dependence that has been captured by means of phenomenological arguments.
+On a broader perspective, our work suggests that switchable substrates offer a new avenue to control thin film dewetting, with obviously relevant implications, for instance, for open microfluidic devices, and paves the way to future studies in this direction, exploiting more complex and dedicated space-time dependencies.
 
-Therefore by varying either $\chi$ or $\lambda$ for a fixed $\Theta$ one can change the intermediate dewetting morphology according to Eq.~(\ref{eq:vel_ratio}).
-Interestingly but not surprising is the fact that the fluid transport is larger for $\Gamma > 1$ than for $\Gamma < 1$.
-
-\section{Summary and Conclusions}\label{sec:sum_conclu}
-To summarize, we have shown numerical simulations of a dewetting thin film on a patterned substrate.
-Generating baseline experiments with time independent patterns of different wavelengths, we than move to a time and spatial varying contact angle field $\theta(\mathbf{x},t)$.
-
-In the next step we define a characteristic pattern velocity $v_0$ which simply is the patterns wavelength $\lambda$ over a characteristic time $t_0$, see Eqs.~(\ref{eq:t0}, \ref{eq:v0}).
-Experiments with increasing pattern velocity $v_{\theta} = \chi v_0$ with $\chi\in\{0,10\}$ show a pronounced change in dewetting morphology from well defined droplets to ligaments.
-With the single balance between pattern velocity $v_{\theta}$ and capillary velocity $U_{\text{Ca}}$ we can identify a parameter that separates the two morphological regimes.
-
-An outlook to this work is to study the transition from ligaments to droplets in a more detailed manner.
-Looking at the edges $\Gamma =1$ with more suitable tools like e.g. bifurcation diagrams will likely be subject to future work.
-But also the update scheme of the contact angle field allows for many applications.
-Among them is the avenue of switchable substrates, where e.g. switching of $\theta(\mathbf{x})$ due to a light source is in the scope of this method.
-
-\begin{acknowledgements}
-The authors acknowledge financial support by the Deutsche Forschungsgemeinschaft (DFG) within the priority program SPP2171 ``Dynamic Wetting of Flexible, Adaptive, and Switchable Substrates'', project HA-4382/11.
-\end{acknowledgements}
-
-%\appendix
-%\section{}\label{app:one}
+\noindent {\it Acknowledgements}.
+SZ and JH acknowledge financial support from the Deutsche Forschungsgemeinschaft (DFG) within the priority program SPP2171 ``Dynamic Wetting of Flexible, Adaptive, and Switchable Substrates'', project HA-4382/11.
 

### 2022-10-03 Notebooks/Moving_wettability.jl

@@ -1,5 +1,5 @@
 ### A Pluto.jl notebook ###
-# v0.17.0
+# v0.19.11
 
 using Markdown
 using InteractiveUtils
@@ -1197,9 +1197,9 @@ version = "0.9.1"
 
 [[Cairo_jll]]
 deps = ["Artifacts", "Bzip2_jll", "Fontconfig_jll", "FreeType2_jll", "Glib_jll", "JLLWrappers", "LZO_jll", "Libdl", "Pixman_jll", "Pkg", "Xorg_libXext_jll", "Xorg_libXrender_jll", "Zlib_jll", "libpng_jll"]
-git-tree-sha1 = "f2202b55d816427cd385a9a4f3ffb226bee80f99"
+git-tree-sha1 = "4b859a208b2397a7a623a03449e4636bdb17bcf2"
 uuid = "83423d85-b0ee-5818-9007-b63ccbeb887a"
-version = "1.16.1+0"
+version = "1.16.1+1"
 
 [[Chain]]
 git-tree-sha1 = "cac464e71767e8a04ceee82a889ca56502795705"
@@ -1298,7 +1298,7 @@ deps = ["Random", "Serialization", "Sockets"]
 uuid = "8ba89e20-285c-5b6f-9357-94700520ee1b"
 
 [[Downloads]]
-deps = ["ArgTools", "LibCURL", "NetworkOptions"]
+deps = ["ArgTools", "FileWatching", "LibCURL", "NetworkOptions"]
 uuid = "f43a241f-c20a-4ad4-852c-f6b1247861c6"
 
 [[EarCut_jll]]
@@ -1330,6 +1330,9 @@ deps = ["Dates", "Mmap", "Printf", "Test", "UUIDs"]
 git-tree-sha1 = "0f5e8d0cb91a6386ba47bd1527b240bd5725fbae"
 uuid = "48062228-2e41-5def-b9a4-89aafe57970f"
 version = "0.9.10"
+
+[[FileWatching]]
+uuid = "7b1f6079-737a-58dc-b8bc-7a2ca5c1b5ee"
 
 [[FixedPointNumbers]]
 deps = ["Statistics"]
@@ -1397,9 +1400,9 @@ version = "0.21.0+0"
 
 [[Glib_jll]]
[…]

### 2022-10-03 Notebooks/structure_analysis.jl

@@ -1,5 +1,5 @@
 ### A Pluto.jl notebook ###
-# v0.17.1
+# v0.19.9
 
 using Markdown
 using InteractiveUtils
@@ -502,36 +502,23 @@ end
 # ╔═╡ 7255922b-0361-4975-9435-9353d1ed553d
 begin
 	vnames = ["vₜ=0" "vₜ=0.1v₀" "vₜ=1v₀" "vₜ=2v₀" "vₜ=4v₀" "vₜ=6v₀" "vₜ=8v₀" "vₜ=10v₀"]
-	lig_delth = plot(lig_deltas[:,9], lig_deltas[:,3], label=vnames[3],
+	Gnames =  ["Γ=0" "Γ=0.15" "Γ=1.5" "Γ=3.0" "Γ=6.0" "Γ=9.0" "Γ=12.0" "Γ=15.0"]
+	lig_delth = plot(lig_deltas[:,9], lig_deltas[:,3], label=Gnames[3],
 		             w = 3, 							# line width
 		             st = :samplemarkers, 				# some recipy stuff
 		             step = 50, 						# density of markers
 		             marker = (8, :auto, 0.6),			# marker size
 	)
-	plot!(lig_deltas[:,9], lig_deltas[:,4], label=vnames[4],
+	plot!(lig_deltas[:,9], lig_deltas[:,4], label=Gnames[4],
 	      w = 3, 							# line width
 		  st = :samplemarkers, 				# some recipy stuff
 		  step = 50, 						# density of markers
 		  marker = (8, :auto, 0.6),			# marker size
 	)
-	# plot!(lig_deltas[:,9], lig_deltas[:,1], label=vnames[3],
-	# 	  w = 3, 							# line width
-	# 	  st = :samplemarkers, 				# some recipy stuff
-	# 	  step = 50, 						# density of markers
-	# 	  marker = (8, :auto, 0.6),			# marker size
-	# 	 )
-	# plot!(lig_deltas[:,9], lig_deltas[:,4], label=vnames[4],
-	# 	  w = 3, 							# line width
-	# 	  st = :samplemarkers, 				# some recipy stuff
-	# 	  step = 50, 						# density of markers
[…]

### 2022-11-25 Reply_RefereeA.tex
Please work

@@
 we assumed that it is a very well established concept employed in any theoretical and computational study of thin liquid film hydrodynamics in the lubrication approximation framework.
-We acknowledge, though, that our assumption was probably erroneous and we give some more explanation. 
-First of all, there is no need (and no way) to go from the disjoining pressure (or the interface potential) to the surface tension, since the two quantities carry complementary physical information. 
+We also assumed that an equation like Eq.(3) (now Eq.(2)), which is approximately twenty-five years old \cite{SCHWARTZ1998173,Mitlin94}, did not deserve a lengthy discussion.
+However, our assumptions were probably erroneous and, therefore, we give, here, some more explanation. 
 Let us, first, make a step back and underline a fundamental difference between the physics of {\it one free interface} and that of a {\it film}:
@@
 We have now fixed this and we apologize for it.
-\JH{Do we need to change more in the paper?}
 }
@@
 and an horizontal length scale $\ell$ (which in our case can be, for example, the pattern wavelength, $\lambda$), where, typically $h_0 \ll \ell$. We chose to use the mean 
-film thickness as a scale, because it is the only relevant length that is common to all the simulations.
+film thickness as a scale, because it is a relevant length that is common to all the simulations.
 In the same spirit, we chose as a reference time $t_0(\theta)$, which is not exactly the capillary time (though being related to it). 
@@
 substrate of contact angle $\theta$. In spite of the substrate being patterned in our case, it is possible to define an effective $t_0$ in terms 
-of the highest contact angle on the pattern, which is what determines the fastest instability, i.e. $t_0(\Theta)$, where 
-$\Theta=\mbox{max}_{\mathbf{x}}\{\theta(\mathbf{x})\}= \theta_0 + \delta \theta$. We deem the so defined characteristic time as an appropriate scale,
+of, e.g., the mean contact angle on the pattern, i.e. $t_0(\theta_0)$. 
+We deem the so defined characteristic time as an appropriate scale,
 since it does not depend on the pattern wavelength and speed. Actually, one can see that it captures well the time scales of the dewetting process
@@
 We recognize, on the other hand that in this perspective the choice of $v_0 = \lambda/t_0$ as a reference velocity looks a bit awkward.
-A more natural candidate would be, indeed, $v_0=h_0/t_0$. We have therefore decided to use the latter in the revised version of the manuscript.
+A more natural candidate would be, indeed, $v_0=\lambda_s/t_0$, where $\lambda_s$ is the spinodal 
+characteristic length corresponding to $\theta_0$. We have therefore decided to use the latter in the revised version of the manuscript.
 We have also rephrased the corresponding paragraph in the {\it Method} section as follows: \\
 
-\textcolor{red}{Hereafter, lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film.
+\textcolor{red}{Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film.
 On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads
-$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$ \cite{Mecke_2005,PhysRevE.100.023108}; in our patterned case, we define 
-$q_0^2=h_{\ast}^{-2}(1-\cos\Theta)f^{\prime}(h_0/h_{\ast})$, where 
-$\Theta = \max_{\mathbf{x}} \theta(\mathbf{x}) = \theta_0 + \delta \theta$.
-A velocity scale also then naturally arises, namely $v_0 = h_0/t_0$.}
+$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$ \cite{Mecke_2005,PhysRevE.100.023108}. In our patterned case, we define 
+$q_0^2=h_{\ast}^{-2}(1-\cos\theta_0)f^{\prime}(h_0/h_{\ast})$.
+Correspondingly, we choose as a velocity scale $v_0 = \lambda_s/t_0$, where $\lambda_s = 2\pi/q_0$.}
 
@@
 the typical velocity is $\sim 1 \mbox{mm}/\mbox{s}$, which gives a Reynolds number $Re \approx 10^{-3}$ \cite{Edwards2016}. 
-In our problem, there are two relevant characteristic velocities: the capillary velocity, $U_{\text{cap}}$, and 
-the pattern speed, $v_{\theta}$. When their ratio $\Gamma = v_{\theta}/U_{\text{cap}} < 1$, the capillary 
-velocity dominates and fixes the Reynolds number which can be, then, evaluated as 
-$Re = U_{\text{cap}}h_0/\nu$ \cite{RevModPhys.69.931}, whereas for $\Gamma > 1$, $Re = v_{\theta} h_0/\nu$. 
+In our problem, there are two relevant characteristic velocities: the retraction speed, 
+$U_{\Theta}$, and 
+the pattern speed, $v_{\theta}$. When their ratio $\Gamma = v_{\theta}/U_{\Theta} < 1$, the retraction speed dominates and fixes the Reynolds number which can be, then, evaluated as 
+$Re = U_{\Theta}h_0/\nu$ \cite{RevModPhys.69.931}, whereas for $\Gamma > 1$, $Re = v_{\theta} h_0/\nu$. 
 In any case, the Reynolds number stays always below $\approx 0.2$.
@@
 {
-We agree that giving the values of surface tension and dynamic viscosity 
-in lattice Boltzmann units was not particularly insightful and could, on 
-the contrary, be misleading, so we have removed the reference to them.
-\JH{But shouldn't we list them somewhere to make sure the paper is reproducible?}
-Concerning length and time scales, as we have discussed in Comment 2, 
-they are made dimensionless 
-with the mean film height and the characteristic time $t_0$. These two quantities 
+Length and time scales, as discussed in Comment 2 and as mentioned also in the original version of the manuscript, 
+are made dimensionless with the mean film height and the characteristic time $t_0$. 
+These two quantities 
 have a clear physical meaning and can, therefore, be used to compare with 
@@
 We have reported these comments on comparison with actual experimental 
-values in the new version of the paper.
+values in the new version of the paper as follows:\\
+\\
+\textcolor{red}{To give a flavour of how to compare with actual experimental systems, let us notice
+first that the expression of $t_0$ depends on the disjoining pressure (through its dependence 
+ $q_0 = \sqrt{\frac{1}{2\gamma} \Pi^{\prime}(h_0)}$) and contains, therefore, energetic details, namely 
+ the Hamaker constant $A_H$.
+Consider, for instance, the problem studied in \cite{becker2003complex}: the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon
+dioxide substrate, where $A_H  = 2.2 \times 10^{-20} \, \text{J}$ and 
+$\gamma = 0.03 \, \text{N}/\text{m}$. Neglecting the short range part of the interface potential
+(such that $\Pi(h)= - \frac{A_H}{6\pi h^3}$)~\cite{Mecke_2005,becker2003complex}, one gets 
+$q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}$.
+The characteristic time of the growth of the instability for these experiments, then, reads \cite{PhysRevLett.99.114503} (the dynamic viscosity of polystyrene is 
+$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$) $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}$ and it can be seen that indeed film rupture occurs within time 
+scales of the order of $t_0$ ($\approx 2 t_0$) \cite{becker2003complex}, as in our 
+simulation for the largest pattern wavelength, $\lambda = L$ (see also our previous work
+with homogeneous substrates~\cite{PhysRevE.104.034801}).} 
+\\
+
+Concerning the values of viscosity and surface tension in lattice Boltzmann units, 
+we acknowledge that giving them in the middle of the discussion on characteristic length and time might  
+have caused confusion. Nevertheless we deem important to provide them, for the sake of readers interested 
+in numerical details of the method as well as for the results reproducibility, therefore we have moved them 
+to a later paragraph, where we added: \\
+\\
+\textcolor{red}{The numerical values in lattice Boltzmann units of the 
+mean film thickness, dynamics viscosity and surface tension are, respectively, $h_0=1$, $\mu=1/6$
+and $\gamma=0.01$.}
 }
@@
 experiments, there are tons of examples in the literature, such as 
-\cite{becker2003complex,fetzer2005new,PhysRevLett.99.114503} to name but a few (see also \cite{craster2009dynamics,RevModPhys.81.739} for reviews). 
+\cite{becker2003complex,fetzer2005new,PhysRevLett.99.114503} to name but a few (see also \cite{craster2009dynamics,RevModPhys.81.739} for reviews). All this literature is cited in the paper.
 }

### 2022-11-25 Reply_RefereeB.tex
Please work

@@
 can be encountered in the literature, depending on the actual interactions
-\cite{PhysRevE.63.011208}. Here, though, our choice of the $(3,2)$ pair of exponents is motivated more by computational reasons rather than by the will of 
+\cite{SCHWARTZ1998173,Mitlin,Teletzke}. Here, though, our choice of the $(3,2)$ pair of exponents is motivated more by computational reasons rather than by the will of 
 simulating a specific liquid with given physical properties.
@@
 is shorter, thus making the overall computational cost more manageable.
-In fact, we have seen in previous study~\cite{PhysRevE.104.034801} that
+In fact, we have seen in a previous study~\cite{PhysRevE.104.034801} that,
 except for this shift of the time scales, the dewetting dynamics remains 
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{Figures/No_vel_render2.png}
-    \caption{}
-    \label{fig:rivulets9_3}
+    \includegraphics[width=0.7\textwidth]{9-3_3-2_comp_Gamma_1-5.pdf}
+    \caption{Minkowski metric $q_2$ as a function of time for 
+    $\Gamma = 1.5$ and $\lambda = 256 h_0$ from two simulations with different disjoining 
+    pressure exponents, $(9,3)$ and $(3,2)$ (as in the paper) respectively.}
+    \label{fig:rivulets_9_3}
 \end{figure}
-We have also added a footnote in the paper commenting on this aspect.
-%We thank the referee for pointing this out.
-%It is true that there are variations concerning the power law model of %the disjoining pressure.
-%A common one to which the referee refers to is $(n,m) = (9,3)$, the %choice we made for the manuscript is $(n,m) = (3,2)$.
-%Both exponent pairs can be found in the literature with even more %variety~\cite{moulton_lega_2013,SCHWARTZ1998173,doi:10.1063/1.4828721,Br%asjen}.
-%The pair $(9,3)$ is the one that can be motivated from functional %derivatives of a Lennard-Jones potential, understandably this could more %appealing to use.
-%We have used both exponent pairs in previous works, see %e.g.~\cite{PhysRevE.104.034801}.
-%In this manuscript we use $(3,2)$ to keep $t_0$ reasonably small and %thus the overall computational cost manageable.
-%Using the pair $(9,3)$ does not alter the results, it effectively shifts %the value of $t_0$ by $\approx 10$.
+We have added the following footnote in the paper commenting on this aspect:\\
+\\
+\textcolor{red}{Expressions of 
+the general form $\Pi(h) = \gamma(1-\cos \theta)(n-1)(m-1)/(n-m)((h_{\ast}/h)^n - (h_{\ast}/h)^m)/h_{\ast}$
+are frequently found in the literature, with various combination of exponents $(n,m)$ ($n>m$), depending
+on the type of interactions among fluid molecules and between fluid and substrate~\cite{SCHWARTZ1998173,Mitlin,Teletzke}. A common choice, for instance, is the pair $(9,3)$, which 
+can be derived from molecular interactions of Lennard-Jones type. Here, the use of the $(3,2)$ pair is motivated mainly by the computational advantage of keeping the characteristic dewetting time 
+shorter. We have run also a single case ($\lambda=256 h_0$ and $\Gamma=1.5$) with $(9,3)$ and observed that 
+the dynamics remains essentially the same, but for a shift in the time scale (see also~\cite{PhysRevE.104.034801}).}
 }
@@
 by the formation of satellite 
-droplets following rivulet breakup~\cite{doi:10.1073/pnas.1820487116} or 
+droplets following rivulet breakup~\cite{peschka2019signatures} or 
 in the rim shape in dewetting by hole nucleation~\cite{fetzer2007quantifying, munch2005lubrication} (see also Fig.~\ref{fig:slip}).
@@
 Our choice of this slip length value was dictated essentially by the need 
-to keep numerically stability even in the strongly time dependent case 
+to keep numerical stability even in the strongly time dependent case 
 (large pattern speed), yet remaining in a small slip regime. 
@@
 picture remains qualitatively the same, we have run a simulation with 
-a slip length five times smaller. The results from the analysis of the Minkowski metric $q_2$, for two cases with low and high pattern speed, reveal that indeed
+a slip length four times smaller. The results from the analysis of the Minkowski metric $q_2$, for two cases with low and high pattern speed, reveal that indeed
 the droplet-to-rivulet transition is confirmed, with $\Gamma$ as the 
@@
     \centering
-    %\includegraphics [width=0.6\textwidth]{slip_measure.pdf}
-    \caption{}
+    \includegraphics [width=0.75\textwidth]{small_slip_two_Gammas.pdf}
+    \caption{Minkowski metric $q_2$ as a function of time for two pattern speeds and slip velocity $\delta = 0.25 h_0$, signalling that the emergence of rivulets for $\Gamma>1$
+    is preserved.}
     \label{fig:q2smallslip}
 \end{figure}
-We have commented on this when we give the value of the slip length, just before 
-equation (5) of the revised manuscript.
-%We thank the referee for bringing this up.
-%Slip and disjoining pressure are an essential part of our model -- as for many %other thin film solvers.
-%Here, they allow us to have moving contact lines with imposed wettability %features.
-%We agree that the value may seem large at first glance.
-%However, as Peschka et al. have shown, the generation of satellite droplets %during rivulet breakup, similar to what we observe, is a no-slip %feature~\cite{doi:10.1073/pnas.1820487116}.
-%We observe the occurrence of satellite droplets in the initial dewetting %period.
-%Increasing the slip length makes those droplets disappear.
-%Another point to mention here is the dependency of the dewetting rim on the %slip.
-%Small to vanishing slip induces a dimple behind the front, while for large slip %lengths this dimple vanishes~\cite{fetzer2007quantifying, %munch2005lubrication}.
-%We checked this behaviour in Fig.~\ref{fig:slip} and found only a small %dependency for slip lengths $\delta < 10$.
-%Spots with a thickness $h \leq h_{\ast}$ are considered dry spots, therefore %the term equilibrium thickness maybe exaturated.
+We have commented on this when we give the value of the slip length, saying that:\\
+\\
+\textcolor{red}{To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$, 
+a value which lies
+within the weak/intermediate slip regime~\cite{peschka2019signatures,fetzer2007quantifying, munch2005lubrication} (we have tried also a value four times smaller, without observing 
+any major qualitative change in the results).} 
 }
@@
 {
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{simple_exp.pdf}
-    \caption{Schematic realization of the experiment described in the manuscript.}
-    \label{fig:experiment_illustration}
-\end{figure}
-We thank the Referee for bringing this up.
-First, we like to note that experiments are not our field of expertise and therefore our envisioned realization could be flawed.
+This is indeed a very interesting comment.
+Although our approach, as theorists/numericists, can be seen more of {\it theoretical
+engineering} and our picture would be, then, unlikely devoid of flaws, we are glad 
+to try to envisage a potential experiment.
+%We thank the Referee for bringing this up.
+%First, we like to note that experiments are not our field of expertise and therefore %our envisioned realization could be flawed.
 
-The idea we have would be to prepare a thin film ($h_0 \approx 4nm$ PS) on a light responsive substrate.
-Dimensions of the substrate should at least be $2$ microns in $x$ and $y$, but larger is fine.
-Arrange several light sources (diodes) in a lattice with a distance of $\lambda \approx 1\mu m$ between them, as displayed in Fig.~\ref{fig:experiment_illustration}.
-It would be perfect if the structure on which the light sources are mounted can be moved in a periodic manner, e.g. wrapped around rolls.
-On the other hand if the array of diodes is large enough, say $100\lambda$ it should be sufficient to move them continuously in one direction. 
-\JH{Could one do this with a DMD as here https://doi.org/10.1021/jp301092y ?
+%The two relevant scales are $t_0$ and $q_0$ (or $2\pi/q_0$).
+%Considering the experimental parameters from~\cite{becker2003complex,PhysRevLett.99.114503} %we get, 
+%\begin{align*}
+%    t_0 &\approx 300\,\text{s},\qquad~~~\qquad \frac{2\pi}{q_0}\approx 
+%    400\,\text{nm}.
+%\end{align*}
+%The time scale is in the region of minutes and the wettability pattern wavelengths are %in the interval $(800\,\text{nm}, 4000\,\text{nm})$. 
+A possibility would be to prepare a thin liquid film on a light responsive substrate~\cite{IchimuraEtAl_Science2000}, under the action of controlled 
+external stimuli (a light emitter). 
+An ideal candidate could be a digital multimirror device (DMD).
+%The dimensions of the substrate should be roughly ten times the maximal pattern wavelength %($\approx 40\,\mu \text{m}$) to account for the periodicity.
+%An ideal control of the external stimuli (light) would be a digital multimirror device (DMD).
+This technology has effectively been used for thin film experiments and additive manufacturing~\cite{doi:10.1021/jp301092y, doi:10.1126/science.aax8760}, it allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$)
+with spatial resolution of $\approx 10 \times 10 \, \mu \text{m}^2$ (the size of a pixel).
+%Using low intensity UV light $365\,\text{nm}$ and blue light %$436$nm~\cite{IchimuraEtAl_Science2000}, we assume that the radiation heat is negligible.
+Considering as a reference, for instance, the system studied in ~\cite{becker2003complex,PhysRevLett.99.114503}, namely a $\sim 4 \, \text{nm}$ thick film 
+of polystyrene deposited on an oxidized silicon wafer,  
+we evaluate the retraction speed 
+$U_{\Theta} = \frac{\Theta^3 \gamma}{9 \mu}$ to be 
+$U_{\Theta} \approx 10^{-2} \, \mu \text{m}/\text{s}$. 
+The (minimum) {\it pattern speed} can be 
+estimated, from the pixel size with frame rate $\sim 1 s^{-1}$, as 
+$v_{\theta} \sim 10 \, \mu \text{m}/\text{s}$, that would give $\Gamma \sim 10^3$, i.e. well within 
+the rivulet regime ($\Gamma > 1$). We note, also, that both $U_{\Theta}$ and $v_{\theta}$ can 
+be widely modulated: the former, playing either with the (temperature and molecular weight dependent) viscosity 
+or tailoring the substrate to making it more hydrophobic (i.e. increasing the contact angle); the latter, tuning
+the spatial resolution and frame rate of the DMD. We expect, thus, the range of achievable $\Gamma$'s to 
+be feasibly extended both to very high ($\Gamma \gg 1$) and very low ($\Gamma \ll 1$) values. 
+We have added the following footnote after the introduction of the contact angle pattern (Eq.(3) in the revised
+manuscript) commenting on this aspect:\\
+\\
+\textcolor{red}{A laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}. 
+The external stimulus, our moving pattern, can be realized by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.} 
+%\JH{Could one do this with a DMD as here https://doi.org/10.1021/jp301092y ?
 Then, one could use a transparent substrate and just do it from the bottom.
@@
 
-In Eq.~(5) we introduce two scales, $t_0$ and $v_0$.
-From experiments of dewetting thin films we know that $t_0$ correlates with the rupture time~\cite{becker2003complex, PhysRevLett.99.114503}.
-Considering the same experimental parameters as above, 
-\begin{align*}
-    t_0 &\approx 1800s,\qquad~~~\qquad \frac{2\pi}{q_0}\approx 400nm, \\
-    \quad v_0 &= \frac{\lambda}{t_0} \approx \frac{1}{900}\frac{\mu m}{s},\quad \quad v_0^{\ast} = \frac{2\pi}{q_0 t_0} \approx \frac{4}{9}\frac{nm}{s},
-\end{align*}
-where we take into account that the rupture time is overestimated by about a factor of two.
-The experiment therefore reduces to the movement of the diodes with a velocity between $[10^{-7}/900, 10^{-5}/900] m/s$. 
-\JH{We need to comment on the experimental realization in the paper. Can we summarize this idea to 1-2 sentences?}
+%\JH{We need to comment on the experimental realization in the paper. Can we summarize this idea to 1-2 sentences?}
 }
@@
 {
-We thank the referee for this comment.
-It is true that we use $q_0$ but never display its actual value.
-Inserting numbers, we have 
-\begin{equation*}
-    q_0 \approx 0.13, \quad \lambda_0 = \frac{2\pi}{q_0} \approx 48.5 \approx \frac{L}{10},  
-\end{equation*}
-therefore $\lambda_0$ is usually smaller than $\lambda$.
-
-No, these simulations do not suffer from finite size effects.
-The honest reason for using multiples of the system size is the ease of the numerical realization.
-Having periodic boundary conditions, a non-multiple wavelength of the system size would induce a discontinuity.
-Given the complexity of the problem, we would like to have as much control of the pattern as possible.
-
-The arbitrary choice we made had however helped us to understand where the boundary between rivulet and droplets appears.
-In Eq.~(11) we use
-\begin{equation*}
-    \Gamma = \frac{v_{\theta}}{U_{\theta}} = \frac{\chi v_0}{U_{\theta}},
-\end{equation*}
-where rivulets appear for $\chi > 1$.
-Substituting the wavelength dependency of $v_0$ with $q_0$,
-\begin{equation*}
-    v_0^{\ast} = \frac{\lambda_0}{t_0} 
-\end{equation*}
-and computing $\Gamma$ in terms of $v_0^{\ast}$ yields
-\begin{equation*}
-    \Gamma = \frac{v_0^{\ast}}{U_{\theta}} = \frac{6\pi h_0^3 q_0^3}{\Theta^3}.
-\end{equation*}
-Relating both with each other yields
-\begin{equation*}
-    \frac{\chi 3\lambda h_0^3 q_0^4}{\Theta^3} = \frac{6\pi h_0^3 q_0^3}{\Theta^3},
-\end{equation*}
-and with most of the terms canceling one obtains
-\begin{equation*}
-    \lambda = \frac{2\pi}{q_0\chi}, 
-\end{equation*}
-which in fact states that if the wavelength $\chi^{-1}2\pi/q_0 < \lambda$ rivulets should form. 
-\JH{Change in paper?}
+We apologize for not giving the value of $q_0 \approx 0.09$ (corresponding to an intrinsic correlation length 
+$\lambda_s = 2\pi/q_0 \approx 70 h_0$), which is now provided in the revised manuscript. We also acknowledge that expressing the wavelength in terms of the size $L$
+was not at all meaningful (no finite size effects are involved).
+In fact, all the lengths (including $L$) are expressed in units of the mean film thickness $h_0$ (and it was so also in the old version of the paper, as commented in the text below eq.
+(4)).
+In order to avoid any confusion, we explicitly write
+the values of all lengths in terms of $h_0$ in the revised version.
 }
@@
 {
-That is indeed a very interesting point.
-In fact, we know the second reference very well from our own work~\cite{PhysRevE.104.034801}.
-It is an appealing model, and we look forward to discussing it in further details in future publications.
-We therefore added the following to the conclusion 
+We guess that the Referee mentions the fact that the Laplacian of the thickness field, $\nabla^2 h$,
+is just a small gradient approximation of the actual local curvature, such that the Laplace contribution to the film pressure should indeed read $\gamma \nabla^2 h/\sqrt{1+|\nabla h|^2}$.
+This is of course correct, but
+the condition $|\nabla h| \ll 1$ holds true in our system, owing to the very same lubrication approximation,
+or, equivalently, to the small contact angles involved; the spinodal wavelength for a homogeneous substrate with
+contact angle $\theta_0$ is $\lambda_s = 2\pi/q_0(\theta_0) \approx 70 h_0$. Therefore, the typical gradients
+of the thickness field can be estimated as $|\nabla h| \sim h_0/\lambda_s \approx 1/70 \ll 1$ (notice 
+that this is, somehow, a worst case scenario since, as we discuss in the text, in the patterned substrates
+$\lambda \stackrel{>}{\sim} \lambda_s$).
+We have added a footnote, when we introduce the film pressure, saying that:\\
 \\
-\textcolor{red}{...}
+\textcolor{red}{The Laplace pressure is the product of the surface tension 
+and the local curvature, therefore its full expression reads $p_L = -\gamma \nabla^2 h/\sqrt{1+|\nabla h|^2}$; 
+since we are in the lubrication regime, though, whereby $|\nabla h| \ll 1$, the denominator can be safely neglected 
+and the pressure reduces to the form involving only the Laplacian of the thickness field~\cite{Benet2014,alizadeh_pahlavan_cueto-felgueroso_hosoi_mckinley_juanes_2018}}, \\
+\\
+citing the related papers as a reference.
 }
-
 \end{itemize}

### 2022-11-25 Reply_RefereeC.tex
Please work

@@
 Lett.} {\bf 126}, 228004 (2021).
-\JH{We have to make sure to strongly doubt the competence of Ref C in the letter to the editor! I guess this will be the only chance for us to get the paper through/}
 }
@@
 The Referee is right. Indeed there are of course several parameters 
-that may influence the physics of the problem. In this (first) contribution, though,
-our aim is to single out the factor determining the novel effect found (the 
-droplet-rivulet duality), namely
-the pattern time dependence (and, hence, the 'wave' speed). 
-One may expect of course that there is a dependence on $\delta \theta$ and $\theta_0$, but this would apply to the static case. 
-A systematic study of the full parameter space would then deserve a dedicated, 
-extended, work that goes somehow beyond the goal of the present paper and a typical PRL contribution.
-Nevertheless, we agree with the Referee that it is appropriate to add further details, therefore we commented in the text 
-on the choice of the checkerboard pattern and of $\theta_0$ and we have added 
-results at changing $\delta \theta$ in a Supplemental Material section.
+that may influence the physics of the problem. 
+Most of them, though, determine the dewetting dynamics even on static heterogeneous surfaces and are not specific of switchable substrates, with time-dependent properties.
+For instance, as correctly identified by the Referee, $\delta \theta$ is one of those 
+parameters.
+Of course, if $\delta \theta \rightarrow 0$ we do not expect to observe rivulets, as much as we do not expect to see arrays of droplets in the static case.
+Indeed, it is a known fact that if $\delta \theta$ were too small, the pattern driving would become too weak and the spinodal instability would set in as the main dewetting mechanism (see, e.g., \cite{KonnurPRL2000}).
+In this contribution, our aim is to single out the factor determining the novel effect found (the droplet-rivulet duality), namely the pattern time-dependence (and, hence, the 'wave' speed).  
+Nevertheless, motivated by the Referee's remark, we have run a test case simulation in 
+the expected rivulet-forming regime ($\Gamma=15$, $\lambda = 256 h_0$), reducing 
+the heterogeneity parameter to $\delta \theta = 5^{\circ}$; 
+the corresponding Minkowski's metric $q_2$ is reported 
+in Fig. \ref{fig:dif_contrast} (together with the case $\delta \theta =10^{\circ}$, used throughout the 
+paper, for comparison): we see that the $q_2$ metric attains the value $q_2 \approx 1$, signalling 
+that the emergence of rivulets is preserved.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.75\textwidth]{gamma15_deltaTheta.pdf}
+    \caption{Comparison of the Minkowski's structure metric $q_2$ for 
+    $\delta\theta = 5^{\circ}, 10^{\circ}$, 
+    $\Gamma = 15$ and $\lambda = 256 h_0$.}
+    \label{fig:dif_contrast}
+\end{figure}
+The rivulets lifetime, though, is shorter than the one for $\delta \theta = 10^{\circ}$.
+We acknowledge, then, that a systematic study of the quantitative dependence of 
+lifetimes on the full parameter space would be important, but we also that such a study
+would deseve a dedicated, extended, work that goes somehow beyond the goal of the present paper and of a typical PRL contribution.
+We have added the new figure and the relative discussion on $\delta \theta$ in the
+Supplementary Material.\\
+We agree, also, that it is appropriate to add further details in the text, therefore we commented in the revised version on the choice of the checkerboard pattern saying that:\\
+\\
+\textcolor{red}{We decorate the substrate with a checkerboard pattern, a common choice that generalizes the broken homogeneity of the stripes to two 
+directions~\cite{KarguptaLangmuir2003,Jalali2018,Nagayama2020,Das2020}.}
 
-{\bf ANDREA: Stefan, we do have these data, right? I think to remember that you ran the simulations}
-
-%The observation of the referee is correct.
-%Indeed, the difference in contact angle amounts to a net force on the liquid.
-%This force can be large enough to actuate %droplets~\cite{doi:10.1021/acs.langmuir.5b02335}.
-
-%The thin film equation is valid only in the regime where $\theta \ll \pi/2$.
-%Choosing $\theta_0 = 20^{\circ}$ and a $\Delta\theta$ of the same magnitude %therefore ensures that we are well within the predictability of our model.
-%Therefore we find that our value of $\theta_0$ is a good compromise to satisfying %the constraints.
-
-%Most of our results are given in terms of $t_0$ and $q_0$.
-%Both contain parameters we choose to set up our numerical experiments. 
-%We supply an equation to calculate them and have no evidence to believe that our theoretical assumptions are incorrect.
-%Could the referee please elaborate why he/she thinks we overlook something?
-
-%\textcolor{pyorange}{Stefan}: I would put the results we have from the x-only %velocity run here.
-%They kind a answer what happens if $\Delta\theta$ becomes larger (rivulets life for %a very short amount of time only).
 }
@@
 I tried hard to understand whether the different cases of $\lambda$ discussed in the paper agree with $2\pi/q_0$ but I was unable to do so. 
+}
+
+\item[ \textbf{{Answer}}]
+{
+We apologize for having lacked thoroughness not providing the numerical 
+value of $q_0$ (of which we reported the expression, though) to compare with the 
+other characteristic scales. Plugging in the values of $\theta_0$,
+$h_0$ and $h_{\ast}$ we get $q_0 \approx 0.09 h_0^{-1}$, whence 
+$\lambda_0 = 2\pi/q_0 \approx 70 h_0$, which corresponds approximately to the shortest pattern wavelength considered. 
+In fact, it is known that for a periodic chemical pattern to be effective in 
+confining droplets over the more hydrophilic patches, the pattern wavelength should 
+not be smaller than the spinodal length scale on the less wettable substrate
+($\lambda_l \approx 50$ in our case) \cite{KarguptaJCP2002,KarguptaLangmuir2000,Nisato1999,Karim1998}.
+We have added comments on this important point in the revised manuscript, when discussing the accessible ranges of pattern wavelength (see also the answer to the comment 5).
+}
+
+\item[ \textbf{\underline{Comment 4.}}]
+{  
 How would the film behave if the dynamics of the film were controlled primarily by spinodal dewetting and not by the contact-angle pattern?
@@
 {
-We apologize for having lacked thoroughness not providing the numerical 
-value of $q_0$ (of which we reported the expression, though) to compare with the 
-other characteristic scales. Plugging in the values of $\Theta = \theta_0 + \delta \theta$, 
-$h_0$ and $h_{\ast}$ we get $q_0 \approx 0.13 h_0^{-1}$, whence 
-$\lambda_0 = 2\pi/q_0 \approx 50 h_0$, so it corresponds approximately to the 
-shortest pattern wavelength considered. The pattern-driven instability overwhelms the spinodal instability because it is active even on an initially
-flat film 
+The pattern-driven instability overwhelms the spinodal instability because it is active even on an initially flat film 
 (i.e., where the initial perturbation $|\delta h_0| \rightarrow 0$) 
@@
 which is also somehow obvious, but it would eventually deprive the pattern of meaning. 
-The emergence of rivulets is observed even for $\lambda \approx \lambda_0$, as we show in a new figure
-(in the Supplemental Material), but their nature is still metastable.
+The emergence of rivulets is observed even for $\lambda \approx \lambda_s$, as we show 
+in Fig. \ref{fig:q2_difflambda}, but their nature is still metastable. 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{lam_256-128-64-51.pdf}
+    \caption{Minkowski's structure metric $q_2$ for three different wavelengths $\lambda=64 h_0$, $\lambda=128 h_0$ and $\lambda=256 h_0$ (as in figure 4 of the main text) and for $\Gamma=1.5$ (all other parameters are as in the main text).
+The time interval during which $q_2 \approx 1$ signals the emergence of rivulets, also for
+    $\lambda = 64 h_0$ which is comparable with the spinodal wavelength $\lambda_s \approx 70 h_0$}
+    \label{fig:q2_difflambda}
+\end{figure}
+We have included also this new figure in the Supplementary Material.
+}
 
-{\bf ANDREA: Stefan your input is needed here, too. Did you make the simulations/figures, with the 
-properly defined $\Gamma$ for $\lambda \approx 2 \pi/q_0$?}
-}
-%We are sorry for creating confusion with the choice of our variables.
-%Let us call $\lambda_0 = 2\pi/q_0$ and rephrase Eq.~(11) in terms of $\lambda_0$ %using $v_0^{\ast}$,
-%\begin{equation*}
-%    v_0^{\ast} = \frac{\lambda_0}{t_0}, 
-%\end{equation*}
-%computing $\Gamma$ in terms of $v_0^{\ast}$ yields
-%\begin{equation*}
-%    \Gamma = \frac{v_0^{\ast}}{U_{\theta}} = \frac{6\pi h_0^3 q_0^3}{\Theta^3}.
-%\end{equation*}
-%Setting this result equal with Eq.~(11) in the manuscript we have
-%\begin{equation*}
-%    \frac{\chi 3\lambda h_0^3 q_0^4}{\Theta^3} = \frac{6\pi h_0^3 q_0^3}{\Theta^3},
-%\end{equation*}
-%with most of the terms canceling
-%\begin{equation*}
-%    \lambda = \frac{2\pi}{q_0\chi}, 
-%\end{equation*}
-%which in fact state that if the wavelength $\chi^{-1}2\pi/q_0 < \lambda$ rivulets %should form. 
-%}
-
-%\item[ \textbf{\underline{Comment 4.}}]
-%{ 
-%How would the film behave if the dynamics of the film were controlled primarily by %spinodal dewetting and not by the contact-angle pattern?
-%}
-
-%\item[ \textbf{{Answer}}]
-%{ 
-%We thank the referee for raising this question.
-%The straightforward answer is we can reproduce the structure factor of a spinodally %dewetting thin film which is given by
-%\begin{equation*}
-%    S(q,t) = S_0 e^{2\omega(q)t},
-%\end{equation*}
-%where is $S_0$ is a constant factor and 
-%\begin{equation*}
-%    \omega(q) = \frac{1}{t_0}\left[2\left(\frac{q}{q_0}\right)^2 - %\left(\frac{q}{q_0}\right)^4\right],
-%\end{equation*}
-%}
-%which can be found in ref.~\cite{PhysRevE.104.034801}, Fig~1.
-%After the generation of droplets, we observe a coarsening scaling according to a %powerlaw.
-
-%\item[ \textbf{\underline{Comment 5.}}]
-%{ 
-%At the same time, if the two characteristic lengthscales ($=$ that of the spinodal %dewetting and that of the patterned substrate) are suitably synchronized, perhaps %rivulets can be rendered permanent rather than transient?
-Or is the dewetting, once %initiated, immanently irreversible even for a
-%time-dependent contact-angle pattern like the one studied in LM17723?
-%}
-
-%\item[ \textbf{{Answer}}]
-%{ 
-%This is an interesting comment.
-%To our understanding, this is not the case as shown in Fig.~5 of the manuscript.
-%We observe a saturation like effect for high pattern velocity.
-%Independent of the velocities, the rivulet admits a varicose instability, which %within our study was not possible to overcome.
-%We hypothesize that another substrate pattern could ensure indefinite rivulet %stability.
-%However as we do not have an intuition, testing N-patterns would be beyond the scope %of this work.
-
-%For further details on the scale synchronization, we would point towards the answer %to Comment 4.
-%}
-
-\item[ \textbf{\underline{Comment 4.}}]
+\item[ \textbf{\underline{Comment 5.}}]
 { 
@@
 For example, the range of achievable wavelengths is limited from below 
-by the need to stay in the lubrication regime, and from above by the size of the system (i.e.
+by the need to stay above the spinodal characteristic length (as explained in the answer to comment 3), and from above by the size of the system (i.e.
 by computing resources).
-We could actually increase the system size and thus push further the upper limit of $\lambda$, but this would only extend the $\lambda^2$ scaling range.
-{\bf ANDREA: May we try one run on a bigger
-system?
-Being 1d it shouldn't take too long, should it?
-It would be also interesting to try a smaller $\Theta$: my handwaving argument on comparing times tells that it should extend a bit the linear, $\sim \lambda$,  scaling range.}
 As for the rivulet lifetimes, analogously, for reasons intrinsic to the stability of the numerical method, much higher pattern speeds cannot be reached
-(in particular one has to comply with the condition $\Gamma \ll 5 \times 10^2$).
+(in particular one has to comply with the condition $\Gamma \ll 5 \times 10^2$). We have added the following
+words of caution in the revised manuscript:\\
+\\
+\textcolor{red}{A few comments on the ranges are noteworthy at this point.
+The wavelengths are obviously limited from above by the size of the computational domain.
+On the other extreme, too short wavelengths are prevented by
+the energetic cost of accomodating droplets on too
+small patches, since the local contact angle (that increases with decreasing $\lambda$) would possibly exceed the largest substrate contact angle, such that the patterning 
+itself would loose meaning. 
+It has been, in fact, proven, both theoretically and experimentally, that 
+for the confinement to be effective, the pattern characteristic size should 
+not be much smaller than the spinodal length scale on the less wettable substrate
+\cite{KarguptaJCP2002,KarguptaLangmuir2000,Nisato1999,Karim1998}.
+We keep, therefore, $\lambda \stackrel{>}{\sim} \lambda_s \approx 70 h_0$. 
+The pattern speeds, on the other hand, are limited by the intrinsic LB bound of the lattice speed of sound
+ \cite{succi}.}\\
+\\
 We tend to disagree, instead, with the comment that any other scaling 
@@
 is 'OK' with the numerical data.
-We think, anyhow, that the Referees' concerns are legitimate and we have, therefore, added 
-comments of caution about the scaling laws in the revised text.
+%We think, anyhow, that the Referees' concerns are legitimate and we have, therefore, added 
+%comments of caution about the scaling laws in the revised text.
 Let us remark, though, that even in real microfluidic systems, 

### 2022-11-25 SuppMat.tex
Please work

\documentclass[amsmath,amssymb,showpacs,prl,superscriptaddress,notitlepage]{revtex4-1}

\usepackage{natbib}

\usepackage{graphics,graphicx,dcolumn,bm,fleqn,epic,eepic,float}
\usepackage{amssymb,amsmath,multirow,rotate,rotating,color}
\usepackage[utf8]{inputenc}

\usepackage[english]{babel}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\tikzset{fontscale/.style = {font=\relsize{#1}}}
\usetikzlibrary{calc}
\usepackage{color}
\usepackage{xcolor}
\usepackage{physics}
%\usepackage[colorlinks=true,linkcolor=blue]{hyperref}%
\hyphenation{title}

\begin{document}

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}
\definecolor{jlblue}{rgb}{0.0,0.6056031611752245,0.9786801175696073}
\definecolor{jlorange}{rgb}{0.8888735002725198,0.43564919034818994,0.2781229361419438}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}

\title{Controlling the dewetting morphologies of thin liquid films by switchable substrates: Supplemental Material}

\author{S. Zitz}
\email{zitz@ruc.dk}
 %Changed order: Work done while Stefan worked in Nuremberg
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
  \affiliation{Department of Chemical and Biological Engineering, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
  \affiliation{IMFUFA, Department of Science and Environment,\\ 
Roskilde University, Postbox 260, DK-4000 Roskilde, Denmark}%
\author{A. Scagliarini}%
\email{andrea.scagliarini@cnr.it}
 \affiliation{Institute for Applied Mathematics "M. Picone" (IAC), 
Consiglio Nazionale delle Ricerche (CNR),\\
Via dei Taurini 19, 00185 Rome, Italy}%
\affiliation{INFN, sezione Roma ``Tor Vergata'', via della Ricerca Scientifica 1, 00133 Rome, Italy}
\author{J. Harting}
\email{j.harting@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
 \affiliation{Department of Chemical and Biological Engineering and Department of Physics, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\date{\today}

\maketitle
\section{Minkowski's structure metric $q_2$ in an extended parameter space}
\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{lam_256-128-64-51.pdf}
    \includegraphics[width=0.4\textwidth]{gamma15_deltaTheta.pdf}
    \caption{LEFT PANEL. Minkowski's structure metric $q_2$ for three different wavelengths $\lambda=64 h_0$, $\lambda=128 h_0$ and 
    $\lambda=256 h_0$ (as in figure 4 of the main text) and for $\Gamma=1.5$ (all other parameters are as in the main text). 
    The time interval during which $q_2 \approx 1$ signals the emergence of rivulets, also for $\lambda = 64 h_0$ which is comparable with the spinodal wavelength $\lambda_s \approx 70 h_0$. RIGHT PANEL. Comparison of the Minkowski's structure metric $q_2$ for $\delta\theta=5^{\circ}$ and $\delta\theta=10^{\circ}$, 
    $\Gamma = 15$ and $\lambda = 256 h_0$.}
    \label{fig:q2_difflambda}
\end{figure}
%\begin{figure}
%    \centering
%    \includegraphics[width=0.7\textwidth]{different_contrasts.pdf}
%    \caption{Comparison of the Minkowski's structure metric $q_2$ for different values of %$\delta\theta$, 
%    $\Gamma = 1.5$ and $\lambda = 256 h_0$.}
%    \label{fig:dif_contrast}
%\end{figure}
\noindent In this section we test the robustness of the observation of the rivulet 
state over a wider parameter space and in particular at changing: 1) the 
pattern wavelength $\lambda$ and 2) the heterogeneity amplitude $\delta \theta$.
To this aim we have run simulations with $\Gamma=1.5$, $\lambda/h_0 = 64, 128$, and with $\Gamma = 15$, $\lambda = 256 h_0$, $\delta \theta = 5^{\circ}$, respectively.
In Fig.~\ref{fig:q2_difflambda} (left panel) we report the measurements of the Minkowski's structure metric $q_2$, whose 
increase from zero up to a plateauing value of $q_2 \approx 1$ indicates the emergence of rivulets, for 
the different $\lambda$'s ($\lambda=256 h_0$, which is the value considered in the main text, is also reported 
for comparison).
Analogously, in the right panel, we report $q_2$ as a function of time for  
$\delta \theta = 5^{\circ}$  and $\delta \theta = 10^{\circ}$ (the value used in the main text). 
We see that the $q_2$ metric attains the value $q_2 \approx 1$, signalling the rivulet state, albeit over 
a time interval shorter than for $\delta \theta = 10^{\circ}$, i.e. the rivulets lifetime decreases with 
$\delta \theta$. This was somehow expected, since obviously (and also in the static case) 
the patterning looses effectiveness as the contact angle mismatch is reduced (see, e.g.~\cite{KonnurPRL2000}).

\newpage

\section{Droplet shape}

\noindent We investigate, here, how the local contact angle of droplets, formed on the more
hydrophilic patches after dewetting, depends on the pattern wavelength. We recall, in fact,
that the patterning is such that the contact angle is not piecewise constant, but varies
with continuity.
The droplet shape is determined by minimization of the total interfacial energy:
\begin{equation}\label{eq:energy}
E = \gamma_{\text{lg}} A_{\text{lg}} + \int_{A_{\text{sl}}} (\gamma_{\text{sl}} - \gamma_{\text{sg}})d \sigma,
\end{equation}
where $A_{\text{lg}}$ and $A_{\text{sl}}$ are the liquid/gas and solid/liquid interface areas, and $\gamma_{\text{lg}}$, $\gamma_{\text{sl}}$,
$\gamma_{\text{sg}}$ are the liquid/gas, solid/liquid and solid/gas interface energies 
per unit area \cite{WuSM2020}. 
In particular, $\gamma_{\text{lg}} \equiv \gamma$ is the surface tension.
Setting $A_{\text{lg}} \equiv A$ and $A_{\text{sl}} \equiv S$, by Young's equation 
$\cos \theta = \frac{\gamma_{\text{sg}} - \gamma_{\text{sl}}}{\gamma}$,
Eq.~(\ref{eq:energy}) can be rewritten as:
\begin{equation}\label{eq:energyoung}
\tilde{E} \equiv \frac{E}{\gamma} = A - \int_S \cos \theta \text{d}x \text{d}y.
\end{equation}
Droplets will form around minima of the contact angle pattern
\begin{equation}\label{eq:contact}
\theta(x,y) = \theta_0 + \delta \theta \sin(q_{\theta} x) \sin (q_{\theta} y) \qquad q_{\theta} = \frac{2\pi}{\lambda},
\end{equation}
namely $(x_n,y_n) = \left((2n+1)\frac{\lambda}{4},(2n+3)\frac{\lambda}{4}\right)$, with
$n=0,\pm 1, \pm 2,\dots$.
If we consider large wavelengths ($q_{\theta} h_0 \ll 1$) and heterogeneity
($\delta \theta \ll 1$) such that the contact angle gradients are small, the
expression (\ref{eq:contact}) can be expanded as
\begin{equation}\label{eq:contact2}
\theta(x,y) \approx \theta_m + \delta \theta q_{\theta}^2 ((x-x_n)^2 + (y - y_n)^2) + o(|\mathbf{x}-\mathbf{x}_n|^2),
\end{equation}
where $\theta_m = \theta_0 - \delta \theta$.
This local radial symmetry allows to approximate the equilibrium droplet shape as a spherical cap of
height $h$ and base radius $a$, whose area is $A = \pi(a^2 + h^2)$; inserting the expression for
$A$ and (\ref{eq:contact2}), neglecting higher than second order terms, in (\ref{eq:energyoung}) gives:
\begin{equation}\label{eq:energy3}
  \tilde{E}(h,a) = \pi(a^2 + h^2) -
  \int_{\mathcal{C}_a(\mathbf{x}_n)} \cos\left[\theta_m + \delta \theta q_{\theta}^2 ((x-x_n)^2 + (y - y_n)^2)\right]
    \text{d}x\text{d}y,
\end{equation}  
where
$\mathcal{C}_a(\mathbf{x}_n)=\{(x,y) \in [0, L]^2|(x-x_n)^2 + (y-y_n)^2 \leq a^2\}$
is the circle of centre $\mathbf{x}_n$ and radius $a$. 
Due to global volume conservation and assuming the droplets to be monodisperse, the droplet
volume is $V_d = h_0L^2/N_d = h_0 \lambda^2/2$,
where $N_d$ is the number of droplets, which equals the number of
minima of (\ref{eq:contact}) in the domain $[0, L]^2$, i.e. $N_d = 2(L/\lambda)^2$.
Enforcing the volume of the spherical cap to be equal to $V_d$ relates $h$ and $a$ by
\begin{equation}
\frac{\pi h}{6}(3a^2 + h^2) \approx \frac{\pi}{2} a^2 h =  \frac{h_0 \lambda^2}{2},    
\end{equation}
in the ``lubrication approximation'' $h \ll a$, whence
\begin{equation}
  h \approx \left(\frac{h_0}{\pi}\right) \left(\frac{\lambda}{a}\right)^2.
\end{equation}  
Inserting the latter expression in (\ref{eq:energy3}) and performing the integral, the
energy (that we indicate now as $E(a)$ to lighten the notation) reads:
\begin{equation}\label{eq:energy4}
  E(a) = \pi \left(a^2 + \frac{h_0^2 \lambda^4}{\pi^2 a^4} \right) - \frac{2\pi}{\delta \theta q_{\theta}^2}
  \left[\sin\left(\theta_m +\frac{\delta \theta}{2}q_{\theta}^2a^2\right) - \sin \theta_m \right].
\end{equation}  
We expand, then, the sine in the second term up to second order in $\delta \theta$ (such that the
energy is first order) and we finally get:
\begin{equation}\label{eq:energyfin}
  E(a) \approx  \pi \left((1-\cos \theta_m) a^2 + \frac{h_0^2 \lambda^4}{\pi^2 a^4} \right)
  + \delta \theta \frac{\pi}{4}\sin \theta_m q_{\theta}^2 a^4 \equiv E_0(a) + \delta \theta E_1(a).
\end{equation}  
The minimum condition $\frac{\partial E}{\partial a} = 0$ \cite{footnote} yields
\begin{equation}\label{eq:minim}
2\pi (1-\cos \theta_m)a^6 - 4h_0 \lambda^4 + \delta \theta \pi^2 \sin \theta_m q_{\theta}^2 a^8 =0.
\end{equation}  
The solution of (\ref{eq:minim}) at zero-th order in $\delta \theta$ is
\begin{equation}\label{eq:a0}
  a_0 = \left[\frac{2h_0\lambda^4}{\pi^2(1-\cos \theta_m)}\right]^{1/6};
\end{equation}
to this order the droplet contact angle $\tan (\theta_d/2) = h/a$ reads
\begin{equation}
\tan \left(\frac{\theta_d}{2}\right) = \frac{h}{a} \approx
\left(\frac{h_0}{\pi}\right)\frac{\lambda^2}{a_0^3} =
\left(\frac{1-\cos \theta_m}{2}\right)^{1/2}\frac{\lambda^2}{(\lambda^{2/3})^3} \approx \frac{\theta_m}{2} \quad \Rightarrow \quad \theta_d \approx \theta_m,
\end{equation}
i.e., it does not depend on $\lambda$. We move, then, to the next order. We seek a solution to
(\ref{eq:minim}) in the form $a = a^{(0)} + \delta \theta a^{(1)} + \dots$, where $a^{(0)} \equiv a_0$;
at the first order in $\delta \theta$ we get:
\begin{equation}
a^{(1)} = -q_{\theta}^2 a_0^3\frac{\sin \theta_m}{6(1-\cos \theta_m)},
\end{equation}
such that the correction to the droplet contact angle provides:
\begin{equation}
  \tan \left(\frac{\theta_d}{2}\right) \approx \frac{h_0 \lambda^2}{\pi a_0^3}
  \left(1+\delta \theta q_{\theta}^2 a_0^2 \frac{\sin \theta_m}{12(1-\cos \theta_m)}\right)
  \quad \Rightarrow \quad \theta_d \approx
  \theta_m \left(1+\frac{\delta \theta}{\theta_m}\frac{b}{\lambda^{2/3}}\right),
\end{equation}  
where $b = \left(\frac{32 h_0^2 \pi^4}{27 \theta_m^5}\right)^{1/3}$,
i.e. the droplet contact angle grows at decreasing pattern wavelength as $\theta_d \sim \lambda^{-2/3}$.

\begin{thebibliography}{99}

\bibitem{KonnurPRL2000} R. Konnur, K. Kargupta and A. Sharma, {\it Phys. Rev. Lett.} {\bf 84}, 931--934 (2000).
\bibitem{WuSM2020} Y. Wu, F. Wang, S. Ma, M. Selzer and B. Nestler,
{\it Soft Matter} {\bf 16}, 6115--6127 (2020).
\bibitem{footnote} It can be easily checked that $\frac{\partial^2 E}{\partial a^2} > 0$ for $a>0$.

\end{thebibliography}

\end{document}

### 2022-11-25 letter_to_editor.tex
Please work

\documentclass[12pt,english]{article}

\usepackage{natbib}

\usepackage{graphics,graphicx,dcolumn,bm,fleqn,epic,eepic,float}
\usepackage{amssymb,amsmath,multirow,rotate,rotating,color}
\usepackage[utf8]{inputenc}

\usepackage[english]{babel}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\tikzset{fontscale/.style = {font=\relsize{#1}}}
\usetikzlibrary{calc}
\makeatother

\newcommand{\figref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\eqnref}[1]{Eq.~(\ref{eq:#1})}
\newcommand{\ts}{\textsuperscript}

\definecolor{tuered}{RGB}{214,0,74}

\newcommand{\todo}[1]{\textbf{\textcolor{tuered}{ TODO: #1}}}

\newcommand{\vectornorm}[1]{\left|\left|#1\right|\right|}

\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\uvec}[1]{\hat{\vec{#1}}}
\newcommand{\tensor}[1]{\mathbf{#1}}

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}

\newcommand{\JH}[1]{\textcolor{blue}{JH: #1}}
\begin{document}

\noindent Dear Editor,\\
\\
\noindent We gratefully thank you for sending us the reports on our manuscript. We are glad to see 
that the Referees agree that the work is interesting and scientifically relevant, with one of them 
(Referee B) inclined to accept it. 
We have made a major revision taking into account all Referees' reasonable remarks, including novel 
results from new simulations and theoretical derivations (see the new Supplementary Material).
Let us, please, stress further how and why we deem that the present paper meets the Physical
Review Letters criteria of impact and innovation (interest, as we already noted, was acknowledged by the Referees).
We have found a novel effect in the dewetting on substrates with properly tailored time dependent properties, and 
suggested how to exploit it to control the dewetting morphologies. 
We think that the impact of this discovery is out of question, as witnessed by the various publications, also on in PRL on dewetting and closely related problems
(just to cite but a few recent ones: J.C. Fern\'andez-Toledano et al, ``How wettability controls nanoprinting'',
{\it Phys. Rev. Lett.} {\bf 124} 224503 (2020); A.A. Pahlavan et al,
``Evaporation of binary-mixture liquid droplets: the formation of picoliter pancakelike shapes'', {\it Phys. Rev. Lett.} {\bf 127}, 024501 (2021); 
C. Clavaud et al, ``Modification of the fluctuation dynamics of ultrathin wetting films'', {\it Phys. Rev. Lett.} {\bf 126}, 228004 (2021)).
What it is questionable here, instead, is the reliability of some of the Referees in judging a paper on microscale multiphase fluid dynamics: one of 
them (Referee A) ignores the concept of disjoining pressure, confuses dewetting with prewetting and concepts from the physics of free surface flows to wetting.
The main ``argument'' of Referee C against publication is that he/she does not see ``how would the transport of a fluid by rivulets 
surpass the transport by droplets''; but the paper is about controlling morphologies, not about transport.
On the other hand, we acknowledge that certain points were not explained and/or discussed adequately, so we have put a lot of effort in clarifying all
doubts and concerns. 
Therefore, we would like to bring to your attention again our manuscript in its new form, confident that it might now be suitable to appear in Physical Review Letters.\\
\\
With our best regards,\\
the Authors.

\end{document}

### 2022-11-25 main_AS_final.tex
Please work

--- 2022-10-03 main_AS.tex
@@
 The mobility function
-\begin{equation}\label{eq:mobility}
-  M_{\delta}(h) = \frac{2h^3 + 6\,\delta\, h^2 + 3\delta^2h}{6\mu},
-\end{equation}  
+$M_{\delta}(h) = \frac{2h^3 + 6\,\delta\, h^2 + 3\delta^2h}{6\mu}$
 depends on the velocity boundary condition at the substrate, parameterized by an effective slip length $\delta$
 (for $\delta \rightarrow 0$ it reduces to the no-slip form $h^3/(3\mu)$). Here, $\mu$ is the fluid dynamic viscosity.
-The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$, 
+The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$~\footnote{\textcolor{red}{The Laplace pressure is the product of the surface tension
+and the local curvature, therefore its full expression reads $p_L = -\gamma \nabla^2 h/\sqrt{1+|\nabla h|^2}$; 
+since we are in the lubrication regime, though, whereby $|\nabla h| \ll 1$, the denominator can be safely neglected 
+and the pressure reduces to the form involving only the Laplacian of the thickness 
+field~\cite{Benet2014,alizadeh_pahlavan_cueto-felgueroso_hosoi_mckinley_juanes_2018}.}}, 
 \textcolor{red}{where $\gamma$ is the surface tension}.
@@
 \end{equation}
-\textcolor{red}{where $f(\xi)=\xi^{-3} - \xi^{-2}$} and $h_{\ast}$ is the height at which the disjoining pressure vanishes and sets the precursor layer thickness.
+\textcolor{red}{where $f(\xi)=\xi^{-3} - \xi^{-2}$} and $h_{\ast}$ is the height at which the disjoining pressure vanishes and sets the precursor layer thickness~\footnote{\textcolor{red}{Expressions of
+the general form $\Pi(h) = \gamma(1-\cos \theta)(n-1)(m-1)/(n-m)((h_{\ast}/h)^n - (h_{\ast}/h)^m)/h_{\ast}$
+are frequently found in the literature, with various combination of exponents $(n,m)$ ($n>m$), depending
+on the type of interactions among fluid molecules and between fluid and substrate~\cite{SCHWARTZ1998173,Mitlin,Teletzke}. A common choice, for instance, is the pair $(9,3)$, which 
+can be derived from molecular interactions of Lennard-Jones type. Here, the use of the $(3,2)$ pair is motivated mainly by the computational advantage of keeping the characteristic dewetting time 
+shorter. We have run also a single case ($\lambda=256 h_0$ and $\Gamma=1.5$) with $(9,3)$ and observed that 
+the dynamics remains essentially the same, but for a shift in the time scale (see also~\cite{PhysRevE.104.034801})}}.
 The time variation of the patterned substrate then enters the model precisely through the disjoining pressure, by making the contact angle space and time dependent, i.e.
 $\theta = \theta(\mathbf{x},t)$.
+\textcolor{red}{We decorate the substrate with a checkerboard pattern, a common choice that generalizes the broken homogeneity of the stripes to two 
+directions~\cite{KarguptaLangmuir2003,Jalali2018,Nagayama2020,Das2020}.}
 In particular, we employ the sinusoidal form
@@
 \end{equation}
-where $q_{\theta} = 2\pi/\lambda$, i.e.~the pattern evolves in time as a plane wave. 
+where $q_{\theta} = 2\pi/\lambda$, i.e.~the pattern evolves in time as a plane wave~\footnote{\textcolor{red}{A laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}.
+The external stimulus, our moving pattern, can be realized by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.}}. 
 We fix the velocity direction to one diagonal, namely $\mathbf{v}_{\theta} = (v_{\theta x},v_{\theta y}) = v_{\theta}(1/\sqrt{2},-1/\sqrt{2})$ (we will return later on the importance of this choice), and we set $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$~\footnote{Notice that, since a typical velocity is such that $v_{\theta} \Delta t \ll \Delta x$ (in one LB time step $\Delta t$ the wave would travel a distance much smaller than a lattice spacing $\Delta x$),
@@
 $(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$ \cite{Mecke_2005,PhysRevE.100.023108}. In our patterned case, we define 
-$q_0^2=h_{\ast}^{-2}(1-\cos\Theta)f^{\prime}(h_0/h_{\ast})$, where 
-$\Theta = \max_{\mathbf{x}} \theta(\mathbf{x}) = \theta_0 + \delta \theta$.
-A velocity scale then also naturally arises, namely $v_0 = h_0/t_0$.}
+  $q_0^2=h_{\ast}^{-2}(1-\cos\theta_0)f^{\prime}(h_0/h_{\ast})$. Correspondingly, we choose as a velocity scale $v_0 = \lambda_s/t_0$, where
+$\lambda_s = 2\pi/q_0$.}
 \begin{figure}
@@
     \includegraphics[width=0.45\textwidth]{Figures/No_vel_render2.png}
-    \caption{Stationary film thickness field ($t>20t_0$) showing the formation of droplets. The color map indicates the contact angle pattern 
+    \caption{Stationary film thickness field ($t>4t_0$) showing the formation of droplets. The color map indicates the contact angle pattern 
     (Eq.~(\ref{eq:sinetheta}) with $v_{\theta}=0$), with lower (higher) values coded in light blue (yellow).
@@
 \end{figure}
+%%%%
+\textcolor{red}{To give a flavour of how to compare with actual experimental systems, let us notice
+first that the expression of $t_0$ depends on the disjoining pressure (through its dependence 
+ $q_0 = \sqrt{\frac{1}{2\gamma} \Pi^{\prime}(h_0)}$) and contains, therefore, energetic details, namely 
+ the Hamaker constant $A_H$.
+Consider, for instance, the problem studied in \cite{becker2003complex}: the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon
+dioxide substrate, where $A_H  = 2.2 \times 10^{-20} \, \text{J}$ and 
+$\gamma = 0.03 \, \text{N}/\text{m}$. Neglecting the short range part of the interface potential
+(such that $\Pi(h)= - \frac{A_H}{6\pi h^3}$)~\cite{Mecke_2005,becker2003complex}, one gets 
+$q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}$.
+The characteristic time of the growth of the instability for these experiments, then, reads \cite{PhysRevLett.99.114503} (the dynamic viscosity of polystyrene is 
+$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$) $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}$ and it can be seen that indeed film rupture occurs within time 
+scales of the order of $t_0$ ($\approx 2 t_0$) \cite{becker2003complex}, as in our 
+simulation for the largest pattern wavelength, $\lambda = L$ (see also our previous work
+with homogeneous substrates~\cite{PhysRevE.104.034801}).}
+%%%%
 All our simulations are run on a bi-periodic square domain of size $L \times L$ with 
-$L/h_0 = 512 $.
-To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$ (see Eqs.~(\ref{eq:mobility})-(\ref{eq:disjoinpressure})).
+$L/h_0 = 512 $. \textcolor{red}{The numerical values in lattice Boltzmann units of the 
+mean film thickness, dynamics viscosity and surface tension are, respectively, $h_0=1$, $\mu=1/6$
+and $\gamma=0.01$.}
+\textcolor{red}{To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$, 
+a value which lies
+within the weak/intermediate slip regime~\cite{peschka2019signatures,fetzer2007quantifying, munch2005lubrication} (we have tried also a value four times smaller, without observing 
+any major qualitative change in the results).} 
 The liquid film is initialized with a height field slightly perturbed around the mean value $h_0$, i.e.
-\begin{equation}\label{eq:hinitial}
-    h(\mathbf{x},0) = h_0 \left[1 + 0.1 \left(\sin\left(\frac{2\pi x}{L}\right)\sin\left(\frac{2\pi y}{L}\right)\right)\right].
-\end{equation}
-Various wavelengths, in the range \textcolor{red}{$\lambda/h_0 \in [50, 512]$}, and velocities, $v_{\theta}/v_0 \in [0, 5 \times 10^3]$, are considered for the wettability pattern, Eq.~(\ref{eq:sinetheta}).
-\textcolor{red}{A comment on the limits of the wavelengths range is noteworthy at this point.
+$h(\mathbf{x},0) = h_0 \left[1 + 0.1 \left(\sin\left(\frac{2\pi x}{L}\right)\sin\left(\frac{2\pi y}{L}\right)\right)\right]$.
+Various wavelengths, in the range \textcolor{red}{$\lambda/h_0 \in [64, 512]$}, and velocities,
+$v_{\theta}/v_0 \in [0, 200]$, are considered for the wettability pattern, Eq.~(\ref{eq:sinetheta}). 
+\textcolor{red}{A few comments on these ranges are noteworthy at this point.
 The wavelengths are obviously limited from above by the size of the computational domain.
-On the other extreme, small values are prevented by
-the limited capability of accomodating droplets of too large height, $h_d$, on too
-small patches, since the local contact angle would possibly exceed the largest substrate contact angle, such that the patterning 
-itself would loose meaning. In fact, if the patches were much smaller than 
-the intrinsic size the droplets would have when formed on a homogeneous substrate
-(fixed by the spinodal characteristic length, $\lambda_s$), the confinement by the 
-pattern would not be effective~\cite{KarguptaPRL2001}
-Besides, the very same lubrication approximation 
-requires small contact angles. 
-We keep, therefore, $\lambda \geq \lambda_s \approx 70 h_0$.}
+On the other extreme, too short wavelengths are prevented by
+the energetic cost of accomodating droplets on too
+small patches, since the local contact angle 
+(that increases with decreasing $\lambda$~\cite{SuppMat}) would possibly exceed the largest substrate contact angle, such that the patterning 
+itself would loose meaning. 
+It has been, in fact, proven, both theoretically and experimentally, that 
+for the confinement to be effective, the pattern characteristic size should 
+not be much smaller than the spinodal length scale on the less wettable substrate
+\cite{KarguptaJCP2002,KarguptaLangmuir2000,Nisato1999,Karim1998}.
+We keep, therefore, $\lambda \stackrel{>}{\sim} \lambda_s \approx 70 h_0$.
+The pattern speeds, on the other hand, are limited by the intrinsic LB bound of the lattice speed of sound
+ \cite{succi}.}\\
 %\textcolor{red}{Hereafter, we will omit 
@@
 The film rupture time, $\tau_r$, is defined as the least $t$ such that $h(\mathbf{x},\tau_r)=h_{\ast}$ (that is, when the free surface "touches" the substrate).
-In Fig.~\ref{fig:model_rt} we report the rupture times, as a function of the wavelength, for stationary ($v_{\theta}=0$) and time-dependent ($v_{\theta}=10 v_0$) patterns. 
+In Fig.~\ref{fig:model_rt} we report the rupture times, as a function of the wavelength, for stationary ($v_{\theta}=0$) and time-dependent ($v_{\theta}=20 v_0$) patterns. 
 It is conveyed that, overall, rupture occurs earlier on the static substrate, suggesting that the time variation tends to stabilize the film, in agreement with linear stability analysis results~\cite{suman2006dynamics}.
-In the stationary case we observe that $\tau_r$ grows linearly with $\lambda$, for short wavelengths, and quadratically for longer $\lambda$.
-Also for $v_{\theta}=10$ the rupture times tend to approach the $\lambda^2$ scaling at large $\lambda$, whereas they tend to saturate to a constant value for short wavelengths.
+We observe that $\tau_r$ grows linearly with $\lambda$, for short wavelengths, and quadratically for longer $\lambda$.
 These facts can be explained, qualitatively, as follows. 
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{Figures/low_lambda_rupture.pdf}
+    \includegraphics[width=0.45\textwidth]{Rupture_times_Gamma_theta_20.pdf}
     %\includegraphics[width=0.45\textwidth]{Figures/Rupture_times_Gamma.pdf}
     % \includegraphics[width=0.45\textwidth]{Figures/rupture_times_all_lambda_v0v10.pdf}
-    \caption{Rupture times, $\tau_r$, as a function of the pattern wavelength $\lambda$, for $v_{\theta}=0$ (\textcolor{jlblue}{$\bullet$}) and $v_{\theta}=10 v_0$ 
+    \caption{Rupture times, $\tau_r$, as a function of the pattern wavelength $\lambda$, for $v_{\theta}=0$ (\textcolor{jlblue}{$\bullet$}) and $v_{\theta}=20 v_0$ 
     (\textcolor{jlorange}{$\star$}).
@@
 \end{figure}
-Interestingly, this phenomenology is qualitatively supported also by the time-dependent case, $v_{\theta} = 10v_0$ (reported in Fig.~\ref{fig:model_rt} as orange stars), \textcolor{red}{but for a smaller prefactor in the linear regime, due to the fact that the pattern wave velocity is large enough to dominate over $U_{\Theta}$ and sets the retraction speed, such that the characteristic time becomes $\tau_R \sim \lambda/v_{\theta}$.}
+%Interestingly, this phenomenology is qualitatively supported %also by the time-dependent case, $v_{\theta} = 20 v_0$ (reported %in Fig.~\ref{fig:model_rt} as orange stars), \textcolor{red}{but %for a smaller prefactor in the linear regime, due to the fact %that the pattern wave velocity is large enough to dominate over %$U_{\Theta}$ and sets the retraction speed, such that the %characteristic time becomes $\tau_R \sim \lambda/v_{\theta}$.}
 %\begin{equation}\label{eq:taur_saturation}
@@
 %\end{equation}
-
 We now focus on the long time dynamics and in particular on the characterization of the dewetting morphologies and how they are affected by the speed of the wettability wave.
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{Figures/static_pattern_test.pdf}
+    \includegraphics[width=0.45\textwidth]{Static_new_theta20t0.pdf}
     \caption{MAIN PANEL: Time evolution of the height fluctuations, $\Delta h(t)$, during the dewetting process on the patterned substrate given by       
     Eq.~(\ref{eq:sinetheta})
-    with $v_{\theta}= 0$ and $\lambda=L$ (blue circles), $\lambda=L/2$ (orange squares) and $\lambda=L/3$ (green stars).
+    with $v_{\theta}= 0$ and $\lambda= 512 h_0$ (blue circles), $\lambda=256 h_0$ (orange squares) and $\lambda=170 h_0$ (green stars).
     The dashed lines indicate the geometrically expected values of the droplet height $h_d$, assuming monodispersity and a perfect spherical cap shape. 
@@
 Notice that the number of droplets converges faster for smaller pattern wavelengths, in line with the observation reported and justified in the previous section that the characteristic dewetting time decreases with the wavelength.\\
-As a further consistency check, we follow the evolution of the quantity
-\begin{equation}\label{eq:deltah}
-  \Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\}-\min_{\mathbf{x}}\{h(\mathbf{x},t)\}.
-\end{equation}
-In the steady state, since the droplets are essentially monodisperse, this observable represents a measure of the mean droplet height $h_d$.
-Assuming that the droplet shape is a spherical cap, it can be estimated as $h_d =\left(\frac{3V_d(1-\cos(\theta_d))}{\pi(2+\cos(\theta_d))}\right)^{1/3}$, where $V_d = \frac{h_0 \lambda^2}{2}$ is the droplet volume and $\theta_d$ is the local contact angle (i.e.
-the value at the contact line).
-This, in turn, depends on $h_d$, through Eq.~(\ref{eq:sinetheta}), thus making the one above an implicit equation to be solved numerically for $h_d = h_d(\lambda)$.
-$\Delta h(t)$ is reported in Fig.~\ref{fig:clusters_v0_sine} for three different wavelengths, $\lambda = \{L/3,L/2,L \}$, together with the theoretically expected $h_d(\lambda)$ (depicted with lines), showing excellent agreement.\\
+\textcolor{red}{In the main panel of the same figure, we see that the height fluctuations $\Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\}-\min_{\mathbf{x}}\{h(\mathbf{x},t)\}$
+grow in time until film rupture and then settle to a constant value, which represents a measure of the mean droplet height $h_d$ (since
+droplets are essentially monodisperse), decreasing with the pattern wavelength (as expectable, due to the decrease of the droplet volume,
+$V_d = \frac{h_0 \lambda^2}{2}$).}\\
+%In the steady state, since the droplets are essentially monodisperse, this observable . Assuming that the droplet shape is a spherical cap,
+%it can be estimated as $h_d =\left(\frac{3V_d(1-\cos(\theta_d))}{\pi(2+\cos(\theta_d))}\right)^{1/3}$, where  is the droplet volume and
+%$\theta_d$ is the local contact angle (i.e. the value at the contact line). This, in turn, depends on $h_d$,
+%through Eq.~(\ref{eq:sinetheta}), thus making the one above an implicit equation to be solved numerically for $h_d = h_d(\lambda)$. 
+%$\Delta h(t)$ is reported in Fig.~\ref{fig:clusters_v0_sine} for three different wavelengths, 
+%$\lambda/h_0 = \{170,256,512 \}$, together with the theoretically expected $h_d(\lambda)$ (depicted with lines),
+%showing excellent agreement.\\
 \begin{figure}
     \centering
-    \includegraphics[width=0.45\textwidth]{Figures/lam_2_inset.pdf}
-    \caption{Time evolution of the second order Minkowski structure metric, $q_2(t)$ (see text for details), for different $\Gamma$ values (see Eq.~(\ref{eq:vel_ratio})), on a substrate with pattern wavelength $\lambda=L/2$.
+    \includegraphics[width=0.45\textwidth]{lam_2_inset_theta20.pdf}
+    \caption{Time evolution of the second order Minkowski structure metric, $q_2(t)$ (see text for details), for different $\Gamma$ values, on a substrate with pattern wavelength $\lambda=256 h_0$.
     The grey-scale insets supply snapshots of the film thickness field associated to the various curves.
@@
 A time-dependent pattern affects the dewetting morphology quite substantially.
-For $v_{\theta} = 0.1 v_0$ we still observe the formation of droplets, similarly to the stationary case ($v_{\theta} = 0$).
+For $v_{\theta} = 0.2 v_0$ we still observe the formation of droplets, similarly to the stationary case ($v_{\theta} = 0$).
 However, these are transported with the contact angle minima, reproducing a somehow similar behaviour recently described in a numerical study of a droplet on a moving wettability step~\cite{GrawitterStark1}.
-If the pattern speed is further increased, for $v_{\theta} = v_0$ we observe the development of rivulet-like structures, aligned with $\mathbf{v}_{\theta}$. 
+If the pattern speed is further increased, for $v_{\theta} = 2 v_0$ we observe the development of rivulet-like structures, aligned with $\mathbf{v}_{\theta}$. 
 The film, in fact, while dewetting in the direction normal to the pattern velocity, is exposed, in the direction of the velocity, to a periodic potential with alternating minima and saddle points, which partially (as we will see) stabilizes the film over ``preferential'' lanes along the diagonals.
@@
     % \includegraphics[width=0.45\textwidth]{Figures/Lig_stab_inset_delta_h.pdf}
-    \includegraphics[width=0.45\textwidth]{Figures/Gamma_riv_lifetime.pdf}
+    \includegraphics[width=0.45\textwidth]{lig_lifetime_Gamma_theta20.pdf}
     \caption{MAIN PANEL: Rivulets life-times, $\tau_{\text{riv}}$, for various $\Gamma$ values.
@@
 The breakup is the result of a varicose mode of the rivulet~\cite{doi:10.1063/1.3211248, PhysRevE.77.061605}, whose wavelength is $\approx\lambda$, such that only $N_{\infty}/2$ droplets are counted after breakup.
-These droplets show a peculiar dynamics, characterized by a periodic sequence of spreading and retraction, driven by the pattern, that we dub "pumping state" (see movie file ligament\_formation\_and\_breakup.mp4 in the Supp.
-Mat.).
+These droplets show a peculiar dynamics, characterized by a periodic sequence of spreading and retraction, driven by the pattern, that we dub "pumping state"~\footnote{see movie file ligament\_formation\_and\_breakup.mp4 in~\cite{SuppMat}.}.
 
@@
 However, as the pattern moves, the instability is tamed due to configurations whereby higher contact angle regions underlie height field maxima, thus tending to revert the fluid flow.
-Heuristically speaking, this means that, if we define $\Delta h(t)$ as in Eq.~(\ref{eq:deltah}), but restricted on the rivulet axis, this should grow exponentially (with a certain growth rate) only when the system is in the unstable configuration.
-Namely, $\Delta h(t) \propto e^{\alpha t}$ (see inset of Fig.~\ref{fig:stab_ligs_lam2}) with a prefactor proportional to the time spent by the rivulet in such a configuration, which goes as $\sim \lambda/v_{\theta}$, therefore $\Delta h(t) \sim \alpha (\lambda/v_{\theta})e^{\alpha t}$.
+Heuristically speaking, this means that, if we evaluate $\Delta h(t)$ restricted on the rivulet axis, this should grow exponentially (with a certain growth rate $\alpha \propto t_0$) only when the system is in the unstable configuration.
+Namely, $\Delta h(t)/\Delta h_0 \propto e^{\alpha t}$ (see inset of Fig.~\ref{fig:stab_ligs_lam2}) with a prefactor proportional to the time spent by the rivulet in such a configuration, which goes as $\sim \lambda/v_{\theta}$, therefore
+$\Delta h(t)/\Delta h_0 \sim \alpha (\lambda/v_{\theta})e^{\alpha t}$. 
 The rivulet life-time can be seen as the rupture time of the structure along its axis, hence such that $\Delta h (\tau_{\text{riv}}) \sim h_0$ \cite{PhysRevE.104.034801}, which yields
 \begin{equation}\label{eq:rivlt}
-    \tau_{\text{riv}} \sim \log(v_{\theta}) \sim \log(\Gamma). 
+    \tau_{\text{riv}} \sim \alpha \log(v_{\theta}) \propto t_0 \log(\Gamma). 
 \end{equation}

### 2022-12-06 Reply_RefereeA.tex
First resubmission

@@
 We are glad to read that he/she finds it of potential interest and in principle 
-suitable of publication, if a thorough revision is performed. 
+suitable for publication, if a thorough revision is performed. 
 We substantially improved and revised the paper for the resubmission and hope that we were able to clarify the obscure points, following all the Referee's remarks.
-Hereafter we provide answers to the various questions/comments raised. All changes in the revised version are reported in red colour.
+Hereafter, we provide answers to the various questions/comments raised. All changes in the revised version are reported in red color.
 
@@
 Thin film hydrodynamics is an intrinsically multiscale phenomenon, whereby there are at least two characteristic length scales: the mean film thickness (let us call it $h_0$)
-and an horizontal length scale $\ell$ (which in our case can be, for example, the pattern wavelength, $\lambda$), where, typically $h_0 \ll \ell$. We chose to use the mean 
+and a horizontal length scale $\ell$ (which in our case can be, for example, the pattern wavelength, $\lambda$), where, typically $h_0 \ll \ell$. We chose to use the mean 
 film thickness as a scale, because it is a relevant length that is common to all the simulations.
@@
 since it does not depend on the pattern wavelength and speed. Actually, one can see that it captures well the time scales of the dewetting process
-from figure 2, where it is shown that the rupture times are of the order of fraction to units of $t_0$.
+from figure 2 of the manuscript, where it is shown that the rupture times for the largest wavelength (which are closer to the homogeneous case) are of the order of $t_0$.
 We recognize, on the other hand that in this perspective the choice of $v_0 = \lambda/t_0$ as a reference velocity looks a bit awkward.
@@
 the phenomenology is complicated by the liquid/substrate (and vapour/substrate) interaction. 
-In such a situation the relevant velocity scale is set by the characteristic speed of contact line 
+In such a situation the relevant velocity scale is set by the characteristic velocity of contact line 
 motion which can be estimated from a viscocapillary balance as 
@@
 in several reviews and undergraduate textbooks, but see, e.g., Craster 
-and Matar \cite{craster2009dynamics}), the disjoining pressure encode 
+and Matar \cite{craster2009dynamics}), the disjoining pressure encodes 
 any kind of interactions among the fluid molecules and the substrate, so 

### 2022-12-06 Reply_RefereeB.tex
First resubmission

@@
 In what follows we provide detailed answers to all Referee's questions and we discuss all the changes made to the manuscript
-(which are typeset in red colour).
+(which are typeset in red color).
 
@@
 a new simulation (for a single pair of pattern wavelength and speed) 
-with the $(9,3)$ exponents and show the results in figure \ref{fig:rivulets_9_3}.
+with the $(9,3)$ exponents and show the results in figure \ref{fig:rivulets_9_3} of this reply.
 \begin{figure}
     \centering
-    \includegraphics[width=0.7\textwidth]{9-3_3-2_comp_Gamma_1-5.pdf}
+    \includegraphics[width=0.7\textwidth]{refBFig_1.pdf}
     \caption{Minkowski metric $q_2$ as a function of time for 
@@
     \centering
-    \includegraphics [width=0.6\textwidth]{slip_measure.pdf}
+    \includegraphics [width=0.6\textwidth]{refBFig_2.pdf}
     \caption{Dewetting rim profiles for various slip parameters $\delta$; notice the presence of a dimple behind the front, typical of 
@@
     \centering
-    \includegraphics [width=0.75\textwidth]{small_slip_two_Gammas.pdf}
+    \includegraphics [width=0.75\textwidth]{refBFig_3.pdf}
     \caption{Minkowski metric $q_2$ as a function of time for two pattern speeds and slip velocity $\delta = 0.25 h_0$, signalling that the emergence of rivulets for $\Gamma>1$
@@
 {
-This is indeed a very interesting comment.
-Although our approach, as theorists/numericists, can be seen more of {\it theoretical
-engineering} and our picture would be, then, unlikely devoid of flaws, we are glad 
-to try to envisage a potential experiment.
+%This is indeed a very interesting comment.
+%Although our approach, as theorists/numericists, can be seen more of {\it theoretical
+%engineering} and our picture would be, then, unlikely devoid of flaws, we are glad 
+%to try to envisage a potential experiment.
+%\JH{??}
 %We thank the Referee for bringing this up.
 %First, we like to note that experiments are not our field of expertise and therefore %our envisioned realization could be flawed.
-
 %The two relevant scales are $t_0$ and $q_0$ (or $2\pi/q_0$).
@@
 %The time scale is in the region of minutes and the wettability pattern wavelengths are %in the interval $(800\,\text{nm}, 4000\,\text{nm})$. 
-A possibility would be to prepare a thin liquid film on a light responsive substrate~\cite{IchimuraEtAl_Science2000}, under the action of controlled 
+A possibility could be to prepare a thin liquid film on a light responsive substrate~\cite{IchimuraEtAl_Science2000}, under the action of controlled 
 external stimuli (a light emitter). 
 An ideal candidate could be a digital multimirror device (DMD).
-%The dimensions of the substrate should be roughly ten times the maximal pattern wavelength %($\approx 40\,\mu \text{m}$) to account for the periodicity.
-%An ideal control of the external stimuli (light) would be a digital multimirror device (DMD).
 This technology has effectively been used for thin film experiments and additive manufacturing~\cite{doi:10.1021/jp301092y, doi:10.1126/science.aax8760}, it allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$)
 with spatial resolution of $\approx 10 \times 10 \, \mu \text{m}^2$ (the size of a pixel).
-%Using low intensity UV light $365\,\text{nm}$ and blue light %$436$nm~\cite{IchimuraEtAl_Science2000}, we assume that the radiation heat is negligible.
 Considering as a reference, for instance, the system studied in ~\cite{becker2003complex,PhysRevLett.99.114503}, namely a $\sim 4 \, \text{nm}$ thick film 
@@
 be feasibly extended both to very high ($\Gamma \gg 1$) and very low ($\Gamma \ll 1$) values. 
-We have added the following footnote after the introduction of the contact angle pattern (Eq.(3) in the revised
-manuscript) commenting on this aspect:\\
-\\
-\textcolor{red}{A laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}. 
-The external stimulus, our moving pattern, can be realized by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.} 
-%\JH{Could one do this with a DMD as here https://doi.org/10.1021/jp301092y ?
-Then, one could use a transparent substrate and just do it from the bottom.
-I think a DMD + light responsive substrate could be a straightforward implementation of a time dependent wettability pattern.}
+We have included these comments in the revised manuscript just before the conclusions.
 
-%\JH{We need to comment on the experimental realization in the paper. Can we summarize this idea to 1-2 sentences?}
 }
@@
 $\lambda \stackrel{>}{\sim} \lambda_s$).
-We have added a footnote, when we introduce the film pressure, saying that:\\
+We have added the following footnote, where we introduce the film pressure:\\
 \\

### 2022-12-06 Reply_RefereeC.tex
First resubmission

@@
 the manuscript well written. Based on his/her important remarks, we have 
-prepared a revised version which we hope will be more convincing also for what concerns
+prepared a revised version which we hope will be convincing also for what concerns
 its impact and interest for the readership of Physical Review Letters.
-Hereafter detailed answers to the various questions raised are reported. All changes 
-in the text are edited in red colour, for the sake of clarity.
+Hereafter, detailed answers to the various questions raised are reported. All changes 
+in the text are edited in red color, for the sake of clarity.
 
@@
 { 
-We say nowhere that the rivulets would transport fluid more efficiently. Actually 
-we do not talk about transport at all. The focus, here, is not on sliding or similar force-driven processes, but 
+It is not stated in the paper that the rivulets would transport fluid more efficiently and actually we do not raise any statements about transport at all.
+The focus of our contribution is not on sliding or similar force-driven processes, but
 on dewetting and, in particular, on the different dewetting {\it morphologies}. 
-It is true, on the other hand, that rivulets are metastable, but, as 
-we showed, their lifetime can be extended increasing the pattern speed
+It is correct, on the other hand, that rivulets are metastable. However, as 
+we showed, their lifetime can be extended by increasing the pattern speed
 (therefore enabling treatment, e.g. curing, of the substrate to 
 fix a given pattern).
-Concerning the breadth and impact to warrant publication in Physical Review Letter, as we think we stressed extensively in the introduction,
-the problem of controlling dewetting, besides being of fundamental interest in Fluid Dynamics (which represents an important part of the 
-PRL readership), it is crucially relevant for applications in 
-nanoprinting (including printable electronics). 
+
+The breadth and impact to warrant publication in Physical Review Letters is in our opinion clearly given: As we stressed extensively in the introduction,
+the problem of controlling dewetting is crucially relevant for applications in 
+nanoprinting (including printable electronics) or the development and application of new smart materials.
+Here, we provide a new approach to control the dewetting morphology by means of switchable substrates which might enable new deposition techniques allowing for unprecedented control.
+Besides the possible applicatio impact, the problem is of fundamental interest in Fluid Dynamics (which represents an important part of the 
+PRL readership).
 This is witnessed by the various publications in PRL on this and closely related problems. To cite but a few recent ones: J.C. Fern\'andez-Toledano et al, "How wettability controls nanoprinting",
@@
 Indeed, it is a known fact that if $\delta \theta$ were too small, the pattern driving would become too weak and the spinodal instability would set in as the main dewetting mechanism (see, e.g., \cite{KonnurPRL2000}).
-In this contribution, our aim is to single out the factor determining the novel effect found (the droplet-rivulet duality), namely the pattern time-dependence (and, hence, the 'wave' speed).  
 Nevertheless, motivated by the Referee's remark, we have run a test case simulation in 
@@
     \centering
-    \includegraphics[width=0.75\textwidth]{gamma15_deltaTheta.pdf}
+    \includegraphics[width=0.75\textwidth]{SupMatFig_2.pdf}
     \caption{Comparison of the Minkowski's structure metric $q_2$ for 
@@
 lifetimes on the full parameter space would be important, but we also that such a study
-would deseve a dedicated, extended, work that goes somehow beyond the goal of the present paper and of a typical PRL contribution.
+would deseve a dedicated and extended work which goes somehow beyond the goal of the present paper and of a typical PRL contribution.
 We have added the new figure and the relative discussion on $\delta \theta$ in the
@@
 
+\JH{Is this enough?}
 }
@@
 {
-We apologize for having lacked thoroughness not providing the numerical 
+We apologize for having lacked thoroughness by not providing the numerical 
 value of $q_0$ (of which we reported the expression, though) to compare with the 
@@
 We have added comments on this important point in the revised manuscript, when discussing the accessible ranges of pattern wavelength (see also the answer to the comment 5).
+
+\JH{Why don't you list the changes anymore in red (here and below)?}
 }
@@
 (i.e., where the initial perturbation $|\delta h_0| \rightarrow 0$) 
-\cite{KonnurPRL2000,KarguptaLangmuir2000,KarguptaPRL2001}; ideally one would need to have a vanishing 
+\cite{KonnurPRL2000,KarguptaLangmuir2000,KarguptaPRL2001}. Ideally, one would need to have a vanishing 
 contact angle mismatch in order to see the spinodal dewetting, namely $\delta \theta \sim \delta h_0/h_0$,
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{lam_256-128-64-51.pdf}
+    \includegraphics[width=0.8\textwidth]{SupMatFig_1.pdf}
     \caption{Minkowski's structure metric $q_2$ for three different wavelengths $\lambda=64 h_0$, $\lambda=128 h_0$ and $\lambda=256 h_0$ (as in figure 4 of the main text) and for $\Gamma=1.5$ (all other parameters are as in the main text).

### 2022-12-06 Old_version/main_AS.tex
First resubmission

--- 2022-10-03 main_AS.tex
@@
 
-\author{S. Zitz}
+\author{S. Zitz}https://www.overleaf.com/project/633b184a2c25f373c2e9f56c
 \email{zitz@ruc.dk}
@@
 (for $\delta \rightarrow 0$ it reduces to the no-slip form $h^3/(3\mu)$). Here, $\mu$ is the fluid dynamic viscosity.
-The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$, 
+The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$~\footnote{\textcolor{red}{The Laplace pressure is the product of the surface tension
+and the local curvature, therefore its full expression reads $p_L = -\gamma \nabla^2 h/\sqrt{1+|\nabla h|^2}$; 
+since we are in the lubrication regime, though, whereby $|\nabla h| \ll 1$, the denominator can be safely neglected 
+and the pressure reduces to the form involving only the Laplacian of the thickness 
+field~\cite{Benet2014,alizadeh_pahlavan_cueto-felgueroso_hosoi_mckinley_juanes_2018}.}}, 
 \textcolor{red}{where $\gamma$ is the surface tension}.
@@
 \end{equation}
-\textcolor{red}{where $f(\xi)=\xi^{-3} - \xi^{-2}$} and $h_{\ast}$ is the height at which the disjoining pressure vanishes and sets the precursor layer thickness.
+\textcolor{red}{where $f(\xi)=\xi^{-3} - \xi^{-2}$} and $h_{\ast}$ is the height at which the disjoining pressure vanishes and sets the precursor layer thickness~\footnote{\textcolor{red}{Expressions of
+the general form $\Pi(h) = \gamma(1-\cos \theta)(n-1)(m-1)/(n-m)((h_{\ast}/h)^n - (h_{\ast}/h)^m)/h_{\ast}$
+are frequently found in the literature, with various combination of exponents $(n,m)$ ($n>m$), depending
+on the type of interactions among fluid molecules and between fluid and substrate~\cite{SCHWARTZ1998173,Mitlin,Teletzke}. A common choice, for instance, is the pair $(9,3)$, which 
+can be derived from molecular interactions of Lennard-Jones type. Here, the use of the $(3,2)$ pair is motivated mainly by the computational advantage of keeping the characteristic dewetting time 
+shorter. We have run also a single case ($\lambda=256$ and $\Gamma=15$) with $(9,3)$ and observed that 
+the dynamics remains essentially the same, but for a shift in the time scale (see also~\cite{PhysRevE.104.034801})}}.
 The time variation of the patterned substrate then enters the model precisely through the disjoining pressure, by making the contact angle space and time dependent, i.e.
@@
 \end{figure}
+%%%%
+\textcolor{red}{To give a flavour of how to compare with actual experimental systems, let us notice
+first that the expression of $t_0$ depends on the disjoining pressure (through its dependence 
+ $q_0 = \sqrt{\frac{1}{2\gamma} \Pi^{\prime}(h_0)}$) and contains, therefore, energetic details, namely 
+ the Hamaker constant $A_H$.
+Consider, for instance, the problem studied in \cite{becker2003complex}: the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon
+dioxide substrate, where $A_H  = 2.2 \times 10^{-20} \, \text{J}$ and 
+$\gamma = 0.03 \, \text{N}/\text{m}$. Neglecting the short range part of the interface potential
+(such that $\Pi(h)= - \frac{A_H}{6\pi h^3}$)~\cite{Mecke_2005,becker2003complex}, one gets 
+$q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}$.
+The characteristic time of the growth of the instability for these experiments, then, reads \cite{PhysRevLett.99.114503} (the dynamic viscosity of polystyrene is 
+$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$) $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}$ and it can be seen that indeed film rupture occurs within time 
+scales of the order of $t_0$ ($\approx 2 t_0$) \cite{becker2003complex}, as in our 
+simulation for the largest pattern wavelength, $\lambda = L$ (see also our previous work
+with homogeneous substrates~\cite{PhysRevE.104.034801}).}~\footnote{\textcolor{red}{A laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}.
+The external stimulus, our moving pattern, can be realized by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.}}
+%%%%
 All our simulations are run on a bi-periodic square domain of size $L \times L$ with 
-$L/h_0 = 512 $.
-To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$ (see Eqs.~(\ref{eq:mobility})-(\ref{eq:disjoinpressure})).
+$L/h_0 = 512 $. \textcolor{red}{The numerical values in lattice Boltzmann units of the 
+mean film thickness, dynamics viscosity and surface tension are, respectively, $h_0=1$, $\mu=1/6$
+and $\gamma=0.01$.}
+\textcolor{red}{To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$, 
+a value which lies
+within the weak/intermediate slip regime~\cite{peschka2019signatures,fetzer2007quantifying, munch2005lubrication} (we have tried also a value four times smaller, without observing 
+any major qualitative change in the results).} 
 The liquid film is initialized with a height field slightly perturbed around the mean value $h_0$, i.e.
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{Figures/low_lambda_rupture.pdf}
+    \includegraphics[width=0.45\textwidth]{Gamma_rupT_with_scaling.pdf}
     %\includegraphics[width=0.45\textwidth]{Figures/Rupture_times_Gamma.pdf}
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{Figures/static_pattern_test.pdf}
+    \includegraphics[width=0.45\textwidth]{Figures/static_pattern_test_no_hd.pdf}
     \caption{MAIN PANEL: Time evolution of the height fluctuations, $\Delta h(t)$, during the dewetting process on the patterned substrate given by       
     Eq.~(\ref{eq:sinetheta})
-    with $v_{\theta}= 0$ and $\lambda=L$ (blue circles), $\lambda=L/2$ (orange squares) and $\lambda=L/3$ (green stars).
+    with $v_{\theta}= 0$ and $\lambda= 512 h_0$ (blue circles), $\lambda=256 h_0$ (orange squares) and $\lambda=170 h_0$ (green stars).
     The dashed lines indicate the geometrically expected values of the droplet height $h_d$, assuming monodispersity and a perfect spherical cap shape. 
@@
 This, in turn, depends on $h_d$, through Eq.~(\ref{eq:sinetheta}), thus making the one above an implicit equation to be solved numerically for $h_d = h_d(\lambda)$.
-$\Delta h(t)$ is reported in Fig.~\ref{fig:clusters_v0_sine} for three different wavelengths, $\lambda = \{L/3,L/2,L \}$, together with the theoretically expected $h_d(\lambda)$ (depicted with lines), showing excellent agreement.\\
+$\Delta h(t)$ is reported in Fig.~\ref{fig:clusters_v0_sine} for three different wavelengths, 
+$\lambda/h_0 = \{170,256,512 \}$, together with the theoretically expected $h_d(\lambda)$ (depicted with lines), showing excellent agreement.\\
 \begin{figure}
@@
     \includegraphics[width=0.45\textwidth]{Figures/lam_2_inset.pdf}
-    \caption{Time evolution of the second order Minkowski structure metric, $q_2(t)$ (see text for details), for different $\Gamma$ values (see Eq.~(\ref{eq:vel_ratio})), on a substrate with pattern wavelength $\lambda=L/2$.
+    \caption{Time evolution of the second order Minkowski structure metric, $q_2(t)$ (see text for details), for different $\Gamma$ values (see Eq.~(\ref{eq:vel_ratio})), on a substrate with pattern wavelength $\lambda=256 h_0$.
     The grey-scale insets supply snapshots of the film thickness field associated to the various curves.

### 2022-12-06 SuppMat.tex
First resubmission

@@
     \centering
-    \includegraphics[width=0.4\textwidth]{lam_256-128-64-51.pdf}
-    \includegraphics[width=0.4\textwidth]{gamma15_deltaTheta.pdf}
+    \includegraphics[width=0.4\textwidth]{SupMatFig_1.pdf}
+    \includegraphics[width=0.4\textwidth]{SupMatFig_2.pdf}
     \caption{LEFT PANEL. Minkowski's structure metric $q_2$ for three different wavelengths $\lambda=64 h_0$, $\lambda=128 h_0$ and 
@@
 with continuity.
-The droplet shape is determined by minimization of the total interfacial energy:
+The droplet shape is determined by minimization of the total interfacial energy
 \begin{equation}\label{eq:energy}
@@
 $\cos \theta = \frac{\gamma_{\text{sg}} - \gamma_{\text{sl}}}{\gamma}$,
-Eq.~(\ref{eq:energy}) can be rewritten as:
+Eq.~(\ref{eq:energy}) can be rewritten as
 \begin{equation}\label{eq:energyoung}
@@
 height $h$ and base radius $a$, whose area is $A = \pi(a^2 + h^2)$; inserting the expression for
-$A$ and (\ref{eq:contact2}), neglecting higher than second order terms, in (\ref{eq:energyoung}) gives:
+$A$ and (\ref{eq:contact2}), neglecting higher than second order terms, in (\ref{eq:energyoung}) gives
 \begin{equation}\label{eq:energy3}
@@
 \begin{equation}
-\frac{\pi h}{6}(3a^2 + h^2) \approx \frac{\pi}{2} a^2 h =  \frac{h_0 \lambda^2}{2},    
+\frac{\pi h}{6}(3a^2 + h^2) \approx \frac{\pi}{2} a^2 h =  \frac{h_0 \lambda^2}{2}    
 \end{equation}
@@
 Inserting the latter expression in (\ref{eq:energy3}) and performing the integral, the
-energy (that we indicate now as $E(a)$ to lighten the notation) reads:
+energy (that we indicate now as $E(a)$ to lighten the notation) reads
 \begin{equation}\label{eq:energy4}
@@
 We expand, then, the sine in the second term up to second order in $\delta \theta$ (such that the
-energy is first order) and we finally get:
+energy is first order) and we finally get
 \begin{equation}\label{eq:energyfin}
@@
 \begin{equation}\label{eq:a0}
-  a_0 = \left[\frac{2h_0\lambda^4}{\pi^2(1-\cos \theta_m)}\right]^{1/6};
+  a_0 = \left[\frac{2h_0\lambda^4}{\pi^2(1-\cos \theta_m)}\right]^{1/6}.
 \end{equation}
-to this order the droplet contact angle $\tan (\theta_d/2) = h/a$ reads
+To this order the droplet contact angle $\tan (\theta_d/2) = h/a$ reads
 \begin{equation}
@@
 i.e., it does not depend on $\lambda$. We move, then, to the next order. We seek a solution to
-(\ref{eq:minim}) in the form $a = a^{(0)} + \delta \theta a^{(1)} + \dots$, where $a^{(0)} \equiv a_0$;
-at the first order in $\delta \theta$ we get:
+(\ref{eq:minim}) in the form $a = a^{(0)} + \delta \theta a^{(1)} + \dots$, where $a^{(0)} \equiv a_0$.
+At the first order in $\delta \theta$ we get
 \begin{equation}
@@
 \end{equation}
-such that the correction to the droplet contact angle provides:
+such that the correction to the droplet contact angle provides
 \begin{equation}

### 2022-12-06 letter_to_editor.tex
First resubmission

@@
 (Referee B) inclined to accept it. 
-We have made a major revision taking into account all Referees' reasonable remarks, including novel 
+\textcolor{red}{In particular, we would like to underline that not a single 
+concern was raised on the physics, namely on the reliability of the results.
+On the other hand, we acknowledge that certain points were not explained and/or discussed adequately, so we have put a lot of effort in clarifying all
+doubts and concerns.}
+We have made, therefore, a major revision taking into account all Referees' reasonable remarks, including novel 
 results from new simulations and theoretical derivations (see the new Supplementary Material).
@@
 We have found a novel effect in the dewetting on substrates with properly tailored time dependent properties, and 
-suggested how to exploit it to control the dewetting morphologies. 
-We think that the impact of this discovery is out of question, as witnessed by the various publications, also on in PRL on dewetting and closely related problems
+suggested how to exploit it to control the dewetting morphologies, 
+\textcolor{red}{even proposing, motivated by one Referee's suggestion, a possible experimental realisation.}
+We think that the impact of this discovery is out of question, as witnessed also by the various publications, also on in PRL on dewetting and closely related problems
 (just to cite but a few recent ones: J.C. Fern\'andez-Toledano et al, ``How wettability controls nanoprinting'',
@@
 C. Clavaud et al, ``Modification of the fluctuation dynamics of ultrathin wetting films'', {\it Phys. Rev. Lett.} {\bf 126}, 228004 (2021)).
-What it is questionable here, instead, is the reliability of some of the Referees in judging a paper on microscale multiphase fluid dynamics: one of 
+What it is somehow questionable, instead, is the trustworthiness of some of the Referees in judging a paper on microscale multiphase fluid dynamics: one of 
 them (Referee A) ignores the concept of disjoining pressure, confuses dewetting with prewetting and concepts from the physics of free surface flows to wetting.
 The main ``argument'' of Referee C against publication is that he/she does not see ``how would the transport of a fluid by rivulets 
-surpass the transport by droplets''; but the paper is about controlling morphologies, not about transport.
-On the other hand, we acknowledge that certain points were not explained and/or discussed adequately, so we have put a lot of effort in clarifying all
-doubts and concerns. 
-Therefore, we would like to bring to your attention again our manuscript in its new form, confident that it might now be suitable to appear in Physical Review Letters.\\
+surpass the transport by droplets''; but the paper is about controlling morphologies, not about transport.\\
+\textcolor{red}{We believe, then, that our manuscript, in its new version, deserves to be brought again to your attention as a contribution to Physical Review Letters, 
+confident that it might be now suitable of publication.}\\
 \\

### 2022-12-06 main_AS_final.tex
First resubmission

@@
 \end{equation}
-by means of a recently developed method, based on a lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801}.
+by means of a recently developed method, based on a lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801, Zitz2022}.
 Eq.~(\ref{eq:thinfilm}) describes, in a lubrication approximation spirit, the evolution of the height field (film thickness) $h(\mathbf{x},t)$, denoting the location of the liquid/air interface.
@@
 \end{equation}
-where $q_{\theta} = 2\pi/\lambda$, i.e.~the pattern evolves in time as a plane wave~\footnote{\textcolor{red}{A laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}.
-The external stimulus, our moving pattern, can be realized by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.}}. 
+where $q_{\theta} = 2\pi/\lambda$, i.e.~the pattern evolves in time as a plane wave.
+\textcolor{red}{A possible laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}.
+The external stimulus, our moving pattern, can be realized, e.g., by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.}
 We fix the velocity direction to one diagonal, namely $\mathbf{v}_{\theta} = (v_{\theta x},v_{\theta y}) = v_{\theta}(1/\sqrt{2},-1/\sqrt{2})$ (we will return later on the importance of this choice), and we set $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$~\footnote{Notice that, since a typical velocity is such that $v_{\theta} \Delta t \ll \Delta x$ (in one LB time step $\Delta t$ the wave would travel a distance much smaller than a lattice spacing $\Delta x$),
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{Figures/No_vel_render2.png}
-    \caption{Stationary film thickness field ($t>4t_0$) showing the formation of droplets. The color map indicates the contact angle pattern 
+    \includegraphics[width=0.45\textwidth]{Figure_1.png}
+    \caption{Stationary film thickness field ($t>t_0$) showing the formation of droplets. The color map indicates the contact angle pattern 
     (Eq.~(\ref{eq:sinetheta}) with $v_{\theta}=0$), with lower (higher) values coded in light blue (yellow).
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{Rupture_times_Gamma_theta_20.pdf}
-    %\includegraphics[width=0.45\textwidth]{Figures/Rupture_times_Gamma.pdf}
-    % \includegraphics[width=0.45\textwidth]{Figures/rupture_times_all_lambda_v0v10.pdf}
+    \includegraphics[width=0.45\textwidth]{Figure_2.pdf}
     \caption{Rupture times, $\tau_r$, as a function of the pattern wavelength $\lambda$, for $v_{\theta}=0$ (\textcolor{jlblue}{$\bullet$}) and $v_{\theta}=20 v_0$ 
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{Static_new_theta20t0.pdf}
+    \includegraphics[width=0.45\textwidth]{Figure_3.pdf}
     \caption{MAIN PANEL: Time evolution of the height fluctuations, $\Delta h(t)$, during the dewetting process on the patterned substrate given by       
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{lam_2_inset_theta20.pdf}
+    \includegraphics[width=0.45\textwidth]{Figure_4.pdf}
     \caption{Time evolution of the second order Minkowski structure metric, $q_2(t)$ (see text for details), for different $\Gamma$ values, on a substrate with pattern wavelength $\lambda=256 h_0$.
@@
     % \includegraphics[width=0.45\textwidth]{Figures/Lig_stab_inset_delta_h.pdf}
-    \includegraphics[width=0.45\textwidth]{lig_lifetime_Gamma_theta20.pdf}
+    \includegraphics[width=0.45\textwidth]{Figure_5.pdf}
     \caption{MAIN PANEL: Rivulets life-times, $\tau_{\text{riv}}$, for various $\Gamma$ values.
@@
 \end{equation}
-This logarithmic dependence is indeed observed in the numerical data as shown in Fig.~\ref{fig:stab_ligs_lam2}.
-
+This logarithmic dependence is indeed observed in the numerical data as shown in Fig.~\ref{fig:stab_ligs_lam2}.\\
+\textcolor{red}{Before concluding, we would like to envisage a possible realization of a 
+dewetting experiment on a switchable substrate of the type modelled by the spatio-temporal 
+contact angle (\ref{eq:sinetheta}). One may think of a thin liquid film cast on a light responsive substrate~\cite{IchimuraEtAl_Science2000}, under the action of controlled 
+external stimuli (a light emitter). 
+An ideal candidate could be a digital multimirror device (DMD).
+This technology has effectively been used for thin film experiments and additive manufacturing~\cite{doi:10.1021/jp301092y, doi:10.1126/science.aax8760}, it allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$)
+with spatial resolution of $\approx 10 \times 10 \, \mu \text{m}^2$ (the size of a pixel).
+Considering as a reference, for instance, the system studied in ~\cite{becker2003complex,PhysRevLett.99.114503}, namely a $\sim 4 \, \text{nm}$ thick film 
+of polystyrene deposited on an oxidized silicon wafer,  
+we evaluate the retraction speed 
+$U_{\Theta} = \frac{\Theta^3 \gamma}{9 \mu}$ to be 
+$U_{\Theta} \approx 10^{-2} \, \mu \text{m}/\text{s}$. 
+The (minimum) {\it pattern speed} can be 
+estimated, from the pixel size with frame rate $\sim 1 s^{-1}$, as 
+$v_{\theta} \sim 10 \, \mu \text{m}/\text{s}$, that would give $\Gamma \sim 10^3$, i.e.
+well within the rivulet regime ($\Gamma > 1$).
+We note, also, that both $U_{\Theta}$ and $v_{\theta}$ can be widely modulated: the former, playing either with the (temperature and molecular weight dependent) viscosity
+or tailoring the substrate to making it more hydrophobic (i.e. increasing the contact angle); the latter, tuning
+the spatial resolution and frame rate of the DMD. We expect, thus, the range of achievable $\Gamma$'s to 
+be feasibly extended both to very high ($\Gamma \gg 1$) and very low ($\Gamma \ll 1$) values.}\\
 \noindent {\it Conclusions.}
@@
 \noindent {\it Acknowledgements}.
-SZ and JH acknowledge financial support from the Deutsche Forschungsgemeinschaft (DFG) within the priority program SPP2171 ``Dynamic Wetting of Flexible, Adaptive, and Switchable Substrates'', project HA-4382/11.
+SZ and JH acknowledge financial support from the Deutsche Forschungsgemeinschaft (DFG) within the priority program SPP2171 ``Dynamic Wetting of Flexible, Adaptive, and Switchable Substrates'', project HA-4382/11 and Project-ID 431791331—CRC1452 ``Catalysis at Liquid Interfaces''.
 

### 2022-12-09 Reply_RefereeA.tex
Resubmission, waiting for Jens

@@
 and both are required to determine the dynamics of dewetting.
-Of course, on the other hand, the disjoining pressure depends also on the surface tension, which appears, in fact, in its expression (3). 
+Of course, on the other hand, the disjoining pressure depends also on the surface tension, which appears, in fact, in its expression (2). 
 We think, though, that part of the confusion might be due to the fact that we do not define the symbol of the surface tension at its first occurrence, but only later. 
@@
 and it can be seen that indeed film rupture occurs within time 
-scales of the order of $t_0$ ($\approx 2 t_0$) \cite{becker2003complex}.
+scales of the order of $t_0$ \cite{becker2003complex}.
 We have reported these comments on comparison with actual experimental 
@@
 $\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$) $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}$ and it can be seen that indeed film rupture occurs within time 
-scales of the order of $t_0$ ($\approx 2 t_0$) \cite{becker2003complex}, as in our 
+scales of the order of $t_0$ \cite{becker2003complex}, as in our 
 simulation for the largest pattern wavelength, $\lambda = L$ (see also our previous work

### 2022-12-09 Reply_RefereeB.tex
Resubmission, waiting for Jens

@@
 external stimuli (a light emitter).
-An ideal candidate could be a digital multimirror device (DMD).
+
+\textcolor{red}{An ideal candidate could be a digital multimirror device (DMD).
 This technology has effectively been used for thin film experiments and additive manufacturing~\cite{doi:10.1021/jp301092y, doi:10.1126/science.aax8760}, it allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$)
@@
 the spatial resolution and frame rate of the DMD. We expect, thus, the range of achievable $\Gamma$'s to 
-be feasibly extended both to very high ($\Gamma \gg 1$) and very low ($\Gamma \ll 1$) values. 
+be feasibly extended both to very high ($\Gamma \gg 1$) and very low ($\Gamma \ll 1$) values.}
+
 We have included these comments in the revised manuscript just before the conclusions.

### 2022-12-09 letter_to_editor.tex
Resubmission, waiting for Jens

@@
 \textcolor{red}{even proposing, motivated by one Referee's suggestion, a possible experimental realisation.}
-We think that the impact of this discovery is out of question, as witnessed also by the various publications, also on in PRL on dewetting and closely related problems
+We think that the impact of this discovery is out of question, as witnessed also by the various publications, also in PRL on dewetting and closely related problems
 (just to cite but a few recent ones: J.C. Fern\'andez-Toledano et al, ``How wettability controls nanoprinting'',

### 2022-12-09 main_AS_final.tex
Resubmission, waiting for Jens

@@
 \definecolor{jlorange}{rgb}{0.8888735002725198,0.43564919034818994,0.2781229361419438}
+\definecolor{jlgreen}{rgb}{0.2422242978521988,0.6432750931576305,0.3044486515341153}
+\definecolor{jlviolet}{rgb}{0.7644401754934356,0.4441117794687767,0.8242975359232758}
 \newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}
@@
 where $q_{\theta} = 2\pi/\lambda$, i.e.~the pattern evolves in time as a plane wave.
-\textcolor{red}{A possible laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}.
-The external stimulus, our moving pattern, can be realized, e.g., by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.}
+% \textcolor{red}{A possible laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}.
+% The external stimulus, our moving pattern, can be realized, e.g., by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.}
 We fix the velocity direction to one diagonal, namely $\mathbf{v}_{\theta} = (v_{\theta x},v_{\theta y}) = v_{\theta}(1/\sqrt{2},-1/\sqrt{2})$ (we will return later on the importance of this choice), and we set $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$~\footnote{Notice that, since a typical velocity is such that $v_{\theta} \Delta t \ll \Delta x$ (in one LB time step $\Delta t$ the wave would travel a distance much smaller than a lattice spacing $\Delta x$),
@@
 $\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$) $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}$ and it can be seen that indeed film rupture occurs within time 
-scales of the order of $t_0$ ($\approx 2 t_0$) \cite{becker2003complex}, as in our 
+scales of the order of $t_0$~\cite{becker2003complex}, as in our 
 simulation for the largest pattern wavelength, $\lambda = L$ (see also our previous work
@@
     Eq.~(\ref{eq:sinetheta})
-    with $v_{\theta}= 0$ and $\lambda= 512 h_0$ (blue circles), $\lambda=256 h_0$ (orange squares) and $\lambda=170 h_0$ (green stars).
-    The dashed lines indicate the geometrically expected values of the droplet height $h_d$, assuming monodispersity and a perfect spherical cap shape. 
+    with $v_{\theta}= 0$ and $\lambda= 512 h_0$ (\textcolor{jlblue}{$\bullet$}), $\lambda=256 h_0$ (\textcolor{jlorange}{$\blacksquare$}) and $\lambda=170 h_0$ (\textcolor{jlgreen}{$\star$}).
+    % The dashed lines indicate the geometrically expected values of the droplet height $h_d$, assuming monodispersity and a perfect spherical cap shape. 
     INSET: Number of droplets, $N(t)$, as a function of time. The three horizontal dashed lines indicate the number of minima of Eq.~(\ref{eq:sinetheta}),
@@
 A time-dependent pattern affects the dewetting morphology quite substantially.
-For $v_{\theta} = 0.2 v_0$ we still observe the formation of droplets, similarly to the stationary case ($v_{\theta} = 0$).
+For $v_{\theta} = 1.7 v_0$ we still observe the formation of droplets, similarly to the stationary case ($v_{\theta} = 0$).
 However, these are transported with the contact angle minima, reproducing a somehow similar behaviour recently described in a numerical study of a droplet on a moving wettability step~\cite{GrawitterStark1}.
-If the pattern speed is further increased, for $v_{\theta} = 2 v_0$ we observe the development of rivulet-like structures, aligned with $\mathbf{v}_{\theta}$. 
+If the pattern speed is further increased, for $v_{\theta} = 17.5 v_0$ we observe the development of rivulet-like structures, aligned with $\mathbf{v}_{\theta}$. 
 The film, in fact, while dewetting in the direction normal to the pattern velocity, is exposed, in the direction of the velocity, to a periodic potential with alternating minima and saddle points, which partially (as we will see) stabilizes the film over ``preferential'' lanes along the diagonals.
@@
 SZ and JH acknowledge financial support from the Deutsche Forschungsgemeinschaft (DFG) within the priority program SPP2171 ``Dynamic Wetting of Flexible, Adaptive, and Switchable Substrates'', project HA-4382/11 and Project-ID 431791331—CRC1452 ``Catalysis at Liquid Interfaces''.
+SZ acknowledge financial support from the Independent Research Fund Denmark through a DFF Sapere Aude Research Leader grant (grant number 9063-00018B).
 

### 2022-12-14 Reply_RefereeB.tex
Resubmission PRL

@@
 since we are in the lubrication regime, though, whereby $|\nabla h| \ll 1$, the denominator can be safely neglected 
-and the pressure reduces to the form involving only the Laplacian of the thickness field~\cite{Benet2014,alizadeh_pahlavan_cueto-felgueroso_hosoi_mckinley_juanes_2018}}, \\
+and the pressure reduces to the form involving only the Laplacian of the thickness field~\cite{Benet2014,juanes2018}}, \\
 \\

### 2022-12-14 coverletter/letter_to_editor.tex
Resubmission PRL

--- 2022-11-25 letter_to_editor.tex
@@
 
+\vspace*{-3cm}
+
 \noindent Dear Editor,\\
-\\
-\noindent We gratefully thank you for sending us the reports on our manuscript. We are glad to see 
+
+we thank you for sending us the constructive reports on our manuscript. We are glad to see 
 that the Referees agree that the work is interesting and scientifically relevant, with one of them 
 (Referee B) inclined to accept it.
-We have made a major revision taking into account all Referees' reasonable remarks, including novel 
+
+In particular, we would like to underline that not a single 
+concern was raised on the physics, namely on the reliability of the results.
+However, we acknowledge that certain points were not explained and/or discussed adequately, and therefore we invested a substantial effort in clarifying all
+doubts and concerns.
+Our major revision takes into account all Referees' reasonable and constructive remarks, including novel 
 results from new simulations and theoretical derivations (see the new Supplementary Material).
+
 Let us, please, stress further how and why we deem that the present paper meets the Physical
 Review Letters criteria of impact and innovation (interest, as we already noted, was acknowledged by the Referees).
-We have found a novel effect in the dewetting on substrates with properly tailored time dependent properties, and 
-suggested how to exploit it to control the dewetting morphologies. 
-We think that the impact of this discovery is out of question, as witnessed by the various publications, also on in PRL on dewetting and closely related problems
+We found a novel effect in the dewetting on substrates with properly tailored time dependent properties and 
+suggest how to exploit it to control the dewetting morphologies. Motivated by a Referee's suggestion, we even propose a possible experimental realisation.
+We think that the impact of this discovery is out of question, as witnessed also by the various recent publications on dewetting and closely related problems, which appeared in PRL 
 (just to cite but a few recent ones: J.C. Fern\'andez-Toledano et al, ``How wettability controls nanoprinting'',
@@
 C. Clavaud et al, ``Modification of the fluctuation dynamics of ultrathin wetting films'', {\it Phys. Rev. Lett.} {\bf 126}, 228004 (2021)).
-What it is questionable here, instead, is the reliability of some of the Referees in judging a paper on microscale multiphase fluid dynamics: one of 
-them (Referee A) ignores the concept of disjoining pressure, confuses dewetting with prewetting and concepts from the physics of free surface flows to wetting.
-The main ``argument'' of Referee C against publication is that he/she does not see ``how would the transport of a fluid by rivulets 
-surpass the transport by droplets''; but the paper is about controlling morphologies, not about transport.
-On the other hand, we acknowledge that certain points were not explained and/or discussed adequately, so we have put a lot of effort in clarifying all
-doubts and concerns. 
-Therefore, we would like to bring to your attention again our manuscript in its new form, confident that it might now be suitable to appear in Physical Review Letters.\\
-\\
+
+We would like to point out a few flaws in the referee reports, which in our opinion have led to a misjudgement of our work.
+Two referees base their criticism on a clear misconception of the aim and background of the paper:
+First, Referee A ignores the concept of disjoining pressure, confuses dewetting with prewetting and concepts from the physics of free surface flows to wetting.
+Second, the main ``argument'' of Referee C against publication is that he/she does not see ``how would the transport of a fluid by rivulets 
+surpass the transport by droplets''. However, as it is clearly stated in the title of our manuscript, the paper is about controlling morphologies, not about transport.
+
+We believe that our manuscript, in its new version, deserves to be brought again to your attention as a contribution to Physical Review Letters and we are 
+confident that it might now be suitable for publication.
+
+~\\
 With our best regards,\\
-the Authors.
+Stefan Zitz, Andrea Scagliarini, and Jens Harting
 

### 2022-12-14 main.tex
Resubmission PRL

@@
 \definecolor{pyred}{HTML}{D62728}
+\definecolor{jlblue}{rgb}{0.0,0.6056031611752245,0.9786801175696073}
+\definecolor{jlorange}{rgb}{0.8888735002725198,0.43564919034818994,0.2781229361419438}
+\definecolor{jlgreen}{rgb}{0.2422242978521988,0.6432750931576305,0.3044486515341153}
+\definecolor{jlviolet}{rgb}{0.7644401754934356,0.4441117794687767,0.8242975359232758}
 \newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}
 
-\title{Switchable substrates and thin film morphologies}
+\title{Controlling the dewetting morphologies of thin liquid films by switchable substrates}
 
 \author{S. Zitz}
-\email{s.zitz@fz-juelich.de}
+\email{zitz@ruc.dk}
+ %Changed order: Work done while Stefan worked in Nuremberg
  \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
-  Forschungszentrum J\"ulich,\\
+  Forschungszentrum J\"ulich,
   F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
   \affiliation{Department of Chemical and Biological Engineering, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
+  \affiliation{IMFUFA, Department of Science and Environment,\\ 
+Roskilde University, Postbox 260, DK-4000 Roskilde, Denmark}%
 \author{A. Scagliarini}%
 \email{andrea.scagliarini@cnr.it}
- \affiliation{Institute for Applied Mathematics "M. Picone" (IAc, \\
-Consiglio Nazionale delle Ricerche,\\
+ \affiliation{Institute for Applied Mathematics "M. Picone" (IAC), 
+Consiglio Nazionale delle Ricerche (CNR),\\
 Via dei Taurini 19, 00185 Rome, Italy}%
@@
 \begin{abstract}
-We study the dynamics of an unstable thin film deposited on a switchable substrate.
-The substrate is itself subject to a time varying function that affects the equilibrium contact angle $\theta_{\text{eq}}(\mathbf{x},t)$.
-We show that in the case of the vanishing time dependence the patterning drives the dewetting process and all fluid is drained into regions of a small $\theta_{\text{eq}}(\mathbf{x})$ and forms droplets.
-Interestingly enough with a time varying equilibrium contact angle we are able to stop the droplet fragmentation and are able to create ligament or rivulet states.
-We show that these rivulet states mainly depend on two quantities, first being the wavelength of the pattern and second being the velocity of the substrate pattern.
-
+\noindent Switchable and adaptive substrates emerged as valuable tools for the control of wetting and actuation of droplet motion.
+Here we report a computational study of the dynamics of an unstable thin liquid film deposited on a switchable substrate, modelled with a space and time varying contact angle.
+With a static pattern, all the fluid is drained into droplets located around contact angle minima, whereas for a sufficiently large rate of wettability variation a state consisting of metastable rivulets is observed.
+A criterion discriminating whether rivulets can be observed or not is identified in terms of a single dimensionless parameter.
+Finally, we show and explain theoretically how the film rupture times, droplet shape and rivulet life time depend on the pattern wavelength and speed.  
 \end{abstract}
@@
 \newcommand{\ts}{\textsuperscript}
-%\tableofcontents
 
-\section{Introduction}\label{sec:intro}
-Thin film flows and especially droplets are an phenomenon we experience in our every day life.
-Rather unknown is the fact that our body uses thin films to cover mucous membranes.
+\noindent {\it Introduction}.
+Wet surfaces and droplets are part of our every-day experience and of numerous industrial processes including coating, tribology, painting and printing, to name but a few~\cite{gross1980fluid,szeri2010fluid,DERYCK1998278,doi:10.1146/annurev.fluid.31.1.347,DASILVASOBRINHO19991204,singh2010inkjet,jo2009evaluation,Wijshoff2010}.
+Moreover, the continuously growing technological interest for lab-on-a-chip devices~\cite{C6LC00387G,Focke} as well as for printable electronics~\cite{Kim_2005, Luechinger_2008} and printable photovoltaics~\cite{Brabec2001,RH20}, whose efficiency relies crucially on a precise control of material deposition upon (de-)wetting of liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e.
+it is not inert but responds dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{ButtEtAl_Langmuir2018,GuoGuo_RSCAdv2016}.
+Several realizations of switchable and adaptive substrates have been proposed~\cite{LiuChem2005, XiCSR2010}, involving smart materials such as polymer brushes~\cite{CohenStuartEtAl_NatMat2010,ayres2007stimuli,YongEtAl_Mat2018}, thermal-responsive hydrogels~\cite{ChenEtAl_SM2010}, light-responsive molecules and microstructures~\cite{IchimuraEtAl_Science2000,DelormeEtAl_Langmuir2005,OscuratoEtAl_AAMI2017} or processes such as electrowetting~\cite{MugeleEtAl_JPCM2005}.
+Probably the simplest, yet non-trivial, modelling of a switchable substrate can be realized by a space and time dependent wettability pattern~\cite{GrawitterStark1}. 
+While a consistent body of theoretical/computational work has been devoted to processes on static heterogeneous substrates, though, the time dependent case is still almost unexplored, with few relevant exceptions focusing on single droplet spreading and sliding~\cite{GrawitterStark1,GrawitterStark2,ThieleHartmann} or limited to an analysis of the linear regime~\cite{suman2006dynamics}.
 
-\begin{equation}\label{eq:thin_film}
-    \partial_t h(\mathbf{x},t) = \nabla\cdot\left(M(h)\nabla p(\mathbf{x},t)\right),
+In this Letter, we study, by means of numerical simulations, the full dewetting dynamics of a thin liquid film deposited over a substrate with a time varying wettability pattern, from rupture to the long time morphology.
+We identify two regimes where the rupture times grow with the pattern wavelength either linearly (on the static pattern) or attain a constant value (in the time dependent case), for short wavelengths, and approach a quadratic law as the wavelength increases.
+These observations are then explained theoretically.
+We show that, by tuning the rate of change (the ``speed'') of the underlying pattern, it is possible to control, to some extent, the dewetting morphology.
+In particular, for large enough pattern speeds, we detect a metastable state, where the film retracts into metastable rivulets, eventually breaking up into multiple droplets.
+We introduce a control parameter that discriminates whether rivulets or just droplets (as in the static situation)
+can be observed and we propose a phenomenological argument that justifies the logarithmic dependence of the rivulets life time on the pattern speed.\\
+\\
+\noindent {\it Method.} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{ReynoldsLubr,RevModPhys.69.931,RevModPhys.81.1131}
+\begin{equation}\label{eq:thinfilm}
+    \partial_t h(\mathbf{x},t) = \nabla\cdot\left(M_{\delta}(h)\nabla p(\mathbf{x},t)\right)
 \end{equation}
-
-\section{Method}\label{sec:method}
-In this section we describe the lattice Boltzmann method (LBM) we use to simulate thin film flows.
-The model we use has been introduced in ref~\cite{PhysRevE.100.033313} and builds upon a class of lattice Boltzmann models originally developed to study the shallow water equations~\cite{Salmon:1999:0022-2402:503, zhou2004lattice, PhysRevE.65.036309}.
-
-\subsection{Thin film lattice Boltzmann method}\label{subsec:LBM}
-The internal degrees of freedom of this method are the distributions functions $f_{\alpha}(\mathbf{x},t)$.
-They describe a distribution of fictitious particles, for this paper we consider the case of $D2Q9$ lattice.
-Such there are nine velocity components ($Q9$) for a simulation in two dimensions ($D2$).
-The dynamics of the thin film and thus the evolution of the distribution functions is given by
-\begin{equation}\label{eq:LBE}
-    \begin{split}
-        &f_{\alpha}(\mathbf{x}+\mathbf{c}^{({\alpha})}\Delta t,t+\Delta t) = \\
-        &\left(1 - \frac{\Delta t}{\tau}\right) f_i(\mathbf{x},t) + \frac{\Delta t}{\tau} f_{\alpha}^{(eq)}(\mathbf{x},t) + \Delta t\mathcal{S}_{\alpha},
-    \end{split}
+by means of a recently developed method, based on a lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801, Zitz2022}.
+Eq.~(\ref{eq:thinfilm}) describes, in a lubrication approximation spirit, the evolution of the height field (film thickness) $h(\mathbf{x},t)$, denoting the location of the liquid/air interface.
+The mobility function
+$M_{\delta}(h) = \frac{2h^3 + 6\,\delta\, h^2 + 3\delta^2h}{6\mu}$
+depends on the velocity boundary condition at the substrate, parameterized by an effective slip length $\delta$
+(for $\delta \rightarrow 0$ it reduces to the no-slip form $h^3/(3\mu)$). Here, $\mu$ is the fluid dynamic viscosity.
+The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$~\footnote{\textcolor{red}{The Laplace pressure is the product of the surface tension
+and the local curvature, therefore its full expression reads $p_L = -\gamma \nabla^2 h/\sqrt{1+|\nabla h|^2}$; 
+since we are in the lubrication regime, though, whereby $|\nabla h| \ll 1$, the denominator can be safely neglected 
+and the pressure reduces to the form involving only the Laplacian of the thickness 
+field~\cite{Benet2014,juanes2018}.}}, 
+\textcolor{red}{where $\gamma$ is the surface tension}.
+The disjoining pressure $\Pi$ can be seen as (minus) the derivative, with respect to the film thickness, of an effective interfacial potential.
+As such, it contains the information on the liquid/solid and solid/gas interactions and, hence, on the wettability, which is parameterized in terms of the contact angle $\theta$~\cite{RevModPhys.81.739, SCHWARTZ1998173}.
+The expression adopted for $\Pi$ is
+\begin{equation}\label{eq:disjoinpressure}
+\textcolor{red}{\Pi(h,\theta) = \frac{2\gamma}{h_{\ast}}(1-\cos(\theta(\mathbf{x},t)))
+  f\left(\frac{h}{h_{\ast}}\right),}
 \end{equation}
-where the $\mathbf{c}^{({\alpha})}$ are the $Q$ discrete lattice velocities that can be computed according to
-\begin{equation}\label{eq:speeds}
-\mathbf{c}^{({\alpha})}  =
-\left\{
-\begin{array}{ll}
-(0,0) & \alpha = 0 \\
-\left[\cos{\frac{(\alpha-1)\pi}{4}}, \sin{\frac{(\alpha-1)\pi}{4}} \right] &  \alpha=1,3,5,7 \\
-\sqrt{2}\left[\cos{\frac{(\alpha-1)\pi}{4}}, \sin{\frac{(\alpha-1)\pi}{4}} \right] & \alpha=2,4,6,8
-\end{array}
-\right.,
+\textcolor{red}{where $f(\xi)=\xi^{-3} - \xi^{-2}$} and $h_{\ast}$ is the height at which the disjoining pressure vanishes and sets the precursor layer thickness~\footnote{\textcolor{red}{Expressions of
+the general form $\Pi(h) = \gamma(1-\cos \theta)(n-1)(m-1)/(n-m)((h_{\ast}/h)^n - (h_{\ast}/h)^m)/h_{\ast}$
+are frequently found in the literature, with various combination of exponents $(n,m)$ ($n>m$), depending
+on the type of interactions among fluid molecules and between fluid and substrate~\cite{SCHWARTZ1998173,Mitlin,Teletzke}. A common choice, for instance, is the pair $(9,3)$, which 
+can be derived from molecular interactions of Lennard-Jones type. Here, the use of the $(3,2)$ pair is motivated mainly by the computational advantage of keeping the characteristic dewetting time 
+shorter. We have run also a single case ($\lambda=256 h_0$ and $\Gamma=1.5$) with $(9,3)$ and observed that 
+the dynamics remains essentially the same, but for a shift in the time scale (see also~\cite{PhysRevE.104.034801})}}.
+The time variation of the patterned substrate then enters the model precisely through the disjoining pressure, by making the contact angle space and time dependent, i.e.
+$\theta = \theta(\mathbf{x},t)$.
+\textcolor{red}{We decorate the substrate with a checkerboard pattern, a common choice that generalizes the broken homogeneity of the stripes to two 
+directions~\cite{KarguptaLangmuir2003,Jalali2018,Nagayama2020,Das2020}.}
+In particular, we employ the sinusoidal form
+\begin{equation}\label{eq:sinetheta}
+   \!\! \theta(\mathbf{x},t) = \theta_0 + \delta\theta\left[\sin\left(q_{\theta} (x+v_{\theta x}t)\right)\sin\left(q_{\theta}(y+v_{\theta y}t)\right)\right],\! 
 \end{equation}
-and $w$ are being the corresponding weights
-\begin{equation}
-w_{\alpha}  =
-\left\{
-\begin{array}{ll}
-\frac{4}{9} & \alpha = 0 \\
-\frac{1}{9} &  \alpha=1,3,5,7 \\
-\frac{1}{36} & \alpha=2,4,6,8
-\end{array}
-\right..
-\end{equation}
-The last term to the right $\mathcal{S}_{\alpha}$ is a (generalized) source term.
-The concrete implementation of the forces depend on the source term as~\cite{https://doi.org/10.1002/fld.4726}
-\begin{equation}\label{eq:source_term}
-    \mathcal{S}_{\alpha} = \frac{3w_{\alpha}}{\mathbf{c}_{\alpha}^2}\mathbf{c}_{\alpha}\cdot\mathbf{F}.
-\end{equation}
-We are able to match thin film dynamics with the usage of a force term comprising of two main components,
-\begin{equation}\label{eq:forcesum}
-    \mathbf{F} = \mathbf{F}_{\mbox{\tiny{fric}}} + \mathbf{F}_{\mbox{\tiny{fric}}}.
-\end{equation}
-The first being a friction force that includes a velocity boundary condition at the fluid substrate interface
-\begin{equation}\label{eq:forcefric}
-\mathbf{F}_{\mbox{\tiny{fric}}} = -\frac{6h\mu\mathbf{u}}{2h^2+6\delta h + 3\delta^2},
-\end{equation}
-where $\mu$ is the viscosity and $\delta$ is a socalled slip length~\cite{RevModPhys.69.931, PhysRevE.100.033313, PhysRevE.63.011208, munch_lubrication_2005}.
-The film thickness $h$ is the zeroth moment of the distribution function $f_{\alpha}$, i.e. $h(\mathbf{x},t) = \sum_{\alpha}f_{\alpha}(\mathbf{x},t)$ see Eq.~(\ref{eq:LBE}). 
-The second component is due to the film pressure and accounts for the minimization of the fluid vapor interface as well as for the wettability of the substrate,
-\begin{equation}\label{eq:forcefilm}
-    \mathbf{F}_{\mbox{\tiny{film}}} = -\frac{\gamma h}{\rho_0}\nabla(\Delta h - \Pi(h)), 
-\end{equation}
-with $\gamma$ being the surface tension and $\rho_0$ being the fluids density, which is $\rho_0 = 1$ and will be dropped in the following. 
-Here we model the interaction between solid substrate and film with a disjoining pressure term~\cite{RevModPhys.69.931, RevModPhys.81.1131, PhysRevE.63.011208},
-\begin{equation}\label{eq:disjoinpressure}
-    \Pi(h) = \kappa(\theta_{\text{eq.}})\left[\left(\frac{h_{\ast}}{h}\right)^n - \left(\frac{h_{\ast}}{h}\right)^m\right],
-\end{equation}
-with $n$ and $m$ being integer exponents, $h_{\ast}$ defines the minimum of the potential and  
-\begin{equation}\label{eq:kappacontactangle}
-    \kappa(\theta_{\text{eq.}}) = (1-\cos\theta_{\text{eq.}}) \frac{(n-1)(m-1)}{(n-m)h_{\ast}},
-\end{equation}
-which therefor allows us to tune the fluids contact angle.
-
-\subsection{Linear stability and characteristic quantities}\label{subsec:charq}
-For further insights concerning the early time dynamics of the unstable film we perform a linear stability analysis.
-Thus we expand Eq.~(\ref{eq:thin_film}) in terms of a small perturbation $\delta h(\mathbf{x},t) = h(\mathbf{x},t) - h_0$ with $h_0$ being the initial mean thickness value at $t=0$.
-Rearanging in terms up to first order in $\delta h$ yields
-\begin{equation}\label{eq:linearstability_realspace}
-    \partial_t \delta h = \frac{h_0^3}{3\mu}\nabla^2\Big(\Pi(\delta h) - \gamma \nabla^2 \delta h\Big).
-\end{equation}
-Under Fourier transformation of $\delta h$ 
-\begin{equation}\label{eq:fouriertrans}
-    \delta h(\mathbf{x},t) = \int(2\pi)^{-2}\delta \tilde{h}(\mathbf{q},t)e^{i\mathbf{q}\cdot\mathbf{x}}\Diff2\mathbf{q},
-\end{equation}
-and upon inserting this in Eq.~(\ref{eq:linearstability_realspace}) we get
-\begin{equation}\label{eq:linearstab_fourier}
-    \partial_t \delta \tilde{h}(\mathbf{q},t) = \omega(\mathbf{q})\delta\tilde{h}(\mathbf{q},t),
-\end{equation}
-with dispersion relation $\omega(q)$, 
-\begin{equation}\label{eq:dispersion}
-    \omega(\mathbf{q}) = \frac{1-(\frac{q^2}{q_0^2}-1)^2}{t_0}.
-\end{equation}
-Both, the maximum of the dispersion relation $q_0$ and therefore the characteristic time $t_0$, are specific to the substrate the film is deposited on and can be calculated using~\cite{PhysRevLett.99.114503}
-\begin{equation}\label{eq:q0_and_t0}
-    q_0^2 = \frac{1}{2}\frac{\partial \Pi(h)}{\partial h}\bigg\rvert_{h=h_0}, \qquad t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}.
-\end{equation}
-
-\subsection{Time and spatially varying contact angles $\theta(\mathbf{x},t)$}\label{subsec:theta}
-To get a spatially varying contact angle $\theta(\mathbf{x},t)$ we make use of the underlying lattice and assign different values for $\theta$ to different grid points. 
-From now on we drop the subscript \textit{eq.} for convenience, such $\theta = \theta_{\text{eq.}}$ unless otherwise stated.
-
-We consider two different patterns and three wavelengths for this manuscript.
-First being the sinusoidal pattern with wavelengths $\lambda \in \{L, L/2, L/3\}$ which is given as
-\begin{equation}\label{eq:sinetheta}
-    \theta_1(x,y) = \theta_0 + \delta\theta\left[\sin\left(\frac{2\pi x}{\lambda}\right)\sin\left(\frac{2\pi y}{\lambda}\right)\right], 
-\end{equation}
-with $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$.
-The second pattern is a triangle wave which accounts for a more realistic szenario, thus $\theta_2$ is given as
-\begin{equation}\label{eq:triangle_wave}
-    \theta_2(x,y) = \theta_0 + \frac{2\delta\theta}{\pi}\left[\sin^{-1}\left(\sin\left(\frac{2\pi x}{\lambda}\right)\right)\sin^{-1}\left(\sin\left(\frac{2\pi y}{\lambda}\right)\right)\right].
-\end{equation}
-
-Both patterns create a contact angle distribution with values within the interval $[10^{\circ}, 30^{\circ}]$ and by definition they are time independent.
-For a time dependent dynamic we simply shift the pattern in a periodic manner, such that
-\begin{equation}\label{eq:theta_shift}
-    \theta(\mathbf{x},t+T) = \theta(\mathbf{x}-\mathbf{1}, t), 
-\end{equation}
-with $T$ being an arbitrary number of lattice Boltzmann time steps.
-
-Due to convenience we artibute a velocity to the pattern which is simply the shift distance divided by the stepping time
-\begin{equation}\label{eq:pattern_speed}
-    v_{\theta} = \frac{\sqrt{2}}{T}.
-\end{equation}
-Since this velocity as well as the choice of $T$ is arbitrary we define a characteristic velocity $v_0$ similar to Eqs.~(\ref{eq:q0_and_t0}) as
-\begin{equation}\label{eq:normvel}
-    v_0 = \frac{\lambda}{t_0(\theta)},
-\end{equation}
-where we use the wavelength $\lambda$ as a characteristic distance and $t_0(30^{\circ})$ as a characteristic time scale.
-Important to note is that $t_0$ is sensitive to the contact angle, and decreases with increasing angle.
-Thus choosing the highest contact angle for the characteristic time scale means $t_0$
-
-\subsection{Numerical system}
-As outlined in Sec.~\ref{subsec:LBM} we simulate the evolution of the thin film with a lattice Boltzmann method.
-For all our simulations we use periodic boundary conditions and a square of size $Lx = Ly = L = 512 \Delta x$ with $\Delta x = \Delta t = 1$.
-The surface tension as well as the viscosity are constants and set to $\gamma = 0.01$ and $\mu = 1/6$.
-Having this value of $\mu$ leaves us with a relaxation time $\tau = 1$.
-To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer and a slip length $\delta = 1$. 
-With this choice of parameters we are almost able simulate a no-slip boundary condition.
-As for our disjoining pressure potential, we use $n=3, m=2$ and $h_{\ast} = 0.07$.
-Inserting this quantities into Eq.~(\ref{eq:q0_and_t0}) with $h_0 = 1$ yields $q_0^2(30^{\circ}) \approx 1.68\cdot10^{-2}$ as well as $t_0(30^{\circ}) \approx 177000\Delta t$ which gives us a first glimpse into relevant time scales.
-
-The initial condition of all our simulations is a film undulated with sinusoidal wave as
-\begin{equation}\label{eq:hinitial}
-    h(\mathbf{x},0) = h_0 \left[1 + \epsilon \left(\sin\left(\frac{2\pi x}{L}\right)\sin\left(\frac{2\pi y}{L}\right)\right)\right],
-\end{equation}
-with $\epsilon = 0.1$ and $h_0 = 1$.
-
-\section{Results}\label{sec:results}
+where $q_{\theta} = 2\pi/\lambda$, i.e.~the pattern evolves in time as a plane wave.
+% \textcolor{red}{A possible laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}.
+% The external stimulus, our moving pattern, can be realized, e.g., by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.}
+We fix the velocity direction to one diagonal, namely $\mathbf{v}_{\theta} = (v_{\theta x},v_{\theta y}) = v_{\theta}(1/\sqrt{2},-1/\sqrt{2})$ (we will return later on the importance of this choice), and we set $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$~\footnote{Notice that, since a typical velocity is such that $v_{\theta} \Delta t \ll \Delta x$ (in one LB time step $\Delta t$ the wave would travel a distance much smaller than a lattice spacing $\Delta x$),
+the time update needs to be interpreted in an integer part sense, that is the pattern is shifted by one $\Delta x$ every $1/v_{\theta x}$ time steps (and equivalently in the $y$-direction).}.
+\textcolor{red}{Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film \cite{Mecke_2005}.
+On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads
+$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$ \cite{Mecke_2005,PhysRevE.100.023108}. In our patterned case, we define 
+  $q_0^2=h_{\ast}^{-2}(1-\cos\theta_0)f^{\prime}(h_0/h_{\ast})$. Correspondingly, we choose as a velocity scale $v_0 = \lambda_s/t_0$, where
+$\lambda_s = 2\pi/q_0$.}
 \begin{figure}
     \centering
-    \includegraphics[width=0.45\textwidth]{Figures/theta_0336.png}
-    \caption{Grayscale image of the height film at $t\approx 2t_0(30^{\circ})$}
-    \label{fig:gray_ligament}
+    \includegraphics[width=0.45\textwidth]{Figure_1.png}
+    \caption{Stationary film thickness field ($t>t_0$) showing the formation of droplets. The color map indicates the contact angle pattern 
+    (Eq.~(\ref{eq:sinetheta}) with $v_{\theta}=0$), with lower (higher) values coded in light blue (yellow).
+    }
+    \label{fig:handtheta}
 \end{figure}
-We now present our results and show how a time dependent wettability or pattern velocity alters the morphology of the film during the dewetting process.
-In the following we consider three time scales.
-First being the initial short time behaviour that ends with the rupture of the film.
-After the film ruptures we observe a transition region that strongly depend on the pattern velocity, see Fig.~\ref{fig:gray_ligament}.
-Lastly we discuss the morphologies of the dewetted film in the long time regime.
+%%%%
+\textcolor{red}{To give a flavour of how to compare with actual experimental systems, let us notice
+first that the expression of $t_0$ depends on the disjoining pressure (through its dependence 
+ $q_0 = \sqrt{\frac{1}{2\gamma} \Pi^{\prime}(h_0)}$) and contains, therefore, energetic details, namely 
+ the Hamaker constant $A_H$.
+Consider, for instance, the problem studied in \cite{becker2003complex}: the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon
+dioxide substrate, where $A_H  = 2.2 \times 10^{-20} \, \text{J}$ and 
+$\gamma = 0.03 \, \text{N}/\text{m}$. Neglecting the short range part of the interface potential
+(such that $\Pi(h)= - \frac{A_H}{6\pi h^3}$)~\cite{Mecke_2005,becker2003complex}, one gets 
+$q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}$.
+The characteristic time of the growth of the instability for these experiments then reads $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}$ (the dynamic viscosity of polystyrene is 
+$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$)~\cite{PhysRevLett.99.114503}. It can be seen that indeed film rupture occurs within time 
+scales of the order of $t_0$~\cite{becker2003complex}, as in our 
+simulation for the largest pattern wavelength, $\lambda = L$ (see also our previous work
+with homogeneous substrates~\cite{PhysRevE.104.034801}).}
+%%%%
+All our simulations are run on a bi-periodic square domain of size $L \times L$ with 
+$L/h_0 = 512 $. \textcolor{red}{The numerical values in lattice Boltzmann units of the 
+mean film thickness, dynamics viscosity and surface tension are, respectively, $h_0=1$, $\mu=1/6$
+and $\gamma=0.01$.}
+\textcolor{red}{To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$, 
+a value which lies
+within the weak/intermediate slip regime~\cite{peschka2019signatures,fetzer2007quantifying, munch2005lubrication} (we have tried also a value four times smaller, without observing 
+any major qualitative change in the results).} 
+The liquid film is initialized with a height field slightly perturbed around the mean value $h_0$, i.e.
+$h(\mathbf{x},0) = h_0 \left[1 + 0.1 \left(\sin\left(\frac{2\pi x}{L}\right)\sin\left(\frac{2\pi y}{L}\right)\right)\right]$.
+Various wavelengths in the range \textcolor{red}{$\lambda/h_0 \in [64, 512]$}, and velocities,
+$v_{\theta}/v_0 \in [0, 200]$, are considered for the wettability pattern, Eq.~(\ref{eq:sinetheta}). 
+\textcolor{red}{A few comments on these ranges are noteworthy at this point.
+The wavelengths are obviously limited from above by the size of the computational domain.
+On the other extreme, too short wavelengths are prevented by
+the energetic cost of accomodating droplets on too
+small patches, since the local contact angle 
+(that increases with decreasing $\lambda$~\cite{SuppMat}) would possibly exceed the largest substrate contact angle, such that the patterning 
+itself would loose meaning. 
+It has been, in fact, proven, both theoretically and experimentally, that 
+for the confinement to be effective, the pattern characteristic size should 
+not be much smaller than the spinodal length scale on the less wettable substrate
+\cite{KarguptaJCP2002,KarguptaLangmuir2000,Nisato1999,Karim1998}.
+We keep, therefore, $\lambda \stackrel{>}{\sim} \lambda_s \approx 70 h_0$.
+The pattern speeds, on the other hand, are limited by the intrinsic LB bound of the lattice speed of sound
+ \cite{succi}.}\\
+%\textcolor{red}{Hereafter, we will omit 
+%to write explicitly the division by $h_0$, $t_0$ and $v_0$ that will be, then, 
+%surmised, unless else specified.}
+\textcolor{red}{The Reynolds number, $Re$, is of the order of $Re \sim 10^{-2}$ on the 
+static substrate and never exceeds the value $\approx 0.2$ in the time dependent case.}
+In Fig.~\ref{fig:handtheta} we show $h(\mathbf{x},t)$ (droplets) and $\theta(\mathbf{x})$ (color coded) for $v_{\theta} = 0$ (i.e.
+the static case) and \textcolor{red}{$\lambda = 256 h_0$}\footnote{For a better visualization we take twice the domain length $L$ and periodically continue the image.}, in late stages of dewetting.
+As expected, droplets form in regions of small contact angles (blue) while the regions of high contact angles (yellow) dewet.\\
 
+\noindent {\it Results.} We start our analysis by studying first how the rupture times depend on the parameters characterising the wettability pattern, namely the wavelength of the contact angle variation, $\lambda$, and wave speed, $v_{\theta}$.
+The film rupture time, $\tau_r$, is defined as the least $t$ such that $h(\mathbf{x},\tau_r)=h_{\ast}$ (that is, when the free surface "touches" the substrate).
+In Fig.~\ref{fig:model_rt} we report the rupture times as a function of the wavelength, for stationary ($v_{\theta}=0$) and time-dependent ($v_{\theta}=20 v_0$) patterns. 
+It is conveyed that, overall, rupture occurs earlier on the static substrate, suggesting that the time variation tends to stabilize the film, in agreement with linear stability analysis results~\cite{suman2006dynamics}.
+We observe that $\tau_r$ grows linearly with $\lambda$, for short wavelengths, and quadratically for longer $\lambda$.
+These facts can be explained, qualitatively, as follows. 
+Let us first notice that the film rupture consists of two processes, each with its associated time scale: the growth of unstable interface perturbation with rate $t_{\theta}^{-1}$, and the retraction of liquid from more hydrophobic regions, occurring in a characteristic time $t_R$.
+For longer pattern wavelengths, the dewetting instability amplifies more slowly, such that it is reasonable to take it as the process determining the rupture time, $\tau_r \sim t_{\theta}$.
+In this case, from the linearized thin-film equation (in one spatial dimension, for simplicity), obtained setting $h=h_0 + \delta h$ with $\delta h \ll h_0$, 
+we can easily see that the exponential growth of the height perturbation is affected by the wettability pattern (variable contact angle) in such a way that $\partial_t (\delta h) \propto (\partial_x^2 (\partial_h\Pi(h_0))) \delta h$.
+Therefore, since the characteristic time $t_{\theta}$ can be estimated dimensionally as $t_{\theta} \sim \delta h/\dot{(\delta h)}$, the rupture time should go as
+\begin{multline}\label{eq:taur_l2}
+    \tau_r \sim t_{\theta} \sim  \delta h/\dot{(\delta h)} \propto \frac{3\mu}{h_0^3}(\partial_x^2 (\partial_h\Pi (h_0)))^{-1} \sim \\
+    t_0 \left(\frac{q_{\theta}}{q_0}\right)^{-2} \propto t_0 q_0^2 \lambda^2.
+\end{multline}
+Conversely, for fast growths ($t_{\theta} \ll t_R$), retraction dominates and fixes the time scale, $\tau_r \sim t_R$. 
+The latter is related to the time the liquid takes to flow out of regions of high contact angle, whose size is $\sim \lambda$. Hence we have 
+\begin{equation}\label{eq:taur_l1}
+ \tau_r \sim \tau_R \propto \frac{1}{U_{\Theta}}\lambda,
+\end{equation}
+where $U_{\Theta}$ is the retraction speed $U_{\Theta} = \frac{\gamma \Theta^3}{9\mu}$~\cite{Edwards2016}, with $\Theta = \max_{\mathbf{x}}\theta(\mathbf{x})$.
 \begin{figure}
     \centering
-    \includegraphics[width=0.45\textwidth]{Figures/makie_try2.png}
-    \caption{Height of the film $h(\mathbf{x},t)$ with the contact angle $\theta(\mathbf{x})$ as contour.}
-    \label{fig:handtheta}
+    \includegraphics[width=0.45\textwidth]{Figure_2.pdf}
+    \caption{Rupture times, $\tau_r$, as a function of the pattern wavelength $\lambda$, for $v_{\theta}=0$ (\textcolor{jlblue}{$\bullet$}) and $v_{\theta}=20 v_0$ 
+    (\textcolor{jlorange}{$\star$}).
+    The continuous and dashed lines indicate the linear, $\sim \lambda$, and quadratic, $\sim \lambda^2$, scaling laws, respectively.
+        }
+    \label{fig:model_rt}
 \end{figure}
-We proceed as follows, first we simulate the dewetting on a given pattern either sinusoidal- or triangular with $\lambda \in \{L, L/2, L/3\}$ without a time dependency in $\theta$.
-In Fig.~\ref{fig:handtheta} we show $h(\mathbf{x},t)$ as surface plot and $\theta_2(\mathbf{x})$ as contour plot for $v_{\theta} = 0, \lambda = L/2$ in late stages of dewetting.
-Subsequent simulations allow for $v_{\theta} \ne 0$.  
-Based on $v_{\theta}$ we can modify the morphology of the dewetting film, in both the intermediate transition regime as well as the late time state.
+%Interestingly, this phenomenology is qualitatively supported %also by the time-dependent case, $v_{\theta} = 20 v_0$ (reported %in Fig.~\ref{fig:model_rt} as orange stars), \textcolor{red}{but %for a smaller prefactor in the linear regime, due to the fact %that the pattern wave velocity is large enough to dominate over %$U_{\Theta}$ and sets the retraction speed, such that the %characteristic time becomes $\tau_R \sim \lambda/v_{\theta}$.}
+%\begin{equation}\label{eq:taur_saturation}
+% \tau_r \sim \tau_R \sim \frac{\lambda}{v_{\theta}} \propto %\frac{\lambda}{v_0} = \lambda\frac{t_0}{\lambda} \sim \mbox{const}.
+%\end{equation}
+We now focus on the long time dynamics and in particular on the characterization of the dewetting morphologies and how they are affected by the speed of the wettability wave.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{Figure_3.pdf}
+    \caption{MAIN PANEL: Time evolution of the height fluctuations, $\Delta h(t)$, during the dewetting process on the patterned substrate given by       
+    Eq.~(\ref{eq:sinetheta})
+    with $v_{\theta}= 0$ and $\lambda= 512 h_0$ (\textcolor{jlblue}{$\bullet$}), $\lambda=256 h_0$ (\textcolor{jlorange}{$\blacksquare$}) and $\lambda=170 h_0$ (\textcolor{jlgreen}{$\star$}).
+    % The dashed lines indicate the geometrically expected values of the droplet height $h_d$, assuming monodispersity and a perfect spherical cap shape. 
+    INSET: Number of droplets, $N(t)$, as a function of time. The three horizontal dashed lines indicate the number of minima of Eq.~(\ref{eq:sinetheta}),
+      which is $2\left(\frac{L}{\lambda}\right)^2$. In the insets, we display snapshots of the stationary droplet states as grey-scale images
+      of the film thickness field, $h(\mathbf{x},t)$.
+      }
+    \label{fig:clusters_v0_sine}
+\end{figure}
+On the stationary substrate, after rupture all the fluid accumulates in droplets centered at contact angle minima.
+Consequently, as we see from the inset of Fig.~\ref{fig:clusters_v0_sine}, where we plot the number of droplets $N(t)$ in time\footnote{A droplet is identified by the set (``cluster'') of points, in the plane, constituting each of the connected components of the set $\{\mathbf{x} \in [0,L]^2 | h(\mathbf{x},t) \geq h_{\ast}$\}; the clusters are determined by means of an algorithm of Hoshen-Kopelman type~\cite{HK}.}, in the steady state ($t \gg t_0$) $N(t)$ attains the value $N_{\infty} = 2(L/\lambda)^2$ (reported as horizontal lines), which equals the minima of Eq.~(\ref{eq:sinetheta}), for $v_{\theta}=0$.
+Notice that the number of droplets converges faster for smaller pattern wavelengths, in line with the observation reported and justified in the previous section that the characteristic dewetting time decreases with the wavelength.\\
+\textcolor{red}{In the main panel of the same figure, we see that the height fluctuations $\Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\}-\min_{\mathbf{x}}\{h(\mathbf{x},t)\}$
+grow in time until film rupture and then settle to a constant value. This represents a measure of the mean droplet height $h_d$ (since
+droplets are essentially monodisperse), decreasing with the pattern wavelength (as expected, due to the decrease of the droplet volume,
+$V_d = \frac{h_0 \lambda^2}{2}$).}\\
+%In the steady state, since the droplets are essentially monodisperse, this observable . Assuming that the droplet shape is a spherical cap,
+%it can be estimated as $h_d =\left(\frac{3V_d(1-\cos(\theta_d))}{\pi(2+\cos(\theta_d))}\right)^{1/3}$, where  is the droplet volume and
+%$\theta_d$ is the local contact angle (i.e. the value at the contact line). This, in turn, depends on $h_d$,
+%through Eq.~(\ref{eq:sinetheta}), thus making the one above an implicit equation to be solved numerically for $h_d = h_d(\lambda)$. 
+%$\Delta h(t)$ is reported in Fig.~\ref{fig:clusters_v0_sine} for three different wavelengths, 
+%$\lambda/h_0 = \{170,256,512 \}$, together with the theoretically expected $h_d(\lambda)$ (depicted with lines),
+%showing excellent agreement.\\
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{Figure_4.pdf}
+    \caption{Time evolution of the second order Minkowski structure metric, $q_2(t)$ (see text for details), for different $\Gamma$ values, on a substrate with pattern wavelength $\lambda=256 h_0$.
+    The grey-scale insets supply snapshots of the film thickness field associated to the various curves.
+    }
+    \label{fig:msm_q2}
+\end{figure}
+A time-dependent pattern affects the dewetting morphology quite substantially.
+For $v_{\theta} = 1.7 v_0$ we still observe the formation of droplets, similarly to the stationary case ($v_{\theta} = 0$).
+However, these are transported with the contact angle minima, reproducing a somehow similar behaviour recently described in a numerical study of a droplet on a moving wettability step~\cite{GrawitterStark1}.
+If the pattern speed is further increased, for $v_{\theta} = 17.5 v_0$ we observe the development of rivulet-like structures, aligned with $\mathbf{v}_{\theta}$. 
+The film, in fact, while dewetting in the direction normal to the pattern velocity, is exposed, in the direction of the velocity, to a periodic potential with alternating minima and saddle points, which partially (as we will see) stabilizes the film over ``preferential'' lanes along the diagonals.
+This makes the chosen velocity direction, $(1/\sqrt{2},-1/\sqrt{2})$ (or, equivalently, the orthogonal one $(1/\sqrt{2},1/\sqrt{2})$), optimal for the formation of rivulets.
 
-\subsection{Early time behaviour}\label{subsec:earlyT}
-Starting with Eq.~(\ref{eq:hinitial}) we have by definition an unstable film that will rupture.
-Taking into account the addition of the wettability gradient we know that fluid will accumulate in regions of high surface energies or small contact angles.
-This is shown in Fig.~\ref{fig:handtheta}, where all droplets are in a region of small contact angles.
+In order to better characterize the various morphologies we apply the theory of Minkowski's functionals. 
+In particular, we employ the second order Minkowski structure metric, $q_2$~\cite{doi:10.1063/1.4774084, Schaller2020}, which can be computed from a Voronoi tessellation of the set of discrete points $(x_i, y_i)$ on the 2D lattice, such that the height field lies above a certain threshold\footnote{The expression is: $q_2 = \frac{1}{N}\sum_j \frac{1}{P_j}|\sum_k L^{(j)}_k e^{2i\phi^{(j)}_k}|$, where the inner sum runs over the edges of length $L^{(j)}_k$, of the $j$-th Voronoi cell, whose perimeter is $P_j$, and $\phi^{(j)}_k$ is the polar angle of the normal to the $k$-th edge.
+The outer sum represents an ensemble average over the $N$ points in the set.}.
+The $q_2$ metric quantifies the degree of anisotropy of the dewetting morphology, so it takes relatively large values if the structures formed display a preferential direction.
+Measuring $q_2$ then enables us to clearly distinguish between the formation of droplets and rivulets: much larger $q_2$ values are attained for the latter type of structure, as we can see in Fig.~\ref{fig:msm_q2}.
+%Our simulations with $\lambda = L/2$ show a clear difference between $v_{\theta} \leq 0.1v_0$ and $v_{\theta}>0.1v_0$  in $q_2$.
+We observe, on the other hand, that such rivulets are metastable and eventually break up into droplets, as indicated by the collapse of $q_2$ at later times. 
+Notice, though, that the $q_2$ signal for any $v_{\theta} >0$ always stays above the one for the static case, suggesting that even the smallest pattern velocity introduces a sizeable deformation of the spherical cap shape.
+\begin{figure}
+    \centering
+    % \includegraphics[width=0.45\textwidth]{Figures/Lig_stab_inset_delta_h.pdf}
+    \includegraphics[width=0.45\textwidth]{Figure_5.pdf}
+    \caption{MAIN PANEL: Rivulets life-times, $\tau_{\text{riv}}$, for various $\Gamma$ values.
+    The dashed line is plotted as a guide to the eye to highlight the logarithmic dependence, in agreement with 
+    the theoretical prediction, Eq.~(\ref{eq:rivlt}).
+    INSET: Height fluctuations, $\Delta h(t)$, vs time, computed along the rivulet axis, for three different $\Gamma$ values.
+    }
+    \label{fig:stab_ligs_lam2}
+\end{figure}
+The breakup is the result of a varicose mode of the rivulet~\cite{doi:10.1063/1.3211248, PhysRevE.77.061605}, whose wavelength is $\approx\lambda$, such that only $N_{\infty}/2$ droplets are counted after breakup.
+These droplets show a peculiar dynamics, characterized by a periodic sequence of spreading and retraction, driven by the pattern, that we dub "pumping state"~\footnote{see movie file ligament\_formation\_and\_breakup.mp4 in~\cite{SuppMat}.}.
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{Figures/All_deltah.pdf}
-    \caption{Evolution of the absolute difference in height, x-axis, up to $t\approx t_0(30^{\circ})$, y-axis, for different wavelengths and patterns.
-    The upper series displays the data for $\theta_1$, see Eq.~(\ref{eq:sinetheta}), and $\lambda = L, L/2, L/3$ in (a,b,c) respectively.
-    Below is the series for $\theta_2$, see Eq.~(\ref{eq:triangle_wave}).
-    Different colors and line styles represent different pattern velocities, all of which are normalized with $v_0$, see Eq.~(\ref{eq:normvel}).}
-    \label{fig:evolutin_deltah}
-\end{figure*}
-Naively one assumes in the early stages of the dewetting process that the perturbations grow exponentially~\cite{RevModPhys.69.931, RevModPhys.81.1131}.
-This is not the case in our simulations as the initial conditions for $h$ as well as $\theta$ are the same, see Eq.~(\ref{eq:hinitial}, \ref{eq:sinetheta}).
-We show this by measuring the time resolved difference between the maximum and the minimum of the height field as, 
-\begin{equation}\label{eq:delta_h_measure}
-    \Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\} - \min_{\mathbf{x}}\{h(\mathbf{x},t)\},    
+We argue that the emergence of rivulets is controlled by the competition of two characteristic velocities: the pattern wave speed, $v_{\theta}$, and the retraction speed, $U_{\Theta}$, introduced in Eq.~(\ref{eq:taur_l1}).
+If $U_{\Theta}$ is large as compared to $v_{\theta}$, the film retraction is faster than the transport due to the contact angle field and thus droplets from.
+However, if $v_{\theta}$ is larger than $U_{\Theta}$, then the retracting film has too little time to form droplets and ends up in the metastable rivulet state.
+\textcolor{red}{It appears, therefore, natural to consider the ratio of these two velocities,
+$\Gamma \equiv v_{\theta}/U_{\theta}$, as the discriminating parameter.}
+%Recalling the expression (\ref{eq:t0}) for the reference velocity $v_0$, %we define the parameter 
+%\begin{equation}\label{eq:vel_ratio}
+%    \Gamma = \frac{v_{\theta}}{U_{\Theta}} = \frac{3\lambda h_0^3 %q_0^4}{\Theta^3}\chi 
+%\end{equation}
+%as the ratio of these two velocities,
+%where $\chi \equiv v_{\theta}/v_0$. 
+We see from Fig.~\ref{fig:msm_q2} that indeed rivulets form only for $\Gamma > 1$. 
+Moreover, the larger $\Gamma$, the more stable the rivulets are; in other words, the rivulet life-time, $\tau_{\text{riv}}$, that can be conventionally taken as the time at which the drop of $q_2$ occurs, grows with $v_{\theta}$ (see Fig.~\ref{fig:stab_ligs_lam2}).
+The rivulet itself is, in fact, prone to dewetting, with the liquid accumulating over patches around contact angle minima.
+However, as the pattern moves, the instability is tamed due to configurations whereby higher contact angle regions underlie height field maxima, thus tending to revert the fluid flow.
+Heuristically speaking, this means that, if we evaluate $\Delta h(t)$ restricted on the rivulet axis, this should grow exponentially (with a certain growth rate $\alpha \propto t_0$) only when the system is in the unstable configuration.
+Namely, $\Delta h(t)/\Delta h_0 \propto e^{\alpha t}$ (see inset of Fig.~\ref{fig:stab_ligs_lam2}) with a prefactor proportional to the time spent by the rivulet in such a configuration, which goes as $\sim \lambda/v_{\theta}$, therefore
+$\Delta h(t)/\Delta h_0 \sim \alpha (\lambda/v_{\theta})e^{\alpha t}$. 
+The rivulet life-time can be seen as the rupture time of the structure along its axis, hence such that $\Delta h (\tau_{\text{riv}}) \sim h_0$ \cite{PhysRevE.104.034801}, which yields
+\begin{equation}\label{eq:rivlt}
+    \tau_{\text{riv}} \sim \alpha \log(v_{\theta}) \propto t_0 \log(\Gamma). 
 \end{equation}
-which is shown in Fig.~\ref{fig:evolutin_deltah}.
-In the upper row we show the data for $\theta_1$ while in the lower we show data for $\theta_2$, see Eqs.~(\ref{eq:sinetheta}-\ref{eq:triangle_wave}).
-Different linestyles and colors represent different $v_{\theta}$'s.
-Each column shows a wavelength $\lambda$, such from the left to the right we have $\lambda = L, L/2, L/3$.
-Clearly one can observe the dip in \ref{fig:evolutin_deltah}(a) as the initial perturbation flattens while flowing into regions of lower contact angle.
-This behaviour however is only pronounced for the time independent pattern and the smallest $v_{\theta}$ we use.
-Thus the film is not spinodally dewetting but rather nucleating in regions of high contact angles.
+This logarithmic dependence is indeed observed in the numerical data as shown in Fig.~\ref{fig:stab_ligs_lam2}.\\
+\textcolor{red}{We would like to envisage a possible realization of a 
+dewetting experiment on a switchable substrate of the type modeled by the spatio-temporal 
+contact angle (\ref{eq:sinetheta}). One may think of a thin liquid film cast on a light responsive substrate~\cite{IchimuraEtAl_Science2000}, under the action of controlled 
+external stimuli (a light emitter). 
+An ideal candidate could be a digital multimirror device (DMD).
+This technology has effectively been used for thin film experiments and additive manufacturing~\cite{doi:10.1021/jp301092y, doi:10.1126/science.aax8760}, it allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$)
+with spatial resolution of $\approx 10 \times 10 \, \mu \text{m}^2$ (the size of a pixel).
+Considering as a reference, for instance, the system studied in ~\cite{becker2003complex,PhysRevLett.99.114503}, namely a $\sim 4 \, \text{nm}$ thick film 
+of polystyrene deposited on an oxidized silicon wafer,  
+we evaluate the retraction speed 
+$U_{\Theta} = \frac{\Theta^3 \gamma}{9 \mu}$ to be 
+$U_{\Theta} \approx 10^{-2} \, \mu \text{m}/\text{s}$. 
+The (minimum) {\it pattern speed} can be 
+estimated from the pixel size with frame rate $\sim 1 s^{-1}$ as 
+$v_{\theta} \sim 10 \, \mu \text{m}/\text{s}$, which  would result in $\Gamma \sim 10^3$, i.e.
+well within the rivulet regime ($\Gamma > 1$).
+We note, also, that both $U_{\Theta}$ and $v_{\theta}$ can be widely modulated: the former, playing either with the (temperature and molecular weight dependent) viscosity
+or tailoring the substrate to making it more hydrophobic (i.e. increasing the contact angle); the latter, tuning
+the spatial resolution and frame rate of the DMD. Thus, we expect the range of achievable $\Gamma$'s to 
+be feasibly extended both to very high ($\Gamma \gg 1$) and very low ($\Gamma \ll 1$) values.}\\
+\noindent {\it Conclusions.}
+We have presented results, from numerical simulations and theoretical analysis, on the dewetting of thin liquid films on a switchable substrate, modelled with a space and time periodically varying contact angle in the thin-film equation.
+Studying how the stability of the film depends on the underlying static pattern, we found that the rupture times grow linearly with the pattern wavelength, for short wavelengths, and quadratically in the long wavelength limit.
+In the time-dependent case, the rupture times are in general longer, indicating an induced greater film stability, and, while the quadratic growth is preserved at long wavelengths, a plateauing behaviour is observed as the wavelength decreases.
+A theoretical explanation has been provided for all these various regimes.
+Furthermore, we showed that, at increasing the wettability wave speed, a transition occurs in the dewetting morphology from a multi-droplet to a metastable multi-rivulet state. 
+A dimensionless parameter, $\Gamma$, controlling the transition has been identified in the ratio of the pattern speed and the typical film retraction speed over the substrate. 
+The rivulets life-time itself grows with the pattern speed, displaying a logarithmic dependence that has been captured by means of phenomenological arguments.
+On a broader perspective, our work suggests that switchable substrates offer a new avenue to control thin film dewetting, with obviously relevant implications, for instance, for open microfluidic devices, and paves the way to future studies in this direction, exploiting more complex and dedicated space-time dependencies.
 
-Interestingly we observe a correlation between the rupture time ($\tau_r$) and the speed ($v_{\theta}$) as well as the wavelength ($\lambda$) of the pattern.
-In agreement with previous studies on time dependent wettability we observe a stabilizing effect due to $v_{\theta}$~\cite{suman2006dynamics}.
-
-\subsection{Intermediate times}\label{subsec:interT}
-After the initial rupture of the film we observe that the dewetting is driven by $\theta(\mathbf{x},t)$.
-
-\subsection{Long time behaviour}\label{subsec:lateT}
-
-\section{Summary and Conclusions}\label{sec:sum_conclu}
-
-\begin{acknowledgements}
-The authors acknowledge financial support by the Deutsche Forschungsgemeinschaft (DFG) within the Cluster of Excellence ''Engineering of Advanced Materials'' (project EXC 315) (Bridge Funding). 
-\end{acknowledgements}
-
-%\appendix
-%\section{}\label{app:one}
+\noindent {\it Acknowledgements}.
+SZ and JH acknowledge financial support from the Deutsche Forschungsgemeinschaft (DFG) within the priority program SPP2171 ``Dynamic Wetting of Flexible, Adaptive, and Switchable Substrates'', project HA-4382/11 and Project-ID 431791331—CRC1452 ``Catalysis at Liquid Interfaces''.
+SZ acknowledges financial support from the Independent Research Fund Denmark through a DFF Sapere Aude Research Leader grant (grant number 9063-00018B).
 

### 2023-04-21 Reply_RefereeA.tex
New replies, thanks for nothing ref A

@@
 we assumed that it is a very well established concept employed in any theoretical and computational study of thin liquid film hydrodynamics in the lubrication approximation framework.
-We also assumed that an equation like Eq.(3) (now Eq.(2)), which is approximately twenty-five years old \cite{SCHWARTZ1998173,Mitlin94}, did not deserve a lengthy discussion.
+We also assumed that an equation like Eq.(3) (now Eq.(2)), which is almost thirty years old \cite{SCHWARTZ1998173,Mitlin94}, did not deserve a lengthy discussion.
 However, our assumptions were probably erroneous and, therefore, we give, here, some more explanation. 
@@
 and both are required to determine the dynamics of dewetting.
+We have included these comments in a section (called 'Disjoining pressure') 
+in the new Supplemental Material.
 Of course, on the other hand, the disjoining pressure depends also on the surface tension, which appears, in fact, in its expression (2). 
@@
 In any case, the Reynolds number stays always below $\approx 0.2$.
-We have added the following comment at the end of the {\it Method} section:\\
+We have added the following comment in the Supplemental Material in the section on 'Parameter values and comparison with experiments':\\
 
@@
 So, we are a bit puzzled about what such literature may have to do with the {\it dewetting} of a film on a chemically patterned substrate.
-Here, in fact, we are modelling a thin liquid film which is
-thermodaynamically stable against evaporation (i.e. one can consider that 
+Here, in fact, we are modelling a thin liquid film which is thermodaynamically stable against evaporation (i.e.
+one can consider that
 we are at a - fixed - temperature below the critical one).
@@
 We have reported these comments on comparison with actual experimental 
-values in the new version of the paper as follows:\\
+values in the new Supplemental Material as follows:\\
 \\

### 2023-04-21 Reply_RefereeB.tex
New replies, thanks for nothing ref A

@@
 \end{figure}
-We have added the following footnote in the paper commenting on this aspect:\\
+We have added the following paragraph in a dedicated section in the new
+Supplemental Material commenting on this aspect:\\
 \\
@@
 \end{figure}
-We have commented on this when we give the value of the slip length, saying that:\\
+We have commented on this when we give the value of the slip length (in the Supplemental Material), saying that:\\
 \\
@@
 A possibility could be to prepare a thin liquid film on a light responsive substrate~\cite{IchimuraEtAl_Science2000}, under the action of controlled 
-external stimuli (a light emitter).
+external stimuli (a light emitter). We have written the following paragraph in the revised
+manuscript, just before the 'Conclusions':
 
-\textcolor{red}{An ideal candidate could be a digital multimirror device (DMD).
+\textcolor{red}{We would like to envisage a possible realization of a 
+dewetting experiment on a switchable substrate of the type modeled by the spatio-temporal 
+contact angle. One may think of a thin liquid film cast on a light responsive substrate~\cite{IchimuraEtAl_Science2000}, under the action of controlled 
+external stimuli (a light emitter). An ideal candidate could be a digital multimirror device (DMD).
 This technology has effectively been used for thin film experiments and additive manufacturing~\cite{doi:10.1021/jp301092y, doi:10.1126/science.aax8760}, it allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$)
@@
 
-We have included these comments in the revised manuscript just before the conclusions.
-
 }
@@
 We apologize for not giving the value of $q_0 \approx 0.09$ (corresponding to an intrinsic correlation length 
-$\lambda_s = 2\pi/q_0 \approx 70 h_0$), which is now provided in the revised manuscript. We also acknowledge that expressing the wavelength in terms of the size $L$
+$\lambda_s = 2\pi/q_0 \approx 70 h_0$), which is now provided in the revised manuscript (in the 
+Supplemental Material). We also acknowledge that expressing the wavelength in terms of the size $L$
 was not at all meaningful (no finite size effects are involved).
@@
 \textcolor{red}{The Laplace pressure is the product of the surface tension 
-and the local curvature, therefore its full expression reads $p_L = -\gamma \nabla^2 h/\sqrt{1+|\nabla h|^2}$; 
-since we are in the lubrication regime, though, whereby $|\nabla h| \ll 1$, the denominator can be safely neglected 
-and the pressure reduces to the form involving only the Laplacian of the thickness field~\cite{Benet2014,juanes2018}}, \\
+and the local curvature. Therefore, its full expression reads $p_L = -\gamma \nabla^2 h/\sqrt{1+|\nabla h|^2}$. 
+In the lubrication regime, whereby $|\nabla h| \ll 1$, the denominator can be safely neglected 
+and the pressure reduces to the form involving only the Laplacian of the thickness 
+field}~\cite{Benet2014,juanes2018}, \\
 \\

### 2023-04-21 Reply_RefereeC.tex
Resubmission PRL
New replies, thanks for nothing ref A

@@
 fix a given pattern).
-
+We think, though, that probably the origin of this misunderstanding could be in the use of 
+the word 'transport' when we introduce the parameter $\Gamma$: we wrote, in fact, 
+"{\it if $U_{\Theta}$ is large as compared to $v_{\theta}$, the film retraction is faster than 
+the transport due to the contact angle field and thus droplets from}". We agree that such use 
+in that context might be not appropriate and misleading and we apologize for it. We have, thus,
+rephrased the sentence as follows:\\
+\\
+\textcolor{red}{If $U_{\Theta}$ is large as compared to $v_{\theta}$, the film retraction is faster than 
+the local contact angle variation and thus droplets form.}\\
+\\
+On the other hand, if the Referee does not believe that it is possible to actually achieve a fast 
+contact angle variation, as compared to a typical retraction speed, i.e.
+a large $\Gamma$, we wrote a new paragraph, just before the conclusions, where we envisage and discuss a possible experimental realization of our numerical setup, that would allow to easily get $\Gamma \approx 10^3$.\\
 The breadth and impact to warrant publication in Physical Review Letters is in our opinion clearly given: As we stressed extensively in the introduction,
@@
 \textcolor{red}{We decorate the substrate with a checkerboard pattern, a common choice that generalizes the broken homogeneity of the stripes to two 
-directions~\cite{KarguptaLangmuir2003,Jalali2018,Nagayama2020,Das2020}.}
+directions~\cite{Jalali2018,Nagayama2020,Das2020}.}
 
-\JH{Is this enough?}
+%\JH{Is this enough?}
 }
@@
 ($\lambda_l \approx 50$ in our case) \cite{KarguptaJCP2002,KarguptaLangmuir2000,Nisato1999,Karim1998}.
-We have added comments on this important point in the revised manuscript, when discussing the accessible ranges of pattern wavelength (see also the answer to the comment 5).
+We have added comments on this important point in the new Supplemental Material, when discussing the accessible ranges of pattern wavelength (see also the answer to the comment 5).
 
-\JH{Why don't you list the changes anymore in red (here and below)?}
+%\JH{Why don't you list the changes anymore in red (here and below)?}
 }
@@
 (in particular one has to comply with the condition $\Gamma \ll 5 \times 10^2$). We have added the following
-words of caution in the revised manuscript:\\
+words of caution in the new Supplemental Material:\\
 \\

### 2023-04-21 Old_version/main-veryold.tex
New replies, thanks for nothing ref A

--- 2022-10-03 Old_version/main.tex
@@
 
-%\title{Switchable substrates and thin film morphologies}
-\title{Spatiotemporal varying contact angles and thin film morphologies}
+\title{Switchable substrates and thin film morphologies}
 
@@
 \email{andrea.scagliarini@cnr.it}
- \affiliation{Institute for Applied Mathematics "M. Picone" (IAC), \\
-Consiglio Nazionale delle Ricerche (CNR),\\
+ \affiliation{Institute for Applied Mathematics "M. Picone" (IAc, \\
+Consiglio Nazionale delle Ricerche,\\
 Via dei Taurini 19, 00185 Rome, Italy}%
@@
 \begin{abstract}
-We study the dynamics of an unstable thin film deposited on a ``switchable`` substrate.
+We study the dynamics of an unstable thin film deposited on a switchable substrate.
 The substrate is itself subject to a time varying function that affects the equilibrium contact angle $\theta_{\text{eq}}(\mathbf{x},t)$.
-We show that in the case of vanishing time dependence the patterning drives the dewetting morphology and that all fluid is drained into regions of small $\theta_{\text{eq}}(\mathbf{x})$ and thus form droplets.
-Interestingly enough with a time varying component to the pattern we are able to stop the droplet fragmentation and create pronounced ligament or rivulet morphologies.
-We show that both the numbers of semistable rivulets as well as number of stable droplets can be controlled with the pattern wavelength ($\lambda$).
-Finally we explain why the ligaments appear with a simple balance between capillary retraction and dynamic wetting.   
+We show that in the case of the vanishing time dependence the patterning drives the dewetting process and all fluid is drained into regions of a small $\theta_{\text{eq}}(\mathbf{x})$ and forms droplets.
+Interestingly enough with a time varying equilibrium contact angle we are able to stop the droplet fragmentation and are able to create ligament or rivulet states.
+We show that these rivulet states mainly depend on two quantities, first being the wavelength of the pattern and second being the velocity of the substrate pattern.
+
 \end{abstract}
@@
 Thin film flows and especially droplets are an phenomenon we experience in our every day life.
-Being it the droplets condensed in the higher atmosphere and return to the ground as rain or the oil film in a pan during cooking. 
-While this two examples are illustrative there are many industrial applications that build on thin film flows. 
-Among them is the wetting of surfaces to reduce wear or put simply, lubrication~\cite{ReynoldsLubr, gross1980fluid, szeri2010fluid}. 
-Coating surfaces to change their physio-chemical properties is another widely used application~\cite{DERYCK1998278, doi:10.1146/annurev.fluid.31.1.347, DASILVASOBRINHO19991204}.
-On the other hand the need to control droplets is at the heart of many printing processes~\cite{singh2010inkjet, jo2009evaluation}.
+Rather unknown is the fact that our body uses thin films to cover mucous membranes.
 
-More recently lab-on-a-chip devices have acquired much attention~\cite{C6LC00387G,Focke}, to work they require precise droplet or film flow on small scales~\cite{stone2004engineering, darhuber2010planar}.
-One possibility for precise control is the local wettability.
-Light switchable substrates~\cite{WANG200718}, therefore substrates that change their wettability based on light illumination, open up a new way to program this devices. 
-From a theoretical point of view the motion of the fluid can be computed using the thin film equation~\cite{RevModPhys.69.931, RevModPhys.81.1131, RevModPhys.81.739, ReynoldsLubr}
 \begin{equation}\label{eq:thin_film}
@@
 \end{equation}
-where $h(\mathbf{x},t)$ is a time and spatial varying film thickness, $M(h)$ is a mobility function and $p(\mathbf{x},t)$ is a pressure term that drives the flow. 
-The function $M(h)$ contains information about the velocity boundary condition at $h=0$, using no-slip yields $M(h) = h^3/(3\mu)$.
-More interesting however is the pressure function $p(\mathbf{x},t)$ as it allows to model the wettability of a given substrate.
-We therefore introduce a time and spatial dependent contact angle to control the local wettability.
-
-Spatial variations of the contact angle are long known and used in previous studies~\cite{PhysRevE.91.023013, Sommers_2006}, time dependency of the contact angle on the other hand is not as well studied~\cite{suman2006dynamics}.
-Switchable substrates are in fact a realization of a time dependent contact angle. 
-Using both spatial as well as time dependent variations of the local contact angle leads to interesting phenomena, for example continuously sliding droplets as shown by Grawitter and Stark~\cite{D0SM02082F}.
-Our approach builds on the same idea but our method, see ref~\cite{PhysRevE.100.033313}, allows us to go beyond a single droplet and to simulate the full dewetting transition starting with a thin film and ending with droplets.
-We use this to show that a time dependent component of the contact angle alter the dewetting morphology quite substantially.
-In the limit of a small time dependency or slow evolution of the contact angle field we qualitatively reproduce the results from Grawitter and Stark, such sliding droplets.
-Speeding up the evolution of the contact angle field introduces a morphological transition, which to the understanding of the authors have not been looked at so far.
-Therefore the purpose of this manuscript is to increase the understanding of the dewetting process.
-Highlighting the fact that a temporal and spatial variation of the contact angle leads to novel morphological structures.
-Thus our finding offer a simple way to understand how time periodic switchable substrates could generate novel dewetting morphologies.
-
-The manuscript is outlined as follows:
-In section~\ref{sec:method} we discuss the numerical method we use to generate our results and introduce relevant scales.
-We then present our results in section~\ref{sec:results}, where we discuss the impact of the time dependency on the contact angle and show that both stability as well as dewetting morphology is affected.
-A short summary of our finding and a closing conclusion with outlook is given in section~\ref{sec:sum_conclu}.
 
 \section{Method}\label{sec:method}
-In this section we describe the lattice Boltzmann method (LBM) we use.
-The LBM is a meso scale approach that builds on the Boltzmann equation to simulate fluid flows~\cite{krueger2017}. 
-It is widely applied to many kinds of flow problems, among them are for instance multi phase and multi component flows~\cite{shan1993lattice,PhysRevE.73.047701}, flows through porous media~\cite{liu2016multiphase} and shallow water flows~\cite{Salmon:1999:0022-2402:503, zhou2004lattice, PhysRevE.65.036309}.
-The model we use has been introduced in ref.~\cite{PhysRevE.100.033313} and builds upon a class of lattice Boltzmann models originally developed to study the shallow water equations~\cite{Salmon:1999:0022-2402:503, PhysRevE.65.036309}.
+In this section we describe the lattice Boltzmann method (LBM) we use to simulate thin film flows.
+The model we use has been introduced in ref~\cite{PhysRevE.100.033313} and builds upon a class of lattice Boltzmann models originally developed to study the shallow water equations~\cite{Salmon:1999:0022-2402:503, zhou2004lattice, PhysRevE.65.036309}.
 
 \subsection{Thin film lattice Boltzmann method}\label{subsec:LBM}
-Since the model has been developed and tested in previous work, ref.~\cite{PhysRevE.100.033313}, we just outline the most important features.
 The internal degrees of freedom of this method are the distributions functions $f_{\alpha}(\mathbf{x},t)$.
-These distribution functions describe the movement of fictitious particles on a lattice, in this case on a $D2Q9$ lattice.
-Meaning there are nine velocity components ($Q9$), directions in which a particle could move, for a simulation in two dimensions ($D2$).
+They describe a distribution of fictitious particles, for this paper we consider the case of $D2Q9$ lattice.
+Such there are nine velocity components ($Q9$) for a simulation in two dimensions ($D2$).
 The dynamics of the thin film and thus the evolution of the distribution functions is given by
@@
 \begin{equation}\label{eq:forcesum}
-    \mathbf{F} = \mathbf{F}_{\mbox{\tiny{fric}}} + \mathbf{F}_{\mbox{\tiny{film}}}.
+    \mathbf{F} = \mathbf{F}_{\mbox{\tiny{fric}}} + \mathbf{F}_{\mbox{\tiny{fric}}}.
 \end{equation}
@@
 The film thickness $h$ is the zeroth moment of the distribution function $f_{\alpha}$, i.e. $h(\mathbf{x},t) = \sum_{\alpha}f_{\alpha}(\mathbf{x},t)$ see Eq.~(\ref{eq:LBE}). 
-The second component is due to the film pressure $p_{\text{film}}$ and accounts for the minimization of the fluid vapor interface as well as for the wettability of the substrate,
-\begin{equation}\label{eq:pressure_here}
-    p_{\text{film}} = -\gamma (\Delta h - \Pi(h)),
+The second component is due to the film pressure and accounts for the minimization of the fluid vapor interface as well as for the wettability of the substrate,
+\begin{equation}\label{eq:forcefilm}
+    \mathbf{F}_{\mbox{\tiny{film}}} = -\frac{\gamma h}{\rho_0}\nabla(\Delta h - \Pi(h)), 
 \end{equation}
-which results in the addition of the force $\mathbf{F}_{\mbox{\tiny{film}}}$
-\begin{equation}\label{eq:forcefilm}
-    \mathbf{F}_{\mbox{\tiny{film}}} = -\frac{h}{\rho_0}\nabla(p_{\text{film}}), 
-\end{equation}
-with $\gamma$ being the (constant) surface tension and $\rho_0$ being the fluids density, which is $\rho_0 = 1$ and will be dropped in the following. 
+with $\gamma$ being the surface tension and $\rho_0$ being the fluids density, which is $\rho_0 = 1$ and will be dropped in the following. 
 Here we model the interaction between solid substrate and film with a disjoining pressure term~\cite{RevModPhys.69.931, RevModPhys.81.1131, PhysRevE.63.011208},
@@
 
-We consider two different patterns and three wavelengths in this manuscript.
-First being the sinusoidal pattern with wavelength $\lambda = L/i$ and wavenumbers $i=[1,2,3]$ which is given as
+We consider two different patterns and three wavelengths for this manuscript.
+First being the sinusoidal pattern with wavelengths $\lambda \in \{L, L/2, L/3\}$ which is given as
 \begin{equation}\label{eq:sinetheta}
@@
 with $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$.
-The second pattern is a triangle wave which accounts for a different kind of pattern gradient.
-To the best of our knowledge contact angles can not be tuned precise enough to match a sinusoidal wave on the scales we are interested in.
-Therefore we assume that a triangle wave with a linear gradient is more likely accessable in experiments, thus $\theta_2$ is given as
+The second pattern is a triangle wave which accounts for a more realistic szenario, thus $\theta_2$ is given as
 \begin{equation}\label{eq:triangle_wave}
-\begin{split}
-    \theta_2(x,y) &= \theta_0 + \\ &\frac{2\delta\theta}{\pi}\left[\sin^{-1}\left(\sin\left(\frac{2\pi x}{\lambda}\right)\right)\sin^{-1}\left(\sin\left(\frac{2\pi y}{\lambda}\right)\right)\right].
-    \end{split}
+    \theta_2(x,y) = \theta_0 + \frac{2\delta\theta}{\pi}\left[\sin^{-1}\left(\sin\left(\frac{2\pi x}{\lambda}\right)\right)\sin^{-1}\left(\sin\left(\frac{2\pi y}{\lambda}\right)\right)\right].
 \end{equation}
@@
 Both patterns create a contact angle distribution with values within the interval $[10^{\circ}, 30^{\circ}]$ and by definition they are time independent.
-To have a time dependency we simply shift the pattern in a periodic manner, such that
+For a time dependent dynamic we simply shift the pattern in a periodic manner, such that
 \begin{equation}\label{eq:theta_shift}
@@
 \end{equation}
-with $T$ being an arbitrary integer number of lattice Boltzmann time steps.
+with $T$ being an arbitrary number of lattice Boltzmann time steps.
 
-Due to convenience we attribute a velocity to the pattern which is the shift distance divided by the stepping time
+Due to convenience we artibute a velocity to the pattern which is simply the shift distance divided by the stepping time
 \begin{equation}\label{eq:pattern_speed}
@@
 \subsection{Numerical system}
-\begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{Figures/makie_try2.png}
-    \caption{Film thickness $h(\mathbf{x},t)$ at a stationary state. 
-    The contact angle $\theta(\mathbf{x})$ is shown as contour.}
-    \label{fig:handtheta}
-\end{figure}
 As outlined in Sec.~\ref{subsec:LBM} we simulate the evolution of the thin film with a lattice Boltzmann method.
 For all our simulations we use periodic boundary conditions and a square of size $Lx = Ly = L = 512 \Delta x$ with $\Delta x = \Delta t = 1$.
-The surface tension as well as the viscosity are constants and set to $\gamma = 0.01$ and $\nu|_{\rho_0=1} = \mu = 1/6$.
+The surface tension as well as the viscosity are constants and set to $\gamma = 0.01$ and $\mu = 1/6$.
 Having this value of $\mu$ leaves us with a relaxation time $\tau = 1$.
-To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer and a slip parameter $\delta = 1$. 
-% With this choice of parameters we are almost able simulate a no-slip boundary condition.
-As for our disjoining pressure potential, see Eq.~(\ref{eq:disjoinpressure}), we use the power pair $n=3, m=2$ and $h_{\ast} = 0.07$.
-Inserting this quantities into Eq.~(\ref{eq:q0_and_t0}) with $h_0 = 1$ yields $q_0^2(30^{\circ}) \approx 1.68\cdot10^{-2}$ as well as $t_0(30^{\circ}) \approx 177000\Delta t$ which gives us a first glimpse into relevant length and time scales.
+To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer and a slip length $\delta = 1$. 
+With this choice of parameters we are almost able simulate a no-slip boundary condition.
+As for our disjoining pressure potential, we use $n=3, m=2$ and $h_{\ast} = 0.07$.
+Inserting this quantities into Eq.~(\ref{eq:q0_and_t0}) with $h_0 = 1$ yields $q_0^2(30^{\circ}) \approx 1.68\cdot10^{-2}$ as well as $t_0(30^{\circ}) \approx 177000\Delta t$ which gives us a first glimpse into relevant time scales.
 
@@
 
-We proceed as follows, first we simulate the dewetting on a given pattern either sinusoidal- or triangular with $\lambda \in \{L, L/2, L/3\}$ without a time dependency in $\theta$.
-In Fig.~\ref{fig:handtheta} we show $h(\mathbf{x},t)$ as surface plot and $\theta_2(\mathbf{x})$ as contour plot for $v_{\theta} = 0, \lambda = L/2$ in late stages of dewetting.
-Droplets form in regions of small contact angles while the regions of high contact angles dewet.
-In subsequent simulations we iterate through $v_{\theta} \in [0.1, 1, 10]v_0$, see Eq.~(\ref{eq:normvel}) .  
-% Based on $v_{\theta}$ we can modify the morphology of the dewetting film, in both the intermediate transition regime as well as the late time state.
-
 \section{Results}\label{sec:results}
-We now present our results and show how a time dependent wettability or pattern velocity alters the dynamics of a dewetting thin film.
-In the following we address the growth of the initial perturbation $\delta h$ and the associated stability of the film.
-Therefor we take a look at the rupture times and show that a pattern velocity $v_{\theta} > 0$ does have a stabilizing effect.
-
-After rupture we discuss the morphological transitions the film undergoes.
-Therefor the transition from a dewetting film to a droplet state.
-This transition is heavily influenced by the concrete choice of $v_{\theta}$ and we observe a strong a alignment with the moving direction of the pattern.
-
-Before we move to the case of $v_{\theta} > 0$ we like to draw a base line with simulation done with a stationary wettability profile.
-
-\subsection{Stationary pattern}\label{subsec:no_vtheta}
-Starting with Eq.~(\ref{eq:hinitial}) we have by definition an unstable film that will rupture.
-This can be easily seen by the fact that the initial perturbation has the smallest possible wavenumber, $q(t=0)=1/L$.
-Taking into account the addition of the wettability gradient we know that fluid will accumulate in regions of high surface energies or small contact angles and flow away from regions of high contact angle.
-An illustrative results of this effect is shown in Fig.~\ref{fig:handtheta}, where all droplets are present in a region of small contact angles.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{Figures/theta_0336.png}
+    \caption{Grayscale image of the height film at $t\approx 2t_0(30^{\circ})$}
+    \label{fig:gray_ligament}
+\end{figure}
+We now present our results and show how a time dependent wettability or pattern velocity alters the morphology of the film during the dewetting process.
+In the following we consider three time scales.
+First being the initial short time behaviour that ends with the rupture of the film.
+After the film ruptures we observe a transition region that strongly depend on the pattern velocity, see Fig.~\ref{fig:gray_ligament}.
+Lastly we discuss the morphologies of the dewetted film in the long time regime.
 
@@
     \centering
-    \includegraphics[width=0.48\textwidth]{Figures/Clusters_novel_sine_picutres.png}
-    \caption{Connected component analysis of the film thickness for different wavelengths $\lambda$ on $\theta_1$ ($\theta_2$ matches qualitatively).
-    On the x-axis we display the number of clusters while on the y-axis we show the time normalized by $t_0$.
-    Different colors and symbols display different wavelengths $\lambda$.
-    The three dashed lines indicate the number of minima the pattern offers.
-    Above the data we show the stationary droplet state as a grayscale image of the film thickness $h(\mathbf{x},t_{\infty})$.}
-    \label{fig:clusters_v0_sine}
+    \includegraphics[width=0.45\textwidth]{Figures/makie_try2.png}
+    \caption{Height of the film $h(\mathbf{x},t)$ with the contact angle $\theta(\mathbf{x})$ as contour.}
+    \label{fig:handtheta}
 \end{figure}
-This is in fact a consistency check of our simulations.
-We expect that in case of $v_{\theta} = 0$ all fluid accumulates in droplets inside the contact angle minima.
-While there are only two minima for $\lambda = L$ there are eight and eighteen minima for $\lambda = L/2$ and $\lambda = L/3$ respectively.
-This can be easily checked with a cluster analysis of the film thickness $h(\mathbf{x},t)$ which we shown in Fig.~\ref{fig:clusters_v0_sine}.
-The height $h(\mathbf{x},t_{\infty})$ is displayed as grayscale images and is linked to stationary state.
-Theoretically we expect therefor two, eight and eighteen clusters (droplets) which we mark with a vertical dashed line in the respective color.
-These values are consistent with our simulations.
-Interesting is the fact that the number of clusters converge faster for smaller wavelengths.
-To some extend this is expected as the distance between contact angle minima is shorter for smaller wavelengths.
+We proceed as follows, first we simulate the dewetting on a given pattern either sinusoidal- or triangular with $\lambda \in \{L, L/2, L/3\}$ without a time dependency in $\theta$.
+In Fig.~\ref{fig:handtheta} we show $h(\mathbf{x},t)$ as surface plot and $\theta_2(\mathbf{x})$ as contour plot for $v_{\theta} = 0, \lambda = L/2$ in late stages of dewetting.
+Subsequent simulations allow for $v_{\theta} \ne 0$.  
+Based on $v_{\theta}$ we can modify the morphology of the dewetting film, in both the intermediate transition regime as well as the late time state.
 
-Another quantity that is easily accessible and helps to check the consistency of our simulations is the height of the droplets $h_d$.
-Given that all the fluid is collected inside the droplets we can make an assumption of the droplets height based on its volume $V_d$  and contact angle $\theta_d$ 
-\begin{equation}\label{eq:height_guess}
-    h_d(V_d, \theta_d) = \left(\frac{3V_d(1-\cos(\theta_d)}{\pi(2+\cos(\theta_d))}\right)^{\frac{1}{3}},
+\subsection{Film rupture}\label{subsec:earlyT}
+Starting with Eq.~(\ref{eq:hinitial}) we have by definition an unstable film that will rupture.
+Taking into account the addition of the wettability gradient we know that fluid will accumulate in regions of high surface energies or small contact angles.
+This is shown in Fig.~\ref{fig:handtheta}, where all droplets are in a region of small contact angles.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{Figures/All_deltah.pdf}
+    \caption{Evolution of the absolute difference in height, x-axis, up to $t\approx t_0(30^{\circ})$, y-axis, for different wavelengths and patterns.
+    The upper series displays the data for $\theta_1$, see Eq.~(\ref{eq:sinetheta}), and $\lambda = L, L/2, L/3$ in (a,b,c) respectively.
+    Below is the series for $\theta_2$, see Eq.~(\ref{eq:triangle_wave}).
+    Different colors and line styles represent different pattern velocities, all of which are normalized with $v_0$, see Eq.~(\ref{eq:normvel}).}
+    \label{fig:evolutin_deltah}
+\end{figure*}
+Naively one assumes in the early stages of the dewetting process that the perturbations grow exponentially~\cite{RevModPhys.69.931, RevModPhys.81.1131}.
+This is not the case in our simulations as the initial conditions for $h$ as well as $\theta$ are the same, see Eq.~(\ref{eq:hinitial}, \ref{eq:sinetheta}).
+We show this by measuring the time resolved difference between the maximum and the minimum of the height field as, 
+\begin{equation}\label{eq:delta_h_measure}
+    \Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\} - \min_{\mathbf{x}}\{h(\mathbf{x},t)\},    
 \end{equation}
-where we implicitly assume that all our droplets have a spherical cap shape.
-Since our method does not allow $h=0$ we measure the droplets height as~\cite{PhysRevE.100.033313}
-\begin{equation}\label{eq:delta_h_measure}
-    \Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\} - \min_{\mathbf{x}}\{h(\mathbf{x},t)\}.
-\end{equation}
-Counting the minima in $\theta(\mathbf{x})$ for the different wavelengths $\lambda$ allows to calculate the volume of a single droplet, see Fig.~\ref{fig:clusters_v0_sine}.
-Since the film dewets regions of high contact angle values we can safely assume that $\theta_d < 20^{\circ}$.
+which is shown in Fig.~\ref{fig:evolutin_deltah}.
+In the upper row we show the data for $\theta_1$ while in the lower we show data for $\theta_2$, see Eqs.~(\ref{eq:sinetheta}-\ref{eq:triangle_wave}).
+Different linestyles and colors represent different $v_{\theta}$'s.
+Each column shows a wavelength $\lambda$, such from the left to the right we have $\lambda = L, L/2, L/3$.
+Clearly one can observe the dip in \ref{fig:evolutin_deltah}(a) as the initial perturbation flattens while flowing into regions of lower contact angle.
+This behaviour however is only pronounced for the time independent pattern and the smallest $v_{\theta}$ we use.
+Thus the film is not spinodally dewetting but rather nucleating in regions of high contact angles.
+
 \begin{figure}
     \centering
-    \includegraphics[width=0.48\textwidth]{Figures/delta_h_v0_sin_with_dash.pdf}
-    \caption{Evolution of $\Delta h$ with time.
-    Different symbols and colors represent different pattern wavelengths.
-    Dashed lines indicate solutions to the cap height $h_d$ given the volume is split equally among the droplets, see Eq.~(\ref{eq:height_guess}).}
-    \label{fig:deltah_v0_sine}
+    \includegraphics[width=0.48\textwidth]{Figures/Rupture_times.pdf}
+    \caption{Collection of rupture times $\tau_r$ grouped according to the substrate pattern.
+    On the x-axis we display the three wavelengths we considere, while on the y-axis we show the time normalized with $t_0(30^{\circ})$.
+    Results are grouped by their pattern, such (lin., sin.) is given by ($\theta_2$, $\theta_1$).
+    }
+    \label{fig:rupture_times}
 \end{figure}
-Measuring the height of the droplets is done in Fig.~\ref{fig:deltah_v0_sine}, where $\Delta h$ is the maximal difference in film thickness which is equivalent to the droplet height $h_d$.
-Focusing on the late time behaviour ($t > 10t_0$) we identify plateauing values. 
-Since the pattern is a continuous functions, see Eq.~(\ref{eq:sinetheta}), we are left with some degree of freedom for the theoretical $\theta_d$ values. 
-As a guide to the eye we introduce three dashed lines.
-The dashed lines are calculated using Eq.~(\ref{eq:height_guess}) with $V_d = (V/2, V/8, V/18)$ and $\theta_d = (14^{\circ}, 16^{\circ}, 17^{\circ})$ for $\lambda=(L,L/2,L/3)$ respectively.
-The values of $\theta_d$ are chosen to fit with the stationary droplet state of our simulations.
-Thus we are fairly confident that our droplet states and thus the simulations capturing the correct physical state.
+Interestingly we observe a correlation between the rupture time ($\tau_r$) and the speed ($v_{\theta}$) as well as the wavelength ($\lambda$) of the pattern.
+Where $\tau_r =\exists~\mathbf{x}:h \le h_{\text{pre.}}$ and $h_{\text{pre.}}$ is the height of the precursor film.
+In agreement with previous studies on time dependent wettability we observe a stabilizing effect due to $v_{\theta}$~\cite{suman2006dynamics}.
+Which can be seen in Fig.~\ref{fig:evolutin_deltah} as large velocities shift the curves to the right and therefore to later times.
+Concerning the pattern we observe the shorter the wavelength the stronger the gradient.
+Thus the rupture time is inverse proportional to the patterns wavelength, as shown in Fig.~\ref{fig:rupture_times}.
+At first glance surprisingly $\theta_1$ yields shorter rupture times than $\theta_2$.
+We attribute this to the gradient of pattern, as  
+\begin{equation}\label{eq:grad_theta}
+    \nabla\theta_i(\mathbf{x},t) \propto \begin{cases}
+    \pm C \quad i=2 \\
+    \cos(\mathbf{x},t) \quad i=1
+    \end{cases},
+\end{equation}
+with $C$ being a constant.
 
-\subsection{Stability}\label{subsec:time_dep_theta}
-\begin{table}[h!]
-  \begin{center}
-    \begin{tabular}{c|c|c|c}
-      \textbf{Pattern} & \textbf{Wavelength} & \textbf{Velocity } & \textbf{Rupture time}\\
-      $\theta(\mathbf{x})$ & $\lambda$ & $v/v_0 $ & $\tau_r/t_0$ \\
-      \hline
-      $\theta_1$ & L & 0 & 1.69\\
-      $\theta_1$ & L & 0.1 & 1.72\\
-      $\theta_1$ & L & 1 & 2.34\\
-      $\theta_1$ & L & 10 & 2.25\\
-      $\theta_1$ & L/2 & 0 & 0.39\\
-      $\theta_1$ & L/2 & 0.1 & 0.39\\
-      $\theta_1$ & L/2 & 1 & 0.59\\
-      $\theta_1$ & L/2 & 10 & 0.68\\
-      $\theta_1$ & L/3 & 0 & 0.23\\
-      $\theta_1$ & L/3 & 0.1 & 0.23\\
-      $\theta_1$ & L/3 & 1 & 0.25\\
-      $\theta_1$ & L/3 & 10 & 0.45\\
-      $\theta_2$ & L & 0 & 1.01\\
-      $\theta_2$ & L & 0.1 & 1.18\\
-      $\theta_2$ & L & 1 & 1.32\\
-      $\theta_2$ & L & 10 & 1.38\\
-      $\theta_2$ & L/2 & 0 & 0.48\\
-      $\theta_2$ & L/2 & 0.1 & 0.48\\
-      $\theta_2$ & L/2 & 1 & 0.7\\
-      $\theta_2$ & L/2 & 10 & 0.76\\
-      $\theta_2$ & L/3 & 0 & 0.39\\
-      $\theta_2$ & L/3 & 0.1 & 0.42\\
-      $\theta_2$ & L/3 & 1 & 0.68\\
-      $\theta_2$ & L/3 & 10 & 0.76\\
-    \end{tabular}
-    \caption{Rupture time values for different pattern wavelengths $\lambda$ and different substrate velocities $v_{\theta}$.
-    Due to sampling we have a systematic error of 0.03 in the rupture times $\tau_r$.}
-    \label{tab:rup_times}
-  \end{center}
-\end{table}
-Taking another look at Fig.~\ref{fig:deltah_v0_sine} reveals that the three wavelengths rupture at different times $\tau_r$.
-The most important values for $\tau_r$ can be found in Tab.~\ref{tab:rup_times}.
-We observe two features, first being that rupture time $\tau_r$ correlates with the wavelength $\lambda$.
-Thus smaller values of $\lambda$ lead to an earlier rupture time.
-Which can be explained with a mismatch of length scales.
-There is a length scale set by Eq.~(\ref{eq:q0_and_t0}) ($\lambda_0 = 2\pi/q_0$) and one that we introduce with the substrates wavelength $\lambda$.
 \begin{figure}
     \centering
-    \includegraphics[width=0.48\textwidth]{Figures/RT_with_v10_inkscape.pdf}
-    \caption{Distribution of rupture times $\tau_r$ normalized by $t_0$ for different pattern wavelengths $\lambda$.
-    The bullets, stars correspond to data from our simulations with $v_{\theta} = 0, 10v_0$.
-    Eq.~(\ref{fig:model_rt}) with $D=2$ is displayed by the blue line.
-    }
-    \label{fig:model_rt}
+    \includegraphics[width = 0.48\textwidth]{Figures/fit_rt.pdf}
+    \caption{Rupture times for different wavelengths and a time independent pattern.
+    The full line is a linear fit.}
+    \label{fig:fit_rt}
 \end{figure}
-Taking this into account we can model the rupture times for the stationary pattern with
-\begin{equation}\label{eq:model_tr}
-    \frac{\tau_r}{t_0} = \frac{1}{2\pi}\frac{L}{\lambda_0 i^D}\approx \xi\frac{\lambda}{\lambda_0},    
+Given the time independent patter we take a closer look into the dependence of the wavelength in Fig~\ref{fig:fit_rt}.
+As stated above the film placed on $\theta_1$ is less stable and ruptures earlier.
+Using a linear fit 
+\begin{equation}\label{eq:lin_fit}
+    y(x) = kx + d,
 \end{equation}
-where $i$ is the wavenumber of the substrates pattern ($\lambda = L/i$) and $D$ is the dimension, thus $D=2$.
+we observe a slope $k = 4.6\cdot 10^{-3}$ for $\theta_1$ and $k= 5.4\cdot 10^{-3}$ for $\theta_2$.
+After the initial rupture events we see, based on $v_{\theta}$, the development of non droplet morphologies.
 
-Interestingly Eq.~(\ref{eq:model_tr}) is able to describe the data for $v_{\theta} = 0.1v_0$ as the difference of rupture times between these two simulations is negligible, see Tab.~\ref{tab:rup_times}.
-Larger $v_{\theta}$ introduce a shift towards longer rupture times, therefore increasing the stability of the film.
-Next we look into the affects of the pattern velocity on the structure formation after the first ruptures.
+\subsection{Morphologies}\label{subsec:interT}
+In the following we discuss how the moving pattern affects the morphology of the retracting film.
+Besides other measures we like to emphasise the relation between $v_{\theta}$ and films dewetting morphology.
+We use Minkovsky functionals to analyze our data, this is done with the software \textbf{papaya2}~\cite{Schaller2020}.
 
-\subsection{Morphologies}\label{subsec:morphme}
-Dewetting morphology is a rather broad topic~\cite{becker2003complex, RevModPhys.81.739, PhysRevLett.99.114503, craster2009dynamics, doi:10.1146/annurev-fluid-011212-140734, alizadeh_pahlavan_cueto-felgueroso_hosoi_mckinley_juanes_2018}, however we know that a simple patterned substrate will ultimately yield droplets in region of small contact angle and is likely to rupture in regions of high contact angle, as shown in Fig.~\ref{fig:handtheta}.
-This is indeed what we observe for the simulation without pattern velocity ($v_{\theta} = 0$).
-\begin{figure}
-    \centering
-    \includegraphics[width=0.48\textwidth]{Figures/q_2_different_v.pdf}
-    \caption{Structure analysis of the pattern $\theta_1,~\lambda=L/2$ using the Minkowski structure metrics (MSM). 
-    The dipole metric $q_2$ is plotted for different pattern velocities $v_{\theta}$, displayed with different colors and symbols.
-    In grayscale we supply the film thickness that is associated with the $q_2$ values.
-    }
-    \label{fig:msm_q2}
-\end{figure}
-Increasing the pattern velocity $v_{\theta} > 0$ does affect the dewetting morphology quite substantially.
-Having $v_{\theta} = 0.1 v_0$ we are still observing the formation of droplets, similar to $v_{\theta} = 0$ but instead of being stationary these droplets, depending on their mass, are transported with the contact angle minima.
-A somewhat similar behaviour was recently described by Grawitter et al.~\cite{D0SM02082F} where they have used a boundary integral method to simulate a droplet on a moving wettability step.
-
-If the pattern velocity is further increased $v_{\theta} > 0.1 v_0$ we observe a transition towards ligament or rivulet like structures.
-These structures are aligned with the update direction of the contact angle field, see Eq.~(\ref{eq:theta_shift}). 
-Thus if
-\begin{equation*}
-    \theta(x,y,t+T) = \theta(x-1, y, t),    
-\end{equation*}
-ligaments form in x direction.
-These ligaments are sensitive to $\lambda$, or more precisely to the contact angle minima since they attract fluid.
-To classify the morphological differences in respect to $v_{\theta}$ we make use of the software \textbf{papaya2}~\cite{Schaller2020}.
-
-We use the film thickness as input and compute different morphological metrics.
-Among the most interesting ones are the Minkowski structure metrics (MSM) $q_s$.
-They can be calculated using~\cite{doi:10.1063/1.4774084}
-\begin{equation}
-    q_s(K) = \frac{|\psi_s(K)|}{\psi_0(K)}, 
-\end{equation}
-where $K$ is the body to we extract morphological information from and $\psi_s$ are the irreducible Minskovski tensors (IMT)
-\begin{equation}
-    \psi_s(K) = \int_{0}^{2\pi}e^{is\phi}\rho_K(\phi) \dd\phi = \sum_{k=-\infty}^{\infty} L_k e^{is\phi_k},
-\end{equation}
-with $L_k$ being the set of normal vectors of the body $K$.
-Within the various $q_s$ we are particularly interested in the dipole one $q_2$. 
-With $q_2$ we can clearly distinguish between a droplet and a ligament structure, as we expect a much larger $q_2$ value for the ligament state which is what we see in Fig.~\ref{fig:msm_q2}.
-Our simulations with $\lambda = L/2$ show a clear difference between $v_{\theta} \leq 0.1v_0$ and $v_{\theta}>0.1v_0$  in $q_2$.
-Interestingly the $q_2$ values for $v_{\theta} = 0.1v_0$ are always larger than those of $v_{\theta} = 0$.
-The difference is due to the distortion of the droplets surface area as it is transported with the pattern.
-
-Moving the contact angle field faster creates metastable ligament states.
-Metastable because this ligaments will eventually break and separate into droplets, as indicated by the dip in $q_2$ for $v_{\theta} = 1, 10 v_0$ in Fig.~\ref{fig:msm_q2}.
-Nevertheless their formation in the first place can be explained with a few relative simple assumptions.
-
-This problem is associated with two time scales, first being $t_0(\theta)$ the time scale of the dewetting thin film, see Eq.~(\ref{eq:q0_and_t0}).
-After the rupture we identify another time scale, the capillary time scale $\tau_{\text{Ca}}$ that drives the rims away from the holes~\cite{Edwardse1600183},
-\begin{equation}\label{eq:cap_time}
-    \tau_{\text{Ca}} = \frac{9\mu l_0}{\gamma \theta^3(\mathbf{x})},
-\end{equation}
-where  $l_0$ is a characteristic length scale of the problem, we use $\lambda/2$ as two rims can travel at most $\lambda/2$ before they merge.
-Knowing this time scales allows us to introduce characteristic velocities.
-One of these two is $v_0$ which is given by Eq.~(\ref{eq:normvel}) and used to normalize $v_{\theta}$.
-The other one is set by $\tau_{\text{Ca}}$ and reads
-\begin{equation}\label{eq:cap_speed}
-    U_{Ca} =\frac{l_0}{\tau_{\text{Ca}}} = \frac{\gamma}{9\mu} \theta^3(\mathbf{x}), 
-\end{equation}
-where $U_{\text{Ca}}$ is the capillary speed.
-The argument is as follows, if $U_{\text{Ca}}$ is large as compared to $v_{\theta}$ than the film retraction is faster than the transport due to the contact angle field and thus droplets from.
-Which explains the cases of $v_{\theta} = 0, 0.1v_0$.
-On the other hand if $v_{\theta}$ is larger than $U_{\text{Ca}}$ than the retracting film has to little time to create stable droplets and ends up in the metastable ligament state.
-Taking a look at our simulations with $v_{\theta} = 1, 10v_0$ we clearly observe ligament states.
-
-In fact the ratio 
-\begin{equation}\label{eq:vel_ratio}
-    \Gamma = \frac{v_{\theta}}{U_{\text{Ca}}} = \frac{3\lambda \chi h_0^3 q_0^4}{\Theta^3}, 
-\end{equation}
-where $\Theta = max(\theta(\mathbf{x}))$ and $v_{\theta} = \chi v_0$, allows to classify the dewetting morphology on a time and space dependent substrate pattern as
-\begin{equation}\label{eq:classify}
-    \Gamma~\begin{cases}
-    < 1\qquad ... \text{droplets} \\
-    > 1\qquad ... \text{ligaments} 
-    \end{cases}.
-\end{equation}
-This explains in fact what we observe and holds true independent of the wettability gradient, i.e. $\theta_1, \theta_2$.
-
-Therefore by varying either $\chi$ or $\lambda$ for a fixed $\Theta$ one can change the intermediate dewetting morphology according to the equation above.
-Interestingly but not surprising is the fact that the fluid transport is larger for $\Gamma > 1$ than for $\Gamma < 1$.
+\subsection{Stability}\label{subsec:lateT}
+As a last point of Sec.~\ref{sec:results} we want to quantify the long time stability of the morphologies we observe.
 
 \section{Summary and Conclusions}\label{sec:sum_conclu}
-To summarize, we have shown numerical simulations of a dewetting thin film on a patterned substrate.
-Generating baseline experiments with time independent patterns of different wavelengths, we than move to a time and spatial varying contact angle field $\theta(\mathbf{x},t)$.
-
-In the next step we define a characteristic pattern velocity $v_0$ which simply is the patterns wavelength $\lambda$ over a characteristic time $t_0$, see Eqs.~(\ref{eq:q0_and_t0}, \ref{eq:normvel}).
-Experiments with increasing pattern velocity $v_{\theta} = \chi v_0$ with $\chi\in\{0,10\}$ show a pronounced change in dewetting morphology from well defined droplets to ligaments.
-With the single balance between pattern velocity $v_{\theta}$ and capillary velocity $U_{\text{Ca}}$ we can identify a parameter that separates the two morphological regimes.
-
-An outlook to this work is to study the transition from ligaments to droplets in a more detailed manner.
-Looking at the edges $\Gamma =1$ with more suitable tools like e.g. bifurcation diagramms will likely be subject to future work.
-But also the update scheme of the contact angle field allows for many applications.
-Among them is the avenue of switchable substrates, where e.g. switching of $\theta(\mathbf{x})$ due to a light source is in the scope of this method.
 
 \begin{acknowledgements}
-The authors acknowledge financial support by the Deutsche Forschungsgemeinschaft (DFG) within the priority program SPP2171 ``Dynamic Wetting of Flexible, Adaptive, and Switchable Substrates'', project HA-4382/11.
+The authors acknowledge financial support by the Deutsche Forschungsgemeinschaft (DFG) within the Cluster of Excellence ''Engineering of Advanced Materials'' (project EXC 315) (Bridge Funding). 
 \end{acknowledgements}

### 2023-04-21 SuppMat.tex
New replies, thanks for nothing ref A

@@
 \maketitle
-\section{Minkowski's structure metric $q_2$ in an extended parameter space}
+
+\section{Disjoining pressure}
+
+\noindent A {\it thin film} is a layer of liquid confined between two interfaces. When the two interfaces are brought close enough to each other (i.e. the layer is {\it thin}), they interact.
+The disjoining pressure stems from this interaction. Thermodynamically, it represents minus the derivative of the free energy difference, per unit surface, between 
+the film and the bulk of the same liquid, or, equivalently, the pressure difference inside and outside the film 
+\cite{Deryaguin1940,DeryaguinChuraev1978}.
+In the case of wetting, there can be three types of interfaces: liquid/air, liquid/solid (the substrate) and, 
+if the substrate is only partially wetted, solid/air, thus giving rise to three-phase contact lines. 
+The usually called "surface tension" is the energy cost of forming a liquid/air interface of unit area. 
+The Young's condition \cite{Young1805,DeGennes1985} connects the surface tension with the analogous energies per unit
+area associated with the solid/liquid ($\gamma_{\text{sl}}$) and solid/gas ($\gamma_{\text{sg}}$) interface as
+\begin{equation}
+\cos \theta = \frac{\gamma_{\text{sg}}-\gamma_{\text{sl}}}{\gamma},
+\end{equation}
+where $\theta$ is the contact angle.
+The expression adopted for $\Pi$, namely
+\begin{equation}
+\Pi(h,\theta) = \frac{2\gamma}{h_{\ast}}(1-\cos(\theta(\mathbf{x},t)))
+  f\left(\frac{h}{h_{\ast}}\right),
+\end{equation}
+where $f(\xi)=\xi^{-3} - \xi^{-2}$ and $h_{\ast}$ is the height at which the disjoining pressure vanishes setting the precursor layer thickness, pertains to a general family of expressions of 
+the form 
+\begin{equation}\label{eq:SMDP}
+\Pi(h) = \gamma(1-\cos \theta)\frac{(n-1)(m-1)}{h_{\ast}(n-m)}\left[\left(\frac{h_{\ast}}{h}\right)^n - 
+\left(\frac{h_{\ast}}{h}\right)^m\right].
+\end{equation}
+Equations like (\ref{eq:SMDP}) are frequently found in the literature, 
+with various combination of exponents $(n,m)$ ($n>m$), depending
+on the type of interactions among fluid molecules and between fluid and substrate~\cite{SCHWARTZ1998173,Mitlin,Teletzke}. A common choice, for instance, is the pair $(9,3)$, which 
+can be derived from molecular interactions of Lennard-Jones type. Here, the use of the $(3,2)$ pair is motivated mainly by the computational advantage of keeping the characteristic dewetting time 
+shorter. We have run also a single case ($\lambda=256 h_0$ and $\Gamma=1.5$) with $(9,3)$ and observed that 
+the dynamics remains essentially the same, but for a shift in the time scale (see also~\cite{PhysRevE.104.034801}).
+
+\section{Parameter values and comparison with experiments}\label{sec:SMparam}
+
+\noindent To give a flavour of how to compare with actual experimental systems, let us notice
+first that the expression of $t_0$ depends on the disjoining pressure (through its dependence 
+ $q_0 = \sqrt{\frac{1}{2\gamma} \Pi^{\prime}(h_0)}$) and contains, therefore, energetic details, namely 
+ the Hamaker constant $A_H$.
+Consider, for instance, the problem studied in \cite{becker2003complex}: the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon
+dioxide substrate, where $A_H  = 2.2 \times 10^{-20} \, \text{J}$ and 
+$\gamma = 0.03 \, \text{N}/\text{m}$. Neglecting the short range part of the interface potential
+(such that $\Pi(h)= - \frac{A_H}{6\pi h^3}$)~\cite{Mecke_2005,becker2003complex}, one gets 
+$q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}$.
+The characteristic time of the growth of the instability for these experiments then reads $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}$ (the dynamic viscosity of polystyrene is 
+$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$)~\cite{PhysRevLett.99.114503}. It can be seen that indeed film rupture occurs within time 
+scales of the order of $t_0$~\cite{becker2003complex}, as in our 
+simulation for the largest pattern wavelength, $\lambda = L$ (see also our previous work
+with homogeneous substrates~\cite{PhysRevE.104.034801}).\\
+All our simulations are run on a bi-periodic square domain of size $L \times L$ with 
+$L/h_0 = 512 $. The numerical values in lattice Boltzmann units of the 
+mean film thickness, dynamics viscosity and surface tension are, respectively, $h_0=1$, $\mu=1/6$
+and $\gamma=0.01$.
+To regularize the contact line divergence~\cite{huh1971hydrodynamic}
+we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$, 
+a value which lies within the weak/intermediate slip regime~\cite{peschka2019signatures,fetzer2007quantifying, munch2005lubrication}
+(we have tried also a value four times smaller, without observing any major qualitative change in the results).
+The liquid film is initialized with a height field slightly perturbed around the mean value $h_0$,
+i.e. $h(\mathbf{x},0) = h_0 \left[1 + 0.1 \left(\sin\left(\frac{2\pi x}{L}\right)\sin\left(\frac{2\pi y}{L}\right)\right)\right]$.
+Various wavelengths in the range $\lambda/h_0 \in [64, 512]$, and velocities,
+$v_{\theta}/v_0 \in [0, 200]$, are considered for the wettability pattern (see main text). 
+A few comments on these ranges are noteworthy at this point. The wavelengths are obviously limited from above by the size of the
+computational domain. On the other extreme, too short wavelengths are prevented by
+the energetic cost of accomodating droplets on too
+small patches, since the local contact angle 
+(that increases with decreasing $\lambda$, see also \ref{sec:SMdroplet}) would possibly exceed the largest substrate contact angle,
+such that the patterning itself would loose meaning. 
+It has been, in fact, proven, both theoretically and experimentally, that 
+for the confinement to be effective, the pattern characteristic size should 
+not be much smaller than the spinodal length scale on the less wettable substrate
+\cite{KarguptaJCP2002,KarguptaLangmuir2000,Nisato1999,Karim1998}.
+We keep, therefore, $\lambda \stackrel{>}{\sim} \lambda_s \approx 70 h_0$.
+The pattern speeds, on the other hand, are limited by the intrinsic LB bound of the lattice speed of sound
+ \cite{succi}.\\
+The Reynolds number, $Re$, is of the order of $Re \sim 10^{-2}$ on the 
+static substrate and never exceeds the value $\approx 0.2$ in the time dependent case.
+
+\section{Minkowski's structure metric $q_2$ in an extended parameter space}\label{sec:SMMinko}
+
 \begin{figure}
@@
 \end{figure}
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=0.7\textwidth]{different_contrasts.pdf}
-%    \caption{Comparison of the Minkowski's structure metric $q_2$ for different values of %$\delta\theta$, 
-%    $\Gamma = 1.5$ and $\lambda = 256 h_0$.}
-%    \label{fig:dif_contrast}
-%\end{figure}
 \noindent In this section we test the robustness of the observation of the rivulet 
@@
 the patterning looses effectiveness as the contact angle mismatch is reduced (see, e.g.~\cite{KonnurPRL2000}).
+%\newpage
 
-\newpage
-
-\section{Droplet shape}
+\section{Droplet shape}\label{sec:SMdroplet}
 
@@
 $\gamma_{\text{sg}}$ are the liquid/gas, solid/liquid and solid/gas interface energies 
-per unit area \cite{WuSM2020}. 
+per unit area \cite{Wu2020}. 
 In particular, $\gamma_{\text{lg}} \equiv \gamma$ is the surface tension.
@@
 \end{equation}  
-The minimum condition $\frac{\partial E}{\partial a} = 0$ \cite{footnote} yields
+The minimum condition $\frac{\partial E}{\partial a} = 0$~\footnote{It can be easily checked that 
+$\frac{\partial^2 E}{\partial a^2} > 0$.} yields  
 \begin{equation}\label{eq:minim}
@@
 
-\begin{thebibliography}{99}
+\bibliography{Ref}
 
-\bibitem{KonnurPRL2000} R. Konnur, K. Kargupta and A. Sharma, {\it Phys. Rev. Lett.} {\bf 84}, 931--934 (2000).
-\bibitem{WuSM2020} Y. Wu, F. Wang, S. Ma, M. Selzer and B. Nestler,
-{\it Soft Matter} {\bf 16}, 6115--6127 (2020).
-\bibitem{footnote} It can be easily checked that $\frac{\partial^2 E}{\partial a^2} > 0$ for $a>0$.
+%\begin{thebibliography}{99}
 
-\end{thebibliography}
+%\bibitem{KonnurPRL2000} R. Konnur, K. Kargupta and A. Sharma, {\it Phys. Rev. Lett.} {\bf %84}, 931--934 (2000).
+%\bibitem{WuSM2020} Y. Wu, F. Wang, S. Ma, M. Selzer and B. Nestler,
+%{\it Soft Matter} {\bf 16}, 6115--6127 (2020).
+%\bibitem{footnote} It can be easily checked that $\frac{\partial^2 E}{\partial a^2} > 0$ %for $a>0$.
+
+%\end{thebibliography}
 

### 2023-04-21 main_shortened.tex
New replies, thanks for nothing ref A

--- 2022-12-14 main.tex
@@
 \newcommand{\JH}[1]{\textcolor{blue}{ JH: #1}}
+\newcommand{\REV}[1]{\textcolor{red}{#1}}
+%\newcommand{\REV}[1]{#1}
 
@@
 \begin{abstract}
-\noindent Switchable and adaptive substrates emerged as valuable tools for the control of wetting and actuation of droplet motion.
-Here we report a computational study of the dynamics of an unstable thin liquid film deposited on a switchable substrate, modelled with a space and time varying contact angle.
-With a static pattern, all the fluid is drained into droplets located around contact angle minima, whereas for a sufficiently large rate of wettability variation a state consisting of metastable rivulets is observed.
-A criterion discriminating whether rivulets can be observed or not is identified in terms of a single dimensionless parameter.
+\noindent Switchable and adaptive substrates emerged as valuable tools for controlling wetting and actuation of droplet motion.
+Here, we report a computational study of the dynamics of an unstable thin liquid film deposited on a switchable substrate, modeled with a space and time varying contact angle.
+With a static pattern, all fluid is drained into droplets located around contact angle minima, whereas for a sufficiently large rate of wettability variation a state consisting of metastable rivulets is observed.
+A criterion discriminating whether rivulets occur or not is identified in terms of a single dimensionless parameter.
 Finally, we show and explain theoretically how the film rupture times, droplet shape and rivulet life time depend on the pattern wavelength and speed.  
@@
 \noindent {\it Introduction}.
-Wet surfaces and droplets are part of our every-day experience and of numerous industrial processes including coating, tribology, painting and printing, to name but a few~\cite{gross1980fluid,szeri2010fluid,DERYCK1998278,doi:10.1146/annurev.fluid.31.1.347,DASILVASOBRINHO19991204,singh2010inkjet,jo2009evaluation,Wijshoff2010}.
-Moreover, the continuously growing technological interest for lab-on-a-chip devices~\cite{C6LC00387G,Focke} as well as for printable electronics~\cite{Kim_2005, Luechinger_2008} and printable photovoltaics~\cite{Brabec2001,RH20}, whose efficiency relies crucially on a precise control of material deposition upon (de-)wetting of liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e.
-it is not inert but responds dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{ButtEtAl_Langmuir2018,GuoGuo_RSCAdv2016}.
-Several realizations of switchable and adaptive substrates have been proposed~\cite{LiuChem2005, XiCSR2010}, involving smart materials such as polymer brushes~\cite{CohenStuartEtAl_NatMat2010,ayres2007stimuli,YongEtAl_Mat2018}, thermal-responsive hydrogels~\cite{ChenEtAl_SM2010}, light-responsive molecules and microstructures~\cite{IchimuraEtAl_Science2000,DelormeEtAl_Langmuir2005,OscuratoEtAl_AAMI2017} or processes such as electrowetting~\cite{MugeleEtAl_JPCM2005}.
+Wet surfaces and droplets are part of our every-day experience and of numerous industrial processes including coating, tribology, painting and printing, to name but a few~\cite{gross1980fluid,szeri2010fluid,doi:10.1146/annurev.fluid.31.1.347,DASILVASOBRINHO19991204,singh2010inkjet,Wijshoff2010}.
+Moreover, the continuously growing interest for lab-on-a-chip devices~\cite{C6LC00387G,Focke} as well as for printable electronics or printable photovoltaics~\cite{Luechinger_2008,RH22}, whose efficiency relies crucially on a precise control of material deposition upon (de-)wetting of liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e.
+it is not inert but responds dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{ButtEtAl_Langmuir2018}.
+Several realizations of switchable and adaptive substrates have been proposed~\cite{XiCSR2010}, involving smart materials such as polymer brushes~\cite{CohenStuartEtAl_NatMat2010}, thermal-responsive hydrogels~\cite{ChenEtAl_SM2010}, light-responsive molecules and microstructures~\cite{IchimuraEtAl_Science2000} or processes such as electrowetting~\cite{MugeleEtAl_JPCM2005}.
 Probably the simplest, yet non-trivial, modelling of a switchable substrate can be realized by a space and time dependent wettability pattern~\cite{GrawitterStark1}. 
-While a consistent body of theoretical/computational work has been devoted to processes on static heterogeneous substrates, though, the time dependent case is still almost unexplored, with few relevant exceptions focusing on single droplet spreading and sliding~\cite{GrawitterStark1,GrawitterStark2,ThieleHartmann} or limited to an analysis of the linear regime~\cite{suman2006dynamics}.
+While a consistent body of theoretical/computational work was devoted to processes on static heterogeneous substrates, the time dependent case is still almost unexplored, with few exceptions focusing on single droplet spreading and sliding~\cite{GrawitterStark1,GrawitterStark2,ThieleHartmann} or limited to analysing the linear regime~\cite{suman2006dynamics}.
 
-In this Letter, we study, by means of numerical simulations, the full dewetting dynamics of a thin liquid film deposited over a substrate with a time varying wettability pattern, from rupture to the long time morphology.
-We identify two regimes where the rupture times grow with the pattern wavelength either linearly (on the static pattern) or attain a constant value (in the time dependent case), for short wavelengths, and approach a quadratic law as the wavelength increases.
+Here, we study by means of numerical simulations the full dewetting dynamics of a thin liquid film deposited on a substrate with a time varying wettability pattern, from rupture to the long time morphology.
+We identify two regimes where the rupture times grow with the pattern wavelength either linearly (on a static pattern) or attain a constant value (in the time dependent case), for short wavelengths, and approach a quadratic law as the wavelength increases.
 These observations are then explained theoretically.
-We show that, by tuning the rate of change (the ``speed'') of the underlying pattern, it is possible to control, to some extent, the dewetting morphology.
+We show that, by tuning the rate of change (the ``speed'') of the underlying pattern, one can control, to some extent, the dewetting morphology.
 In particular, for large enough pattern speeds, we detect a metastable state, where the film retracts into metastable rivulets, eventually breaking up into multiple droplets.
-We introduce a control parameter that discriminates whether rivulets or just droplets (as in the static situation)
-can be observed and we propose a phenomenological argument that justifies the logarithmic dependence of the rivulets life time on the pattern speed.\\
+We introduce a control parameter to discriminate whether rivulets or just droplets (as in the static situation)
+can be observed and propose a phenomenological argument to justify the logarithmic dependence of the rivulets life time on the pattern speed.\\
 \\
-\noindent {\it Method.} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{ReynoldsLubr,RevModPhys.69.931,RevModPhys.81.1131}
+\noindent {\it Method.} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{RevModPhys.69.931,RevModPhys.81.1131}
 \begin{equation}\label{eq:thinfilm}
@@
 \end{equation}
-by means of a recently developed method, based on a lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801, Zitz2022}.
+by means of a recently developed lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801, Zitz2022}.
 Eq.~(\ref{eq:thinfilm}) describes, in a lubrication approximation spirit, the evolution of the height field (film thickness) $h(\mathbf{x},t)$, denoting the location of the liquid/air interface.
@@
 (for $\delta \rightarrow 0$ it reduces to the no-slip form $h^3/(3\mu)$). Here, $\mu$ is the fluid dynamic viscosity.
-The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$~\footnote{\textcolor{red}{The Laplace pressure is the product of the surface tension
-and the local curvature, therefore its full expression reads $p_L = -\gamma \nabla^2 h/\sqrt{1+|\nabla h|^2}$; 
-since we are in the lubrication regime, though, whereby $|\nabla h| \ll 1$, the denominator can be safely neglected 
+The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$ \REV{($\gamma$ is the surface tension)}~\footnote{\REV{The Laplace pressure is the product of the surface tension
+and the local curvature. Therefore, its full expression reads $p_L = -\gamma \nabla^2 h/\sqrt{1+|\nabla h|^2}$. 
+In the lubrication regime, whereby $|\nabla h| \ll 1$, the denominator can be safely neglected 
 and the pressure reduces to the form involving only the Laplacian of the thickness 
-field~\cite{Benet2014,juanes2018}.}}, 
-\textcolor{red}{where $\gamma$ is the surface tension}.
+field}~\cite{Benet2014,juanes2018}}.
 The disjoining pressure $\Pi$ can be seen as (minus) the derivative, with respect to the film thickness, of an effective interfacial potential.
@@
 \begin{equation}\label{eq:disjoinpressure}
-\textcolor{red}{\Pi(h,\theta) = \frac{2\gamma}{h_{\ast}}(1-\cos(\theta(\mathbf{x},t)))
+\REV{\Pi(h,\theta) = \frac{2\gamma}{h_{\ast}}(1-\cos(\theta(\mathbf{x},t)))
   f\left(\frac{h}{h_{\ast}}\right),}
 \end{equation}
-\textcolor{red}{where $f(\xi)=\xi^{-3} - \xi^{-2}$} and $h_{\ast}$ is the height at which the disjoining pressure vanishes and sets the precursor layer thickness~\footnote{\textcolor{red}{Expressions of
-the general form $\Pi(h) = \gamma(1-\cos \theta)(n-1)(m-1)/(n-m)((h_{\ast}/h)^n - (h_{\ast}/h)^m)/h_{\ast}$
-are frequently found in the literature, with various combination of exponents $(n,m)$ ($n>m$), depending
-on the type of interactions among fluid molecules and between fluid and substrate~\cite{SCHWARTZ1998173,Mitlin,Teletzke}. A common choice, for instance, is the pair $(9,3)$, which 
-can be derived from molecular interactions of Lennard-Jones type. Here, the use of the $(3,2)$ pair is motivated mainly by the computational advantage of keeping the characteristic dewetting time 
-shorter. We have run also a single case ($\lambda=256 h_0$ and $\Gamma=1.5$) with $(9,3)$ and observed that 
-the dynamics remains essentially the same, but for a shift in the time scale (see also~\cite{PhysRevE.104.034801})}}.
-The time variation of the patterned substrate then enters the model precisely through the disjoining pressure, by making the contact angle space and time dependent, i.e.
-$\theta = \theta(\mathbf{x},t)$.
-\textcolor{red}{We decorate the substrate with a checkerboard pattern, a common choice that generalizes the broken homogeneity of the stripes to two 
-directions~\cite{KarguptaLangmuir2003,Jalali2018,Nagayama2020,Das2020}.}
+\REV{where $f(\xi)=\xi^{-3} - \xi^{-2}$}. $h_{\ast}$ is the height at which $\Pi$ vanishes and sets the precursor layer thickness~\cite{SuppMat}.
+The time variation of the patterned substrate enters the model through the disjoining pressure, by making the contact angle space and time dependent, i.e.~$\theta = \theta(\mathbf{x},t)$.
+\REV{We decorate the substrate with a checkerboard pattern, a common choice that generalizes the broken homogeneity of the stripes to two 
+directions~\cite{Jalali2018,Nagayama2020,Das2020}.}
 In particular, we employ the sinusoidal form
@@
 where $q_{\theta} = 2\pi/\lambda$, i.e.~the pattern evolves in time as a plane wave.
-% \textcolor{red}{A possible laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}.
+% \REV{A possible laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}. 
 % The external stimulus, our moving pattern, can be realized, e.g., by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.}
-We fix the velocity direction to one diagonal, namely $\mathbf{v}_{\theta} = (v_{\theta x},v_{\theta y}) = v_{\theta}(1/\sqrt{2},-1/\sqrt{2})$ (we will return later on the importance of this choice), and we set $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$~\footnote{Notice that, since a typical velocity is such that $v_{\theta} \Delta t \ll \Delta x$ (in one LB time step $\Delta t$ the wave would travel a distance much smaller than a lattice spacing $\Delta x$),
-the time update needs to be interpreted in an integer part sense, that is the pattern is shifted by one $\Delta x$ every $1/v_{\theta x}$ time steps (and equivalently in the $y$-direction).}.
-\textcolor{red}{Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film \cite{Mecke_2005}.
+We fix the velocity direction to one diagonal, namely $\mathbf{v}_{\theta} = (v_{\theta x},v_{\theta y}) = v_{\theta}(1/\sqrt{2},-1/\sqrt{2})$ (we will return later on the importance of this choice), and we set $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$~\footnote{Since a typical velocity is such that $v_{\theta} \Delta t \ll \Delta x$ (in one time step $\Delta t$ the wave would travel a distance much smaller than a lattice spacing $\Delta x$),
+the time update needs to be interpreted in an integer part sense, that is the pattern is shifted by $\Delta x$ every $1/v_{\theta x}$ $\Delta t$ (equivalently in $y$-direction).}.
+\REV{Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film~\cite{Mecke_2005}.
 On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads
-$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$ \cite{Mecke_2005,PhysRevE.100.023108}. In our patterned case, we define 
+$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$~\cite{Mecke_2005,PhysRevE.100.023108}. In our patterned case, we define 
   $q_0^2=h_{\ast}^{-2}(1-\cos\theta_0)f^{\prime}(h_0/h_{\ast})$. Correspondingly, we choose as a velocity scale $v_0 = \lambda_s/t_0$, where
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{Figure_1.png}
+    \includegraphics[width=0.4\textwidth]{Figure_1.png}
     \caption{Stationary film thickness field ($t>t_0$) showing the formation of droplets. The color map indicates the contact angle pattern 
-    (Eq.~(\ref{eq:sinetheta}) with $v_{\theta}=0$), with lower (higher) values coded in light blue (yellow).
+    (Eq.~(\ref{eq:sinetheta}) with $v_{\theta}=0$), with lower (higher) values in light blue (yellow).
     }
@@
 %%%%
-\textcolor{red}{To give a flavour of how to compare with actual experimental systems, let us notice
-first that the expression of $t_0$ depends on the disjoining pressure (through its dependence 
- $q_0 = \sqrt{\frac{1}{2\gamma} \Pi^{\prime}(h_0)}$) and contains, therefore, energetic details, namely 
- the Hamaker constant $A_H$.
-Consider, for instance, the problem studied in \cite{becker2003complex}: the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon
-dioxide substrate, where $A_H  = 2.2 \times 10^{-20} \, \text{J}$ and 
-$\gamma = 0.03 \, \text{N}/\text{m}$. Neglecting the short range part of the interface potential
-(such that $\Pi(h)= - \frac{A_H}{6\pi h^3}$)~\cite{Mecke_2005,becker2003complex}, one gets 
-$q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}$.
-The characteristic time of the growth of the instability for these experiments then reads $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}$ (the dynamic viscosity of polystyrene is 
-$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$)~\cite{PhysRevLett.99.114503}. It can be seen that indeed film rupture occurs within time 
-scales of the order of $t_0$~\cite{becker2003complex}, as in our 
-simulation for the largest pattern wavelength, $\lambda = L$ (see also our previous work
-with homogeneous substrates~\cite{PhysRevE.104.034801}).}
-%%%%
-All our simulations are run on a bi-periodic square domain of size $L \times L$ with 
-$L/h_0 = 512 $. \textcolor{red}{The numerical values in lattice Boltzmann units of the 
-mean film thickness, dynamics viscosity and surface tension are, respectively, $h_0=1$, $\mu=1/6$
-and $\gamma=0.01$.}
-\textcolor{red}{To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$, 
-a value which lies
-within the weak/intermediate slip regime~\cite{peschka2019signatures,fetzer2007quantifying, munch2005lubrication} (we have tried also a value four times smaller, without observing 
-any major qualitative change in the results).} 
-The liquid film is initialized with a height field slightly perturbed around the mean value $h_0$, i.e.
-$h(\mathbf{x},0) = h_0 \left[1 + 0.1 \left(\sin\left(\frac{2\pi x}{L}\right)\sin\left(\frac{2\pi y}{L}\right)\right)\right]$.
-Various wavelengths in the range \textcolor{red}{$\lambda/h_0 \in [64, 512]$}, and velocities,
-$v_{\theta}/v_0 \in [0, 200]$, are considered for the wettability pattern, Eq.~(\ref{eq:sinetheta}). 
-\textcolor{red}{A few comments on these ranges are noteworthy at this point.
-The wavelengths are obviously limited from above by the size of the computational domain.
-On the other extreme, too short wavelengths are prevented by
-the energetic cost of accomodating droplets on too
-small patches, since the local contact angle 
-(that increases with decreasing $\lambda$~\cite{SuppMat}) would possibly exceed the largest substrate contact angle, such that the patterning 
-itself would loose meaning. 
-It has been, in fact, proven, both theoretically and experimentally, that 
-for the confinement to be effective, the pattern characteristic size should 
-not be much smaller than the spinodal length scale on the less wettable substrate
-\cite{KarguptaJCP2002,KarguptaLangmuir2000,Nisato1999,Karim1998}.
-We keep, therefore, $\lambda \stackrel{>}{\sim} \lambda_s \approx 70 h_0$.
-The pattern speeds, on the other hand, are limited by the intrinsic LB bound of the lattice speed of sound
- \cite{succi}.}\\
-%\textcolor{red}{Hereafter, we will omit 
-%to write explicitly the division by $h_0$, $t_0$ and $v_0$ that will be, then, 
-%surmised, unless else specified.}
-\textcolor{red}{The Reynolds number, $Re$, is of the order of $Re \sim 10^{-2}$ on the 
-static substrate and never exceeds the value $\approx 0.2$ in the time dependent case.}
-In Fig.~\ref{fig:handtheta} we show $h(\mathbf{x},t)$ (droplets) and $\theta(\mathbf{x})$ (color coded) for $v_{\theta} = 0$ (i.e.
-the static case) and \textcolor{red}{$\lambda = 256 h_0$}\footnote{For a better visualization we take twice the domain length $L$ and periodically continue the image.}, in late stages of dewetting.
-As expected, droplets form in regions of small contact angles (blue) while the regions of high contact angles (yellow) dewet.\\
+Fig.~\ref{fig:handtheta} shows $h(\mathbf{x},t)$ (droplets) and $\theta(\mathbf{x})$ (color coded) for $v_{\theta} = 0$ (i.e.
+the static case) and \REV{$\lambda = 256 h_0$}~\footnote{For a better visualization double the domain length $L$ and periodically continue the image.}, in late stages of dewetting.
+As expected, droplets form in regions of small contact angles (blue) while the regions of high contact angles (yellow) dewet.
 
-\noindent {\it Results.} We start our analysis by studying first how the rupture times depend on the parameters characterising the wettability pattern, namely the wavelength of the contact angle variation, $\lambda$, and wave speed, $v_{\theta}$.
-The film rupture time, $\tau_r$, is defined as the least $t$ such that $h(\mathbf{x},\tau_r)=h_{\ast}$ (that is, when the free surface "touches" the substrate).
+\noindent {\it Results.} We first investigate how the rupture times depend on the parameters characterizing the wettability pattern, namely the wavelength of the contact angle variation, $\lambda$, and wave speed, $v_{\theta}$.
+The film rupture time, $\tau_r$, is defined as the least $t$ such that $h(\mathbf{x},\tau_r)=h_{\ast}$ (i.e., when the free surface "touches" the substrate).
 In Fig.~\ref{fig:model_rt} we report the rupture times as a function of the wavelength, for stationary ($v_{\theta}=0$) and time-dependent ($v_{\theta}=20 v_0$) patterns. 
 It is conveyed that, overall, rupture occurs earlier on the static substrate, suggesting that the time variation tends to stabilize the film, in agreement with linear stability analysis results~\cite{suman2006dynamics}.
-We observe that $\tau_r$ grows linearly with $\lambda$, for short wavelengths, and quadratically for longer $\lambda$.
-These facts can be explained, qualitatively, as follows. 
-Let us first notice that the film rupture consists of two processes, each with its associated time scale: the growth of unstable interface perturbation with rate $t_{\theta}^{-1}$, and the retraction of liquid from more hydrophobic regions, occurring in a characteristic time $t_R$.
-For longer pattern wavelengths, the dewetting instability amplifies more slowly, such that it is reasonable to take it as the process determining the rupture time, $\tau_r \sim t_{\theta}$.
+We observe that $\tau_r$ grows linearly with $\lambda$ for short wavelengths and quadratically for longer $\lambda$.
+These facts can be qualitatively explained as follows. 
 In this case, from the linearized thin-film equation (in one spatial dimension, for simplicity), obtained setting $h=h_0 + \delta h$ with $\delta h \ll h_0$, 
@@
 The latter is related to the time the liquid takes to flow out of regions of high contact angle, whose size is $\sim \lambda$. Hence we have 
+ %\tau_r \sim \tau_R \propto \frac{1}{U_{\Theta}}\lambda,
 \begin{equation}\label{eq:taur_l1}
- \tau_r \sim \tau_R \propto \frac{1}{U_{\Theta}}\lambda,
+ \tau_r \sim \tau_R \propto U_{\Theta}^{-1}\lambda,
 \end{equation}
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{Figure_2.pdf}
+    \includegraphics[width=0.4\textwidth]{Figure_2.pdf}
     \caption{Rupture times, $\tau_r$, as a function of the pattern wavelength $\lambda$, for $v_{\theta}=0$ (\textcolor{jlblue}{$\bullet$}) and $v_{\theta}=20 v_0$ 
@@
 \end{figure}
-%Interestingly, this phenomenology is qualitatively supported %also by the time-dependent case, $v_{\theta} = 20 v_0$ (reported %in Fig.~\ref{fig:model_rt} as orange stars), \textcolor{red}{but %for a smaller prefactor in the linear regime, due to the fact %that the pattern wave velocity is large enough to dominate over %$U_{\Theta}$ and sets the retraction speed, such that the %characteristic time becomes $\tau_R \sim \lambda/v_{\theta}$.}
-%\begin{equation}\label{eq:taur_saturation}
-% \tau_r \sim \tau_R \sim \frac{\lambda}{v_{\theta}} \propto %\frac{\lambda}{v_0} = \lambda\frac{t_0}{\lambda} \sim \mbox{const}.
-%\end{equation}
-We now focus on the long time dynamics and in particular on the characterization of the dewetting morphologies and how they are affected by the speed of the wettability wave.
+We now focus on the long time dynamics, the characterization of the dewetting morphologies and how they are affected by the speed of the wettability wave.
 \begin{figure}
     \centering
-    \includegraphics[width=0.45\textwidth]{Figure_3.pdf}
+    \includegraphics[width=0.4\textwidth]{Figure_3.pdf}
     \caption{MAIN PANEL: Time evolution of the height fluctuations, $\Delta h(t)$, during the dewetting process on the patterned substrate given by       
@@
     INSET: Number of droplets, $N(t)$, as a function of time. The three horizontal dashed lines indicate the number of minima of Eq.~(\ref{eq:sinetheta}),
-      which is $2\left(\frac{L}{\lambda}\right)^2$. In the insets, we display snapshots of the stationary droplet states as grey-scale images
+      which is $2\left(\frac{L}{\lambda}\right)^2$. The napshots depict the stationary droplet states as grey-scale images
       of the film thickness field, $h(\mathbf{x},t)$.
@@
 \end{figure}
-On the stationary substrate, after rupture all the fluid accumulates in droplets centered at contact angle minima.
-Consequently, as we see from the inset of Fig.~\ref{fig:clusters_v0_sine}, where we plot the number of droplets $N(t)$ in time\footnote{A droplet is identified by the set (``cluster'') of points, in the plane, constituting each of the connected components of the set $\{\mathbf{x} \in [0,L]^2 | h(\mathbf{x},t) \geq h_{\ast}$\}; the clusters are determined by means of an algorithm of Hoshen-Kopelman type~\cite{HK}.}, in the steady state ($t \gg t_0$) $N(t)$ attains the value $N_{\infty} = 2(L/\lambda)^2$ (reported as horizontal lines), which equals the minima of Eq.~(\ref{eq:sinetheta}), for $v_{\theta}=0$.
+On the stationary substrate, after rupture all fluid accumulates in droplets centered at contact angle minima.
+Consequently, as seen from the inset of Fig.~\ref{fig:clusters_v0_sine}, where we plot the number of droplets $N(t)$ vs.
+time~\footnote{A droplet is identified by the set (``cluster'') of points, in the plane, constituting each of the connected components of the set $\{\mathbf{x} \in [0,L]^2 | h(\mathbf{x},t) \geq h_{\ast}$\}.
+% the clusters are determined by means of a Hoshen-Kopelman algorithm~\cite{HK}.
+},
+in the steady state ($t \gg t_0$) $N(t)$ attains the value $N_{\infty} = 2(L/\lambda)^2$ (horizontal lines), which equals the minima of Eq.~(\ref{eq:sinetheta}), for $v_{\theta}=0$.
 Notice that the number of droplets converges faster for smaller pattern wavelengths, in line with the observation reported and justified in the previous section that the characteristic dewetting time decreases with the wavelength.\\
-\textcolor{red}{In the main panel of the same figure, we see that the height fluctuations $\Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\}-\min_{\mathbf{x}}\{h(\mathbf{x},t)\}$
+\REV{In the main panel of the figure, the height fluctuations $\Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\}-\min_{\mathbf{x}}\{h(\mathbf{x},t)\}$
 grow in time until film rupture and then settle to a constant value. This represents a measure of the mean droplet height $h_d$ (since
-droplets are essentially monodisperse), decreasing with the pattern wavelength (as expected, due to the decrease of the droplet volume,
+droplets are essentially monodisperse), decreasing with the pattern wavelength (as expected, due to a decreasing droplet volume,
 $V_d = \frac{h_0 \lambda^2}{2}$).}\\
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{Figure_4.pdf}
-    \caption{Time evolution of the second order Minkowski structure metric, $q_2(t)$ (see text for details), for different $\Gamma$ values, on a substrate with pattern wavelength $\lambda=256 h_0$.
-    The grey-scale insets supply snapshots of the film thickness field associated to the various curves.
-    }
+    \includegraphics[width=0.4\textwidth]{Figure_4.pdf}
+    \caption{Time evolution of the second order Minkowski structure metric, $q_2(t)$, for different $\Gamma$ values, on a substrate with pattern wavelength $\lambda=256 h_0$.
+    The grey-scale insets supply snapshots of the corresponding film thickness fields.}
     \label{fig:msm_q2}
@@
     % \includegraphics[width=0.45\textwidth]{Figures/Lig_stab_inset_delta_h.pdf}
-    \includegraphics[width=0.45\textwidth]{Figure_5.pdf}
-    \caption{MAIN PANEL: Rivulets life-times, $\tau_{\text{riv}}$, for various $\Gamma$ values.
-    The dashed line is plotted as a guide to the eye to highlight the logarithmic dependence, in agreement with 
+    \includegraphics[width=0.4\textwidth]{Figure_5.pdf}
+    \caption{MAIN PANEL: Rivulet life-times, $\tau_{\text{riv}}$, for various $\Gamma$.
+    The dashed line is a guide to the eye to highlight the logarithmic dependence, in agreement with 
     the theoretical prediction, Eq.~(\ref{eq:rivlt}).
-    INSET: Height fluctuations, $\Delta h(t)$, vs time, computed along the rivulet axis, for three different $\Gamma$ values.
+    INSET: Height fluctuations, $\Delta h(t)$, vs time, along the rivulet axis, for three different $\Gamma$.
     }
@@
 The breakup is the result of a varicose mode of the rivulet~\cite{doi:10.1063/1.3211248, PhysRevE.77.061605}, whose wavelength is $\approx\lambda$, such that only $N_{\infty}/2$ droplets are counted after breakup.
-These droplets show a peculiar dynamics, characterized by a periodic sequence of spreading and retraction, driven by the pattern, that we dub "pumping state"~\footnote{see movie file ligament\_formation\_and\_breakup.mp4 in~\cite{SuppMat}.}.
+These droplets show a peculiar dynamics, characterized by a periodic sequence of spreading and retraction, driven by the pattern, that we dub "pumping state"~\footnote{see movie ligament\_formation\_and\_breakup.mp4 in~\cite{SuppMat}.}.
 
 We argue that the emergence of rivulets is controlled by the competition of two characteristic velocities: the pattern wave speed, $v_{\theta}$, and the retraction speed, $U_{\Theta}$, introduced in Eq.~(\ref{eq:taur_l1}).
-If $U_{\Theta}$ is large as compared to $v_{\theta}$, the film retraction is faster than the transport due to the contact angle field and thus droplets from.
+\REV{If $U_{\Theta}$ is large as compared to $v_{\theta}$, the film retraction is faster than 
+the local contact angle variation and thus droplets form.}
+%the transport due to the contact angle field and thus droplets from. 
 However, if $v_{\theta}$ is larger than $U_{\Theta}$, then the retracting film has too little time to form droplets and ends up in the metastable rivulet state.
-\textcolor{red}{It appears, therefore, natural to consider the ratio of these two velocities,
+\REV{It appears, therefore, natural to consider the ratio of these two velocities,
 $\Gamma \equiv v_{\theta}/U_{\theta}$, as the discriminating parameter.}
@@
 However, as the pattern moves, the instability is tamed due to configurations whereby higher contact angle regions underlie height field maxima, thus tending to revert the fluid flow.
-Heuristically speaking, this means that, if we evaluate $\Delta h(t)$ restricted on the rivulet axis, this should grow exponentially (with a certain growth rate $\alpha \propto t_0$) only when the system is in the unstable configuration.
+Heuristically speaking, this means that, if we evaluate $\Delta h(t)$ restricted on the rivulet axis, it should grow exponentially (with a certain growth rate $\alpha \propto t_0$) only when the system is in the unstable configuration.
 Namely, $\Delta h(t)/\Delta h_0 \propto e^{\alpha t}$ (see inset of Fig.~\ref{fig:stab_ligs_lam2}) with a prefactor proportional to the time spent by the rivulet in such a configuration, which goes as $\sim \lambda/v_{\theta}$, therefore
 $\Delta h(t)/\Delta h_0 \sim \alpha (\lambda/v_{\theta})e^{\alpha t}$. 
-The rivulet life-time can be seen as the rupture time of the structure along its axis, hence such that $\Delta h (\tau_{\text{riv}}) \sim h_0$ \cite{PhysRevE.104.034801}, which yields
+The rivulet life-time can be seen as the rupture time of the structure along its axis, hence such that $\Delta h (\tau_{\text{riv}}) \sim h_0$~\cite{PhysRevE.104.034801}, which yields
 \begin{equation}\label{eq:rivlt}
@@
 This logarithmic dependence is indeed observed in the numerical data as shown in Fig.~\ref{fig:stab_ligs_lam2}.\\
-\textcolor{red}{We would like to envisage a possible realization of a 
+\REV{We envisage a possible realization of a 
 dewetting experiment on a switchable substrate of the type modeled by the spatio-temporal 
@@
 An ideal candidate could be a digital multimirror device (DMD).
-This technology has effectively been used for thin film experiments and additive manufacturing~\cite{doi:10.1021/jp301092y, doi:10.1126/science.aax8760}, it allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$)
+This technology was effectively used for thin film experiments and additive manufacturing~\cite{doi:10.1021/jp301092y, doi:10.1126/science.aax8760}.
+It allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$)
 with spatial resolution of $\approx 10 \times 10 \, \mu \text{m}^2$ (the size of a pixel).
-Considering as a reference, for instance, the system studied in ~\cite{becker2003complex,PhysRevLett.99.114503}, namely a $\sim 4 \, \text{nm}$ thick film 
+Considering as a reference, for instance, the system studied in~\cite{becker2003complex,PhysRevLett.99.114503}, namely a $\sim 4 \, \text{nm}$ thick film 
 of polystyrene deposited on an oxidized silicon wafer,  
@@
 well within the rivulet regime ($\Gamma > 1$).
-We note, also, that both $U_{\Theta}$ and $v_{\theta}$ can be widely modulated: the former, playing either with the (temperature and molecular weight dependent) viscosity
-or tailoring the substrate to making it more hydrophobic (i.e. increasing the contact angle); the latter, tuning
+Also, both $U_{\Theta}$ and $v_{\theta}$ can be widely modulated: the former by varying the (temperature and molecular weight dependent) viscosity
+or tailoring the substrate to make it more hydrophobic (i.e. increasing the contact angle); the latter, by tuning
 the spatial resolution and frame rate of the DMD. Thus, we expect the range of achievable $\Gamma$'s to 
@@
 \noindent {\it Conclusions.}
-We have presented results, from numerical simulations and theoretical analysis, on the dewetting of thin liquid films on a switchable substrate, modelled with a space and time periodically varying contact angle in the thin-film equation.
-Studying how the stability of the film depends on the underlying static pattern, we found that the rupture times grow linearly with the pattern wavelength, for short wavelengths, and quadratically in the long wavelength limit.
-In the time-dependent case, the rupture times are in general longer, indicating an induced greater film stability, and, while the quadratic growth is preserved at long wavelengths, a plateauing behaviour is observed as the wavelength decreases.
-A theoretical explanation has been provided for all these various regimes.
+We presented numerical simulations and a theoretical analysis on the dewetting of thin liquid films on a switchable substrate, modelled with a space and time periodically varying contact angle in the thin-film equation.
+Studying how the film stability depends on the underlying static pattern, we found that the rupture times grow linearly with the pattern wavelength, for short wavelengths, and quadratically in the long wavelength limit.
+In the time-dependent case, the rupture times are generally longer, indicating an induced greater film stability, and, while the quadratic growth is preserved at long wavelengths, a plateauing behavior was observed as the wavelength decreases.
+A theoretical explanation was provided for all these regimes.
 Furthermore, we showed that, at increasing the wettability wave speed, a transition occurs in the dewetting morphology from a multi-droplet to a metastable multi-rivulet state. 
-A dimensionless parameter, $\Gamma$, controlling the transition has been identified in the ratio of the pattern speed and the typical film retraction speed over the substrate. 
-The rivulets life-time itself grows with the pattern speed, displaying a logarithmic dependence that has been captured by means of phenomenological arguments.
+A dimensionless parameter, $\Gamma$, controlling the transition was identified in the ratio of the pattern speed and the typical film retraction speed. 
+The rivulets life-time itself grows with the pattern speed, displaying a logarithmic dependence that was captured by means of phenomenological arguments.
 On a broader perspective, our work suggests that switchable substrates offer a new avenue to control thin film dewetting, with obviously relevant implications, for instance, for open microfluidic devices, and paves the way to future studies in this direction, exploiting more complex and dedicated space-time dependencies.
@@
 \noindent {\it Acknowledgements}.
-SZ and JH acknowledge financial support from the Deutsche Forschungsgemeinschaft (DFG) within the priority program SPP2171 ``Dynamic Wetting of Flexible, Adaptive, and Switchable Substrates'', project HA-4382/11 and Project-ID 431791331—CRC1452 ``Catalysis at Liquid Interfaces''.
-SZ acknowledges financial support from the Independent Research Fund Denmark through a DFF Sapere Aude Research Leader grant (grant number 9063-00018B).
+We acknowledge financial support from the German Research Foundation (DFG) (priority program SPP2171 / project HA-4382/11 and Project-ID 431791331—CRC1452), and from the Independent Research Fund Denmark (grant 9063-00018B).
 

### 2023-07-10 Reply_RefereeC.tex
More excitement

@@
 Here, we provide a new approach to control the dewetting morphology by means of switchable substrates which might enable new deposition techniques allowing for unprecedented control.
-Besides the possible applicatio impact, the problem is of fundamental interest in Fluid Dynamics (which represents an important part of the 
+Besides the possible application impact, the problem is of fundamental interest in Fluid Dynamics (which represents an important part of the 
 PRL readership).

### 2023-07-10 Reply_RefereeD.tex
More excitement

\documentclass[12pt,english]{article}

\usepackage{natbib}

\usepackage{graphics,graphicx,dcolumn,bm,fleqn,epic,eepic,float}
\usepackage{amssymb,amsmath,multirow,rotate,rotating,color}
\usepackage[utf8]{inputenc}

\usepackage[english]{babel}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\tikzset{fontscale/.style = {font=\relsize{#1}}}
\usetikzlibrary{calc}
\makeatother

\newcommand{\figref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\eqnref}[1]{Eq.~(\ref{eq:#1})}
\newcommand{\ts}{\textsuperscript}

\definecolor{tuered}{RGB}{214,0,74}

\newcommand{\todo}[1]{\textbf{\textcolor{tuered}{ TODO: #1}}}

\newcommand{\vectornorm}[1]{\left|\left|#1\right|\right|}

\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\uvec}[1]{\hat{\vec{#1}}}
\newcommand{\tensor}[1]{\mathbf{#1}}

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}

\newcommand{\JH}[1]{\textcolor{blue}{JH: #1}}
\begin{document}

% The referee PDF has been generated from r239 -- 'make referee' command.

\section*{Reply to Referee D}
We thank the new Referee for his/her report on our manuscript. 
We are glad to read that he/she acknowledges the systematic approach we have taken and are happy to see that he/she finds our results interesting. 
We understand the concerns raised by the referee and wonder why the recommendation is negative.
Hereafter, we provide answers to the various questions/comments raised. 
All changes in the revised version are reported in red color.

\begin{itemize}

\item[ \textbf{\underline{Comment 1.}}]
{ 
It seems impossible to construct the patterns described by Equation (3) in a real experiment, making the current research less attractive.
}

\item[ \textbf{{Answer}}]
{
We are very surprised by this comment and do not agree with the assessment of the referee. 
% For static patterns it is well known how to pattern substrates in a precise and controlled way using, e.g. lithography. 
It is true that the novelty of the here presented work is strongly coupled to both the temporal and spatial component of the wettability and as such to equation (3).
We are certain that switchable substrates offer this kind of control, and for photoswitchable materials spiropyran should be a perfect candidate~\cite{keyvanradSpiropyranbasedAdvancedPhotoswitchable2022}.

Not only do we present our numerical and analytical results in this letter, but we also supply a potential realization for a lab experiment. 
We provide a schematic setup using a photoswitchable substrate and a digital multimirror device. 
As we are well aware that this is not sufficient for an experiment we further identify the relevant length and time scales for both the film and the apparatus.
% Hoping that experimentalist confirm transition we supply a way to construct pattern Eq. (3).
%We therefore have a hard time understanding why it seems impossible to the referee to construct the pattern equation %(3), as he/she does not give us a single increment of information in his/her comment.
}

\item[ \textbf{\underline{Comment 2.}}] 
{
This work is completed based on the lubrication equation for the thin film flows.
From theoretical/mathematical aspects, the novelty of this work, in my opinion, is not strong enough. 
This lubrication model employed here was proposed about 40 years ago and has been thoroughly investigated~\cite{RevModPhys.69.931, RevModPhys.81.1131}. 
There are no improvements or modifications to the model in this work. 
Additionally, the model for the rupture time comes from a simple scaling analysis, which is not very convincing when considering other classical mathematical approaches, such as linear instability analysis, similarity transformation, etc.
}

\item[ \textbf{{Answer}}]
{
We find it puzzling that the referee arrives at the conclusion that working on a theory which is 40 years old does not allow for the novelty attributed to PRL.
A good counter example is: M. Wilczek et al,
"Sliding drops – ensemble statistics from single drop bifurcations", {\it Phys. Rev. Lett.} {\bf 119}, 204501 (2017), in which Wilczek et al. use the thin film equation on an inclined plane.
Similar arguments can be constructed about the Navier-Stokes equation which is 180 years old, still there is a lot of research dedicated towards, e.g., turbulence.

We agree with the referee that from a purely mathematical point of view lubrication theory is untouched.
There is no reason or advantage to add artificial complexity to a well working theory which has been developed for this singular purpose.
However, when it comes to dynamical effects and non-linear behaviour of thin films, the topic this letter revolves around, mathematics falls short and only numerical simulation or experiments offer deeper insights.
To this end we use a novel lattice Boltzmann method to solve the thin film equation at time scales where initial small perturbations have grown to similar sizes as $h_0$.
It is well known that the rupture time marks the end of the linear regime and linear instability theory as suggested by the referee, which has been used in our previous work~\cite{PhysRevE.104.034801}. 
However, it is simply not able to explain our observations, as $\tau_r$ is dependent on $\Gamma$. 
Observing the morphological transition during dewetting from a droplet to a rivulet state is by no means covered by linear analysis. 
It is a dynamic competition of relevant velocities which to the best knowledge of the authors has not been discussed or shown before.
}

\item[ \textbf{\underline{Comment 3.}}]
{
The mobility function in Equation (1) contains the slip length $\delta$. 
But there is no further information in this work. 
Is the slip length of a patterned substrate constant? 
If so, it is quite unreasonable because a hydrophobic substrate is usually more slippery than a hydrophilic one. 
Moreover, previous works had showed that the slip is crucial to the interface dynamics~\cite{zhang_sprittles_lockerby_2021, zhao_zhang_si_2023}, which should be taken into account.
}

\item[ \textbf{{Answer}}]
{
We thank the Referee for this remark. It is of course true that the slip length 
should grow with the contact angle, on the other hand it is also well known that for low contact angle (actually up to $\sim 100^{\circ}$) such dependence is relatively weak, 
as it is expected to follow the relation $\lambda \propto (1+\cos \theta)^{-2}$
(see Huang et al, {\it Phys. Rev. Lett.} {\bf 101}, 226101 (2008)). The ratio 
of the slip lengths on patches of contanct angles $10^{\circ}$ and $30^{\circ}$
(which is the case of our simulations), for instance, would then be $\approx 0.9$,
such that the slip length should have been varied between $0.95$ and $1.05$ lbu. We deemed, then, reasonable to consider it approximately constant. 
To provide an idea, in reference~\cite{zhang_sprittles_lockerby_2021}, for instance, effects were appreciated 
when slip lengths were two to five times larger than each other. 
}

\item[ \textbf{\underline{Comment 4.}}]
{ 
According to the fluid parameters displayed in the supplemental material, the initial thickness of the film is about 4 nm. 
Thermal fluctuations had been shown to play a key role at this scale~\cite{PhysRevLett.99.114503, PhysRevLett.126.228004}, which might affect the dynamics seriously.
}

\item[ \textbf{{Answer}}]
{
We do not say that the initial film thickness is $4 \, \text{nm}$, we just considered the experiments of Becker et al (where the initial film thickness is $\sim 4 \, \text{nm}$) as a reference, since they provide experimental values of quantities (such as the Hamaker constant) that enable the comparison 
with the numerics. The effect of thermal fluctuations, though, is known to essentially show up as 
a speed up of the growth of the instability, but it does not affect particularly the dewetting morphology (as shown in Becker et al, {\it Nat. Mater.} {\bf 2}, 59--63, who made direct comparison 
of the numerical simulation of a lubrication model with experiments).
}

\end{itemize}

\bibliographystyle{abbrv}
\bibliography{Ref}

\end{document}

### 2023-07-10 main_exciting.tex
More excitement

--- 2023-04-21 main_shortened.tex
@@
 Here, we report a computational study of the dynamics of an unstable thin liquid film deposited on a switchable substrate, modeled with a space and time varying contact angle.
-With a static pattern, all fluid is drained into droplets located around contact angle minima, whereas for a sufficiently large rate of wettability variation a state consisting of metastable rivulets is observed.
+With a static pattern, all fluid is drained into droplets located around contact angle minima, whereas for a sufficiently large rate of wettability variation a \REV{novel topological transition as such the occurence of a metastable rivulet state is observed}.
 A criterion discriminating whether rivulets occur or not is identified in terms of a single dimensionless parameter.

### 2023-11-14 Reply_Referees-2nd.tex
Work on references and some minor editing.

--- 2023-07-10 Reply_RefereeD.tex
@@
 
+\section*{Reply to 2nd report of Referee A}
+\begin{itemize}  
+\item[ \textbf{\underline{Comment 1.}}]
+{ 
+With the response given to me by the authors, I think that the
+research is probably valid. I still fail to see why I should be very
+excited about the findings, so I think that the manuscript would be
+better suited for a more specialized journal, like Physical Review E.
+The reason for this recommendation is that I can see what the
+patterned substrate does, in the sense that the wavelength competes
+with that of dewetting. That droplets can mode (move?) in a wettability
+gradient has also been clear for a long time. I am wondering what is
+there to be gained by combining these ideas, and I find the rationale
+of the manuscript to be not extremely compelling. In my view, this
+would be needed for PRL. I am sorry that I cannot be more positive.
+}
+
+\item[ \textbf{{Answer}}]
+{
+We thank the Referee for taking his/her time to read our 
+manuscript once more and his/her subjective opinion. It is good to read 
+that he/she thinks that our research is valid and that he/she finds the 
+manuscript compelling to some extent.
+
+We would like to stress once more that 
+the main focus of the paper is not dewetting or droplet movement in a {\it stationary} wettability gradient, but dewetting 
+morphologies over a {\it switchable substrate}, i.e. whose wettability pattern 
+is {\it time-dependent}, whence the title 'Controlling dewetting morphologies of 
+thin liquid films by switchable substrates'.
+
+In this context, we have discovered a new morphological transition which we have fully characterized 
+and theoretically interpreted in terms of the control parameter $\Gamma$.
+This said, even in the static pattern case, although the dewetting problem was previously studied, we 
+reported for the first time, together with a theoretical explanation, the absolutely non-trivial dependence of the rupture times on the pattern wavelength.
+Let us remark again that a precise control of the liquid-vapour interface during a dynamic process is a highly desired feature for many applications. We did show that the liquid-vapour 
+interface even after dewetting can be increased significantly due to
+the discussed morphological transition.
+
+More in general wetting, dewetting and the dynamics of droplets on complex substrates 
+are state-of-the-art topics of interest to a broad audience, both in experimental and 
+theoretical fluid physics. 
+We also believe that a deeper understanding of the dynamics of thin liquid films on switchable 
+substrates, which this paper contributes to, is essential to push forward the current reach of microfluidic technology.
+}
+\end{itemize}
+
+\section*{Reply to 2nd report of Referee B}
+\begin{itemize}  
+\item[ \textbf{\underline{Comment 1.}}]
+{ 
+The authors have responded convincingly to my concerns and addressed
+them sufficiently in the revised manuscript. I have also read the
+responses to the other reports and I feel that the authors have
+addressed these recommendations adequately. In view of this, and my
+original favorable assessment, I recommend this manuscript for
+publication.
+
+I have just two remarks that are related to concern 5 in my original
+report:
+
+1. Most importantly, the authors have added footnote 26, which I find
+to be unnecessary, but notice that the curvature has the denominator
+to the power of 3/2.
+
+2. The film thickness-dependent surface tension referred to in my
+report is not related to the trivial small gradient approximation, and
+appears whether the slope is small or large. It is a second order
+effect which decays according to d Pi(h) / dh, and could be relevant
+for thin films and helpful to address the contact line problem, but
+not likely to change the results qualitatively.
+}
+
+\item[ \textbf{{Answer}}]
+{
+We thank the Referee for clarifying his/her question and we apologize for 
+our misunderstanding (we have removed the footnote and the comment on 
+the small gradient expansion of the curvature). 
+We agree that, when thermal fluctuations are relevant,
+it is important to consider the renormalized (i.e., height dependent) surface tension,
+especially when it comes to the initial growth of the instability. It is, on the other 
+hand, reasonable to expect, as the Referee remarks, that in terms of the dewetting 
+morphology (and its relation to the time-dependence of the substrate) the phenomenology
+should not be particularly affected, at least qualitatively.
+}
+\end{itemize}
+
 \section*{Reply to Referee D}
@@
 There is no reason or advantage to add artificial complexity to a well working theory which has been developed for this singular purpose.
-However, when it comes to dynamical effects and non-linear behaviour of thin films, the topic this letter revolves around, mathematics falls short and only numerical simulation or experiments offer deeper insights.
+However, when it comes to dynamical effects and non-linear behaviour of thin films, the topic this letter revolves around, mathematics falls short and only numerical simulation or experiments offer deeper insights, which we also outline in the first paragraph of the introduction:
+
+``\textit{While a consistent body of theoretical/computational work was devoted to processes on static heterogeneous substrates, the time dependent case is still almost unexplored, with few exceptions focusing on single droplet spreading and sliding~\cite{GrawitterStark1,GrawitterStark2,ThieleHartmann} or limited to analysing the linear regime~\cite{suman2006dynamics}.}''
+
 To this end we use a novel lattice Boltzmann method to solve the thin film equation at time scales where initial small perturbations have grown to similar sizes as $h_0$.

### 2023-11-14 main_exciting.tex
Work on references and some minor editing.

--- 2023-04-21 main_shortened.tex
@@
-\documentclass[twocolumn,amsmath,amssymb,showpacs,prl,superscriptaddress]{revtex4-1} %,preprint,jcp,nofootinbib
+\documentclass[amsmath,amssymb,showpacs,prf,superscriptaddress, longbibliography]{revtex4-1} %,preprint,jcp,nofootinbib
+%\documentclass[twocolumn,amsmath,amssymb,showpacs,prl,superscriptaddress]{revtex4-1} %,preprint,jcp,nofootinbib
 
@@
 \usepackage{xcolor}
+\usepackage[normalem]{ulem}
 \usepackage{physics}
@@
  \noindent Switchable and adaptive substrates emerged as valuable tools for controlling wetting and actuation of droplet motion. 
-Here, we report a computational study of the dynamics of an unstable thin liquid film deposited on a switchable substrate, modeled with a space and time varying contact angle.
-With a static pattern, all fluid is drained into droplets located around contact angle minima, whereas for a sufficiently large rate of wettability variation a state consisting of metastable rivulets is observed.
+ Here, we report a computational study of the dynamics of an unstable thin liquid film deposited on a switchable substrate, modelled with a space and time varying contact angle.
+ \REV{For a sufficiently large rate of wettability variation a novel topological transition appears. 
+ Instead of breakup into droplets, as expected for a substrate with multiple wetting minima, a metastable rivulet state emerges}. 
  A criterion discriminating whether rivulets occur or not is identified in terms of a single dimensionless parameter.
-Finally, we show and explain theoretically how the film rupture times, droplet shape and rivulet life time depend on the pattern wavelength and speed.  
+ Finally, we show and \REV{derive} theoretically how the film rupture times, droplet shape and rivulet life time depend on the pattern wavelength and speed.
 \end{abstract}
@@
 \noindent {\it Introduction}.
-Wet surfaces and droplets are part of our every-day experience and of numerous industrial processes including coating, tribology, painting and printing, to name but a few~\cite{gross1980fluid,szeri2010fluid,doi:10.1146/annurev.fluid.31.1.347,DASILVASOBRINHO19991204,singh2010inkjet,Wijshoff2010}.
-Moreover, the continuously growing interest for lab-on-a-chip devices~\cite{C6LC00387G,Focke} as well as for printable electronics or printable photovoltaics~\cite{Luechinger_2008,RH22}, whose efficiency relies crucially on a precise control of material deposition upon (de-)wetting of liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e.
-it is not inert but responds dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{ButtEtAl_Langmuir2018}.
-Several realizations of switchable and adaptive substrates have been proposed~\cite{XiCSR2010}, involving smart materials such as polymer brushes~\cite{CohenStuartEtAl_NatMat2010}, thermal-responsive hydrogels~\cite{ChenEtAl_SM2010}, light-responsive molecules and microstructures~\cite{IchimuraEtAl_Science2000} or processes such as electrowetting~\cite{MugeleEtAl_JPCM2005}.
-Probably the simplest, yet non-trivial, modelling of a switchable substrate can be realized by a space and time dependent wettability pattern~\cite{GrawitterStark1}. 
-While a consistent body of theoretical/computational work was devoted to processes on static heterogeneous substrates, the time dependent case is still almost unexplored, with few exceptions focusing on single droplet spreading and sliding~\cite{GrawitterStark1,GrawitterStark2,ThieleHartmann} or limited to analysing the linear regime~\cite{suman2006dynamics}.
+Wet surfaces and droplets are part of our every-day experience and of numerous industrial processes including coating, tribology, painting and printing, to name but a few~\cite{grossFluidFilmLubrication1980a,szeriFluidFilmLubrication2010,quereFluidCoatingFiber1999,dasilvasobrinhoStudyDefectsUltrathin1999,singhInkjetPrintingProcess2010,wijshoffDynamicsPiezoInkjet2010}.
+Moreover, the continuously growing interest for lab-on-a-chip devices~\cite{samieiReviewDigitalMicrofluidics2016,fockeLabonaFoilMicrofluidicsThin2010} as well as for printable electronics or printable photovoltaics~\cite{luechingerGraphenestabilizedCopperNanoparticles2008,ronsinFormationCrystallineBulk2022}, whose efficiency relies crucially on a precise control of material deposition upon (de-)wetting of liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e.
+it is not inert but responds dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{buttAdaptiveWettingAdaptation2018}.
+Several realizations of switchable and adaptive substrates have been proposed~\cite{xinReversiblySwitchableWettability2010}, involving smart materials such as polymer brushes~\cite{stuartEmergingApplicationsStimuliresponsive2010}, thermal-responsive hydrogels~\cite{chenThermalresponsiveHydrogelSurface2010}, light-responsive molecules and microstructures~\cite{ichimuraLightDrivenMotionLiquids2000} or processes such as electrowetting~\cite{mugeleElectrowettingConvenientWay2005}.
+\REV{Modelling the coupled problem of a thin film on a switchable substrate and as such the dynamics of the substarte as well as the film is far from being trivial but can be realized by a space and time dependent wettability pattern~\cite{grawitterSteeringDropletsSubstrates2021}}.
+While a consistent body of theoretical/computational work was devoted to processes on static heterogeneous substrates, the time dependent case is still almost unexplored, with few exceptions focusing on single droplet spreading and sliding~\cite{grawitterSteeringDropletsSubstrates2021,grawitterDropletsSubstratesOscillating2021,thieleGradientDynamicsModel2020} or limited to analysing the linear regime~\cite{sumanDynamicsThinLiquid2006}.
 
-Here, we study by means of numerical simulations the full dewetting dynamics of a thin liquid film deposited on a substrate with a time varying wettability pattern, from rupture to the long time morphology.
+Here, we study by means of numerical simulations the full dewetting dynamics of a thin liquid film deposited on a substrate with a time varying wettability pattern, starting from \REV{a slightly perturbed film to film} rupture \REV{and} long time morphology.
 We identify two regimes where the rupture times grow with the pattern wavelength either linearly (on a static pattern) or attain a constant value (in the time dependent case), for short wavelengths, and approach a quadratic law as the wavelength increases.
 These observations are then explained theoretically. 
-We show that, by tuning the rate of change (the ``speed'') of the underlying pattern, one can control, to some extent, the dewetting morphology.
-In particular, for large enough pattern speeds, we detect a metastable state, where the film retracts into metastable rivulets, eventually breaking up into multiple droplets.
+\REV{We show that, by tuning the rate of change of the underlying pattern, one can control the dewetting morphology, which is often desired in microfluidic applications.} 
+In particular, for large enough pattern speeds, we detect a \REV{novel} state, where the film retracts into metastable rivulets, eventually breaking up into multiple droplets.
 We introduce a control parameter to discriminate whether rivulets or just droplets (as in the static situation)
@@
 \\
-\noindent {\it Method.} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{RevModPhys.69.931,RevModPhys.81.1131}
+\noindent {\it Method.} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{oronLongscaleEvolutionThin1997,crasterDynamicsStabilityThin2009}
 \begin{equation}\label{eq:thinfilm}
@@
 \end{equation}
-by means of a recently developed lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801, Zitz2022}.
+by means of a recently developed lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,zitzLatticeBoltzmannSimulations2021, zitzSwalbeJlLattice2022a}.
 Eq.~(\ref{eq:thinfilm}) describes, in a lubrication approximation spirit, the evolution of the height field (film thickness) $h(\mathbf{x},t)$, denoting the location of the liquid/air interface.
@@
 (for $\delta \rightarrow 0$ it reduces to the no-slip form $h^3/(3\mu)$). Here, $\mu$ is the fluid dynamic viscosity.
-The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$ \REV{($\gamma$ is the surface tension)}~\footnote{\REV{The Laplace pressure is the product of the surface tension
-and the local curvature. Therefore, its full expression reads $p_L = -\gamma \nabla^2 h/\sqrt{1+|\nabla h|^2}$. 
-In the lubrication regime, whereby $|\nabla h| \ll 1$, the denominator can be safely neglected 
-and the pressure reduces to the form involving only the Laplacian of the thickness 
-field}~\cite{Benet2014,juanes2018}}.
+The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$ ($\gamma$ is the surface tension)
 The disjoining pressure $\Pi$ can be seen as (minus) the derivative, with respect to the film thickness, of an effective interfacial potential.
-As such, it contains the information on the liquid/solid and solid/gas interactions and, hence, on the wettability, which is parameterized in terms of the contact angle $\theta$~\cite{RevModPhys.81.739, SCHWARTZ1998173}.
+As such, it contains the information on the liquid/solid and solid/gas interactions and, hence, on the wettability, which is parameterized in terms of the contact angle $\theta$~\cite{bonnWettingSpreading2009, schwartzSimulationDropletMotion1998}.
 The expression adopted for $\Pi$ is
 \begin{equation}\label{eq:disjoinpressure}
-\REV{\Pi(h,\theta) = \frac{2\gamma}{h_{\ast}}(1-\cos(\theta(\mathbf{x},t)))
-  f\left(\frac{h}{h_{\ast}}\right),}
+\Pi(h,\theta) = \frac{2\gamma}{h_{\ast}}(1-\cos(\theta(\mathbf{x},t)))
+  f\left(\frac{h}{h_{\ast}}\right),
 \end{equation}
-\REV{where $f(\xi)=\xi^{-3} - \xi^{-2}$}. $h_{\ast}$ is the height at which $\Pi$ vanishes and sets the precursor layer thickness~\cite{SuppMat}.
+where $f(\xi)=\xi^{-3} - \xi^{-2}$. $h_{\ast}$ is the height at which $\Pi$ vanishes and sets the precursor layer thickness~\cite{SuppMat}.
 The time variation of the patterned substrate enters the model through the disjoining pressure, by making the contact angle space and time dependent, i.e.~$\theta = \theta(\mathbf{x},t)$.
-\REV{We decorate the substrate with a checkerboard pattern, a common choice that generalizes the broken homogeneity of the stripes to two 
-directions~\cite{Jalali2018,Nagayama2020,Das2020}.}
+We decorate the substrate with a checkerboard pattern, a common choice that generalizes the broken homogeneity of the stripes to two 
+directions~\cite{jalaliFabricationCharacterizationScalable2018,nagayamaIntermediateWettingState2020,dasSurfacedirectedSpinodalDecomposition2020}.
 In particular, we employ the sinusoidal form
@@
 where $q_{\theta} = 2\pi/\lambda$, i.e.~the pattern evolves in time as a plane wave.
-% \REV{A possible laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}. 
-% The external stimulus, our moving pattern, can be realized, e.g., by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.}
 We fix the velocity direction to one diagonal, namely $\mathbf{v}_{\theta} = (v_{\theta x},v_{\theta y}) = v_{\theta}(1/\sqrt{2},-1/\sqrt{2})$ (we will return later on the importance of this choice), and we set $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$~\footnote{Since a typical velocity is such that $v_{\theta} \Delta t \ll \Delta x$ (in one time step $\Delta t$ the wave would travel a distance much smaller than a lattice spacing $\Delta x$),
 the time update needs to be interpreted in an integer part sense, that is the pattern is shifted by $\Delta x$ every $1/v_{\theta x}$ $\Delta t$ (equivalently in $y$-direction).}.
-\REV{Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film~\cite{Mecke_2005}.
-On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads
-$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$~\cite{Mecke_2005,PhysRevE.100.023108}. In our patterned case, we define 
+Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film~\cite{meckeThermalFluctuationsThin2005}.
+On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads $(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$~\cite{meckeThermalFluctuationsThin2005,zhangMolecularSimulationThin2019}.
+In our patterned case, we define 
   $q_0^2=h_{\ast}^{-2}(1-\cos\theta_0)f^{\prime}(h_0/h_{\ast})$. Correspondingly, we choose as a velocity scale $v_0 = \lambda_s/t_0$, where
-$\lambda_s = 2\pi/q_0$.}
+$\lambda_s = 2\pi/q_0$.
 \begin{figure}
@@
 Fig.~\ref{fig:handtheta} shows $h(\mathbf{x},t)$ (droplets) and $\theta(\mathbf{x})$ (color coded) for $v_{\theta} = 0$ (i.e.
-the static case) and \REV{$\lambda = 256 h_0$}~\footnote{For a better visualization double the domain length $L$ and periodically continue the image.}, in late stages of dewetting.
+the static case) and $\lambda = 256 h_0$~\footnote{For a better visualization double the domain length $L$ and periodically continue the image.}, in late stages of dewetting.
 As expected, droplets form in regions of small contact angles (blue) while the regions of high contact angles (yellow) dewet.
@@
 In Fig.~\ref{fig:model_rt} we report the rupture times as a function of the wavelength, for stationary ($v_{\theta}=0$) and time-dependent ($v_{\theta}=20 v_0$) patterns. 
-It is conveyed that, overall, rupture occurs earlier on the static substrate, suggesting that the time variation tends to stabilize the film, in agreement with linear stability analysis results~\cite{suman2006dynamics}.
+It is conveyed that, overall, rupture occurs earlier on the static substrate, suggesting that the time variation tends to stabilize the film, in agreement with linear stability analysis results~\cite{sumanDynamicsThinLiquid2006}.
 We observe that $\tau_r$ grows linearly with $\lambda$ for short wavelengths and quadratically for longer $\lambda$.
@@
 Therefore, since the characteristic time $t_{\theta}$ can be estimated dimensionally as $t_{\theta} \sim \delta h/\dot{(\delta h)}$, the rupture time should go as
-\begin{multline}\label{eq:taur_l2}
-    \tau_r \sim t_{\theta} \sim  \delta h/\dot{(\delta h)} \propto \frac{3\mu}{h_0^3}(\partial_x^2 (\partial_h\Pi (h_0)))^{-1} \sim \\
-    t_0 \left(\frac{q_{\theta}}{q_0}\right)^{-2} \propto t_0 q_0^2 \lambda^2.
-\end{multline}
+\begin{equation}\label{eq:taur_l2}
+    \tau_r \sim t_{\theta} \sim  \delta h/\dot{(\delta h)} \propto \frac{3\mu}{h_0^3}(\partial_x^2 (\partial_h\Pi (h_0)))^{-1} \sim t_0 \left(\frac{q_{\theta}}{q_0}\right)^{-2} \propto t_0 q_0^2 \lambda^2.
+\end{equation}
 Conversely, for fast growths ($t_{\theta} \ll t_R$), retraction dominates and fixes the time scale, $\tau_r \sim t_R$. 
 The latter is related to the time the liquid takes to flow out of regions of high contact angle, whose size is $\sim \lambda$. Hence we have 
- %\tau_r \sim \tau_R \propto \frac{1}{U_{\Theta}}\lambda,
 \begin{equation}\label{eq:taur_l1}
@@
 \end{equation}
-where $U_{\Theta}$ is the retraction speed $U_{\Theta} = \frac{\gamma \Theta^3}{9\mu}$~\cite{Edwards2016}, with $\Theta = \max_{\mathbf{x}}\theta(\mathbf{x})$.
+where $U_{\Theta}$ is the retraction speed $U_{\Theta} = \frac{\gamma \Theta^3}{9\mu}$~\cite{edwardsNotSpreadingReverse2016a}, with $\Theta = \max_{\mathbf{x}}\theta(\mathbf{x})$.
 \begin{figure}
@@
     with $v_{\theta}= 0$ and $\lambda= 512 h_0$ (\textcolor{jlblue}{$\bullet$}), $\lambda=256 h_0$ (\textcolor{jlorange}{$\blacksquare$}) and $\lambda=170 h_0$ (\textcolor{jlgreen}{$\star$}). 
-    % The dashed lines indicate the geometrically expected values of the droplet height $h_d$, assuming monodispersity and a perfect spherical cap shape. 
     INSET: Number of droplets, $N(t)$, as a function of time. The three horizontal dashed lines indicate the number of minima of Eq.~(\ref{eq:sinetheta}),
-      which is $2\left(\frac{L}{\lambda}\right)^2$. The napshots depict the stationary droplet states as grey-scale images
+      which is $2\left(\frac{L}{\lambda}\right)^2$. The snapshots depict the stationary droplet states as grey-scale images
       of the film thickness field, $h(\mathbf{x},t)$.
@@
 time~\footnote{A droplet is identified by the set (``cluster'') of points, in the plane, constituting each of the connected components of the set $\{\mathbf{x} \in [0,L]^2 | h(\mathbf{x},t) \geq h_{\ast}$\}.
-% the clusters are determined by means of a Hoshen-Kopelman algorithm~\cite{HK}.
 },
@@
 Notice that the number of droplets converges faster for smaller pattern wavelengths, in line with the observation reported and justified in the previous section that the characteristic dewetting time decreases with the wavelength.\\
-\REV{In the main panel of the figure, the height fluctuations $\Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\}-\min_{\mathbf{x}}\{h(\mathbf{x},t)\}$
-grow in time until film rupture and then settle to a constant value. This represents a measure of the mean droplet height $h_d$ (since
-droplets are essentially monodisperse), decreasing with the pattern wavelength (as expected, due to a decreasing droplet volume,
-$V_d = \frac{h_0 \lambda^2}{2}$).}\\
-%In the steady state, since the droplets are essentially monodisperse, this observable . Assuming that the droplet shape is a spherical cap,
-%it can be estimated as $h_d =\left(\frac{3V_d(1-\cos(\theta_d))}{\pi(2+\cos(\theta_d))}\right)^{1/3}$, where  is the droplet volume and
-%$\theta_d$ is the local contact angle (i.e. the value at the contact line). This, in turn, depends on $h_d$,
-%through Eq.~(\ref{eq:sinetheta}), thus making the one above an implicit equation to be solved numerically for $h_d = h_d(\lambda)$. 
-%$\Delta h(t)$ is reported in Fig.~\ref{fig:clusters_v0_sine} for three different wavelengths, 
-%$\lambda/h_0 = \{170,256,512 \}$, together with the theoretically expected $h_d(\lambda)$ (depicted with lines),
-%showing excellent agreement.\\
+In the main panel of the figure, the height fluctuations $\Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\}-\min_{\mathbf{x}}\{h(\mathbf{x},t)\}$ grow in time until film rupture and then settle to a constant value.
+This represents a measure of the mean droplet height $h_d$ (since droplets are essentially monodisperse), decreasing with the pattern wavelength (as expected, due to a decreasing droplet volume,
+$V_d = \frac{h_0 \lambda^2}{2}$).\\
 \begin{figure}
@@
 For $v_{\theta} = 1.7 v_0$ we still observe the formation of droplets, similarly to the stationary case ($v_{\theta} = 0$). 
-However, these are transported with the contact angle minima, reproducing a somehow similar behaviour recently described in a numerical study of a droplet on a moving wettability step~\cite{GrawitterStark1}.
+However, these are transported with the contact angle minima, reproducing a somehow similar behaviour recently described in a numerical study of a droplet on a moving wettability step~\cite{grawitterSteeringDropletsSubstrates2021}.
 If the pattern speed is further increased, for $v_{\theta} = 17.5 v_0$ we observe the development of rivulet-like structures, aligned with $\mathbf{v}_{\theta}$. 
@@
 In order to better characterize the various morphologies we apply the theory of Minkowski's functionals. 
-In particular, we employ the second order Minkowski structure metric, $q_2$~\cite{doi:10.1063/1.4774084, Schaller2020}, which can be computed from a Voronoi tessellation of the set of discrete points $(x_i, y_i)$ on the 2D lattice, such that the height field lies above a certain threshold\footnote{The expression is: $q_2 = \frac{1}{N}\sum_j \frac{1}{P_j}|\sum_k L^{(j)}_k e^{2i\phi^{(j)}_k}|$, where the inner sum runs over the edges of length $L^{(j)}_k$, of the $j$-th Voronoi cell, whose perimeter is $P_j$, and $\phi^{(j)}_k$ is the polar angle of the normal to the $k$-th edge.
+In particular, we employ the second order Minkowski structure metric, $q_2$~\cite{mickelShortcomingsBondOrientational2013, schallerPapaya22DIrreducible2020}, which can be computed from a Voronoi tessellation of the set of discrete points $(x_i, y_i)$ on the 2D lattice, such that the height field lies above a certain threshold\footnote{The expression is: $q_2 = \frac{1}{N}\sum_j \frac{1}{P_j}|\sum_k L^{(j)}_k e^{2i\phi^{(j)}_k}|$, where the inner sum runs over the edges of length $L^{(j)}_k$, of the $j$-th Voronoi cell, whose perimeter is $P_j$, and $\phi^{(j)}_k$ is the polar angle of the normal to the $k$-th edge.
 The outer sum represents an ensemble average over the $N$ points in the set.}.
@@
 Measuring $q_2$ then enables us to clearly distinguish between the formation of droplets and rivulets: much larger $q_2$ values are attained for the latter type of structure, as we can see in Fig.~\ref{fig:msm_q2}.
-%Our simulations with $\lambda = L/2$ show a clear difference between $v_{\theta} \leq 0.1v_0$ and $v_{\theta}>0.1v_0$  in $q_2$.
 We observe, on the other hand, that such rivulets are metastable and eventually break up into droplets, as indicated by the collapse of $q_2$ at later times. 
@@
     \centering
-    % \includegraphics[width=0.45\textwidth]{Figures/Lig_stab_inset_delta_h.pdf}
     \includegraphics[width=0.4\textwidth]{Figure_5.pdf}
@@
 \end{figure}
-The breakup is the result of a varicose mode of the rivulet~\cite{doi:10.1063/1.3211248, PhysRevE.77.061605}, whose wavelength is $\approx\lambda$, such that only $N_{\infty}/2$ droplets are counted after breakup.
+The breakup is the result of a varicose mode of the rivulet~\cite{diezBreakupFluidRivulets2009a, mechkovStabilityLiquidRidges2008}, whose wavelength is $\approx\lambda$, such that only $N_{\infty}/2$ droplets are counted after breakup.
 These droplets show a peculiar dynamics, characterized by a periodic sequence of spreading and retraction, driven by the pattern, that we dub "pumping state"~\footnote{see movie ligament\_formation\_and\_breakup.mp4 in~\cite{SuppMat}.}.
@@
 We argue that the emergence of rivulets is controlled by the competition of two characteristic velocities: the pattern wave speed, $v_{\theta}$, and the retraction speed, $U_{\Theta}$, introduced in Eq.~(\ref{eq:taur_l1}).
-\REV{If $U_{\Theta}$ is large as compared to $v_{\theta}$, the film retraction is faster than 
-the local contact angle variation and thus droplets form.}
-%the transport due to the contact angle field and thus droplets from. 
+If $U_{\Theta}$ is large as compared to $v_{\theta}$, the film retraction is faster than the local contact angle variation and thus droplets form. 
 However, if $v_{\theta}$ is larger than $U_{\Theta}$, then the retracting film has too little time to form droplets and ends up in the metastable rivulet state. 
-\REV{It appears, therefore, natural to consider the ratio of these two velocities,
-$\Gamma \equiv v_{\theta}/U_{\theta}$, as the discriminating parameter.}
-%Recalling the expression (\ref{eq:t0}) for the reference velocity $v_0$, %we define the parameter 
-%\begin{equation}\label{eq:vel_ratio}
-%    \Gamma = \frac{v_{\theta}}{U_{\Theta}} = \frac{3\lambda h_0^3 %q_0^4}{\Theta^3}\chi 
-%\end{equation}
-%as the ratio of these two velocities,
-%where $\chi \equiv v_{\theta}/v_0$. 
+It appears, therefore, natural to consider the ratio of these two velocities, $\Gamma \equiv v_{\theta}/U_{\theta}$, as the discriminating parameter.
 We see from Fig.~\ref{fig:msm_q2} that indeed rivulets form only for $\Gamma > 1$. 
@@
 $\Delta h(t)/\Delta h_0 \sim \alpha (\lambda/v_{\theta})e^{\alpha t}$. 
-The rivulet life-time can be seen as the rupture time of the structure along its axis, hence such that $\Delta h (\tau_{\text{riv}}) \sim h_0$~\cite{PhysRevE.104.034801}, which yields
+The rivulet life-time can be seen as the rupture time of the structure along its axis, hence such that $\Delta h (\tau_{\text{riv}}) \sim h_0$~\cite{zitzLatticeBoltzmannSimulations2021}, which yields
 \begin{equation}\label{eq:rivlt}
@@
 This logarithmic dependence is indeed observed in the numerical data as shown in Fig.~\ref{fig:stab_ligs_lam2}.\\
-\REV{We envisage a possible realization of a 
-dewetting experiment on a switchable substrate of the type modeled by the spatio-temporal 
-contact angle (\ref{eq:sinetheta}). One may think of a thin liquid film cast on a light responsive substrate~\cite{IchimuraEtAl_Science2000}, under the action of controlled 
-external stimuli (a light emitter). 
+We envisage a possible realization of a dewetting experiment on a switchable substrate of the type modelled by the spatio-temporal contact angle (\ref{eq:sinetheta}). 
+One may think of a thin liquid film cast on a light responsive substrate~\cite{ichimuraLightDrivenMotionLiquids2000}, under the action of controlled external stimuli (a light emitter). 
 An ideal candidate could be a digital multimirror device (DMD).
-This technology was effectively used for thin film experiments and additive manufacturing~\cite{doi:10.1021/jp301092y, doi:10.1126/science.aax8760}.
-It allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$)
-with spatial resolution of $\approx 10 \times 10 \, \mu \text{m}^2$ (the size of a pixel).
-Considering as a reference, for instance, the system studied in~\cite{becker2003complex,PhysRevLett.99.114503}, namely a $\sim 4 \, \text{nm}$ thick film 
-of polystyrene deposited on an oxidized silicon wafer,  
-we evaluate the retraction speed 
-$U_{\Theta} = \frac{\Theta^3 \gamma}{9 \mu}$ to be 
-$U_{\Theta} \approx 10^{-2} \, \mu \text{m}/\text{s}$. 
-The (minimum) {\it pattern speed} can be 
-estimated from the pixel size with frame rate $\sim 1 s^{-1}$ as 
-$v_{\theta} \sim 10 \, \mu \text{m}/\text{s}$, which  would result in $\Gamma \sim 10^3$, i.e.
+This technology was effectively used for thin film experiments and additive manufacturing~\cite{vieyrasalasActiveControlEvaporative2012, sahaScalableSubmicrometerAdditive2019}. 
+It allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$) with spatial resolution of $\approx 10 \times 10 \, \mu \text{m}^2$ (the size of a pixel).
+Considering as a reference, for instance, the system studied in~\cite{beckerComplexDewettingScenarios2003,fetzerThermalNoiseInfluences2007}, namely a $\sim 4 \, \text{nm}$ thick film of polystyrene deposited on an oxidized silicon wafer, we evaluate the retraction speed $U_{\Theta} = \frac{\Theta^3 \gamma}{9 \mu}$ to be $U_{\Theta} \approx 10^{-2} \, \mu \text{m}/\text{s}$.
+The (minimum) {\it pattern speed} can be estimated from the pixel size with frame rate $\sim 1 s^{-1}$ as $v_{\theta} \sim 10 \, \mu \text{m}/\text{s}$, which would result in $\Gamma \sim 10^3$, i.e.
 well within the rivulet regime ($\Gamma > 1$).
-Also, both $U_{\Theta}$ and $v_{\theta}$ can be widely modulated: the former by varying the (temperature and molecular weight dependent) viscosity
-or tailoring the substrate to make it more hydrophobic (i.e. increasing the contact angle); the latter, by tuning
-the spatial resolution and frame rate of the DMD. Thus, we expect the range of achievable $\Gamma$'s to 
-be feasibly extended both to very high ($\Gamma \gg 1$) and very low ($\Gamma \ll 1$) values.}\\
+Also, both $U_{\Theta}$ and $v_{\theta}$ can be widely modulated: the former by varying the (temperature and molecular weight dependent) viscosity or tailoring the substrate to make it more hydrophobic (i.e.
+increasing the contact angle); the latter, by tuning the spatial resolution and frame rate of the DMD.
+Thus, we expect the range of achievable $\Gamma$'s to be feasibly extended both to very high ($\Gamma \gg 1$) and very low ($\Gamma \ll 1$) values.\\
 \noindent {\it Conclusions.}
-We presented numerical simulations and a theoretical analysis on the dewetting of thin liquid films on a switchable substrate, modelled with a space and time periodically varying contact angle in the thin-film equation.
+\REV{We presented numerical simulations and a theoretical analysis on the dewetting of thin liquid films on a switchable substrate.}
 Studying how the film stability depends on the underlying static pattern, we found that the rupture times grow linearly with the pattern wavelength, for short wavelengths, and quadratically in the long wavelength limit.
 In the time-dependent case, the rupture times are generally longer, indicating an induced greater film stability, and, while the quadratic growth is preserved at long wavelengths, a plateauing behavior was observed as the wavelength decreases.
-A theoretical explanation was provided for all these regimes.
 Furthermore, we showed that, at increasing the wettability wave speed, a transition occurs in the dewetting morphology from a multi-droplet to a metastable multi-rivulet state. 
-A dimensionless parameter, $\Gamma$, controlling the transition was identified in the ratio of the pattern speed and the typical film retraction speed. 
+\REV{We find that this surprising morphological transition can be described with a single dimensionless parameter, $\Gamma$.}
+\REV{Considering only the ratio of the pattern speed and the typical film retraction speed.} 
 The rivulets life-time itself grows with the pattern speed, displaying a logarithmic dependence that was captured by means of phenomenological arguments.
@@
 
-\bibliography{Ref}
+\bibliography{betterRef}
 

### 2024-02-15 SuppMat.tex
Resubmission PRFluids

@@
-\documentclass[amsmath,amssymb,showpacs,prl,superscriptaddress,notitlepage]{revtex4-1}
-
-\usepackage{natbib}
+\documentclass[amsmath,amssymb,showpacs,prf,superscriptaddress,notitlepage,longbibliography]{revtex4-1}
+%\documentclass[amsmath,amssymb,showpacs,prl,superscriptaddress,notitlepage]{revtex4-1}
+\bibstyle{apsrev}
 
@@
 The disjoining pressure stems from this interaction. Thermodynamically, it represents minus the derivative of the free energy difference, per unit surface, between 
-the film and the bulk of the same liquid, or, equivalently, the pressure difference inside and outside the film 
-\cite{Deryaguin1940,DeryaguinChuraev1978}.
-In the case of wetting, there can be three types of interfaces: liquid/air, liquid/solid (the substrate) and, 
-if the substrate is only partially wetted, solid/air, thus giving rise to three-phase contact lines. 
+the film and the bulk of the same liquid, or, equivalently, the pressure difference inside and outside the film \cite{Deryaguin1940,DeryaguinChuraev1978}.
+In the case of wetting, there can be three types of interfaces: liquid/air, liquid/solid (the substrate) and, if the substrate is only partially wetted, solid/air, thus giving rise to three-phase contact lines.
 The usually called "surface tension" is the energy cost of forming a liquid/air interface of unit area. 
-The Young's condition \cite{Young1805,DeGennes1985} connects the surface tension with the analogous energies per unit
-area associated with the solid/liquid ($\gamma_{\text{sl}}$) and solid/gas ($\gamma_{\text{sg}}$) interface as
+The Young's condition \cite{youngIIIEssayCohesion1805,degennesWettingStaticsDynamics1985} connects the surface tension with the analogous energies per unit area associated with the solid/liquid ($\gamma_{\text{sl}}$) and solid/gas ($\gamma_{\text{sg}}$) interface as
 \begin{equation}
@@
 \end{equation}
-where $f(\xi)=\xi^{-3} - \xi^{-2}$ and $h_{\ast}$ is the height at which the disjoining pressure vanishes setting the precursor layer thickness, pertains to a general family of expressions of 
-the form 
+where $f(\xi)=\xi^{-3} - \xi^{-2}$ and $h_{\ast}$ is the height at which the disjoining pressure vanishes setting the precursor layer thickness, pertains to a general family of expressions of the form
 \begin{equation}\label{eq:SMDP}
@@
 \end{equation}
-Equations like (\ref{eq:SMDP}) are frequently found in the literature, 
-with various combination of exponents $(n,m)$ ($n>m$), depending
-on the type of interactions among fluid molecules and between fluid and substrate~\cite{SCHWARTZ1998173,Mitlin,Teletzke}. A common choice, for instance, is the pair $(9,3)$, which 
-can be derived from molecular interactions of Lennard-Jones type. Here, the use of the $(3,2)$ pair is motivated mainly by the computational advantage of keeping the characteristic dewetting time 
-shorter. We have run also a single case ($\lambda=256 h_0$ and $\Gamma=1.5$) with $(9,3)$ and observed that 
-the dynamics remains essentially the same, but for a shift in the time scale (see also~\cite{PhysRevE.104.034801}).
+Equations like (\ref{eq:SMDP}) are frequently found in the literature, with various combination of exponents $(n,m)$ ($n>m$), depending on the type of interactions among fluid molecules and between fluid and substrate~\cite{schwartzSimulationDropletMotion1998,mitlinDewettingSolidSurface1993,teletzkeHowLiquidsSpread1987}.
+A common choice, for instance, is the pair $(9,3)$, which can be derived from molecular interactions of Lennard-Jones type. 
+Here, the use of the $(3,2)$ pair is motivated mainly by the computational advantage of keeping the characteristic dewetting time shorter. 
+We have run also a single case ($\lambda=256 h_0$ and $\Gamma=1.5$) with $(9,3)$ and observed that the dynamics remains essentially the same, but for a shift in the time scale (see also~\cite{zitzLatticeBoltzmannSimulations2021}).
 
@@
 
-\noindent To give a flavour of how to compare with actual experimental systems, let us notice
-first that the expression of $t_0$ depends on the disjoining pressure (through its dependence 
- $q_0 = \sqrt{\frac{1}{2\gamma} \Pi^{\prime}(h_0)}$) and contains, therefore, energetic details, namely 
- the Hamaker constant $A_H$.
-Consider, for instance, the problem studied in \cite{becker2003complex}: the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon
-dioxide substrate, where $A_H  = 2.2 \times 10^{-20} \, \text{J}$ and 
-$\gamma = 0.03 \, \text{N}/\text{m}$. Neglecting the short range part of the interface potential
-(such that $\Pi(h)= - \frac{A_H}{6\pi h^3}$)~\cite{Mecke_2005,becker2003complex}, one gets 
-$q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}$.
-The characteristic time of the growth of the instability for these experiments then reads $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}$ (the dynamic viscosity of polystyrene is 
-$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$)~\cite{PhysRevLett.99.114503}. It can be seen that indeed film rupture occurs within time 
-scales of the order of $t_0$~\cite{becker2003complex}, as in our 
-simulation for the largest pattern wavelength, $\lambda = L$ (see also our previous work
-with homogeneous substrates~\cite{PhysRevE.104.034801}).\\
-All our simulations are run on a bi-periodic square domain of size $L \times L$ with 
-$L/h_0 = 512 $. The numerical values in lattice Boltzmann units of the 
-mean film thickness, dynamics viscosity and surface tension are, respectively, $h_0=1$, $\mu=1/6$
-and $\gamma=0.01$.
-To regularize the contact line divergence~\cite{huh1971hydrodynamic}
-we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$, 
-a value which lies within the weak/intermediate slip regime~\cite{peschka2019signatures,fetzer2007quantifying, munch2005lubrication}
-(we have tried also a value four times smaller, without observing any major qualitative change in the results).
-The liquid film is initialized with a height field slightly perturbed around the mean value $h_0$,
-i.e. $h(\mathbf{x},0) = h_0 \left[1 + 0.1 \left(\sin\left(\frac{2\pi x}{L}\right)\sin\left(\frac{2\pi y}{L}\right)\right)\right]$.
-Various wavelengths in the range $\lambda/h_0 \in [64, 512]$, and velocities,
-$v_{\theta}/v_0 \in [0, 200]$, are considered for the wettability pattern (see main text). 
-A few comments on these ranges are noteworthy at this point. The wavelengths are obviously limited from above by the size of the
-computational domain. On the other extreme, too short wavelengths are prevented by
-the energetic cost of accomodating droplets on too
-small patches, since the local contact angle 
-(that increases with decreasing $\lambda$, see also \ref{sec:SMdroplet}) would possibly exceed the largest substrate contact angle,
-such that the patterning itself would loose meaning. 
-It has been, in fact, proven, both theoretically and experimentally, that 
-for the confinement to be effective, the pattern characteristic size should 
-not be much smaller than the spinodal length scale on the less wettable substrate
-\cite{KarguptaJCP2002,KarguptaLangmuir2000,Nisato1999,Karim1998}.
+\noindent To give a flavour of how to compare with actual experimental systems, let us notice first that the expression of $t_0$ depends on the disjoining pressure (through its dependence $q_0 = \sqrt{\frac{1}{2\gamma} \Pi^{\prime}(h_0)}$) and contains, therefore, energetic details, namely the Hamaker constant $A_H$.
+Consider, for instance, the problem studied in \cite{beckerComplexDewettingScenarios2003}: the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon dioxide substrate, where $A_H  = 2.2 \times 10^{-20} \, \text{J}$ and $\gamma = 0.03 \, \text{N}/\text{m}$.
+Neglecting the short range part of the interface potential (such that $\Pi(h)= - \frac{A_H}{6\pi h^3}$)~\cite{meckeThermalFluctuationsThin2005,beckerComplexDewettingScenarios2003}, one gets $q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}$.
+The characteristic time of the growth of the instability for these experiments then reads $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}$ (the dynamic viscosity of polystyrene is $\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$)~\cite{fetzerThermalNoiseInfluences2007}.
+It can be seen that indeed film rupture occurs within time scales of the order of $t_0$~\cite{beckerComplexDewettingScenarios2003}, as in our simulation for the largest pattern wavelength, $\lambda = L$ (see also our previous work with homogeneous substrates~\cite{zitzLatticeBoltzmannSimulations2021}).\\
+All our simulations are run on a bi-periodic square domain of size $L \times L$ with $L/h_0 = 512 $. 
+The numerical values in lattice Boltzmann units of the mean film thickness, dynamics viscosity and surface tension are, respectively, $h_0=1$, $\mu=1/6$ and $\gamma=0.01$.
+To regularize the contact line divergence~\cite{huhHydrodynamicModelSteady1971} we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$, a value which lies within the weak/intermediate slip regime~\cite{peschkaSignaturesSlipDewetting2019,fetzerQuantifyingHydrodynamicSlip2007, munchLubricationModelsSmall2005a} (we have tried also a value four times smaller, without observing any major qualitative change in the results).
+The liquid film is initialized with a height field slightly perturbed around the mean value $h_0$, i.e.
+$h(\mathbf{x},0) = h_0 \left[1 + 0.1 \left(\sin\left(\frac{2\pi x}{L}\right)\sin\left(\frac{2\pi y}{L}\right)\right)\right]$.
+Various wavelengths in the range $\lambda/h_0 \in [64, 512]$, and velocities, $v_{\theta}/v_0 \in [0, 200]$, are considered for the wettability pattern (see main text). 
+A few comments on these ranges are noteworthy at this point. The wavelengths are obviously limited from above by the size of the computational domain. 
+On the other extreme, too short wavelengths are prevented by the energetic cost of accommodating droplets on too small patches, since the local contact angle (that increases with decreasing $\lambda$, see also Sec.~\ref{sec:SMdroplet}) would possibly exceed the largest substrate contact angle, such that the patterning itself would lose meaning.
+It has been, in fact, proven, both theoretically and experimentally, that for the confinement to be effective, the pattern characteristic size should not be much smaller than the spinodal length scale on the less wettable substrate~\cite{karguptaMorphologicalSelforganizationDewetting2002,karguptaInstabilityPatternFormation2000,nisatoExcitationSurfaceDeformation1999,karimPhaseSeparationUltrathin1998}.
 We keep, therefore, $\lambda \stackrel{>}{\sim} \lambda_s \approx 70 h_0$.
-The pattern speeds, on the other hand, are limited by the intrinsic LB bound of the lattice speed of sound
- \cite{succi}.\\
-The Reynolds number, $Re$, is of the order of $Re \sim 10^{-2}$ on the 
-static substrate and never exceeds the value $\approx 0.2$ in the time dependent case.
+The pattern speeds, on the other hand, are limited by the intrinsic LB bound of the lattice speed of sound~\cite{succiLatticeBoltzmannEquation2001}.\\
+The Reynolds number, $Re$, is of the order of $Re \sim 10^{-2}$ on the static substrate and never exceeds the value $\approx 0.2$ in the time dependent case.
 
-\section{Minkowski's structure metric $q_2$ in an extended parameter space}\label{sec:SMMinko}
+\section{Minkowski's structure metric \texorpdfstring{$q_2$}{second metric} in an extended parameter space}\label{sec:SMMinko}
 
@@
 \end{figure}
-\noindent In this section we test the robustness of the observation of the rivulet 
-state over a wider parameter space and in particular at changing: 1) the 
-pattern wavelength $\lambda$ and 2) the heterogeneity amplitude $\delta \theta$.
+\noindent In this section we test the robustness of the observation of the rivulet state over a wider parameter space and in particular at changing: 1) the pattern wavelength $\lambda$ and 2) the heterogeneity amplitude $\delta \theta$.
 To this aim we have run simulations with $\Gamma=1.5$, $\lambda/h_0 = 64, 128$, and with $\Gamma = 15$, $\lambda = 256 h_0$, $\delta \theta = 5^{\circ}$, respectively.
-In Fig.~\ref{fig:q2_difflambda} (left panel) we report the measurements of the Minkowski's structure metric $q_2$, whose 
-increase from zero up to a plateauing value of $q_2 \approx 1$ indicates the emergence of rivulets, for 
-the different $\lambda$'s ($\lambda=256 h_0$, which is the value considered in the main text, is also reported 
-for comparison).
-Analogously, in the right panel, we report $q_2$ as a function of time for  
-$\delta \theta = 5^{\circ}$  and $\delta \theta = 10^{\circ}$ (the value used in the main text). 
-We see that the $q_2$ metric attains the value $q_2 \approx 1$, signalling the rivulet state, albeit over 
-a time interval shorter than for $\delta \theta = 10^{\circ}$, i.e. the rivulets lifetime decreases with 
-$\delta \theta$. This was somehow expected, since obviously (and also in the static case) 
-the patterning looses effectiveness as the contact angle mismatch is reduced (see, e.g.~\cite{KonnurPRL2000}).
+In Fig.~\ref{fig:q2_difflambda} (left panel) we report the measurements of the Minkowski's structure metric $q_2$, whose increase from zero up to a plateauing value of $q_2 \approx 1$ indicates the emergence of rivulets, for the different $\lambda$'s ($\lambda=256 h_0$, which is the value considered in the main text, is also reported for comparison).
+Analogously, in the right panel, we report $q_2$ as a function of time for $\delta \theta = 5^{\circ}$  and $\delta \theta = 10^{\circ}$ (the value used in the main text). 
+We see that the $q_2$ metric attains the value $q_2 \approx 1$, signalling the rivulet state, albeit over a time interval shorter than for $\delta \theta = 10^{\circ}$, i.e.
+the rivulets' lifetime decreases with $\delta \theta$.
+This was somehow expected, since obviously (and also in the static case) the patterning looses effectiveness as the contact angle mismatch is reduced (see, e.g.~\cite{konnurInstabilityMorphologyThin2000}).
 %\newpage
@@
 
-\noindent We investigate, here, how the local contact angle of droplets, formed on the more
-hydrophilic patches after dewetting, depends on the pattern wavelength. We recall, in fact,
-that the patterning is such that the contact angle is not piecewise constant, but varies
-with continuity.
+\noindent We investigate, here, how the local contact angle of droplets, formed on the more hydrophilic patches after dewetting, depends on the pattern wavelength. 
+We recall, in fact, that the patterning is such that the contact angle is not piecewise constant, but varies with continuity.
 The droplet shape is determined by minimization of the total interfacial energy
@@
 \end{equation}
-where $A_{\text{lg}}$ and $A_{\text{sl}}$ are the liquid/gas and solid/liquid interface areas, and $\gamma_{\text{lg}}$, $\gamma_{\text{sl}}$,
-$\gamma_{\text{sg}}$ are the liquid/gas, solid/liquid and solid/gas interface energies 
-per unit area \cite{Wu2020}. 
+where $A_{\text{lg}}$ and $A_{\text{sl}}$ are the liquid/gas and solid/liquid interface areas, and $\gamma_{\text{lg}}$, $\gamma_{\text{sl}}$, $\gamma_{\text{sg}}$ are the liquid/gas, solid/liquid and solid/gas interface energies per unit area~\cite{wuHowChemicalPatterns2020}.
 In particular, $\gamma_{\text{lg}} \equiv \gamma$ is the surface tension.
-Setting $A_{\text{lg}} \equiv A$ and $A_{\text{sl}} \equiv S$, by Young's equation 
-$\cos \theta = \frac{\gamma_{\text{sg}} - \gamma_{\text{sl}}}{\gamma}$,
-Eq.~(\ref{eq:energy}) can be rewritten as
+Setting $A_{\text{lg}} \equiv A$ and $A_{\text{sl}} \equiv S$, by Young's equation $\cos \theta = \frac{\gamma_{\text{sg}} - \gamma_{\text{sl}}}{\gamma}$, Eq.~(\ref{eq:energy}) can be rewritten as
 \begin{equation}\label{eq:energyoung}
@@
 \end{equation}
-namely $(x_n,y_n) = \left((2n+1)\frac{\lambda}{4},(2n+3)\frac{\lambda}{4}\right)$, with
-$n=0,\pm 1, \pm 2,\dots$.
-If we consider large wavelengths ($q_{\theta} h_0 \ll 1$) and heterogeneity
-($\delta \theta \ll 1$) such that the contact angle gradients are small, the
-expression (\ref{eq:contact}) can be expanded as
+namely $(x_n,y_n) = \left((2n+1)\frac{\lambda}{4},(2n+3)\frac{\lambda}{4}\right)$, with $n=0,\pm 1, \pm 2,\dots$.
+If we consider large wavelengths ($q_{\theta} h_0 \ll 1$) and heterogeneity ($\delta \theta \ll 1$) such that the contact angle gradients are small, the expression (\ref{eq:contact}) can be expanded as
 \begin{equation}\label{eq:contact2}
@@
 where $\theta_m = \theta_0 - \delta \theta$.
-This local radial symmetry allows to approximate the equilibrium droplet shape as a spherical cap of
-height $h$ and base radius $a$, whose area is $A = \pi(a^2 + h^2)$; inserting the expression for
-$A$ and (\ref{eq:contact2}), neglecting higher than second order terms, in (\ref{eq:energyoung}) gives
+This local radial symmetry allows to approximate the equilibrium droplet shape as a spherical cap of height $h$ and base radius $a$, whose area is $A = \pi(a^2 + h^2)$; inserting the expression for $A$ and (\ref{eq:contact2}), neglecting higher than second order terms, in (\ref{eq:energyoung}) gives
 \begin{equation}\label{eq:energy3}
@@
 where
-$\mathcal{C}_a(\mathbf{x}_n)=\{(x,y) \in [0, L]^2|(x-x_n)^2 + (y-y_n)^2 \leq a^2\}$
-is the circle of centre $\mathbf{x}_n$ and radius $a$. 
-Due to global volume conservation and assuming the droplets to be monodisperse, the droplet
-volume is $V_d = h_0L^2/N_d = h_0 \lambda^2/2$,
-where $N_d$ is the number of droplets, which equals the number of
-minima of (\ref{eq:contact}) in the domain $[0, L]^2$, i.e. $N_d = 2(L/\lambda)^2$.
+$\mathcal{C}_a(\mathbf{x}_n)=\{(x,y) \in [0, L]^2|(x-x_n)^2 + (y-y_n)^2 \leq a^2\}$ is the circle of centre $\mathbf{x}_n$ and radius $a$. 
+Due to global volume conservation and assuming the droplets to be monodisperse, the droplet volume is $V_d = h_0L^2/N_d = h_0 \lambda^2/2$, where $N_d$ is the number of droplets, which equals the number of minima of (\ref{eq:contact}) in the domain $[0, L]^2$, i.e.
+$N_d = 2(L/\lambda)^2$.
 Enforcing the volume of the spherical cap to be equal to $V_d$ relates $h$ and $a$ by
@@
 \end{equation}  
-Inserting the latter expression in (\ref{eq:energy3}) and performing the integral, the
-energy (that we indicate now as $E(a)$ to lighten the notation) reads
+Inserting the latter expression in (\ref{eq:energy3}) and performing the integral, the energy (that we indicate now as $E(a)$ to lighten the notation) reads
 \begin{equation}\label{eq:energy4}
@@
 \end{equation}  
-We expand, then, the sine in the second term up to second order in $\delta \theta$ (such that the
-energy is first order) and we finally get
+We expand, then, the sine in the second term up to second order in $\delta \theta$ (such that the energy is first order) and we finally get
 \begin{equation}\label{eq:energyfin}
@@
 \end{equation}
-i.e., it does not depend on $\lambda$. We move, then, to the next order. We seek a solution to
-(\ref{eq:minim}) in the form $a = a^{(0)} + \delta \theta a^{(1)} + \dots$, where $a^{(0)} \equiv a_0$.
+i.e., it does not depend on $\lambda$. 
+We move, then, to the next order. 
+We seek a solution to (\ref{eq:minim}) in the form $a = a^{(0)} + \delta \theta a^{(1)} + \dots$, where $a^{(0)} \equiv a_0$.
 At the first order in $\delta \theta$ we get
@@
 \end{equation}  
-where $b = \left(\frac{32 h_0^2 \pi^4}{27 \theta_m^5}\right)^{1/3}$,
-i.e. the droplet contact angle grows at decreasing pattern wavelength as $\theta_d \sim \lambda^{-2/3}$.
+where $b = \left(\frac{32 h_0^2 \pi^4}{27 \theta_m^5}\right)^{1/3}$, i.e. the droplet contact angle grows at decreasing pattern wavelength as $\theta_d \sim \lambda^{-2/3}$.
 
-\bibliography{Ref}
+\bibliography{betterRef}
 
@@
 
-%\bibitem{KonnurPRL2000} R. Konnur, K. Kargupta and A. Sharma, {\it Phys. Rev. Lett.} {\bf %84}, 931--934 (2000).
+%\bibitem{konnurInstabilityMorphologyThin2000} R. Konnur, K. Kargupta and A. Sharma, {\it Phys. Rev. Lett.} {\bf %84}, 931--934 (2000).
 %\bibitem{WuSM2020} Y. Wu, F. Wang, S. Ma, M. Selzer and B. Nestler,

### 2024-02-15 main_exciting.tex
Resubmission PRFluids

@@
 Several realizations of switchable and adaptive substrates have been proposed~\cite{xinReversiblySwitchableWettability2010}, involving smart materials such as polymer brushes~\cite{stuartEmergingApplicationsStimuliresponsive2010}, thermal-responsive hydrogels~\cite{chenThermalresponsiveHydrogelSurface2010}, light-responsive molecules and microstructures~\cite{ichimuraLightDrivenMotionLiquids2000} or processes such as electrowetting~\cite{mugeleElectrowettingConvenientWay2005}.
-\REV{Modelling the coupled problem of a thin film on a switchable substrate and as such the dynamics of the substarte as well as the film is far from being trivial but can be realized by a space and time dependent wettability pattern~\cite{grawitterSteeringDropletsSubstrates2021}}.
+\REV{Modelling the coupled problem of a thin film on a switchable substrate and as such the dynamics of the substrate as well as the film is far from being trivial but can be realized by a space and time dependent wettability pattern~\cite{grawitterSteeringDropletsSubstrates2021}}.
 While a consistent body of theoretical/computational work was devoted to processes on static heterogeneous substrates, the time dependent case is still almost unexplored, with few exceptions focusing on single droplet spreading and sliding~\cite{grawitterSteeringDropletsSubstrates2021,grawitterDropletsSubstratesOscillating2021,thieleGradientDynamicsModel2020} or limited to analysing the linear regime~\cite{sumanDynamicsThinLiquid2006}.
 
-Here, we study by means of numerical simulations the full dewetting dynamics of a thin liquid film deposited on a substrate with a time varying wettability pattern, starting from \REV{a slightly perturbed film to film} rupture \REV{and} long time morphology.
+Here, we study by means of numerical simulations the full dewetting dynamics of a thin liquid film deposited on a substrate with a time varying wettability pattern, starting from \REV{a slightly perturbed film to film} rupture \REV{and} long time morphology~\cite{konnurInstabilityMorphologyThin2000}.
 We identify two regimes where the rupture times grow with the pattern wavelength either linearly (on a static pattern) or attain a constant value (in the time dependent case), for short wavelengths, and approach a quadratic law as the wavelength increases.
@@
 \\
-\noindent {\it Method.} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{oronLongscaleEvolutionThin1997,crasterDynamicsStabilityThin2009}
+\noindent {\it Method.} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{oronLongscaleEvolutionThin1997, crasterDynamicsStabilityThin2009}
 \begin{equation}\label{eq:thinfilm}
@@
 \end{equation}
-by means of a recently developed lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,zitzLatticeBoltzmannSimulations2021, zitzSwalbeJlLattice2022a}.
+by means of a recently developed lattice Boltzmann (LB) scheme~\cite{zitzLatticeBoltzmannMethod2019,zitzLatticeBoltzmannSimulations2021, zitzSwalbeJlLattice2022a}.
 Eq.~(\ref{eq:thinfilm}) describes, in a lubrication approximation spirit, the evolution of the height field (film thickness) $h(\mathbf{x},t)$, denoting the location of the liquid/air interface. 
-The mobility function
-$M_{\delta}(h) = \frac{2h^3 + 6\,\delta\, h^2 + 3\delta^2h}{6\mu}$
-depends on the velocity boundary condition at the substrate, parameterized by an effective slip length $\delta$
-(for $\delta \rightarrow 0$ it reduces to the no-slip form $h^3/(3\mu)$). Here, $\mu$ is the fluid dynamic viscosity.
-The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$ ($\gamma$ is the surface tension)
-The disjoining pressure $\Pi$ can be seen as (minus) the derivative, with respect to the film thickness, of an effective interfacial potential.
-As such, it contains the information on the liquid/solid and solid/gas interactions and, hence, on the wettability, which is parameterized in terms of the contact angle $\theta$~\cite{bonnWettingSpreading2009, schwartzSimulationDropletMotion1998}.
+The mobility function $M_{\delta}(h) = \frac{2h^3 + 6\,\delta\, h^2 + 3\delta^2h}{6\mu}$ depends on the velocity boundary condition at the substrate, parameterized by an effective slip length $\delta$ (for $\delta \rightarrow 0$ it reduces to the no-slip form $h^3/(3\mu)$)~\cite{huhHydrodynamicModelSteady1971, peschkaSignaturesSlipDewetting2019, fetzerQuantifyingHydrodynamicSlip2007, munchLubricationModelsSmall2005a}.
+Here, $\mu$ is the fluid dynamic viscosity.
+The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$ ($\gamma$ is the surface tension~\cite{wuHowChemicalPatterns2020}).
+The disjoining pressure $\Pi$ can be seen as (minus) the derivative, with respect to the film thickness, of an effective interfacial potential~\cite{Deryaguin1940, DeryaguinChuraev1978}.
+As such, it contains the information on the liquid/solid and solid/gas interactions and, hence, on the wettability, which is parameterized in terms of the contact angle $\theta$~\cite{bonnWettingSpreading2009, schwartzSimulationDropletMotion1998, youngIIIEssayCohesion1805, degennesWettingStaticsDynamics1985}.
 The expression adopted for $\Pi$ is
@@
 \end{equation}
-where $f(\xi)=\xi^{-3} - \xi^{-2}$. $h_{\ast}$ is the height at which $\Pi$ vanishes and sets the precursor layer thickness~\cite{SuppMat}.
+where $f(\xi)=\xi^{-3} - \xi^{-2}$.
+$h_{\ast}$ is the height at which $\Pi$ vanishes and sets the precursor layer thickness~\cite{SuppMat, schwartzSimulationDropletMotion1998, mitlinDewettingSolidSurface1993, teletzkeHowLiquidsSpread1987}.
 The time variation of the patterned substrate enters the model through the disjoining pressure, by making the contact angle space and time dependent, i.e.~$\theta = \theta(\mathbf{x},t)$.
 We decorate the substrate with a checkerboard pattern, a common choice that generalizes the broken homogeneity of the stripes to two 
-directions~\cite{jalaliFabricationCharacterizationScalable2018,nagayamaIntermediateWettingState2020,dasSurfacedirectedSpinodalDecomposition2020}.
+directions~\cite{jalaliFabricationCharacterizationScalable2018, nagayamaIntermediateWettingState2020, dasSurfacedirectedSpinodalDecomposition2020}.
 In particular, we employ the sinusoidal form
@@
 Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film~\cite{meckeThermalFluctuationsThin2005}.
-On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads $(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$~\cite{meckeThermalFluctuationsThin2005,zhangMolecularSimulationThin2019}.
+On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads $(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$~\cite{meckeThermalFluctuationsThin2005, zhangMolecularSimulationThin2019}.
 In our patterned case, we define 
@@
 
-\noindent {\it Results.} We first investigate how the rupture times depend on the parameters characterizing the wettability pattern, namely the wavelength of the contact angle variation, $\lambda$, and wave speed, $v_{\theta}$.
+\noindent {\it Results.} We first investigate how the rupture times depend on the parameters characterizing the wettability pattern, namely the wavelength of the contact angle variation, $\lambda$, and wave speed, $v_{\theta}$~\cite{karguptaMorphologicalSelforganizationDewetting2002,karguptaInstabilityPatternFormation2000,nisatoExcitationSurfaceDeformation1999,karimPhaseSeparationUltrathin1998, succiLatticeBoltzmannEquation2001}.
 The film rupture time, $\tau_r$, is defined as the least $t$ such that $h(\mathbf{x},\tau_r)=h_{\ast}$ (i.e., when the free surface "touches" the substrate).
@@
 Conversely, for fast growths ($t_{\theta} \ll t_R$), retraction dominates and fixes the time scale, $\tau_r \sim t_R$. 
-The latter is related to the time the liquid takes to flow out of regions of high contact angle, whose size is $\sim \lambda$. Hence we have 
+The latter is related to the time the liquid takes to flow out of regions of high contact angle, whose size is $\sim \lambda$. 
+Hence we have 
 \begin{equation}\label{eq:taur_l1}
